### 2024-06-04 preregistration/README.qmd
intermediate version of preregistration

---
title: ""
format: 
  html:
    template: ../assets/template.html
    css: ../assets/style.css
bibliography: ../manuscript/references.bib
csl: ../manuscript/apa.csl
link-citations: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(lmerTest)
library(flextable)
library(ftExtra)

load("../analysis_objects/dag_localtests.RData")

# set up flextable for tables
set_flextable_defaults(
  font.family = "Times", 
  font.size = 10,
  font.color = "black",
  line_spacing = 1,
  padding.bottom = 1, 
  padding.top = 1,
  padding.left = 1,
  padding.right = 1
)
```

*Last updated `r format(Sys.time(), "on %A, %B %d, %Y at %I:%M %p")`*

## Overview

The independent measures used in this study are drawn from studies that were previously conducted in the Longitudinal Internet studies for the Social Sciences (LISS) panel. These data are openly available after signing an agreement with LISS. The dependent measures will be part of a new data collection in the LISS panel which is scheduled between May 2024 and July 2024.

As this study partly uses secondary data, this preregistration document is based on the template by @akker_2021. Where necessary, we make clear which decisions relate to/are based on secondary data, and which decisions relate to/are based on newly collected data.

There are six parts, which you can jump to following the links below:

-   [Part 1 - Study information]
-   [Part 2 - Data Description]
-   [Part 3 - Variables]
-   [Part 4 - Knowledge of Data]
-   [Part 5 - Analyses]
-   [Part 6 - Statement of Integrity]

## Part 1 - Study Information

### Q1: Preliminary title DONE

"Attention shifting and inhibition abilities among people from adverse conditions"

### Q2: Authors DONE

-   Stefan Vermeent^1,2^
-   Anna-Lena Schubert^3^
-   Jean-Louis van Gelder^2^
-   Willem E. Frankenhuis^2,4^

^1^Utrecht University, Utrecht, The Netherlands

^2^Max Planck Institute for the Study of Crime, Security and Law, Freiburg, Germany

^3^University of Mainz, Mainz, Germany

^3^University of Amsterdam, Amsterdam, The Netherlands

### Q3: Research Questions DONE

**R1:** How is adversity associated with general speed of processing across all cognitive tasks?

**R2:** How is adversity associated with inhibition and attention shifting abilities after accounting for general speed of processing?

**R3:** How is adversity associated with task-general and task-specific response caution?

### Q4: Hypotheses DONE

Data patterns that would be consistent with hypotheses about abilities generated from a **deficit framework**:

A1. Adversity is negatively associated with general speed of processing.

A2. Adversity is negatively associated with either or both latent estimates of inhibition and attention shifting, after accounting for general speed of processing.

A3. If latent estimates of inhibition and attention shifting are highly correlated, adversity is negatively associated with a general EF factor spanning both.

Data patterns that would be consistent with hypotheses about abilities generated from **adaptation frameworks**:

B1. Adversity is positively associated with the latent estimate of attention shifting *regardless of the association with inhibition*, after accounting for general speed of processing. A positive association with inhibition would be unexpected yet interesting, and would require revising existing adaptation theoretical frameworks.

B2. The association between adversity and the latent estimate of attention shifting is practically equivalent to zero *and* the association between adversity and the latent estimate of inhibition is negative.

Finally, we have the following hypothesis about response caution:

C1. Adversity is associated with more task-general response caution.

## Part 2 - Data Description

### Q5: Dataset DONE

Our study will include 1000 participants who were randomly sampled from the LISS panel (Scherpenzeel, 2011). The LISS panel is a representative probability sample of roughly 5,000 Dutch households (\~7,500 individuals) drawn from the population register by Statistics Netherlands on an invite-only basis. Households without a computer or internet connection are provided with these facilities by LISS. Each year, participants complete the same core battery of questionnaires about---among other topics---their financial situation in the past year. In addition, participants can complete additional online questionnaires every month, with variable content. The current study integrates two data sources. First, our sample of 1000 participants will participate in a new LISS study between May 2024 and July 2024 (hereafter referred to as 'newly collected data'), in which we will include a measure of neighborhood threat, three cognitive tasks measuring attention shifting, two cognitive tasks measuring inhibition, and one cognitive task measuring processing speed. Second, we will access data that were previously collected in LISS (hereafter referred to as 'the LISS archive').

### Q6: Public Availability DONE

All datasets are available in the LISS data archive. Researchers who want to access the data are required to sign a statement confirming that information about individual persons, households, etc., will not be released to others (go to <https://statements.centerdata.nl> for more information).

### Q7: Data Access DONE

Data can be accessed through the following links.

-   [Convential and Computer Crime Victimization study](https://www.dataarchive.lissdata.nl/study-units/view/25), doi: https://doi.org/10.17026/dans-zch-j8xt
-   [Economic Situation: Income](https://www.dataarchive.lissdata.nl/study-units/view/35), doi: https://doi.org/10.17026/dans-24y-dkqk
-   [Background variables](https://www.dataarchive.lissdata.nl/study-units/view/322), doi: https://doi.org/10.57990/qn3k-as78

After the new data have been collected, they will become available in the general overview at <https://www.dataarchive.lissdata.nl/study-units/view/1>

### Q8: Date of Download DONE

-   Stefan Vermeent (lead author and data analyst)
    -   For a full overview of when specific independent variables in the LISS data archive were previously accessed, see the 'Overview of project milestones' section at <https://github.com/StefanVermeent/liss_ef_2024>.
-   Anna-Lena Schubert will not access the data.
-   Jean-Louis van Gelder will not access the data.
-   Willem E. Frankenhuis will not access the data.

### Q9: Data Collection DONE

General information about recruitment of participants into the LISS study can be found at <https://www.lissdata.nl/methodology>.

Our study will include 1,000 participants who were randomly sampled from the LISS panel. The requirements will be that they are between 20 and 55 years old and have completed at least 1 wave of the Conventional and Computer Crime Victimization study (see Q10).

### Q10: Codebooks DONE

Detailed codebooks for existing data in the LISS archive can be downloaded at the following links:

-   [Convential and Computer Crime Victimization study](https://www.dataarchive.lissdata.nl/study-units/view/25), doi: https://doi.org/10.17026/dans-zch-j8xt
-   [Economic Situation: Income](https://www.dataarchive.lissdata.nl/study-units/view/35), doi: https://doi.org/10.17026/dans-24y-dkqk
-   [Background variables](https://www.dataarchive.lissdata.nl/study-units/view/322), doi: https://doi.org/10.57990/qn3k-as78

Codebooks for the newly collected data will become available later at <https://github.com/StefanVermeent/liss_ef_2024>

## Part 3 - Variables

### Q11: Manipulated Variables: DONE

**Not applicable**

### Q12: Measured Variables

#### Potential confounders

We identified potential confounders of the empirical estimands (i.e., the association between self-reported threat and deprivation with cognitive outcomes) using a Directed Acyclic Graph (DAG).
We analyzed the following variables both conceptually (i.e., based on previous literature and/or our own assumptions) and, in cases of doubt, statistically: (1) age, (2) sex, (3) past adversity exposure (material deprivation and threat combined), (4) recent threat exposure, and (5) recent material deprivation.
Statistical analysis of potential confounders was based on data from a previous LISS study (see Q17 and Q18) using the `dagitty` R package [@textor_2016].
Specifically, we used the `dagitty::localTests()` function to establish whether DAG-implied conditional independencies between certain key variables were supported by the data [@ankan_2021].
The outcome in these conceptual DAGs was working memory capacity, measured on a latent level in the previous LISS study.
Code for these analyses can be found under `scripts/00_dag`.

Figure 1A shows the initially conceptualized DAG. 
Black lines depict our estimands, and red lines depict potential confounding paths.
Solid lines depict paths that we were reasonably sure about given previous literature.
Dashed lines depict paths that we were less sure about, and therefore decided to statistically test as outlined above.

(1) **Age.** *Conceptual reasoning*: Age has a causal effect on processing speed [@salthouse_2016; @salthouse_2019] and on response caution [@starns_2010]. Older adults may on average be more financially stable.

We know of no evidence to consider age as a precedent for threat exposure, especially because we measure recent adversity and not cumulative lifetime adversity (which is likely related to age as older individuals on average will have more cumulative exposure to adverse events).

(2) **Sex.**

(4) **Childhood adversity.** *Conceptual reasoning*:

childhood adversity on education: @fry_2018
Education on poverty: @hofmarcher_2021

```{r}
#| label: Figure1
#| fig-width: 6.25
#| fig-height: 7
#| echo: false
#| dpi: 600
#| out-width: 6in
#| fig-cap: |
#|   **Figure 1.** Directed Acyclic Graph (DAG) depicting the empirical estimands (black lines) and presumed confounding effects (red lines). Age, education level, and childhood adversity are all conceptualized as confounding variables, as they are expected to 
#|   Importantly, we consider deprivation to be a confounder in the effect of threat on cognitive ability, and conversely, we consider threat to be a mediator of the effect of deprivation on cognitive ability [@ning_2023]. From this, it follows that deprivation should be adjusted for when estimating the effect of threat on cognitive ability. However, threat should *not* be adjusted for when estimating the effect of deprivation on cognitive ability.

knitr::include_graphics("figures/fig1.png")
```

```{r}
#| echo: false
table1 <- 
  tibble(
    `(Potential) confounder` = c(
      "Age",
      "Education",
      "Sex",
      "Childhood adversity",
      "Recent material deprivation",
      "Recent threat"
    ),
    `Causal effect on cognitive performance` = c(
      "@salthouse_2016; \n @salthouse_2019; \n @starns_2010",
      "@lovden_2020",
      "@ning_2023", 
      "@bos_2009; \n @goodman_2019",
      "*Estimand of interest*",
      "*Estimand of interest*"
    ),
    `Causal effect on recent threat` = c(
      "Unclear",
      "Unclear",
      "@ning_2023", 
      "@bywaters_2016",
      "@bywaters_2016; @ning_2023; ",
      ""
    ),
    `Causal effect on recent material deprivation` = c(
      "Older adults may have more financial resources",
      "@fry_2018",
      "@ning_2023", 
      "@bywaters_2016",
      "",
      ""
    ),
    `Statistical (non-)support` = c(
      paste0("Conditional on childhood adversity and recent deprivation, \n *r* = ", localTest_dag03 |> pull(estimate), " SE = ", localTest_dag03 |> pull(std.error), ", 95% CI = ", localTest_dag03 |> pull(ci), "."),

      paste0("Conditional on childhood adversity, recent deprivation, and age, \n *r* = ", localTest_dag01 |> pull(estimate), " SE = ", localTest_dag01 |> pull(std.error), ", 95% CI = ", localTest_dag01 |> pull(ci), "."),
      "", 
      "",
      paste0("conditional on childhood adversity and age, deprivation is a significant confounder of threat \n *r* = ", localTest_dag04 |> pull(estimate), " SE = ", localTest_dag04 |> pull(std.error), ", 95% CI = ", localTest_dag04 |> pull(ci), "."),
      ""
    ),
    `Confounder in threat model` = c(
      "Yes",
      "No",
      "Yes", 
      "Yes",
      "Yes",
      ""
    ),
    `Confounder in deprivation model` = c(
      "Yes",
      "Yes",
      "Yes", 
      "Yes",
      "",
      "No, is a mediator"
    )
  ) |> 
  flextable() |> 
  ftExtra::colformat_md()

table1
```

#### Independent Variables

Our plan for IV computation is identical to that in a previous project using LISS data (see Q17 and Q18). While we know how the different subscales correlated with each other in that project, the current project will consist of a (partially) different subsample of the LISS panel. While we expect that IV computation will be identical to that project, we leave open the possibility that correlations could fall out differently, thus requiring different approaches.

*Perceived resource scarcity (Mean).* This is a measure of perceived deprivation. It will consist of

*Perceived resource scarcity (Coefficient of variation).* This is a measure of unpredictability

*Income-to-needs ratio (Mean).* This is a measure of objective deprivation.

*Income-to-needs ration (Coefficient of variation).* This is a measure of unpredictability

*Neighborhood threat.* This is a measure of neighborhood threat.

#### Dependent Variables DONE

The exact nature and number of dependent variables will depend on the final latent model that will be selected (see Part 5, Q19). As a rule, we will estimate associations between each measure of adversity and each latent factor that is estimated.

### Q14: Missing data DONE

We will use Full Information Maximum Likelihood (FIML) to handle missing data.

### Q15: Outliers

For each cognitive task, we will remove trials with: (1) Response times (RTs) \< 250 ms; (2) RTs \> 3 SD above the participant-level average log-transformed mean RT, separately for different task conditions (e.g., congruent and incongruent). 
Finally, if a participant performs at chance level on a task, we will exclude task data of that particular task only. 
Chance level will be determined using the accuracy rate at the 97.5% tail of a binomial distribution if a participant would be purely guessing.

### Q16: Sample Weights DONE

**Not applicable**

## Part 4 - Knowledge of Data

### Q17: Relevant Publications DONE

All authors have previously used the same independent adversity measures to measure their associations with working memory performance. The Registered Report associated with this project has received Stage 1 In-Principle Acceptance at *Peer Community in Registered Reports* (<https://osf.io/dp7wc>). The sample in this study consisted of \~800 people from the LISS panel. An unknown portion of these people is also expected to participate in the current study.

### Q18: Prior Knowledge

In a previous study (mentioned in Q17) we used the same independent adversity measures. From this study, we know that the perceived scarcity measures correlated \> .60 with each other, allowing computation of uniformly weighted averages. We also know that correlations among the three threat measures, a well as the perceived scarcity coefficients of variation, were low, thus requiring PCA. Although the current study will consist of a (partially) different sample, we expect to reach the same conclusions here.

From this study, we know that none of the adversity measures was significantly associated with working memory capacity or working memory updating. The inhibition and attention shifting data have not yet been collected at the time of writing this preregistration. Therefore, we do not know how the adversity measures are associated with inhibition and attention shifting.

## Part 5 - Analyses

### Q19: Hypotheses -\> Statistical Tests DONE

**Step 1: DDM estimation.** We will fit all DDM models as hierarchical Bayesian models in JAGS using the `runjags` package [@denwood_2016].

For the Posner Task, we will estimate a single drift rate, non-decision time, and boundary separation across all trials:

```{r}
#| eval: false
"model {
  #likelihood function
  for (t in 1:nTrials) {
    y[t] ~ dwiener(alpha[subject[t]], 
                   tau[subject[t]], 
                   0.5, 
                   delta[subject[t]])
  }

  for (s in 1:nSubjects) {
    tau[s]  ~ dnorm(muTau, precTau) T(.0001, 1)
    delta[s] ~ dnorm(muDelta, precDelta) T(-10, 10)
    alpha[s]  ~ dnorm(muAlpha, precAlpha) T(.1, 5)
  }

  #priors
  muTau ~ dunif(.0001, 1)
  muDelta ~ dunif(-10, 10)
  muAlpha~ dunif(.1, 5)

  precAlpha  ~ dgamma(.001, .001)
  precTau ~ dgamma(.001, .001)
  precDelta ~ dgamma(.001, .001)
}"
```

For all other tasks, we will estimate drift rate and non-decision time separately for each task condition (e.g., incongruent and congruent), while keeping boundary separation equal across conditions:

```{r}
#| eval: false
"model {
  #likelihood function
  for (t in 1:nTrials) {
    y[t] ~ dwiener(alpha[subject[t]], 
                   tau[condition[t], subject[t]], 
                   0.5, 
                   delta[condition[t], subject[t]])
  }
  for (s in 1:nSubjects) {
    for (c in 1:nCon) {
      tau[c, s]  ~ dnorm(muTau[c], precTau) T(.0001, 1)
      delta[c, s] ~ dnorm(muDelta[c] , precDelta) T(-10, 10)
    }
    alpha[s]  ~ dnorm(muAlpha, precAlpha) T(.1, 5)
  }
  #priors
  for (c in 1:nCon){ 
    muTau[c] ~ dunif(.0001, 1)
    muDelta[c] ~ dunif(-10, 10)
  } 
  muAlpha~ dunif(.1, 5)

  precAlpha  ~ dgamma(.001, .001)
  precTau ~ dgamma(.001, .001)
  precDelta ~ dgamma(.001, .001)
}"
```

Each model will be fitted using three Markov Chain Monte Carlo (MCMC) chains. Each chain will start with 2,000 burn-in samples, followed by 10,000 additional samples. Of these 10,000 samples, every 10th sample will be retained to limit autocorrelation, resulting in a total of 3,000 posterior samples across all chains.

Model fit will be assessed in three ways. First, we will visually inspect the traces of the group-level parameter estimates, which should not contain large jumps or drifts. Second, we will calculate the Gelman-Rubin statistic (R\^), which should be lower than 1.1 [@gelman_1992]. Third, we will conduct simulation-based parameter recovery. From each participant's DDM parameter estimates, we will simulate 500 RTs and corresponding accuracies. We will calculate Pearson correlations between simulated and observed RTs (at the 25th, 50th, and 75th percentile) as well as simulated and observed accuracy rates. RT correlations need to be .80 or higher to indicate good fit. For accuracy rates, we will use the same cut-off, unless there are ceiling effects (\> 95%), which is likely for congruent/repeat trials. Thus, if accuracy for one condition shows poor recovery, but accuracy for the other condition and all RT quantiles are recovered well, we will still consider this good model fit.

**Step 2: SEM.** We will construct the final SEM sequentially. First, we will optimize the fit of the drift rate, boundary separation, and non-decision time sub-models. Second, we will combine these three models into a single measurement model. Third, we will add the regression paths between measures of adversity and the latent factors.

**Drift rate sub-model.** The baseline drift rate model will be a bi-factor model which accounts for general processing speed. A latent processing speed factor will load on all the manifest drift rates. A latent inhibition factor will load on the manifest drift rates of incongruent trials of the Flanker and Simon Task. A latent Attention Shifting Task will load on the manifest drift rates of switch trials of the Color-shape, Global-local, and Animacy-size Task.

```{r}
#| label: Figure2
#| fig-width: 6.25
#| fig-height: 7
#| echo: false
#| dpi: 600
#| out-width: 6in
#| fig-cap: |
#|   **Figure 2.** Overview of the drift rate models. Blue ovals represent latent factors that will be use as dependent variables. Rectangles represent manifest variables. Bidirectional arrows represent (co-)variances. Unidirectional arrows represent factor loadings. Covariances in red are constrained to 1. If a constrained version of the model provides better fit, the final model will collapse the latent factors into a single factor.

knitr::include_graphics("figures/fig2.png")
```

<br>

We will first fit the model as specified in Figure 2A (see `mod_v1a` in the code block below). Then, we will fit a second version of this model in which a single common EF factor is estimated, covering both inhibition and attention shifting tasks. 
Using `lavaan::anova`, we will then test whether the second, more constrained model leads to worse model fit, qualified as a significant chi squared change test, and an AIC value > 10. 
If so, we will include submodel 2 in our final model.

```{r}
#| eval: false

# Drift rate model 1 (see Figure 1A)

mod_v1a <- '
  # Latent factors
  ps   =~ 1*pos_v + flanker_inc_v + flanker_con_v + simon_inc_v + simon_con_v + cs_sw_v + cs_rep_v + gl_sw_v + gl_rep_v + as_sw_v + as_rep_v
  inh  =~ 1*flanker_inc_v + simon_inc_v
  as =~ 1*cs_sw_v + gl_sw_v + as_sw_v

  # Covariances
  inh  ~~ as + 0*ps
  as ~~ 0*ps
'

fit_v1a <- lavaan::sem(
  model = mod_v1a, 
  data = data
)

# Drift rate model 1, constrained version (see Figure 1B)

mod_v1b <- '
  # Latent factors
  ps_v =~ 1*pos_v + flanker_inc_v + flanker_con_v + simon_inc_v + simon_con_v + cs_sw_v + cs_rep_v + gl_sw_v + gl_rep_v + ansi_sw_v + ansi_rep_v
  ef_v =~ 1*flanker_inc_v + simon_inc_v + cs_sw_v + gl_sw_v + ansi_sw_v

 # Covariances
 ps_v ~~ 0*ef_v
'

fit_v1b <- lavaan::sem(
  model = mod_v1b, 
  data = data
)

lavaan::anova(fit_v1a, fit_v1b)
```

**Boundary separation sub-model.** Like the drift rate model, the baseline boundary separation model will be a bi-factor model containing a general boundary separation factor, as well as inhibition- and attention shifting-specific factors (see Figure 3A). We will use the same procedure to choose between the baseline model, and a second model estimating a common boundary separation factor.

```{r}
#| label: Figure3
#| fig-width: 6.25
#| fig-height: 7
#| dpi: 600
#| echo: false
#| out-width: 6in
#| fig-cap: |
#|   **Figure 3.** Overview of the boundary separation models. Blue ovals represent latent factors that will be use as dependent variables. Rectangles represent manifest variables. Bidirectional arrows represent (co-)variances. Unidirectional arrows represent factor loadings. Covariances in red are constrained to 1. If a constrained version of the model provides better fit, the final model will collapse the latent factors into a single factor.

knitr::include_graphics("figures/fig3.png")
```

**Non-decision time sub-model.** The baseline model for the non-decision time, as well as the model selectin procedure, will be identical to the drift rate and boundary separation models (see Figure 3A).

```{r}
#| label: Figure4
#| fig-width: 6.25
#| fig-height: 7
#| dpi: 600
#| echo: false
#| out-width: 6in
#| fig-cap: |
#|   **Figure 4.** Overview of the drift rate models. Blue ovals represent latent factors that will be use as dependent variables. Rectangles represent manifest variables. Bidirectional arrows represent (co-)variances. Unidirectional arrows represent factor loadings. Covariances in red are constrained to 1. If a constrained version of the model provides better fit, the final model will collapse the latent factors into a single factor.

knitr::include_graphics("figures/fig4.png")
```

<br>

**Full measurement model.** After determining the final sub-model for all three DDM parameters, we will combine them all into a single model. 
We expect that we will have to estimate covariances between latent factors of different DDM parameters that share at least one manifest variable (e.g., inhibition drift rates and general boundary separation). 
However, we foresee that we might have to add or remove covariances to improve model fit.
These steps will be taken before accessing and adding the independent variables.

**Final model.** After finalizing the full measurement model, we will add the adversity measures as predictors to the model. 
Each adversity measure will be regressed on each latent factor. 
We will estimate covariances between each pair of adversity measures.

### Q20: Predicted effect sizes DONE

We do not have specific predictions for effect sizes but we deem standardized regression coefficients for interaction effects = .10 (or higher) and -.10 (or lower) meaningful. These effects would indicate that an effect of adversity on a a particular aspect of performance score is meaningfully more negative or more positive.

Effects between -.10 and .10 are not of interest for determining differences between adversity and aspects of performance. However, we are interested in determining if simple effects falling between -.10 and .10 are consistent with an effect of 0. For this scenario, we will use Two One-Sided T-tests (TOST) equivalence testing with -.10 and .10 as bounds.

### Q21: Statistical Power DONE

A power analysis based on @Kretzschmar_2019 indicated that with an alpha of α = 0.05, we will have \> 90% power to detect small effect sizes (β = 0.1) with a sample size ranging between N = 730 and N = 980 when the reliability of the measures is at least moderate (0.6 -- 0.7). Therefore, we aim for a total sample size of N = 1,000.

### Q22: Inferential Criteria DONE

We will use an alpha level of .05 throughout. p-values will be corrected for multiple testing using the false discovery rate. We will do so separately for tests involving drift rates, non-decision times, and boundary separations, as we have different hypotheses for each of these parameters.

We will test for three types of associations between different types of adversity and cognitive ability (as indexed by the drift rates): enhanced ability, impaired ability, or intact ability. Enhanced cognitive ability is defined as a *positive* association between adversity and drift rate. Impaired cognitive ability is defined as a *negative* association between adversity and drift rate. intact ability is defined as an association that is practically equivalent to zero. We define practical equivalence as a standardized effect falling between -0.10 and 0.10. We will use two one-sided tests (TOST) equivalence testing against these bounds to test for practical equivalence.

### Q23: Assumption Violations/Model Non-Convergence DONE

In case of non-normally distributed variables, we will use

If any assumptions are violated, we will update the preregistration and/or report deviations from the preregistration. However, we do not anticipate any serious violations.

### Q24: Reliability and Robustness Testing

NA

### Q25: Exploratory Analyses DONE

We do not preregister any exploratory analyses.

## Part 6 - Statement of Integrity

We state that we filled out this preregistration to the best of our knowledge and that no other preregistration exists pertaining to the same hypotheses and dataset.

## References

<div id="refs"></div>

### 2024-06-04 materials/liss/I_flanker/2_flanker_procedure.js
Task materials

@@ -81,15 +81,3 @@ var flanker_test_procedure02 = {
   randomize_order: true,
   repetitions: 4,
 };
-
-
-var flanker_pp_feedback = {
-  type: jsPsychSurveyText,
-  questions: [
-    {prompt: 'Heeft u feedback op dit spel (bijvoorbeeld: duidelijkheid van de instructies; moeilijkheidsgraad; iets wat niet goed leek te werken)?<br>We zouden het op prijs stellen als u het ons hieronder laat weten!', name: 'user_feedback', rows:5, required: false},
-  ],
-  data: {
-    variable: "pp_feedback",
-    task: "flanker_feedback"
-  }
-}

### 2024-06-04 materials/liss/I_simon/2_simon_procedure.js

@@ -142,13 +142,3 @@ var simon_test_procedure02 = {
   ]
 }
 
-var simon_pp_feedback = {
-  type: jsPsychSurveyText,
-  questions: [
-    {prompt: 'Heeft u feedback op dit spel (bijvoorbeeld: duidelijkheid van de instructies; moeilijkheidsgraad; iets wat niet goed leek te werken)?<br>We zouden het op prijs stellen als u het ons hieronder laat weten!', name: 'user_feedback', rows:5, required: false},
-  ],
-  data: {
-    variable: "pp_feedback",
-    task: "simon_feedback"
-  }
-}

### 2024-06-04 materials/liss/P_posner/2_posner_procedure.js

@@ -141,14 +141,3 @@ var posner_test_procedure02 = {
     {stim1: 'F', stim2: 'f', condition: 'same',      correct_response: 'A'},
     ]
 }
-
-var posner_pp_feedback = {
-  type: jsPsychSurveyText,
-  questions: [
-    {prompt: 'Heeft u feedback op dit spel (bijvoorbeeld: duidelijkheid van de instructies; moeilijkheidsgraad; iets wat niet goed leek te werken)?<br>We zouden het op prijs stellen als u het ons hieronder laat weten!', name: 'user_feedback', rows:5, required: false},
-  ],
-  data: {
-    variable: "pp_feedback",
-    task: "posner_feedback"
-  }
-}

### 2024-06-04 materials/liss/S_animacysize/2_animacysize_procedure.js

@@ -120,13 +120,3 @@ var animacysize_interblock = {
   }
 }
 
-var animacysize_pp_feedback = {
-  type: jsPsychSurveyText,
-  questions: [
-    {prompt: 'Heeft u feedback op dit spel (bijvoorbeeld: duidelijkheid van de instructies; moeilijkheidsgraad; iets wat niet goed leek te werken)?<br>We zouden het op prijs stellen als u het ons hieronder laat weten!', name: 'user_feedback', rows:5, required: false},
-  ],
-  data: {
-    variable: "pp_feedback",
-    task: "animacysize_feedback"
-  }
-}

### 2024-06-04 materials/liss/S_colorshape/2_colorshape_procedure.js

@@ -11,7 +11,11 @@ var colorshape_test = {
   prompt: "<div style='width: 600px; height:50px;'>" + prompt_yellow + prompt_tri + prompt_circle + prompt_blue + "</div><br><br>" +
           "<div style='width: 600px;'><h1 style='float: left; margin:0;'>A</h1><h1 style='float: right; margin:0;'>L</h1></div>",
   data: {
-    stimulus: function(){return jsPsych.timelineVariable('stim_chr')}
+    stimulus: function(){return jsPsych.timelineVariable('stim_chr')},
+    rule: function(){return jsPsych.timelineVariable('data')['rule']},
+    condition: function(){return jsPsych.timelineVariable('data')['type']},
+    variable: function(){return jsPsych.timelineVariable('data')['variable']},
+    task: function(){return jsPsych.timelineVariable('data')['task']},
   }
 }
 
@@ -108,13 +112,3 @@ var colorshape_interblock = {
 }
 
 
-var colorshape_pp_feedback = {
-  type: jsPsychSurveyText,
-  questions: [
-    {prompt: 'Heeft u feedback op dit spel (bijvoorbeeld: duidelijkheid van de instructies; moeilijkheidsgraad; iets wat niet goed leek te werken)?<br>We zouden het op prijs stellen als u het ons hieronder laat weten!', name: 'user_feedback', rows:5, required: false},
-  ],
-  data: {
-    variable: "pp_feedback",
-    task: "colorshape_feedback"
-  }
-}

### 2024-06-04 materials/liss/S_globallocal2/2_globallocal2_procedure.js

@@ -70,39 +70,39 @@ var globallocal2_target = {
 var globallocal2_procedure01 = {
 timeline: [globallocal2_fixation, globallocal2_target],
   timeline_variables: [
-    {stimulus: SQsq, stim_chr: 'SQsq', key_answer: 'A', data: {rule: 'global', type: 'first', congruency: 'congruent', variable: 'globallocal2_01', task: 'globallocal'}},
-    {stimulus: RECsq, stim_chr: 'RECsq', key_answer: 'A', data: {rule: 'local', type: 'switch', congruency: 'incongruent', variable: 'globallocal2_01', task: 'globallocal'}},
-    {stimulus: SQrec, stim_chr: 'SQrec', key_answer: 'A', data: {rule: 'global', type: 'switch', congruency: 'incongruent', variable: 'globallocal2_01', task: 'globallocal'}},
-    {stimulus: SQrec, stim_chr: 'SQrec', key_answer: 'L', data: {rule: 'local', type: 'switch', congruency: 'incongruent', variable: 'globallocal2_01', task: 'globallocal'}},
-    {stimulus: SQsq, stim_chr: 'SQsq', key_answer: 'A', data: {rule: 'local', type: 'repeat', congruency: 'congruent', variable: 'globallocal2_01', task: 'globallocal'}},
-    {stimulus: RECsq, stim_chr: 'RECsq', key_answer: 'A', data: {rule: 'local', type: 'repeat', congruency: 'incongruent', variable: 'globallocal2_01', task: 'globallocal'}},
-    {stimulus: RECrec, stim_chr: 'RECrec', key_answer: 'L', data: {rule: 'global', type: 'switch', congruency: 'congruent', variable: 'globallocal2_01', task: 'globallocal'}},
-    {stimulus: SQrec, stim_chr: 'SQrec', key_answer: 'A', data: {rule: 'global', type: 'repeat', congruency: 'incongruent', variable: 'globallocal2_01', task: 'globallocal'}},
-    {stimulus: RECrec, stim_chr: 'RECrec', key_answer: 'L', data: {rule: 'global', type: 'repeat', congruency: 'congruent', variable: 'globallocal2_01', task: 'globallocal'}},
-    {stimulus: SQsq, stim_chr: 'SQsq', key_answer: 'A', data: {rule: 'local', type: 'switch', congruency: 'congruent', variable: 'globallocal2_01', task: 'globallocal'}},
[…]

### 2024-06-04 manuscript/manuscript.qmd
intermediate version of manuscript

---
bibliography: references.bib
csl: apa.csl
format: 
  docx:
    reference-doc: reference-doc.docx
output:
  officedown::rdocx_document:
    page_margins:
      bottom: 1
      footer: 0
      gutter: 0
      header: 0.5
      left: 1
      right: 1
      top: 1
    plots:
      align: center
      caption:
        pre: 'Figure '
        sep: '. '
        style: Image Caption
    tables:
      caption:
        pre: 'Table '
        sep: '. '
        style: Table Caption
  pdf_document: default
  word_document: default
editor: 
  markdown: 
    wrap: sentence
---

```{r include = FALSE}
library(flextable)
library(stringr)
library(dplyr)
library(tidyr)
library(tibble)
library(lavaan)
library(officer)

knitr::opts_chunk$set(
  echo = F,
  fig.align = "center",
  fig.pos = "!t", 
  out.extra = "",
  fig.show = "asis",
  message = FALSE,
  tab.topcaption = T,
  warning = FALSE
)

# set up flextable for tables
set_flextable_defaults(
  font.family = "Times", 
  font.size = 10,
  font.color = "black",
  line_spacing = 1,
  padding.bottom = 1, 
  padding.top = 1,
  padding.left = 1,
  padding.right = 1
)
```

#### **Attention shifting and inhibition abilities among people living in adverse conditions**

<br>

#### Stefan Vermeent^1,2^, Anna-Lena Schubert^3^, Jaap J. A. Denissen^1^, & Willem E. Frankenhuis^1,2,4^

#### ^1^ Department of Psychology, Utrecht University, Utrecht, The Netherlands

#### ^2^ Max Planck Institute for the Study of Crime, Security, and Law, Freiburg, Germany

#### ^3^ Department of Psychology, University of Mainz, Mainz, Germany

#### ^4^ Evolutionary and Population Biology, Institute for Biodiversity and Ecosystem Dynamics, University of Amsterdam, Amsterdam, the Netherlands

<br>

# Corresponding author

Correspondence concerning this article should be addressed to Stefan Vermeent, Department of Psychology, Utrecht University, Heidelberglaan 1, 3584 CS Utrecht, The Netherlands. Email: p.c.s.vermeent@gmail.com / p.c.s.vermeent@uu.nl.

# Data Availability

All scripts and materials needed to reproduce the findings are available on the article's Github repository ([https://stefanvermeent.github.io/liss_wm_profiles_2023/](https://stefanvermeent.github.io/liss_wm_profiles_2023/)).
We also include instructions on how to reproduce each step of our analyses. 
In this paper, we make use of data from the LISS panel (Longitudinal Internet studies for the Social Sciences) managed by the non-profit research institute Centerdata (Tilburg University, the Netherlands). 
All datasets are available in the LISS data archive. 
Researchers who want to access the data are required to sign a statement confirming that information about individual persons, households, etc., will not be released to others (go to [https://statements.centerdata.nl](https://statements.centerdata.nl) for more information).

# Funding statement
This research was conducted in whole or in part using ODISSEI, the Open Data Infrastructure for Social Science and Economic Innovations(<https://ror.org/03m8v6t10>).
WEF’s contributions have been supported by the Dutch Research Council (V1.Vidi.195.130) and the James S. McDonnell Foundation (https://doi.org/10.37717/220020502).

# Disclosures

We declare no conflicts of interest.

# Ethics Approval Statement

This study was approved by the Ethics Review Board of the Faculty of Social & Behavioural Sciences of Utrecht University (FETC20-490) and the Ethics committee for research in the Sciences and Life Sciences of the University of Amsterdam (FNWI-41_2023).

\pagebreak

Exposure to adversity (i.e., prolonged exposure to intense stress) is associated with robust, persisting changes in executive functioning (EF) [@ellis_2022; @farah_2006; @sheridan_2022; @sheridan_2014].
Given the importance of EF---a set of abilities involved in the top-down, goal-directed regulation of behavior [@miyake_2000; @rey_mermet_2019]---much research has been devoted to understanding which EF abilities are most strongly associated with adversity.
In some cases, these insights have already been used to screen for early EF deficits, and to develop interventions to help bridge performance gaps [@distefano_2021].

A lot of what we know about the association between adversity and EF stems from analyzing raw performance, most often response times (RT) or accuracy rates.
In addition, it is common for studies to only include a single task per EF ability.
For example, previous studies have inferred difficulties in ignoring distractions (inhibition) in people from adversity using a single inhibition task [e.g., @farah_2006; @fields_2021; @mezzacappa_2004; @noble_2005].
The same thing is true for observed adaptive benefits in the ability to shift attention between tasks (attention shifting) [@fields_2021; @howard_2020; @nweze_2021; @young_2022; @mittal_2015].
In effect, an important assumption underlying many deficit and adaptation findings is that raw task performance offers a reliable, relatively pure estimate of the EF ability [@white_2022].

Unfortunately, an increasing number of psychometric investigations into EF do not support this assumption [@krumm_2009; @loffler_2024; @rey_mermet_2019; @rouder_2019; @stahl_2014].
A central issue is that raw task performance is not task-pure.
Aside from the EF ability needed to solve the task, a person's RT is also influenced by other factors such as response caution (i.e., their relative emphasis on speed versus accuracy).
Accounting for these factors using cognitive modeling, a recent study found that children with more household threat were slower on EF tasks partly because they were more cautious [@vermeent_2024a].

Aside from different processing stages, performance across EF tasks is also influenced by general processing speed [@frischkorn_2019; @lerche_2020; @loffler_2024].

[Might not be any ]
Even more problematic are recent studies suggesting that performance on EF tasks taps into little more than general processing speed.
For example, after accounting for other factors such as response caution, another study found that a common (latent) EF factor was fully explained by performance on really basic processing speed tasks [@loffler_2024].
A few studies have focused on this distinction in the context of adversity and/or socioeconomical status.
on a raw performance level, lower SES was found to be negatively correlated with a task-general factor, with additional associations on the task-specific level.
In another

# # Implications and potential solutions for adversity research

An important implication of this work is that standard assessments---based on raw performance of single EF tasks---may be ill-suited at estimating associations between adversity and specific EF abilities.
Patterns that may look like deficits (or adaptations) in specific abilities may not be driven by these abilities at all.
For example, if adversity is associated with slower task-general processing speed, analyses focused on individual EF tasks could make it seem like many different abilities impaired, instead of a single, general process.

Fortunately, these challenges can be tackled using a combination of cognitive modeling and structural equation modeling.
As a first step, ability-relevant processing should be separated from ability-irrelevant processes, such as response caution or the time taken to prepare for a task [...].
One effective way of doing this is through the use of cognitive models like the Drift Diffusion Model [DDM\; @forstmann_2016; @ratcliff_1998; @ratcliff_2008; @wagenmakers_2009].
The DDM broadly divides the time taken on speeded EF tasks into a decision-making stage and a non-decision-making stage [see Figure 1].
In the decision-making stage, people accumulate evidence for one of two responses (e.g., left or right button), until they have enough evidence for one response over the other.
The amount of evidence a person requires can vary between people (as well as between tasks within the same person) [XXX], and is modeled as the distance between the decision boundaries for each respective response.
Finally, the DDM models two non-decision stages: one at the start of the trial (e.g., task preparations, stimulus encoding) and one at the end of the trial (i.e., response execution).

When applied to empirical performance data, the DDM distinguishes between three parameters mapping onto distinct cognitive processes.
The *drift rate* reflects the rate at which the evidence accumulation process drifts towards the decision boundary, and thus provides a measure of the speed/efficiency of information processing.
The *boundary separation* reflects the width between the two decision boundaries, and provides a measure of a person's response caution.
Finally, the *non-decision time* reflects a combination of preparation time (e.g., stimulus encoding) as well as execution time (e.g., time spent pressing the button).

Researchers have argued that the drift rate on EF tasks is a more valid measure of EF ability, at least to the extent to which the tasks are reliable measures of EF ability [@loffler_2024; @vermeent_2024a; @weigard_2021].
However, drift rates are still influenced at least partly by general speed of processing.
Theoretically, this should be the part of the variance that is shared across drift rates of all tasks, and of conditions both including an EF component (e.g., incongruent trials) and excluding an EF component (e.g., congruent trials) [@loffler_2024].
This part of the variance can be estimated on a latent level using SEM.
Any remaining task-specific variance then contains EF-specific variance.

As mentioned above, a recent investigation did not find significant associations between adversity and task-specific EF variances [@vermeent_2024a].
This could mean one of two things.
First, task-specific variance may reliably reflect specific EF abilities, and adversity may not be associated with these abilities.
Second, task-specific variance may *not* reliably reflect specific EF abilities.
That is, after accounting for general speed of processing, little more than measurement error may remain.
Testing these hypotheses requires sampling two or more tasks that are supposed to measure the same ability, and using latent modeling to investigate whether they form a consistent latent factor after accounting for general speed of processing.

# The current study

# Methods

## Participants

Participants were 1,000 Dutch people randomly drawn from the Longitudinal Internet studies for the Social Sciences (LISS) panel [@scherpenzeel_2011].
The LISS panel is an invite-only, representative probability sample of the Dutch population consisting of roughly 7,500 individuals across 5,000 households.
LISS participants complete a yearly core battery of questionnaires about various key domains, among which their financial situation over the past year.
In addition, they can participate in further monthly studies covering a wide range of different topics.
The current study was fielded between May and June 2024.
Participants were able to complete the cognitive tasks across two or more sessions to increase participation rates.
People were eligible for participation if they (1) were between 18 and 55 years old; (2) if they had completed at least one wave of a study in the LISS archive assessing exposure to crime (see below); (3) if they had agreed to linking their LISS data to government microdata (not relevant here).
Participants were excluded if they interacted with other browser tabs mid-trial on any of the cognitive tasks (but not during breaks or in-between tasks). 
If performance on a specific task was at or below chance level, we excluded the data for that task but retained other task data of that participant.

## Adversity measures

We preregistered our approach to compute adversity composites based on observed correlations between measures.
In short, if all measures of an adversity type correlated .60 or higher (indicating "strong" correlations), we calculated a uniformly weighted average.
If one or more correlations were lower than .60, we used Principal Component Analysis (PCA) to the separate measures and extracted only the first principal component score.
Perceived scarcity and income-to-needs were considered as two separate types of material deprivation (and unpredictability), regardless of their correlation.

### Neighborhood threat

*Perceived neighborhood crime.* Participants completed the Neighborhood Violence Scale [NVS\; @frankenhuis_2018; @frankenhuis_2020] in the current study.
The NVS included seven items about current perceived neighborhood threat (e.g., “Where I live, it is important to be able to defend yourself against physical harm”), on a scale of 1 ("Completely disagree") to 7 ("Completely agree"). 
All items of the NVS were averaged together and standardized.
We also included four items on perceived neighborhood threat from the LISS archive (six waves: <https://doi.org/10.17026/dans-zch-j8xt>), in which participant reported how often it happens that they 1) “avoid certain areas in your place of residence because you perceive them as unsafe”, 2) “do not respond to a call at the door because you feel that it is unsafe”, 3) “leave valuable items at home to avoid theft or robbery in the street?”, 4) “make a detour, by car or on foot, to avoid unsafe areas?” on a scale of 1 (“(Almost) never”), 2 (“Sometimes”), or 3 (“Often”). 
We recoded these items so that "(Almost) never" corresponded to zero.
We summed the items within each wave, and then calculated an average across waves for which participants had data.

*Crime victimization.* We used seven items from the LISS archive (six waves: <https://doi.org/10.17026/dans-zch-j8xt>) in which participants reported whether or not they were the victim of eight types of crime in the last two years: (1) burglary or attempted burglary; (2) theft from their car; (3) theft of their wallet or purse, handbag, or other personal possession; (4) wreckage of their car or other private property; (5) intimidation by any other means; (6) maltreatment of such serious nature that it required medical attention; (7) maltreatment that did not require medical attention. 
As the last wave occurred in 2018, we included the same items in the current study asking about people's crime exposure in the last two years.
We computed the total number of distinct crimes that participants were exposed to at any moment in time (a 'variety score'\; @sweeten_2012).

*Neighborhood threat composite.* Correlations among the two measures of perceived neighborhood crime and crime victimization were low.
Therefore, we used PCA to extract the first principal component score.
This score accounted for XXX% of the variance in the three measures.

### Material deprivation

*Perceived scarcity (mean).* Measures of perceived scarcity were derived from the LISS archive, using the yearly recurring core study on household and personal income (16 waves: <https://doi.org/10.57990/1gr4-bf42>).
First, participants reported how difficult it currently is to live off the income of their household, on a scale from 0 (very hard) to 10 (very easy). 
Second, participants reported which of the following best applied to their current financial situation: (1) “we are accumulating debt”; (2) “we are somewhat eating into savings”; (3) “we are just managing to make ends meet”; (4) “we have a little bit of money to spare”; (5) “we have a lot of money to spare”. 
Third, participants reported which of the following applied to their current financial situation (0 = no, 1 = yes): (1) “having trouble making ends meet”; (2) unable to quickly replace things that break”; (3) “having to lend money for necessary expenditures”; (4) “running behind in paying rent/mortgage or general utilities”; (5) “debt collector/bailiff at the door in the last month”; (6) “received financial support from family or friends in the last month”.
Responses were coded so that higher scores indicated more perceived scarcity.

We first averaged each measure separately across waves, and then scaled them.
As all item correlations were > .60, we computed a uniformly weighted average.

*Income-to-needs (mean).* The income-to-needs ratio was derived from monthly self-reported net household income in the LISS archive (<https://doi.org/10.57990/qn3k-as78>).
Reported income of zero was set to missing, as this could indicate an unwillingness to report on income rather than a true income of zero.
If net household income data was missing (or zero) for an entire year, we imputed, if possible, the reported yearly net household income from the core study on income (<https://doi.org/10.57990/1gr4-bf42>), dividing by 12 to obtain monthly estimates.
We then divided the income by the poverty threshold set by Statistics Netherlands (CBS) [@brakel_2023; CBS, personal communication, December 15, 2023].
For households with more than three children, we used the threshold for a family with three children, as CBS does not calculate thresholds for families larger than three children.
For the same reason, income of households with more than two adults was compared to the threshold for households with two adults.
We computed the average income-to-needs ratio across years, reverse coding it so that higher scores indicated households further below the threshold.

### Unpredictability

*Perceived scarcity (SD/mean).* Unpredictability over time was computed using the coefficient of variation, which is the within-person standard deviation across years divided by the mean across years (@key_2017; @liu_2022; @ugarte_2023; @walasek_2024; @young_2020).
This measure accounts for the fact that the mean and standard deviation tend to be highly positively correlated.
Unpredictability in perceived scarcity was derived from the same measures as the mean level of perceived scarcity (see above).
We first computed the standard deviation for each measure separately across years, and then scaled them.
We then computed a uniformly weighted average, which was divided by the mean level of perceived scarcity across years.

*Income-to-needs (SD/mean).* Unpredictability over time in the income-to-needs ratio was also computed using the coefficient of variation.
We first computed the within-person standard deviation of the income-to-needs ratio across years, and then divided it by the mean income-to-needs ratio across years.

### Cognitive measures

All cognitive tasks were programmed in jsPsych 7.3 [@deleeuw_2015].
At the start of the session, participants entered fullscreen mode to avoid distractions from other browser tabs.
The tasks were presented against a light-gray background.

*Flanker Task.* This task measures inhibition of distractor interference [...]
On each trial, participants saw five arrows side-by-side horizontally, pointing either left or right.
Their task was to indicate the direction of the central arrow.
The arrows were randomly presented 300 pixels above or below the center of the screen.
On 50% of the trials, all arrows pointed in the same direction (congruent trials).
On the other half, the arrows surrounding the central arrow pointed in the opposite direction (incongruent trials).
Participants first completed eight practice trials, followed by two test blocks of 32 trials each, for a total of 64 trials.

*Simon Task.* This task measures inhibition of prepotent responses [...].
On each trial, participants saw the word "LEFT" or "RIGHT" (printed in Dutch), presented either on the left or right side of the screen.
Their task was to press the 'A' key if the word was "LEFT" and the 'L' key if the word was "RIGHT", regardless of the location on the screen.
on 50% of the trials, the word matched the location (e.g., the word "LEFT" presented on the left side; congruent trials).
On the other half, the word did not match the location (e.g., the word "LEFT" presented on the right side; incongruent trials).
Participants first completed eight practice trials, followed by two test blocks of 32 trials each, for a total of 64 trials.

*Color-Shape Task.* This task measures the ability to shift attention between different task goals [XXX].
On each trial, participants saw a square or a circle in the center of the screen that was either blue or yellow.
Depending on the task rule printed above the stimulus, their task was to classify the stimulus based on it's shape or color.
On 50% of the trials, the rule was the same as on the preceding trial (repeat trials).
On the other half, the rule was different than on the preceding trial (switch trials).
The same stimulus was never presented more than twice in a row, and there were never more than three repeat or switch trials in a row. 
Participants first completed eight practice trials, followed by two test blocks of 32 trials each, for a total of 64 trials.

*Animacy-Size Task.* This task measures the ability to shift attention between different task goals [XXX].
On each trial, participants saw a single noun (in Dutch) referring to an animal or object [adopted from XXX].
Depending on the task rule printed above the noun, their task was to classify it based on whether it referred to a living or non-living thing (e.g., wasp vs. piano), or whether it referred to something that was smaller or larger than a soccer ball (e.g., ring vs. dolphin).
On 50% of the trials, the rule was the same as on the preceding trial (repeat trials).
On the other half, the rule was different than on the preceding trial (switch trials).
There were never more than three repeat or switch trials in a row.
Participants first completed eight practice trials, followed by two test blocks of 32 trials each, for a total of 64 trials.

*Global-local Task.* This task measures the ability to shift attention between task goals [XXX]
Stimuli were adapted from [...].
On each trial, participants saw a large square or rectangle which was composed of 16 small squares or rectangles.
The stimulus where flanked on both side by a drawing of an elephant or mouse, which was presented 1,000 ms prior to the appearance of the stimulus.
If the stimulus was flanked by the image of an elephant (50% of trials), participants had to indicate whether the global image was a square or rectangle.
If the stimulus was flanked by the image of a mouse (50% of trials), participants had to indicate whether the local images were squares or rectangles.
On 50 % of the trials, the rule was the same as on the preceding trial (repeat trials).
On the other half, the rule was different than on the preceding trial (switch trials).
Finally, the stimuli were congruent on 50% of the trials (e.g., large square consisting of small squares) and incongruent on the other half (e.g., large square consisting of small rectangles).
Congruency, task rule (switch vs. repeat) or focus (global vs. local) where never repeated more than three times in a row.
Participants first completed eight practice trials, followed by two test blocks of 32 trials each, for a total of 64 trials.

*Posner Task.* This task measures basic speed of processing.
On each trial, participants saw two letters in the center of the screen, drawn from the set A, B, F, H, Q, a, b, f, h , q.
Their task was to indicate whether the letters were the same (e.g., "AA", "bB") or different (e.g., "AQ", "Fh").
On 50% of the trials, the letters were the same, and on the other half they were different.
Participants first completed 8 practice trials, followed by two test blocks of 32 trials each, for a total of 64 trials.

### Covariates

We preregistered a Directed Acyclic Graph (DAG) to inform the inclusion of covariates to the model.
Based on this DAG, we decided to control for age (in years), the quadratic effect of age, education level, and childhood adversity.
Childhood adversity was the unweighted average across two scales measuring childhood threat and deprivation: (1) the Neighborhood Violence Scale [XXX], which consists of seven items (e.g., "Physical fights were common in the neighborhood where I grew up"), and (2) four items measuring access to housing, food, clothing, and medicine.

## Procedure

We obtained ethical approval from the first author's institutional ethical board. 
The study was fielded on the LISS platform, and participants could only complete the study on a laptop or desktop PC.
Participants started with the six cognitive tasks, in randomized order.
After each task, they were asked if they wanted to continue to the next task or conclude the session. 
If they choose for the latter, they were able to continue with the tasks at another time.
The cognitive tasks took around 25 minutes to complete.
After completing all cognitive tasks, participants completed questionnaires about past and present neighborhood threat and material deprivation, exposure to crime victimization in the past two years, trait impulsivity (not considered here) and trait future orientation (not considered here).
Finally, they answered a few standard LISS evaluation questions about their experiences with the study, with the opportunity to provide written feedback.

## Analysis plan

### Data cleaning

For all tasks, we removed trials with RTs < 250ms and trials with RTs more than 3.2 SD above the intra-individual log-transformed mean RT.

### DDM estimation

We used an Hierarchical Bayesian implementation of the DDM (HDDM) [@vandekerckhove_2011; @wiecki_2013].
The benefit of HDDM over traditional approaches is that the model uses information of the whole sample to inform individual parameter estimation.
The model will be applied to each task separately.
Across all tasks, the bias parameter was held constant, as there is no reason for participants to favor one response over the other across trials.
For the Posner Task, we estimated a single drift rate, boundary separation, and non-decision time for each participant.
For all other tasks, drift rates and non-decision times were estimated separately for congruent (repeat) and incongruent (switch) trials.
Boundary separation was always fixed across conditions, following the assumption that people cannot anticipate the condition of the next trial, and hence cannot adjust their strategy on a trial-by-trial basis.

The DDM models were fit using the *runjags* package [@denwood_2016].
Each model was fit with three Markov Chain Monte Carlo (MCMC) chains.
We used 2,000 burn-in samples and 10,000 additional samples, of which every 10th sample was retained, resulting in a total of 3,000 posterior samples.
See the supplemental materials for more information about model convergence and fit assessment.

### SEM

### 2024-06-04 manuscript/old_snippets.qmd
intermediate version of manuscript

---
title: "old snippets"
format: html
editor: visual
---

The question of how exposure to adversity (i.e., prolonged exposure to intense stress) is associated with executive functioning (EF) has inspired a large, robust body of research over the last few decades [@mclaughlin_2019; @sheridan_2022].
This literature indicates that the association depends on the dimension of adversity as well as the specific EF ability under investigation, among other factors [...].
On the one hand, adversity exposure---especially a lack of access to basic resources---tends to be associated with lowered performance on a wide range of EF tasks [@farah_2006; @sheridan_2022; @sheridan_2014].
Aside from such *deficits* in EF, there is emerging evidence that exposure to adversity could enhance certain EF abilities through *developmental adaptations* to the environment [@ellis_2021; @frankenhuis_2013; @frankenhuis_2020].

Both types of findings are generally based on raw performance differences, such as response times (RT) or accuracy, and often using a single task per EF ability.

Instead, it has proven difficult to reliably measure specific EF abilities, for two related reasons.
First, performance differences may arise because of factors other than inhibition or shifting ability, such as a person's level of response caution [XXX].
Second, adversity may be associated with general speed of processing shared across all EF tasks, instead of with specific EF abilities [XXX].
In this article, we leverage insights from psychometric research into EF to more precisely estimate how adversity is associated with two EF abilities: attention shifting and inhibition.

People who report more exposure to adversity (i.e., prolonged exposure to intense stress) tend to show lowered performance on a wide range of executive functioning tasks [EF\; @johnson_2021; @lund_2020; @lund_2022; @roos_2016].
EF refers to a set of abilities involved in the top-down, goal-directed regulation of behavior [@miyake_2000; @rey_mermet_2019].
Negative associations between adversity exposure and EF performance may already arise in early childhood, and tend to persist into adulthood [@hostinar_2013; @mclaughlin_2019; @sheridan_2022; @shields_2017].
Researchers working on a *deficit* literature have generally interpreted these associations as impairments in specific EF abilities [e.g., inhibition\; @fields_2021; @mezzacappa_2004; @mittal_2015] as well as impairments in brain networks supporting EF more generally [...].

Recent evidence suggests, however, that adversity might be associated with separate *components* of EF in different, sometimes even opposing, ways.
Two notable examples are inhibition [the ability to ignore task-irrelevant distractions\; ...] and attention shifting [the ability to flexibly shift attention between different task goals\; ...].
Performance on inhibition tasks is often lowered in people from adversity or from lower socioeconomic backgrounds [@fields_2021; @mittal_2015; but see @vermeent_2024a], attention shifting is sometimes found to be enhanced or intact
[@mittal_2015; @fields_2021; @young_2022].
Researchers drawing on this *adaptation* literature have generally interpreted these associations as evidence that attention shifting offers an adaptive advantage in unpredictable environments, while inhibition does not [@ellis_2022; @frankenhuis_2016; @frankenhuis_2020; @mittal_2015].

The validity of inferring cognitive deficits or adaptations based on raw performance differences hinges on the task's ability to reliably measure the actual EF ability.
That is, in order to infer impaired (or enhanced) ability from raw performance differences, we need to be sure that raw performance indeed reliably reflects differences in the ability.
Emerging psychometric evidence increasingly shows that this is not the case [@loffler_2024; @vermeent_2024a; @weigard_2021].
First, performance differences may arise because of factors other than inhibition or shifting ability, such as a person's level of response caution.
Second, adversity may be associated with general speed of processing shared across all EF tasks, instead of with specific inhibition or shifting specifically.
In this article, we take a psychometric approach to address these issues.

A second problem is that performance on tasks that rely on speed---which is true for many inhibition and attention shifting paradigms---is influenced by general speed of processing [@frischkorn_2019; @lerche_2020; @loffler_2024].
If adversity is associated with slower general speed of processing, this will propagate across performance on many different tasks.
If EF abilities are assessed in isolation using one or a small number of tasks, lowered general speed of processing could be mistaken as several deficits in distinct EF abilities.
Evidence for this possibility comes from a recent study documenting an association between exposure to household threat and task-general speed of processing, but not with specific EF abilities [@vermeent_2024a].
Likewise, young adults who reported more childhood exposure to violence and unpredictability showed lower quality of stimulus processing on the Flanker Task, but did not differ in their ability to inhibit distractions [@vermeent_2024b].

Several studies, some of which using more advanced cognitive modeling approaches, show that performance across a broad set of EF tasks is best explained by a single general factor reflecting more basic processing speed/efficiency. [@loffler_2024; @vermeent_2024a; @weigard_2021].
Even when tasks are supposed to measure the same EF ability, most of what is shared between them instead reflects mere processing speed [@hedge_2021].

This is especially worrying for findings relating to inhibition, as previous research has often failed to find a coherent latent factor underlying different inhibition paradigms [@krumm_2009; @rey_mermet_2019; @rouder_2019; @stahl_2014].
This begs the question whether negative associations between adversity and performance on inhibition tasks are truly the result of lowered inhibition, with some work suggesting they are not [@vermeent_2024b].

Thus, our conclusions about how adversity is associated with inhibition will be influenced strongly by the specific task that is used.
In this article, we focus on three issues that result from an overreliance on specific paradigms coupled with a lack of comparison across paradigms.
First, task paradigms differ in several ways, which could lead to differen
Second, performance on most measures of inhibition and attention shifting is confounded with general processing speed, which is often not accounted for.
Third, performance on these measures not just influenced by EF ability, but also by other, ability-irrelevant processes.

### 2024-06-05 preregistration/README.qmd
intermediate commit of preregistration

@@
   html:
+    css: D:/repositories/liss_ef_2024/assets/style.css
     template: ../assets/template.html
-    css: ../assets/style.css
 bibliography: ../manuscript/references.bib
@@
 link-citations: true
+editor: 
+  markdown: 
+    wrap: sentence
 ---
@@
 library(ftExtra)
+library(tinytable)
 
@@
 *Last updated `r format(Sys.time(), "on %A, %B %d, %Y at %I:%M %p")`*
+
+```{r}
+#| include: false
+
+pagedown::chrome_print(input = "preregistration/README.html", output = "preregistration/preregistration.pdf")
+```
+
+[click here to download a pdf version](https://github.com/Schecher1/Minecraft-Server-Creator/blob/master/README.md)
 
@@
 -   Anna-Lena Schubert^3^
--   Jean-Louis van Gelder^2^
+-   Jean-Louis van Gelder^2^?
 -   Willem E. Frankenhuis^2,4^
@@
 
-### Q3: Research Questions DONE
+### Q3: Research Questions
 
-**R1:** How is adversity associated with general speed of processing across all cognitive tasks?
+-   **R1:** How is adversity associated with general speed of processing across all cognitive tasks?
 
-**R2:** How is adversity associated with inhibition and attention shifting abilities after accounting for general speed of processing?
+-   **R2:** How is adversity associated with inhibition and attention shifting abilities after accounting for general speed of processing?
 
-**R3:** How is adversity associated with task-general and task-specific response caution?
+-   **R3:** How is adversity associated with task-general and task-specific response caution?
 
-### Q4: Hypotheses DONE
+### Q4: Hypotheses
 
@@
 
-A1. Adversity is negatively associated with general speed of processing.
+-   **A1.** Adversity is negatively associated with general speed of processing.
 
-A2. Adversity is negatively associated with either or both latent estimates of inhibition and attention shifting, after accounting for general speed of processing.
+-   **A2.** Adversity is negatively associated with either or both latent estimates of inhibition and attention shifting, after accounting for general speed of processing.
 
-A3. If latent estimates of inhibition and attention shifting are highly correlated, adversity is negatively associated with a general EF factor spanning both.
+-   **A3.** If latent estimates of inhibition and attention shifting are highly correlated, adversity is negatively associated with a general EF factor spanning both.
 
@@
 
-B1.
-Adversity is positively associated with the latent estimate of attention shifting *regardless of the association with inhibition*, after accounting for general speed of processing.
+-   **B1.** Adversity is positively associated with the latent estimate of attention shifting *regardless of the association with inhibition*, after accounting for general speed of processing.
 A positive association with inhibition would be unexpected yet interesting, and would require revising existing adaptation theoretical frameworks.
 
-B2.
-The association between adversity and the latent estimate of attention shifting is practically equivalent to zero *and* the association between adversity and the latent estimate of inhibition is negative.
+-   **B2.** The association between adversity and the latent estimate of attention shifting is practically equivalent to zero *and* the association between adversity and the latent estimate of inhibition is negative.
 
@@
 
-C1. Adversity is associated with more task-general response caution.
+-   **C1.** Adversity is associated with more task-general response caution.
+
+Go back to [Overview].
 
@@
 
-### Q5: Dataset DONE
+### Q5: Dataset
 
-Our study will include 1000 participants who were randomly sampled from the LISS panel (Scherpenzeel, 2011).
-The LISS panel is a representative probability sample of roughly 5,000 Dutch households (\~7,500 individuals) drawn from the population register by Statistics Netherlands on an invite-only basis.
+The LISS panel is a representative probability sample of roughly 5,000 Dutch households (\~7,500 individuals) drawn from the population register by Statistics Netherlands on an invite-only basis [@scherpenzeel_2011].
 Households without a computer or internet connection are provided with these facilities by LISS.
@@
 In addition, participants can complete additional online questionnaires every month, with variable content.
+
 The current study integrates two data sources.
-First, our sample of 1000 participants will participate in a new LISS study between May 2024 and July 2024 (hereafter referred to as 'newly collected data'), in which we will include a measure of neighborhood threat, three cognitive tasks measuring attention shifting, two cognitive tasks measuring inhibition, and one cognitive task measuring processing speed.
+First, we will sample 1,000 participants in a new LISS study, collected between in May and June 2024 (hereafter referred to as 'newly collected data'), in which we will include measures of neighborhood threat and deprivation, three cognitive tasks measuring attention shifting, two cognitive tasks measuring inhibition, and one cognitive task measuring processing speed.
 Second, we will access data that were previously collected in LISS (hereafter referred to as 'the LISS archive').
+This preregistration is submitted before completion of (and access to) the newly collected data.
 
-### Q6: Public Availability DONE
+### Q6: Public Availability
 
-All datasets are available in the LISS data archive.
+All previously collected LISS data are available in the LISS data archive.
 Researchers who want to access the data are required to sign a statement confirming that information about individual persons, households, etc., will not be released to others (go to <https://statements.centerdata.nl> for more information).
 
-### Q7: Data Access DONE
+### Q7: Data Access
 
@@
 
-After the new data have been collected, they will become available in the general overview at <https://www.dataarchive.lissdata.nl/study-units/view/1>
+After the new data have been collected, they will become available in the general overview at <https://www.dataarchive.lissdata.nl/study-units/view/1>.
 
-### Q8: Date of Download DONE
+### Q8: Date of Download
 
@@
 
-### Q9: Data Collection DONE
+### Q9: Data Collection TBD (fix inclusion criteria)
 
 General information about recruitment of participants into the LISS study can be found at <https://www.lissdata.nl/methodology>.
+The inclusion criteria for the new data collection are that people are between 20 and 55 years old and have completed at least 1 wave of the Conventional and Computer Crime Victimization study (see Q10).
 
-Our study will include 1,000 participants who were randomly sampled from the LISS panel.
-The requirements will be that they are between 20 and 55 years old and have completed at least 1 wave of the Conventional and Computer Crime Victimization study (see Q10).
-
-### Q10: Codebooks DONE
+### Q10: Codebooks
 
@@
 
-Codebooks for the newly collected data will become available later at <https://github.com/StefanVermeent/liss_ef_2024>
+Codebooks for the newly collected data will become available later at <https://www.dataarchive.lissdata.nl/study-units/view> and <https://github.com/StefanVermeent/liss_ef_2024>.
+
+Go back to [Overview].
 
@@
 
-### Q11: Manipulated Variables: DONE
+### Q11: Manipulated Variables
 
@@
 
-We identified potential confounders of the empirical estimands (i.e., the association between self-reported threat and deprivation with cognitive outcomes) using a Directed Acyclic Graph (DAG).
-We analyzed the following variables both conceptually (i.e., based on previous literature and/or our own assumptions) and, in cases of doubt, statistically: (1) age, (2) sex, (3) past adversity exposure (material deprivation and threat combined), (4) recent threat exposure, and (5) recent material deprivation.
+We used Directed Acyclic Graphs (DAG) to identify potential confounders of the empirical estimands (i.e., the association between self-reported threat and deprivation with cognitive outcomes).
+For more information on DAGs, see @rohrer_2018.
+
+We analyzed the following potential confounders both conceptually (i.e., based on previous literature and/or our own assumptions) and, in cases of doubt, statistically (see below): (1) age [@salthouse_2016; @salthouse_2019; @starns_2010], (2) biological sex [@ning_2023], (3) childhood adversity exposure (material deprivation and threat combined) [@bos_2009; @goodman_2019], and (4) recent material deprivation as a confounder for the effect of threat on cognitive performance [@bywaters_2016; @ning_2023].
+
+**Estimand 1: Material deprivation -\> Cognitive performance.** See Figure 2A for a visualization of the DAG for material deprivation.
+All models involving material deprivation as the main predictor will control for these four covariates.
+In line with recent work, we consider material deprivation to precede threat exposure [@bywaters_2016; @ning_2023].
+For this reason, we do not include threat exposure in models where we are interested in the total effect of material deprivation on cognitive performance.
+
+**Estimand 2: Threat exposure -\> Cognitive performance.** See Figure 2B for a visualization of the DAG for threat exposure.
+We considered the same set of confounders as for material deprivation.
+However, in the case of threat exposure, we were conceptually less sure about the potential effects of age and education on threat exposure.
+We also wanted to gather more empirical support for the assumption that material deprivation should be considered as a confounder of the effect of threat exposure on cognitive performance [@bywaters_2016; @ning_2023].
+
+Table 1 offers an overview of conceptual and statistical evidence for and against each potential confounder.
 Statistical analysis of potential confounders was based on data from a previous LISS study (see Q17 and Q18) using the `dagitty` R package [@textor_2016].
-Specifically, we used the `dagitty::localTests()` function to establish whether DAG-implied conditional independencies between certain key variables were supported by the data [@ankan_2021].
+Specifically, we used the `dagitty::localTests()` function to establish whether DAG-implied conditional independencies between certain key variables were supported by these data [@ankan_2021].
 The outcome in these conceptual DAGs was working memory capacity, measured on a latent level in the previous LISS study.
+For each of the three confounders under investigation, we removed the causal pathway between that variable and threat exposure from the DAG, and then tested whether this imposed independence (conditional on the other causal effects in the model) was consistent with the data.
 Code for these analyses can be found under `scripts/00_dag`.
-
-Figure 1A shows the initially conceptualized DAG. 
-Black lines depict our estimands, and red lines depict potential confounding paths.
-Solid lines depict paths that we were reasonably sure about given previous literature.
-Dashed lines depict paths that we were less sure about, and therefore decided to statistically test as outlined above.
-
-(1) **Age.** *Conceptual reasoning*: Age has a causal effect on processing speed [@salthouse_2016; @salthouse_2019] and on response caution [@starns_2010].
-Older adults may on average be more financially stable.
-
-We know of no evidence to consider age as a precedent for threat exposure, especially because we measure recent adversity and not cumulative lifetime adversity (which is likely related to age as older individuals on average will have more cumulative exposure to adverse events).
-
-(2) **Sex.**
-
-(4) **Childhood adversity.** *Conceptual reasoning*:
-
-childhood adversity on education: @fry_2018
-Education on poverty: @hofmarcher_2021
 
@@
 #| fig-cap: |
-#|   **Figure 1.** Directed Acyclic Graph (DAG) depicting the empirical estimands (black lines) and presumed confounding effects (red lines).
-Age, education level, and childhood adversity are all conceptualized as confounding variables, as they are expected to
-#|   Importantly, we consider deprivation to be a confounder in the effect of threat on cognitive ability, and conversely, we consider threat to be a mediator of the effect of deprivation on cognitive ability [@ning_2023].
+#|   **Figure 1.** Directed Acyclic Graph (DAG) depicting the empirical estimands (blue lines) and presumed confounding effects (red lines).
+Importantly, we consider deprivation to be a confounder in the effect of threat on cognitive ability, and conversely, we consider threat to be a mediator of the effect of deprivation on cognitive ability [@ning_2023].
 From this, it follows that deprivation should be adjusted for when estimating the effect of threat on cognitive ability.
@@
 ```{r}
-#| echo: false
+#| include: false
+
 table1 <- 
@@
       "Childhood adversity",
-      "Recent material deprivation",
-      "Recent threat"
+      "Recent material deprivation"
     ),
     `Causal effect on cognitive performance` = c(
-      "@salthouse_2016; \n @salthouse_2019; \n @starns_2010",
+      "@salthouse_2016 \n @salthouse_2019 \n @starns_2010",
       "@lovden_2020",
       "@ning_2023", 
-      "@bos_2009; \n @goodman_2019",
-      "*Estimand of interest*",
+      "@bos_2009 \n @goodman_2019",
       "*Estimand of interest*"
@@
     `Causal effect on recent threat` = c(
-      "Unclear",
-      "Unclear",
+      "*Unclear*",
+      "*Unclear*",
       "@ning_2023", 
       "@bywaters_2016",
-      "@bywaters_2016; @ning_2023; ",
-      ""
+      "@bywaters_2016 \n @ning_2023"
     ),
-    `Causal effect on recent material deprivation` = c(
-      "Older adults may have more financial resources",
-      "@fry_2018",
-      "@ning_2023", 
-      "@bywaters_2016",
-      "",
-      ""
-    ),
-    `Statistical (non-)support` = c(
+    `Evidence for conditional independence` = c(
       paste0("Conditional on childhood adversity and recent deprivation, \n *r* = ", localTest_dag03 |> pull(estimate), " SE = ", localTest_dag03 |> pull(std.error), ", 95% CI = ", localTest_dag03 |> pull(ci), "."),
@@
       "",
-      paste0("conditional on childhood adversity and age, deprivation is a significant confounder of threat \n *r* = ", localTest_dag04 |> pull(estimate), " SE = ", localTest_dag04 |> pull(std.error), ", 95% CI = ", localTest_dag04 |> pull(ci), "."),
-      ""
+      paste0("conditional on childhood adversity and age, deprivation is a significant confounder of threat \n *r* = ", localTest_dag04 |> pull(estimate), " SE = ", localTest_dag04 |> pull(std.error), ", 95% CI = ", localTest_dag04 |> pull(ci), ".")
     ),
@@
       "Yes",
-      "Yes",
-      ""
+      "Yes"
     ),
-    `Confounder in deprivation model` = c(
-      "Yes",
-      "Yes",
-      "Yes", 
-      "Yes",
-      "",
-      "No, is a mediator"
+  ) |> 
+  flextable() |> 
+  ftExtra::colformat_md() |> 
+  mk_par(
+    j = 4,
+    value = as_paragraph(
+      gg_chunk(value = localTest_plot_list, width = 1, height = 1)
     )
   ) |> 
-  flextable() |> 
-  ftExtra::colformat_md()
+  add_header_row(
+    values = " ",
+    colwidths = 5
+  ) |> # Add a new header row on top. We can use this new row to add the title
+  compose(
+    i = 1, j = 1,
+    as_paragraph(as_b("Table 1. "), "Conceptual and statistical confounder analysis for the threat exposure model."),
+    part = "header"
+  ) |>
+  add_footer_row(
+    values = " ",
+    colwidths = 5
+  ) %>% 
+  add_footer_row(
+    values = " ",
+    colwidths = 5
+  ) %>% 
+  compose(
+    i = 1, j = 1, 
+    as_paragraph(as_i("Note: "), "the second and third columns present conceptual support (based on previous literature and/or our own assumptions) in favor of a causal effect between the confounder and threat and cognitive performance.
+The embedded plots in column 4 present the outcome of statistical analyses that tested against independence between the confounder and threat exposure, conditional on other variables associated with both [see @ankan_2021 and the main text for more details]."),
+    part = "footer"
+  ) |> 
+  autofit()
+```
+
+```{r}
+#| echo: false
 
@@
 ```
+
+We found evidence against conditional independence between age and threat.
+We also found evidence against independence between material deprivation and threat exposure, conditional on age and childhood adversity.
+Finally, we did not find evidence against conditional independence between education and threat.
+Therefore, all models involving recent threat exposure as the main predictor will control for age, sex, childhood adversity, and recent material deprivation.
 
@@
 
-*Perceived resource scarcity (Mean).* This is a measure of perceived deprivation. It will consist of
+*Perceived resource scarcity (Mean).* This is a measure of perceived deprivation.
+It will consist of
 
@@
 
+Go back to [Overview].
+
 ## Part 4 - Knowledge of Data
@@
 Therefore, we do not know how the adversity measures are associated with inhibition and attention shifting.
+
+Go back to [Overview].
 
@@
 Then, we will fit a second version of this model in which a single common EF factor is estimated, covering both inhibition and attention shifting tasks.
-Using `lavaan::anova`, we will then test whether the second, more constrained model leads to worse model fit, qualified as a significant chi squared change test, and an AIC value > 10. 
+Using `lavaan::anova`, we will then test whether the second, more constrained model leads to worse model fit, qualified as a significant chi squared change test, and an AIC value \> 10.
 If so, we will include submodel 2 in our final model.
@@
 
-A power analysis based on @Kretzschmar_2019 indicated that with an alpha of α = 0.05, we will have \> 90% power to detect small effect sizes (β = 0.1) with a sample size ranging between N = 730 and N = 980 when the reliability of the measures is at least moderate (0.6 -- 0.7).
+A power analysis based on @kretzschmar_2019 indicated that with an alpha of α = 0.05, we will have \> 90% power to detect small effect sizes (β = 0.1) with a sample size ranging between N = 730 and N = 980 when the reliability of the measures is at least moderate (0.6 -- 0.7).
 Therefore, we aim for a total sample size of N = 1,000.
@@
 
-If any assumptions are violated, we will update the preregistration and/or report deviations from the preregistration. However, we do not anticipate any serious violations.
+If any assumptions are violated, we will update the preregistration and/or report deviations from the preregistration.
+However, we do not anticipate any serious violations.
 
@@
 
+Go back to [Overview].
+
 ## Part 6 - Statement of Integrity
@@
 
-<div id="refs"></div>
+::: {#refs}
+:::

### 2024-06-05 preregistration/README.qmd
README files

@@
     template: ../assets/template.html
+    keep_md: true
 bibliography: ../manuscript/references.bib
@@
 
-```{r}
-#| include: false
-
-pagedown::chrome_print(input = "preregistration/README.html", output = "preregistration/preregistration.pdf")
-```
-
-[click here to download a pdf version](https://github.com/Schecher1/Minecraft-Server-Creator/blob/master/README.md)
+[click here to download a pdf version](https://github.com/StefanVermeent/liss_ef_2024/blob/main/preregistration/preregistration.pdf)
 
@@
 These data are openly available after signing an agreement with LISS.
-The dependent measures will be part of a new data collection in the LISS panel which is scheduled between May 2024 and July 2024.
+The dependent measures will be part of a new data collection in the LISS panel which is scheduled in May and June 2024.
 

### 2024-06-05 preregistration/README.qmd
add pdf download link to prereg

@@
   html:
-    css: D:/repositories/liss_ef_2024/assets/style.css
+    css: ../assets/style.css
     template: ../assets/template.html
-    keep_md: true
 bibliography: ../manuscript/references.bib

### 2024-07-16 scripts/02_analyze_data.R
new scripts

@@ -143,7 +143,7 @@ globallocal2_clean <- globallocal2_data |>
   filter(variable != "end", condition != "first") |>
   group_by(id, condition) |>
   mutate(
-    ex_fast = ifelse(rt < 250, TRUE, FALSE),
+    ex_fast = ifelse(rt < 1250, TRUE, FALSE),
     ex_slow = ifelse(scale(rt) |> as.numeric() > 3.2 | rt > 5000, TRUE, FALSE)
   ) |>
   group_by(id) |>

### 2024-07-16 preregistration/README.qmd
qmd and html version of final preregistration

@@
 library(ftExtra)
-library(tinytable)
 
@@
 
-The independent measures used in this study are drawn from studies that were previously conducted in the Longitudinal Internet studies for the Social Sciences (LISS) panel.
-These data are openly available after signing an agreement with LISS.
-The dependent measures will be part of a new data collection in the LISS panel which is scheduled in May and June 2024.
+This study will be based on data collected within the Longitudinal Internet Studies for the Social Sciences (LISS) panel.
+Most of the independent measures will be drawn from previous LISS panel studies, which are openly available in their data archive after signing an agreement with LISS.
+Some independent measures and all dependent measures will be part of a new data collection in the LISS panel which is scheduled in May and June 2024.
 
@@
 
-### Q1: Preliminary title DONE
+### Q1: Preliminary title
 
@@
 
-### Q2: Authors DONE
+### Q2: Authors
 
@@
 -   Anna-Lena Schubert^3^
--   Jean-Louis van Gelder^2^?
 -   Willem E. Frankenhuis^2,4^
@@
 
--   **R1:** How is adversity associated with general speed of processing across all cognitive tasks?
+-   **R1:** How is adversity associated with general speed of processing that is shared across all cognitive tasks?
 
@@
 
--   **R3:** How is adversity associated with task-general and task-specific response caution?
+-   **R3:** How is adversity associated with task-general and domain-specific response caution?
 
@@
 
-Data patterns that would be consistent with hypotheses about abilities generated from a **deficit framework**:
+Results would be consistent with hypotheses generated from a **deficit framework** if:
 
@@
 
-Data patterns that would be consistent with hypotheses about abilities generated from **adaptation frameworks**:
+Results would be consistent with hypotheses generated from **adaptation frameworks** if:
 
@@
 The current study integrates two data sources.
-First, we will sample 1,000 participants in a new LISS study, collected between in May and June 2024 (hereafter referred to as 'newly collected data'), in which we will include measures of neighborhood threat and deprivation, three cognitive tasks measuring attention shifting, two cognitive tasks measuring inhibition, and one cognitive task measuring processing speed.
+First, we will sample 1,000 participants in a new LISS study, collected between May and June 2024 (hereafter referred to as 'newly collected data'), in which we will include measures of neighborhood threat and deprivation, three cognitive tasks measuring attention shifting, two cognitive tasks measuring inhibition, and one cognitive task measuring processing speed.
 Second, we will access data that were previously collected in LISS (hereafter referred to as 'the LISS archive').
@@
 -   Anna-Lena Schubert will not access the data.
--   Jean-Louis van Gelder will not access the data.
 -   Willem E. Frankenhuis will not access the data.
 
-### Q9: Data Collection TBD (fix inclusion criteria)
+### Q9: Data Collection
 
 General information about recruitment of participants into the LISS study can be found at <https://www.lissdata.nl/methodology>.
-The inclusion criteria for the new data collection are that people are between 20 and 55 years old and have completed at least 1 wave of the Conventional and Computer Crime Victimization study (see Q10).
+The inclusion criteria for the new data collection are that people are between 16 and 55 years old and have given permission for their data to be linked to data from the Dutch Central Bureau of Statistics (CBS\; not relevant for this study).
 
@@
 
+#### Independent Variables
+
+The IVs included in this study are identical to a subset of IVs used in a previous project using LISS data (see Q17 and Q18).
+Each IV consists of several indicators.
+If all indicators correlate .60 or higher, we will compute a uniformly weighted average.
+If indicators correlate less than .60, we will use Principal Component Analysis (PCA) to extract only the first component.
+
+*Material deprivation.* Consists of three measures taken from the LISS archive, measured yearly between 2008 and 2023 ([https://doi.org/10.57990/1gr4-bf42](https://doi.org/10.57990/1gr4-bf42)).
+First, participants answered how hard or easy it currently is to live off the income of their household, on a scale from 0 (very hard) to 10 (very easy) (reverse-coded).
+Second, participants indicated which of the following best applied to their current situation: (1) "we are accumulating debt"; (2) "we are somewhat eating into savings"; (3) "we are just managing to make ends meet"; (4) "we have a little bit of money to spare"; (5) "we have a lot of money to spare" (reverse-coded).
+Third, participants indicated which of the following issues they were currently confronted with (0 = no, 1 = yes): (1) "having trouble making ends meet"; (2) unable to quickly replace things that break"; (3) "having to lend money for necessary expenditures"; (4) "running behind in paying rent/mortgage or general utilities"; (5) "debt collector/bailiff at the door in the last month"; (6) "received financial support from family or friends in the last month".
+
+We will separately average each item across time, before combining them into a single composite.
+
+*Threat.* Consists of two measures taken from the LISS archive, measured across six waves between 2008 and 2018 ([https://doi.org/10.17026/dans-zch-j8xt](https://doi.org/10.17026/dans-zch-j8xt)), and one measure included in the newly collected data.
+
+First, participants answered four questions about the safety of their neighborhood (e.g., "[How often do you] avoid certain areas in your place of residence because you perceive them as unsafe?"), on a scale of 1 ("(Almost) never), 2 ("Sometimes"), or 3 ("Often").
+We will sum the responses within each non-missing wave, and calculate an average across waves. 
+Second, participants indicated whether they had been the victim of eight types of crime in the two years prior (0 = no, 1 = yes).
+We will include seven items concerning exposure to crime: (1) burglary or attempted burglary; (2) theft from their car; (3) theft of their wallet or purse, handbag, or other personal possession; (4) wreckage of their car or other private property; (5) intimidation by any other means; (6) maltreatment of such serious nature that it required medical attention; (7) maltreatment that did not require medical attention.
+We will compute a variety score by summing the exposures to *unique* types of crime across all waves.
+If a participant reports exposure to the same crime multiple times, this will count as one unique exposure [@sweeten_2012]..
+
+Third, we included the Neighborhood Violence Scale [NVS\; @frankenhuis_2018; @frankenhuis_2020] in the newly collected data, consisting of seven items measuring perceived exposure to neighborhood violence (e.g., "Crime is common in the neighborhood where I live"; "Where I live, it is important to be able to defend yourself against physical harm").
+Participants answered these questions on a scale of 1 ("Completely disagree") to 7 ("Completely agree").
+We will compute an average of the seven items.
+
 #### Potential confounders
 
-We used Directed Acyclic Graphs (DAG) to identify potential confounders of the empirical estimands (i.e., the association between self-reported threat and deprivation with cognitive outcomes).
+We used Directed Acyclic Graphs (DAG) to identify potential confounders of the key estimands (i.e., the association between self-reported threat and deprivation with cognitive outcomes).
+A DAG is a visual overview of our assumptions about how variables are causally related.
+They are graphs consisting of nodes (variables) and directed arrows (causal pathways).
+An arrow between two variables represents the assumption that experimentally manipulating the variable at the origin of the arrow will change the variable at the end of the arrow (but not the other way around).
+DAGs help in the identification of variables that need to be adjusted for in the statistical models (i.e., confounders with arrows to both the main predictor and the outcome), and, importantly, which variables should not be adjusted for (i.e., colliders and mediators).
 For more information on DAGs, see @rohrer_2018.
 
-We analyzed the following potential confounders both conceptually (i.e., based on previous literature and/or our own assumptions) and, in cases of doubt, statistically (see below): (1) age [@salthouse_2016; @salthouse_2019; @starns_2010], (2) biological sex [@ning_2023], (3) childhood adversity exposure (material deprivation and threat combined) [@bos_2009; @goodman_2019], and (4) recent material deprivation as a confounder for the effect of threat on cognitive performance [@bywaters_2016; @ning_2023].
+We constructed separate DAGs for threat and deprivation.
+As potential confounders of the effect of threat and deprivation on cognitive performance, we examined (1) age [@salthouse_2016; @salthouse_2019; @starns_2010], (2) education [@hofmarcher_2021], (3) sex [@ning_2023], and (4) childhood adversity exposure (material deprivation and threat combined) [@bos_2009; @goodman_2019].
+We also examined evidence for potential causal associations between threat and deprivation (more information below).
+Examinations were based on previous literature and/or our own assumptions.
+
+In the case of threat exposure, we were theoretically less sure about the potential effects of age and education.
+We also wanted to gather more empirical support for the assumption that material deprivation precedes threat exposure [@bywaters_2016; @lacey_2022; @ning_2023].
+Therefore, aside from the theoretical considerations outlined above, we also statistically tested support for parts of the DAG for threat involving age, education, and deprivation using data from a previous LISS study (see Q17 and Q18), using working memory capacity as the outcome variable in the DAG.
+Specifically, we used the `dagitty::localTests()` function to establish whether DAG-implied independencies between certain key variables (conditional on the other causal pathways) were supported by these data [@ankan_2021; @textor_2016].
+
+We removed the causal pathway between a single variable (e.g., age) and threat exposure from the DAG, and then tested whether this imposed independence (conditional on the other causal effects in the model) was consistent with the data.
+If the test was statistically significant, there was a non-zero correlation between the two variables, even after accounting for other variables related to both.
+As most variables were not normally distributed, we used LOESS (locally estimated scatterplot) regression with 5,000 bootstrap samples.
+The models did not include sex, as a combination of categorical and continuous predictors would not allow us to account for non-normally distributed data.
+Code for these analyses can be found at [https://github.com/StefanVermeent/liss_ef_2024/tree/main/scripts/00_dag.R](https://github.com/StefanVermeent/liss_ef_2024/tree/main/scripts/00_dag.R).
 
 **Estimand 1: Material deprivation -\> Cognitive performance.** See Figure 2A for a visualization of the DAG for material deprivation.
-All models involving material deprivation as the main predictor will control for these four covariates.
-In line with recent work, we consider material deprivation to precede threat exposure [@bywaters_2016; @ning_2023].
-For this reason, we do not include threat exposure in models where we are interested in the total effect of material deprivation on cognitive performance.
+On theoretical grounds, we will control for all four variables in models involving material deprivation as the main predictor.  
+Some recent work suggests that material deprivation precedes exposure to adversities such as threat [@bywaters_2016; @lacey_2022; @ning_2023].
+If so, threat exposure is a mediator of the effect of material deprivation on cognitive performance, and should therefore not be controlled for when testing the total effect of material deprivation.
+However, establishing this causal relationship is difficult in our study for two reasons.
+First, The studies listed above all focused on childhood adversity, while we focus on adversity in adulthood.
+Second, our measures of threat and deprivation were measured at the same time, making it difficult to establish causality.
+Therefore, we will present two sets of results for the association between material deprivation and cognitive performance:
+
+(1) The effect of material deprivation WITHOUT including threat exposure, and including age, education, sex, childhood adversity.
+*\*We consider these the primary results, in line with the DAG in Figure 2A\**.
+(2) The effect of material deprivation WITH threat exposure included. *\*We consider this a robustness check to assess the effect of our causal assumption\**.
 
@@
 We considered the same set of confounders as for material deprivation.
-However, in the case of threat exposure, we were conceptually less sure about the potential effects of age and education on threat exposure.
-We also wanted to gather more empirical support for the assumption that material deprivation should be considered as a confounder of the effect of threat exposure on cognitive performance [@bywaters_2016; @ning_2023].
+Table 1 offers an overview of conceptual and statistical evidence for and against each potential confounder in the threat models.
+We found evidence against independence between age and threat, suggesting that older people reported less exposure to threat.
+We also found evidence against independence between material deprivation and threat exposure, conditioned on age and childhood adversity, suggesting that people with more material deprivation also reported more exposure to threat.
+Finally, we did not find evidence against independence between education and threat, suggesting that people with lower education did not report more exposure to threat.
+Therefore, all models involving recent threat exposure as the main predictor will control for age, sex, childhood adversity, and recent material deprivation (but not for education).
 
-Table 1 offers an overview of conceptual and statistical evidence for and against each potential confounder.
-Statistical analysis of potential confounders was based on data from a previous LISS study (see Q17 and Q18) using the `dagitty` R package [@textor_2016].
-Specifically, we used the `dagitty::localTests()` function to establish whether DAG-implied conditional independencies between certain key variables were supported by these data [@ankan_2021].
-The outcome in these conceptual DAGs was working memory capacity, measured on a latent level in the previous LISS study.
-For each of the three confounders under investigation, we removed the causal pathway between that variable and threat exposure from the DAG, and then tested whether this imposed independence (conditional on the other causal effects in the model) was consistent with the data.
-Code for these analyses can be found under `scripts/00_dag`.
+Similar to the deprivation models, we will present two sets of results for the association between threat and cognitive performance.
+
+(1) The effect of threat WITH material deprivation included as confounder. *\*We consider these the primary results, in line with the DAG in Figure 2B\**.
+(2) The effect of threat WITHOUT material deprivation included. *\*We consider this a robustness check to assess the effect of our causal assumption\**.
 
@@
 #| label: Figure1
-#| fig-width: 6.25
-#| fig-height: 7
 #| echo: false
 #| dpi: 600
-#| out-width: 6in
+#| out-width: 450
 #| fig-cap: |
 #|   **Figure 1.** Directed Acyclic Graph (DAG) depicting the empirical estimands (blue lines) and presumed confounding effects (red lines).
-Importantly, we consider deprivation to be a confounder in the effect of threat on cognitive ability, and conversely, we consider threat to be a mediator of the effect of deprivation on cognitive ability [@ning_2023].
+We consider deprivation to be a confounder in the effect of threat on cognitive ability, and conversely, we consider threat to be a mediator of the effect of deprivation on cognitive ability [@bywaters_2016; @lacey_2022; @ning_2023].
 From this, it follows that deprivation should be adjusted for when estimating the effect of threat on cognitive ability.
@@
 
-We found evidence against conditional independence between age and threat.
-We also found evidence against independence between material deprivation and threat exposure, conditional on age and childhood adversity.
-Finally, we did not find evidence against conditional independence between education and threat.
-Therefore, all models involving recent threat exposure as the main predictor will control for age, sex, childhood adversity, and recent material deprivation.
+#### Dependent Variables
 
-#### Independent Variables
+The exact nature and number of dependent variables will depend on the final latent model that we will select (see [Part 5 - Analyses]).
+As a rule, we will estimate associations between each measure of adversity and each latent factor that is estimate.
 
-Our plan for IV computation is identical to that in a previous project using LISS data (see Q17 and Q18).
-While we know how the different subscales correlated with each other in that project, the current project will consist of a (partially) different subsample of the LISS panel.
-While we expect that IV computation will be identical to that project, we leave open the possibility that correlations could fall out differently, thus requiring different approaches.
+We will account for two potential sources of measurement error: (1) environmental noise and (2) state anxiety.
+Both measures are repeated after each cognitive task.
+We measure environmental noise with a single item, rated on a scale of 1 to 5: "How much noise was there in your environment during the reaction game?"
+We measure state anxiety with the shortened version of the State-Trait Anxiety Inventory [STAI\; @bij_2003; @marteau_1992], which asks participants how calm, tense, upset, relaxed, content, and worried they currently feel, on a scale of 1 ("not al all") to 4 ("very much").
+We will recode (when necessary) and then sum the answers so that higher scores reflect more state anxiety.
+For each task, we will then calculate the deviation from the intraindividual mean anxiety score.
+We will account for noise and anxiety by residualizing them out of all the manifest drift diffusion estimates (e.g., for incongruent drift rate on the Flanker trial: `lm(flanker_v_inc ~ noise_flanker + stai_diff_flanker) |> residuals()`).
 
-*Perceived resource scarcity (Mean).* This is a measure of perceived deprivation.
-It will consist of
-
-*Perceived resource scarcity (Coefficient of variation).* This is a measure of unpredictability
-
-*Income-to-needs ratio (Mean).* This is a measure of objective deprivation.
-
-*Income-to-needs ration (Coefficient of variation).* This is a measure of unpredictability
-
-*Neighborhood threat.* This is a measure of neighborhood threat.
-
-#### Dependent Variables DONE
-
-The exact nature and number of dependent variables will depend on the final latent model that will be selected (see Part 5, Q19).
-As a rule, we will estimate associations between each measure of adversity and each latent factor that is estimated.
-
-### Q14: Missing data DONE
+### Q14: Missing data
 
@@
 Finally, if a participant performs at chance level on a task, we will exclude task data of that particular task only.
-Chance level will be determined using the accuracy rate at the 97.5% tail of a binomial distribution if a participant would be purely guessing.
+We will determine chance level using the accuracy rate at the 97.5% tail of a binomial distribution if a participant would be purely guessing.
 
-### Q16: Sample Weights DONE
+### Q16: Sample Weights
 
@@
 
-### Q17: Relevant Publications DONE
+### Q17: Relevant Publications
 
@@
 
-From this study, we know that none of the adversity measures was significantly associated with working memory capacity or working memory updating.
+From this study, we know that none of the adversity measures were significantly associated with working memory capacity or working memory updating.
 The inhibition and attention shifting data have not yet been collected at the time of writing this preregistration.
@@
 
-### Q19: Hypotheses -\> Statistical Tests DONE
+### Q19: Hypotheses -\> Statistical Tests
 
@@
 
-For all other tasks, we will estimate drift rate and non-decision time separately for each task condition (e.g., incongruent and congruent), while keeping boundary separation equal across conditions:
+For all other tasks, we will estimate drift rate, boundary separation, and non-decision time separately for each task condition (e.g., incongruent and congruent):
 
@@
   for (t in 1:nTrials) {
-    y[t] ~ dwiener(alpha[subject[t]], 
+    y[t] ~ dwiener(alpha[condition[t], subject[t]], 
                    tau[condition[t], subject[t]], 
@@
       delta[c, s] ~ dnorm(muDelta[c] , precDelta) T(-10, 10)
+      alpha[c, s] ~ dnorm(muAlpha[c], precAlpha) T(.1, 5)
     }
-    alpha[s]  ~ dnorm(muAlpha, precAlpha) T(.1, 5)
+
   }
@@
     muDelta[c] ~ dunif(-10, 10)
+    muAlpha[c] ~ dunif(.1, 5) 
   }
-  muAlpha~ dunif(.1, 5)
 
@@
 
-Each model will be fitted using three Markov Chain Monte Carlo (MCMC) chains.
+We will fit each model using three Markov Chain Monte Carlo (MCMC) chains.
 Each chain will start with 2,000 burn-in samples, followed by 10,000 additional samples.
-Of these 10,000 samples, every 10th sample will be retained to limit autocorrelation, resulting in a total of 3,000 posterior samples across all chains.
+Of these 10,000 samples, we will retain every 10th sample to limit autocorrelation, resulting in a total of 3,000 posterior samples across all chains.
 
-Model fit will be assessed in three ways.
+We will assess model fit in three ways.
 First, we will visually inspect the traces of the group-level parameter estimates, which should not contain large jumps or drifts.
@@
 A latent Attention Shifting Task will load on the manifest drift rates of switch trials of the Color-shape, Global-local, and Animacy-size Task.
+Finally, we will specify residual covariances between indicators of the same task (e.g., congruent and incongruent conditions of the Flanker Task).
 
@@
 #| label: Figure2
-#| fig-width: 6.25
-#| fig-height: 7
 #| echo: false
 #| dpi: 600
-#| out-width: 6in
+#| out-width: 450
 #| fig-cap: |
 #|   **Figure 2.** Overview of the drift rate models.
-Blue ovals represent latent factors that will be use as dependent variables.
+Blue ovals represent latent factors that we will use as dependent variables.
 Rectangles represent manifest variables.
@@
 Using `lavaan::anova`, we will then test whether the second, more constrained model leads to worse model fit, qualified as a significant chi squared change test, and an AIC value \> 10.
-If so, we will include submodel 2 in our final model.
+If so, we will include sub-model 2 in our final model.
 
@@
 **Boundary separation sub-model.** Like the drift rate model, the baseline boundary separation model will be a bi-factor model containing a general boundary separation factor, as well as inhibition- and attention shifting-specific factors (see Figure 3A).
-We will use the same procedure to choose between the baseline model, and a second model estimating a common boundary separation factor.
+We will use the same procedure to choose between the baseline model and a second model estimating a common boundary separation factor.
 
@@
 #| label: Figure3
-#| fig-width: 6.25
-#| fig-height: 7
 #| dpi: 600
 #| echo: false
-#| out-width: 6in
+#| out-width: 450
 #| fig-cap: |
@@
 
-**Non-decision time sub-model.** The baseline model for the non-decision time, as well as the model selectin procedure, will be identical to the drift rate and boundary separation models (see Figure 3A).
+**Non-decision time sub-model.** The baseline model for the non-decision time, as well as the model selection procedure, will be identical to the drift rate and boundary separation models (see Figure 3A).
 
@@
 #| label: Figure4
-#| fig-width: 6.25
-#| fig-height: 7
 #| dpi: 600
 #| echo: false
-#| out-width: 6in
+#| out-width: 450
 #| fig-cap: |
@@
 **Full measurement model.** After determining the final sub-model for all three DDM parameters, we will combine them all into a single model.
-We expect that we will have to estimate covariances between latent factors of different DDM parameters that share at least one manifest variable (e.g., inhibition drift rates and general boundary separation).
-However, we foresee that we might have to add or remove covariances to improve model fit.
-These steps will be taken before accessing and adding the independent variables.
+We will specify residual covariances between the task-general latent factors (processing speed, general caution, and general non-decision time), as well as between the domain-specific latent factors (separately for inhibition and attention shifting, unless a common factors is favored instead).
 
-**Final model.** After finalizing the full measurement model, we will add the adversity measures as predictors to the model.
+**Final model.** After finalizing the full measurement model, we will construct two versions of the final model: one to estimate the association between material deprivation and DDM parameters, and one to estimate the association between threat exposure and DDM parameters (see [Potential confounders] for more information on the set of predictors that we will include to each model]).
 Each adversity measure will be regressed on each latent factor.
-We will estimate covariances between each pair of adversity measures.
+We will not report regression coefficients of control variables to prevent their over-interpretation.
 
-### Q20: Predicted effect sizes DONE
+### Q20: Predicted effect sizes
 
-We do not have specific predictions for effect sizes but we deem standardized regression coefficients for interaction effects = .10 (or higher) and -.10 (or lower) meaningful.
-These effects would indicate that an effect of adversity on a a particular aspect of performance score is meaningfully more negative or more positive.
+We do not have specific predictions for effect sizes but we deem standardized regression coefficients for effects .10 (or higher) and -.10 (or lower) as practically meaningful.
+Effects between -.10 and .10 are not of interest for determining differences between adversity and aspects of performance.
+However, we are interested in determining if effects falling between -.10 and .10 could be considered to be practically equivalent to zero.
+We will test this using Two One-Sided T-Tests (TOST) equivalence testing, using -.10 and .10 as bounds.
 
-Effects between -.10 and .10 are not of interest for determining differences between adversity and aspects of performance.
-However, we are interested in determining if simple effects falling between -.10 and .10 are consistent with an effect of 0.
-For this scenario, we will use Two One-Sided T-tests (TOST) equivalence testing with -.10 and .10 as bounds.
-
-### Q21: Statistical Power DONE
+### Q21: Statistical Power
 
@@
 
-### Q22: Inferential Criteria DONE
+### Q22: Inferential Criteria
 
-We will use an alpha level of .05 throughout.
-p-values will be corrected for multiple testing using the false discovery rate.
+We will use an alpha level of .05 throughout, and correct p-values for multiple testing using the false discovery rate.
 We will do so separately for tests involving drift rates, non-decision times, and boundary separations, as we have different hypotheses for each of these parameters.
@@
 Impaired cognitive ability is defined as a *negative* association between adversity and drift rate.
-intact ability is defined as an association that is practically equivalent to zero.
-We define practical equivalence as a standardized effect falling between -0.10 and 0.10.
-We will use two one-sided tests (TOST) equivalence testing against these bounds to test for practical equivalence.
+intact ability is defined as an association that is practically equivalent to zero (see Q20).
+Any (practically equivalent) association between adversity and boundary separation or non-decision time will not be interpreted as a difference in cognitive ability, but rather a difference in strategies (in the case of boundary separation) or a difference in preparation/execution speed (in the case of non-decision time).
 
-### Q23: Assumption Violations/Model Non-Convergence DONE
+### Q23: Assumption Violations/Model Non-Convergence
 
-In case of non-normally distributed variables, we will use
+In case of non-normally distributed variables, we will use maximum likelihood with robust (Huber-White) standard errors and a scaled test statistic (using the "MLR" setting in `lavaan`).
+We are likely to violate the assumption of independence, as some participants are nested within households.
+Such clustering will be accounted for in `lavaan` by adding the household ID as a cluster (`cluster = 'nohouse_encr'`).
+We do not anticipate any other serious assumption violations, but will report on them if necessary.
 
-If any assumptions are violated, we will update the preregistration and/or report deviations from the preregistration.
-However, we do not anticipate any serious violations.
+It is possible that some preregistered SEM models will not converge or have suboptimal fit.
+If so, we will use modification indices to re-specify the model.
+We will take these steps before accessing the independent variables.
 
@@
 
-NA
+See [Q19: Hypotheses -\> Statistical Tests] for DDM model fit assessments.
 
-### Q25: Exploratory Analyses DONE
+### Q25: Exploratory Analyses
 

### 2024-09-12 manuscript/manuscript.qmd
intermediate changes to manuscript

@@
 
+source('../scripts/0_functions/general-functions.R')
+
+load("../data/exclusions.RData")
+load("../analysis_objects/results_exploratory.RData")
+
 knitr::opts_chunk$set(
@@
 
-#### Stefan Vermeent^1,2^, Anna-Lena Schubert^3^, Jaap J. A. Denissen^1^, & Willem E. Frankenhuis^1,2,4^
+#### Stefan Vermeent^1,2^, Anna-Lena Schubert^3^, & Willem E. Frankenhuis^2,4^
 
@@
 
-All scripts and materials needed to reproduce the findings are available on the article's Github repository ([https://stefanvermeent.github.io/liss_wm_profiles_2023/](https://stefanvermeent.github.io/liss_wm_profiles_2023/)).
-We also include instructions on how to reproduce each step of our analyses. 
+All scripts and materials needed to reproduce the findings are available on the article's Github repository ([https://stefanvermeent.github.io/liss_ef_2024/](https://stefanvermeent.github.io/liss_ef_2024/)).
 In this paper, we make use of data from the LISS panel (Longitudinal Internet studies for the Social Sciences) managed by the non-profit research institute Centerdata (Tilburg University, the Netherlands).
-All datasets are available in the LISS data archive. 
+All datasets are available in the LISS data archive ([https://www.dataarchive.lissdata.nl/study-units/view/1](https://www.dataarchive.lissdata.nl/study-units/view/1)). 
 Researchers who want to access the data are required to sign a statement confirming that information about individual persons, households, etc., will not be released to others (go to [https://statements.centerdata.nl](https://statements.centerdata.nl) for more information).
@@
 # Funding statement
-This research was conducted in whole or in part using ODISSEI, the Open Data Infrastructure for Social Science and Economic Innovations(<https://ror.org/03m8v6t10>).
+This research was conducted in part using ODISSEI, the Open Data Infrastructure for Social Science and Economic Innovations (<https://ror.org/03m8v6t10>).
 WEF’s contributions have been supported by the Dutch Research Council (V1.Vidi.195.130) and the James S. McDonnell Foundation (https://doi.org/10.37717/220020502).
@@
 
-Exposure to adversity (i.e., prolonged exposure to intense stress) is associated with robust, persisting changes in executive functioning (EF) [@ellis_2022; @farah_2006; @sheridan_2022; @sheridan_2014].
-Given the importance of EF---a set of abilities involved in the top-down, goal-directed regulation of behavior [@miyake_2000; @rey_mermet_2019]---much research has been devoted to understanding which EF abilities are most strongly associated with adversity.
-In some cases, these insights have already been used to screen for early EF deficits, and to develop interventions to help bridge performance gaps [@distefano_2021].
+People who live in adverse conditions tend to show lowered performance on executive functioning tasks [xxx].
+For example, exposure to deprivation has been linked to slower performance on inhibition tasks [@farah_2006; @fields_2021; @mezzacappa_2004; @noble_2005].
+This is typically interpreted as reflecting an impaired ability to maintain attentional focus on a central goal while ignoring distractions.
+These findings fit into a broader deficit framework, which emphasizes the ways in which adversity may hamper cognitive development across many different domains [xxx].
 
-A lot of what we know about the association between adversity and EF stems from analyzing raw performance, most often response times (RT) or accuracy rates.
-In addition, it is common for studies to only include a single task per EF ability.
-For example, previous studies have inferred difficulties in ignoring distractions (inhibition) in people from adversity using a single inhibition task [e.g., @farah_2006; @fields_2021; @mezzacappa_2004; @noble_2005].
-The same thing is true for observed adaptive benefits in the ability to shift attention between tasks (attention shifting) [@fields_2021; @howard_2020; @nweze_2021; @young_2022; @mittal_2015].
-In effect, an important assumption underlying many deficit and adaptation findings is that raw task performance offers a reliable, relatively pure estimate of the EF ability [@white_2022].
+In contrast, adaptation frameworks posit that people from adversity may develop specialized cognitive abilities through developmental adaptations in response to environmental challenges [@ellis_2022; @frankenhuis_2013; @frankenhuis_2020].
+This may result in enhanced or intact performance in specific EF abilities in relation to specific types of adversity.
+Consistent with these predictions, some studies have found intact or even enhanced performance in attention shifting, which is the ability to dynamically switch attention between two tasks [@fields_2021; @howard_2020; @nweze_2021; @young_2022; @mittal_2015].
+Thus, an important aim of both frameworks is to understand how adversity exposure is associated with differences in specific EF abilities.
 
-Unfortunately, an increasing number of psychometric investigations into EF do not support this assumption [@krumm_2009; @loffler_2024; @rey_mermet_2019; @rouder_2019; @stahl_2014].
-A central issue is that raw task performance is not task-pure.
-Aside from the EF ability needed to solve the task, a person's RT is also influenced by other factors such as response caution (i.e., their relative emphasis on speed versus accuracy).
-Accounting for these factors using cognitive modeling, a recent study found that children with more household threat were slower on EF tasks partly because they were more cautious [@vermeent_2024a].
+However, both deficit and adaptation interpretations suffer from a *performance-ability gap*.
+That is, lower task performance (e.g., slower response speed, lower accuracy) does not necessarily reflect lower cognitive ability.
+There are two reasons for this.
+First, raw performance *within* individual EF tasks is influenced by several processes besides the ability of interest, such as response caution and the speed of response execution [@hedge_2021; @ratcliff_2008; @stahl_2014].
+Second, raw performance *across* EF tasks is influenced by general processes, such as basic speed of processing [@loffler_2024].
+It therefore remains an open question to what extent adversity is associated with specific EF abilities after accounting for other sources of individual differences.
+In this study, we use cognitive modeling to more precisely estimate associations between adversity and two EF abilities: inhibition and attention shifting.
 
-Aside from different processing stages, performance across EF tasks is also influenced by general processing speed [@frischkorn_2019; @lerche_2020; @loffler_2024].
+## Separating EF ability from other decision-making processes
 
-[Might not be any ]
-Even more problematic are recent studies suggesting that performance on EF tasks taps into little more than general processing speed.
-For example, after accounting for other factors such as response caution, another study found that a common (latent) EF factor was fully explained by performance on really basic processing speed tasks [@loffler_2024].
-A few studies have focused on this distinction in the context of adversity and/or socioeconomical status.
-on a raw performance level, lower SES was found to be negatively correlated with a task-general factor, with additional associations on the task-specific level.
-In another
+Although mean RTs are routinely used as a direct proxy for cognitive ability, it is well-known that they confound multiple stages of processing, from initial preparations, to the active processing of task-relevant information, to deciding on and executing a response [@ratcliff_1978; @forstmann_2016; @wagenmakers_2009].
+In addition, people differ in how cautiously they make decisions.
+More cautious people tend to respond more slowly, but this does not necessarily indicate a lower ability.
+Differentiating between response caution and EF abilities on the basis of raw performance is difficult, especially because the trade-off between RTs and accuracy is often subtle [@domingue_2022].
+Thus, investigating how adversity is associated with inhibition and attention shifting requires isolating the factors that contribute to differences in RTs.
 
-# # Implications and potential solutions for adversity research
+One way of doing this is through the use of cognitive models like the Drift Diffusion Model [DDM\; @forstmann_2016; @ratcliff_1998; @ratcliff_2008; @wagenmakers_2009].
+The DDM assumes that people accumulate information for one of two responses (e.g., left or right button), until they have enough evidence for one response over the other (see Figure 1 for a visual overview of the DDM).
+When applied to trial-level RTs and accuracies, the DDM estimates four parameters that map onto theoretically meaningful cognitive processes.
+The *drift rate* reflects the rate at which the evidence accumulation process drifts towards the decision boundary, and thus provides a measure of the speed/efficiency of information processing.
+A higher drift rate leads to faster responses and higher accuracy (all else being equal).
+It is generally assumed that individual differences in EF abilities are captured in the drift rate (although not exclusively, see below\; @loffler_2024; @vermeent_2024a; @weigard_2021).
+The *boundary separation* reflects the width between the two decision boundaries, and provides a measure of a person's response caution.
+The *non-decision time* reflects a combination of preparation time (e.g., stimulus encoding) as well as execution time (e.g., time spent pressing the button).
+Finally, the *starting-point* (not considered here) allows for initial bias for one of two responses.
 
-An important implication of this work is that standard assessments---based on raw performance of single EF tasks---may be ill-suited at estimating associations between adversity and specific EF abilities.
-Patterns that may look like deficits (or adaptations) in specific abilities may not be driven by these abilities at all.
-For example, if adversity is associated with slower task-general processing speed, analyses focused on individual EF tasks could make it seem like many different abilities impaired, instead of a single, general process.
+# Accounting for task-impurity
 
-Fortunately, these challenges can be tackled using a combination of cognitive modeling and structural equation modeling.
-As a first step, ability-relevant processing should be separated from ability-irrelevant processes, such as response caution or the time taken to prepare for a task [...].
-One effective way of doing this is through the use of cognitive models like the Drift Diffusion Model [DDM\; @forstmann_2016; @ratcliff_1998; @ratcliff_2008; @wagenmakers_2009].
-The DDM broadly divides the time taken on speeded EF tasks into a decision-making stage and a non-decision-making stage [see Figure 1].
-In the decision-making stage, people accumulate evidence for one of two responses (e.g., left or right button), until they have enough evidence for one response over the other.
-The amount of evidence a person requires can vary between people (as well as between tasks within the same person) [XXX], and is modeled as the distance between the decision boundaries for each respective response.
-Finally, the DDM models two non-decision stages: one at the start of the trial (e.g., task preparations, stimulus encoding) and one at the end of the trial (i.e., response execution).
+The EF literature has a long history of accounting for task impurity by using structural equation modeling (SEM) [@miyake_2000].
+Specific features of EF tasks may give rise to individual differences that do not purely reflect the EF ability.
+For example, content differences (e.g., numeric vs. figural, abstract vs. concrete) have been shown to affect cognitive performance (e.g., @duquennois_2022; @lerche_2020; @young_2022).
+SEM enables separating such method variance and variance associated with non-executive abilities from domain-specific variance (e.g., inhibition, shifting) (@miyake_2012).
+This requires including at least two or more different tasks that measure the same EF ability.
+The most widely used latent factor model of EF distinguishes between inhibition, attention shifting, and working memory updating, although several studies were unable to recover this factor structure, especially with regard to inhibition as a unitary and separate factor [@karr_2018; @rey_mermet_2018; @stahl_2014; @vonbastian_2020].
 
-When applied to empirical performance data, the DDM distinguishes between three parameters mapping onto distinct cognitive processes.
-The *drift rate* reflects the rate at which the evidence accumulation process drifts towards the decision boundary, and thus provides a measure of the speed/efficiency of information processing.
-The *boundary separation* reflects the width between the two decision boundaries, and provides a measure of a person's response caution.
-Finally, the *non-decision time* reflects a combination of preparation time (e.g., stimulus encoding) as well as execution time (e.g., time spent pressing the button).
+Studies that combine SEM with cognitive modeling have found that shared variance across EF tasks largely reflects processing speed.
+One study was able to replicate the three-factor structure of EF when using raw performance measures.
+However, when modeling drift rates, all measures loaded on a single latent factor, which was fully explained by processing speed [@loffler_2024].
+Another investigation using an extension of the DDM for inhibition tasks found that shared variance across nine different tasks mostly reflected processing speed and strategy differences, not inhibition ability [@hedge_2021].
 
-Researchers have argued that the drift rate on EF tasks is a more valid measure of EF ability, at least to the extent to which the tasks are reliable measures of EF ability [@loffler_2024; @vermeent_2024a; @weigard_2021].
-However, drift rates are still influenced at least partly by general speed of processing.
-Theoretically, this should be the part of the variance that is shared across drift rates of all tasks, and of conditions both including an EF component (e.g., incongruent trials) and excluding an EF component (e.g., congruent trials) [@loffler_2024].
-This part of the variance can be estimated on a latent level using SEM.
-Any remaining task-specific variance then contains EF-specific variance.
-
-As mentioned above, a recent investigation did not find significant associations between adversity and task-specific EF variances [@vermeent_2024a].
-This could mean one of two things.
-First, task-specific variance may reliably reflect specific EF abilities, and adversity may not be associated with these abilities.
-Second, task-specific variance may *not* reliably reflect specific EF abilities.
-That is, after accounting for general speed of processing, little more than measurement error may remain.
-Testing these hypotheses requires sampling two or more tasks that are supposed to measure the same ability, and using latent modeling to investigate whether they form a consistent latent factor after accounting for general speed of processing.
+This poses a challenge for adversity research, which often investigates associations between adversity and EF abilities based on a small set of tasks.
+Two recent studies suggest that this may overestimate the extent to which adversity lowers specific EF abilities.
+First, socioeconomic status (which is correlated with, but not the same as adversity) was mostly negatively associated with a task-general factor based on raw performance, while task-specific associations for the most part did not differ meaningfully from zero [@bignardi_2024].
+Second, household threat (but not material deprivation) was negatively associated with task-general drift rates, but not with task-specific variance on three EF tasks [@vermeent_2024a].
+However, both studies used data sets that only included a single task per ability, which may have contributed to the lack of task-specific associations.
+In this study, we built on this work to investigate how adversity is associated with differences in task-general speed, shifting and inhibition ability, and response caution.
 
 # The current study
+
+We analyzed performance on two inhibition tasks, three attention shifting tasks, and one basic processing speed task in relation to adversity exposure in a sample of adults.
+Using a combination of DDM and SEM, our aim was to answer three central questions.
+First, how is adversity associated with general speed of processing that is shared across all cognitive tasks?
+Second, how is adversity associated with inhibition and attention shifting ability after accounting for general speed of processing?
+Third, how is adversity associated with a general tendency for response caution (i.e., across all tasks and conditions), as well as with response caution associated with EF performance (i.e., in task conditions including an EF component)?
+We focused on two self-reported types of adversity: material deprivation (a lack of access to material resources) and threat (the potential for harm imposed by others).
+Both forms of adversity have been associated with both lowered and enhanced performance in EF [e.g., @fields_2021; @schafer_2022; @sheridan_2022; @vermeent_2024a; @young_2022], and are central to recent dimensional models of adversity [@mclaughlin_2021; @sheridan_2014].
+
+We preregistered potential data patterns that we considered consistent with hypotheses generated from deficit and adaptation frameworks.
+Any negative association between adversity and either general speed of processing or domain-specific processing would be consistent with deficit frameworks.
+Associations with response caution would not be consistent with deficit frameworks, as they reflect strategies rather than cognitive abilities.
+Two patterns would be consistent with adaptation frameworks. 
+First, if adversity is positively associated with attention shifting-related processing, regardless of the association with inhibition-related processing.
+Second, if adversity is associated with intact attention shifting-related processing *and* negatively associated with inhibition-related processing.
+If adversity is associated with intact performance in both abilities, this would suggest that adversity exposure is unrelated to the development of these abilities.
 
@@
 
-Participants were 1,000 Dutch people randomly drawn from the Longitudinal Internet studies for the Social Sciences (LISS) panel [@scherpenzeel_2011].
+The study was completed by a total of `r exclusions$sample$finish` participants who were drawn from the Dutch Longitudinal Internet Studies for the Social Sciences (LISS) panel [@scherpenzeel_2011].
 The LISS panel is an invite-only, representative probability sample of the Dutch population consisting of roughly 7,500 individuals across 5,000 households.
@@
 In addition, they can participate in further monthly studies covering a wide range of different topics.
-The current study was fielded between May and June 2024.
+The current study was fielded between May and August 2024.
 Participants were able to complete the cognitive tasks across two or more sessions to increase participation rates.
-People were eligible for participation if they (1) were between 18 and 55 years old; (2) if they had completed at least one wave of a study in the LISS archive assessing exposure to crime (see below); (3) if they had agreed to linking their LISS data to government microdata (not relevant here).
-Participants were excluded if they interacted with other browser tabs mid-trial on any of the cognitive tasks (but not during breaks or in-between tasks). 
+People were eligible for participation if they were between 18 and 55 years old, and if they had agreed to linking their LISS data to government microdata (not relevant here).
 If performance on a specific task was at or below chance level, we excluded the data for that task but retained other task data of that participant.
+The final sample after exclusions consisted of `r exclusions$sample$final` participants.
 
@@
 
-*Perceived neighborhood crime.* Participants completed the Neighborhood Violence Scale [NVS\; @frankenhuis_2018; @frankenhuis_2020] in the current study.
+Participants completed the Neighborhood Violence Scale [NVS\; @frankenhuis_2018; @frankenhuis_2020] in the current study.
 The NVS included seven items about current perceived neighborhood threat (e.g., “Where I live, it is important to be able to defend yourself against physical harm”), on a scale of 1 ("Completely disagree") to 7 ("Completely agree").
@@
 
-*Crime victimization.* We used seven items from the LISS archive (six waves: <https://doi.org/10.17026/dans-zch-j8xt>) in which participants reported whether or not they were the victim of eight types of crime in the last two years: (1) burglary or attempted burglary; (2) theft from their car; (3) theft of their wallet or purse, handbag, or other personal possession; (4) wreckage of their car or other private property; (5) intimidation by any other means; (6) maltreatment of such serious nature that it required medical attention; (7) maltreatment that did not require medical attention.
+We used seven items from the LISS archive (six waves: <https://doi.org/10.17026/dans-zch-j8xt>) in which participants reported whether or not they were the victim of eight types of crime in the last two years: (1) burglary or attempted burglary; (2) theft from their car; (3) theft of their wallet or purse, handbag, or other personal possession; (4) wreckage of their car or other private property; (5) intimidation by any other means; (6) maltreatment of such serious nature that it required medical attention; (7) maltreatment that did not require medical attention.
 As the last wave occurred in 2018, we included the same items in the current study asking about people's crime exposure in the last two years.
@@
 
-*Neighborhood threat composite.* Correlations among the two measures of perceived neighborhood crime and crime victimization were low.
+Correlations among the two measures of perceived neighborhood crime and crime victimization were low.
 Therefore, we used PCA to extract the first principal component score.
@@
 
-### Material deprivation
+### Perceived scarcity
 
-*Perceived scarcity (mean).* Measures of perceived scarcity were derived from the LISS archive, using the yearly recurring core study on household and personal income (16 waves: <https://doi.org/10.57990/1gr4-bf42>).
+Measures of perceived scarcity were derived from the LISS archive, using the yearly recurring core study on household and personal income (16 waves: <https://doi.org/10.57990/1gr4-bf42>).
 First, participants reported how difficult it currently is to live off the income of their household, on a scale from 0 (very hard) to 10 (very easy). 
@@
 
-*Income-to-needs (mean).* The income-to-needs ratio was derived from monthly self-reported net household income in the LISS archive (<https://doi.org/10.57990/qn3k-as78>).
-Reported income of zero was set to missing, as this could indicate an unwillingness to report on income rather than a true income of zero.
-If net household income data was missing (or zero) for an entire year, we imputed, if possible, the reported yearly net household income from the core study on income (<https://doi.org/10.57990/1gr4-bf42>), dividing by 12 to obtain monthly estimates.
-We then divided the income by the poverty threshold set by Statistics Netherlands (CBS) [@brakel_2023; CBS, personal communication, December 15, 2023].
-For households with more than three children, we used the threshold for a family with three children, as CBS does not calculate thresholds for families larger than three children.
-For the same reason, income of households with more than two adults was compared to the threshold for households with two adults.
-We computed the average income-to-needs ratio across years, reverse coding it so that higher scores indicated households further below the threshold.
-
-### Unpredictability
-
-*Perceived scarcity (SD/mean).* Unpredictability over time was computed using the coefficient of variation, which is the within-person standard deviation across years divided by the mean across years (@key_2017; @liu_2022; @ugarte_2023; @walasek_2024; @young_2020).
-This measure accounts for the fact that the mean and standard deviation tend to be highly positively correlated.
-Unpredictability in perceived scarcity was derived from the same measures as the mean level of perceived scarcity (see above).
-We first computed the standard deviation for each measure separately across years, and then scaled them.
-We then computed a uniformly weighted average, which was divided by the mean level of perceived scarcity across years.
-
-*Income-to-needs (SD/mean).* Unpredictability over time in the income-to-needs ratio was also computed using the coefficient of variation.
-We first computed the within-person standard deviation of the income-to-needs ratio across years, and then divided it by the mean income-to-needs ratio across years.
-
 ### Cognitive measures
@@
 
-*Flanker Task.* This task measures inhibition of distractor interference [...]
+*Flanker Task.* This task measures inhibition of distractor interference [@eriksen_1974].
 On each trial, participants saw five arrows side-by-side horizontally, pointing either left or right.
@@
 
-*Simon Task.* This task measures inhibition of prepotent responses [...].
+*Simon Task.* This task measures inhibition of prepotent responses [@simon_1969].
 On each trial, participants saw the word "LEFT" or "RIGHT" (printed in Dutch), presented either on the left or right side of the screen.
@@
 
-*Color-Shape Task.* This task measures the ability to shift attention between different task goals [XXX].
+*Color-Shape Task.* This task measures the ability to shift attention between different task goals [@miyake_2004].
 On each trial, participants saw a square or a circle in the center of the screen that was either blue or yellow.
@@
 
-*Animacy-Size Task.* This task measures the ability to shift attention between different task goals [XXX].
-On each trial, participants saw a single noun (in Dutch) referring to an animal or object [adopted from XXX].
+*Animacy-Size Task.* This task measures the ability to shift attention between different task goals [@arrington_2004].
+On each trial, participants saw a single noun (in Dutch) referring to an animal or object [adopted from @braem_2017].
 Depending on the task rule printed above the noun, their task was to classify it based on whether it referred to a living or non-living thing (e.g., wasp vs.
@@
 
-*Global-local Task.* This task measures the ability to shift attention between task goals [XXX]
-Stimuli were adapted from [...].
+*Global-local Task.* This task measures the ability to shift attention between task goals.
+Stimuli were adapted from @huizinga_2010.
 On each trial, participants saw a large square or rectangle which was composed of 16 small squares or rectangles.
@@
 
-We preregistered a Directed Acyclic Graph (DAG) to inform the inclusion of covariates to the model.
-Based on this DAG, we decided to control for age (in years), the quadratic effect of age, education level, and childhood adversity.
-Childhood adversity was the unweighted average across two scales measuring childhood threat and deprivation: (1) the Neighborhood Violence Scale [XXX], which consists of seven items (e.g., "Physical fights were common in the neighborhood where I grew up"), and (2) four items measuring access to housing, food, clothing, and medicine.
+We used Directed Acyclic Graphs (DAG) to identify potential confounders of the key estimands (i.e., the association between self-reported threat and deprivation with cognitive outcomes).
+A DAG is a visual overview of our assumptions about how variables are causally related.
+They are graphs consisting of nodes (variables) and directed arrows (causal pathways).
+An arrow between two variables represents the assumption that experimentally manipulating the variable at the origin of the arrow will change the variable at the end of the arrow (but not the other way around).
+DAGs help in the identification of variables that need to be adjusted for in the statistical models (i.e., confounders with arrows to both the main predictor and the outcome), and, importantly, which variables should not be adjusted for (i.e., colliders and mediators).
+For more information on DAGs, see @rohrer_2018.
+
+We preregistered separate DAGs for threat and deprivation.
+As potential confounders of the effect of threat and deprivation on cognitive performance, we examined (1) age [@salthouse_2016; @salthouse_2019; @starns_2010], (2) education [@hofmarcher_2021], (3) sex [@ning_2023], and (4) childhood adversity exposure (material deprivation and threat combined) [@bos_2009; @goodman_2019].
+We also examined evidence for potential causal associations between threat and deprivation.
+Examinations were based on previous literature and/or our own assumptions.
+
+In the case of threat exposure, we were theoretically less sure about the potential effects of age and education.
+Therefore, we also statistically tested support for parts of the DAG for threat involving age, education, and deprivation using data from a previous LISS study [@vermeent_2024c], using working memory capacity as the outcome variable in the DAG.
+Specifically, we used the *dagitty* R package [@ankan_2021; @textor_2016] to test whether DAG-implied independencies between key variables (conditional on the other causal pathways) were supported by these data (see the preregistration for more details).
+
+Figure 2A depicts the final preregistered DAGs for material deprivation and threat exposure.
+For material deprivation, we decided to control for age, education, sex, and childhood adversity exposure on theoretical grounds.
+For threat exposure, we decided to control for age, sex, and childhood adversity, but *not* for education.
+We did not find theoretical support for an causal effect of education on threat exposure.
+In addition, our statistical analysis based on previously collected data did not provide evidence against independence between education and threat, suggesting that people with lower education did not report more exposure to threat.
+
+Some recent work suggests that material deprivation precedes exposure to adversities such as threat [@bywaters_2016; @lacey_2022; @ning_2023].
+If so, threat exposure is a mediator of the effect of material deprivation on cognitive performance, and should therefore not be controlled for when testing the total effect of material deprivation.
+However, establishing this causal relationship is difficult in our study for two reasons.
+First, The studies listed above all focused on childhood adversity, while we focus on adversity in adulthood.
+Second, our measures of threat and deprivation were measured at the same time, making it difficult to establish causality.
+Therefore, we will present two sets of results for the associations between material deprivation and threat exposure with cognitive performance.
+For material deprivation, our primary model did *not* include threat exposure as a confounder.
+A secondary model *did* include threat exposure, which we considered this a robustness check to assess the effect of our causal assumption.
+For threat exposure, our primary model included material deprivation as a confounder.
+A secondary model did *not* include material deprivation.
 
@@
 For all tasks, we removed trials with RTs < 250ms and trials with RTs more than 3.2 SD above the intra-individual log-transformed mean RT.
+In addition, if participants performed at chance level on a particular task, we excluded the data for that task only.
+The cut-off for chance performance was determined using the accuracy at the 97.5 % tail of a binomial distribution in the case of pure guessing.
 
@@
 The benefit of HDDM over traditional approaches is that the model uses information of the whole sample to inform individual parameter estimation.
-The model will be applied to each task separately.
-Across all tasks, the bias parameter was held constant, as there is no reason for participants to favor one response over the other across trials.
+The HDDM was applied to each task separately.
+Across all tasks, the bias parameter was held constant, as this parameter is uninformative when the decision boundaries are coded as 'correct' and 'incorrect' (as opposed to mapping on the response options).
 For the Posner Task, we estimated a single drift rate, boundary separation, and non-decision time for each participant.
-For all other tasks, drift rates and non-decision times were estimated separately for congruent (repeat) and incongruent (switch) trials.
-Boundary separation was always fixed across conditions, following the assumption that people cannot anticipate the condition of the next trial, and hence cannot adjust their strategy on a trial-by-trial basis.
+For all other tasks, drift rates, boundary separation, and non-decision times were estimated separately for congruent (repeat) and incongruent (switch) trials.
 
@@
 We used 2,000 burn-in samples and 10,000 additional samples, of which every 10th sample was retained, resulting in a total of 3,000 posterior samples.
-See the supplemental materials for more information about model convergence and fit assessment.
+See the supplemental materials for more information about model convergence and fit assessment.'
+
+We accounted for two potential sources of measurement error in the DDM estimates: (1) environmental noise and (2) state anxiety.
+Both measures were repeated after each cognitive task.
+We measured environmental noise with a single item, rated on a scale of 1 to 5: "How much noise was there in your environment during the reaction game?"
+We measured state anxiety with the shortened version of the State-Trait Anxiety Inventory [STAI\; @bij_2003; @marteau_1992], which asks participants how calm, tense, upset, relaxed, content, and worried they currently felt, on a scale of 1 ("not al all") to 4 ("very much").
+We recoded (when necessary) and then summed the answers with higher scores reflecting more state anxiety.
+For each task, we will then calculate the deviation from the intraindividual mean anxiety score.
+We residualized variance associated with noise and anxiety out of all the drift diffusion estimates using linear regression.
 
 ### SEM
+
+We constructed the final SEM sequentially.
+First, we optimized the fit of the drift rate, boundary separation, and non-decision time sub-models.
+Second, we combined these three models into a single measurement model.
+Third, we added the regression paths between measures of adversity and the latent factors.
+To assess goodness-of-fit, we used the root mean square error of approximation (RMSEA) and the comparative fit index (CFI). 
+CFI values >.0.90 (>0.95) and RMSEA values <0.08 (≤0.06) were interpreted as acceptable (good) fit.
+
+Each sub-model was a bi-factor model, containing as manifest variables the parameter estimates of both conditions for each task.
+Within each sub-model, we estimated covariances between the conditions of each task.
+We estimated a general factor accounting for all manifest variables.
+A second domain-specific inhibition factor loaded on the parameter estimates of incongruent trials of the Flanker and Simon task.
+A third domain-specific Attention Shifting factor loaded on the parameter estimates of switch trials of the Color-shape, Global-local, and Animacy-size task.
+These two factors were allowed to covary.
+For each sub-model, we compared the fit of this model to a version with a common EF factor loading on incongruent/switch trials of all tasks.
+We deemed the second model a better fit when we observed a significant chi squared change test and an AIC value \> 10.
+
+For the drift rate model, we interpreted the general factor as basic speed of processing, and the domain-specific factors as reflecting specific inhibition and attention-shifting ability (or common EF in the case of model 2).
+For the boundary separation/non-decision time models, we interpreted the general factor as general response caution/speed of non-decision processes, and the domain-specific factors as reflecting response caution/speed of non-decision processes specific to conflict trials.
+
+After optimizing each sub-model, we combined them into a single measurement model.
+The general latent factors (processing speed, general caution, and general non-decision time) were allowed to covary, as well as the domain-specific latent factors (separately for inhibition and attention shifting, unless a common factor was favored).
+
+Finally, we constructed two versions of the final model, one to estimate the association between material deprivation and the outcome measures, and one to estimate the association between threat exposure and the outcome measures.
+In both models, the adversity measure was regressed on each latent factor, together with the control variables (see the Control Variables section for more details).
+We do not report regression coefficients of control variables to prevent their over-interpretation.
+
+# Results
+
+## Model fit
+
+### DDM convergence and fit
+
+The DDM converged normally for all tasks except for the Color-shape task, which showed large jumps in the traces.
+We solved this by increasing burnin and the number of posterior samples.
+In addition, model fit was above the preregistered cut-off of *r* = .80 for all tasks (all *r*s > XXX).
+(see the supplemental materials for a full overview of model convergence and fit results).
+
+### SEM
+
+As preregistered, we started with optimizing fit in SEM models for each DDM parameter separately before combining them all into a single model.
+For drift rates, both the model specifying separate inhibition and attention shifting factors (model 1) as well as the model specifying a common EF factor (model 2) provided a good fit to the data (model 1: CFI = `r fit_v1_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_v1_fitstats[['rmsea.robust']] |> round(2)` [`r fit_v1_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_v1_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]; model 2: CFI = `r fit_v2_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_v2_fitstats[['rmsea.robust']] |> round(2)` [`r fit_v2_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_v2_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]).
+Model comparison showed that model 2 led only to a slightly deteriorated fit, which was below the preregistered cut-off (AIC = `r fit_v_compare$AIC |> diff() |> formatC(digits = 2, width = 3, flag = "0", format = 'f')`).
+Therefore, we selected the drift rate model with a general processing speed factor loading on all manifest drift rates and a common EF factor loading on drift rates of conflict trials.
+For boundary separation, both the model specifying separate factors for response caution on conflict trials of inhibition and attention shifting tasks (model 1) as well as the model specifying a common factor for response caution on conflict trials (model 2) provided a good fit to the data (model 1: CFI = `r fit_a1_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_a1_fitstats[['rmsea.robust']] |> round(2)` [`r fit_a1_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_a1_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]; model 2: CFI = `r fit_a2_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_a2_fitstats[['rmsea.robust']] |> round(2)` [`r fit_a2_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_a2_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]).
+Model comparison showed that model 2 led only to a slightly deteriorated fit, which was below the preregistered cut-off (AIC = `r fit_a_compare$AIC |> diff() |> formatC(digits = 2, width = 3, flag = "0", format = 'f')`).
+Therefore, we selected the boundary separation model with a general response caution factor loading on all manifest boundary separations and a common response caution factor loading on boundary separations of conflict trials.
+For non-decision time, the two preregistered models did not converge.
+A third model only including a general non-decision time factor did converge and provided a good fit to the data, CFI = `r fit_t3_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_t3_fitstats[['rmsea.robust']] |> round(2)` [`r fit_t3_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_t3_fitstats[['rmsea.ci.upper.robust']] |> round(2)`].
+The final model combining all three DDM parameters provided a good fit, although it required the addition of covariances between residual variances of manifest DDM parameters of the same task, CFI = `r fit_meas_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_meas_fitstats[['rmsea.robust']] |> round(2)` [`r fit_meas_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_meas_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]
+
+## Preregistered analyses
+
+## Non-preregistered analyses
+
+In our primary models, we found evidence suggesting an (indirect) association between childhood adversity and general speed of processing (note that in our DAGs, the effect of childhood adversity on cognitive processes is mediated by recent adversity).
+To explore potential *direct* effects of childhood adversity on task-general and domain-specific cognitive processes, we fitted an exploratory SEM model in which we evaluated associations between cognitive processes and childhood adversity *without including recent adversity*, and controlling for sex.
+We decided to include childhood exposure to threat and deprivation separately, in contrast to the primary model which included a composite of the two, consistent with the main analyses focusing on recent exposure to threat and deprivation.
+The results are summarised in Figure XX.
+
+For childhood exposure to deprivation, we found a significant negative association with general processing speed, $\beta$ = `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(est.std) |> round(2)`, SE = `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(se) |> round(2)`, 95% CI = [`r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(ci.lower) |> round(2)`, `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(ci.upper) |> round(2)`] *p* `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(pvalue_adj) |> format_p(pvalue = _)`.
+People who reported higher levels of childhood exposure to deprivation processed information more slowly across tasks.
+In addition, we found a significant positive association between childhood exposure to deprivation and non-decision time, $\beta$ = `r expl_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(est.std) |> round(2)`, SE = `r expl_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(se) |> round(2)`, 95% CI = [`r expl_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(ci.lower) |> round(2)`, `r expl_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(ci.upper) |> round(2)`] *p* `r expl_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(pvalue_adj) |> format_p(pvalue = _)`.
+People who reported higher levels of childhood exposure to deprivation took more time encoding stimulus information and/or executing responses across tasks
+We did not find significant associations between childhood exposure to deprivation and general response caution, nor to domain-specific factors.
+Equivalence tests revealed that none of the associations fell within the region of practical equivalence.
+
+For childhood exposure to threat, we found a significant negative association with general processing speed, $\beta$ = `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(est.std) |> round(2)`, SE = `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(se) |> round(2)`, 95% CI = [`r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(ci.lower) |> round(2)`, `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(ci.upper) |> round(2)`] *p* `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(pvalue_adj) |> format_p(pvalue = _)`.
+People who reported higher levels of childhood exposure to threat processed information more slowly across tasks.
+We did not find significant associations between childhood exposure to threat and general response caution or general non-decision time, nor to domain-specific factors.
+Equivalence tests revealed that only the association between childhood exposure to threat and general response caution fell within the region of practical equivalence (*p* `r expl_eqtests |> filter(lhs == "gen_a", rhs == "child_thr") |> pull(eq_pvalue) |> format_p(pvalue = _)`).
+
+# Discussion
+
+The goal of this study was to investigate how exposure to two types of adversity---material deprivation and threat---in adulthood is associated to inhibition and attention shifting ability.
+Participants completed two inhibition tasks, three attention shifting tasks, and one basic processing speed task.
+First, we used DDM to
+
+\pagebreak
+
+## References
+
+::: {#refs}
+:::

### 2024-09-12 manuscript/manuscript.qmd
trying to reestablish git

@@
 
-The goal of this study was to investigate how exposure to two types of adversity---material deprivation and threat---in adulthood is associated to inhibition and attention shifting ability.
+The goal of this study was to XXXinvestigate how exposure to two types of adversity---material deprivation and threat---in adulthood is associated to inhibition and attention shifting ability.
 Participants completed two inhibition tasks, three attention shifting tasks, and one basic processing speed task.

### 2024-09-12 README.qmd
reconnect repo

---
title: ""
format: 
  html:
    css: assets/style.css
    template: assets/template.html
bibliography: manuscript/references.bib
csl: manuscript/apa.csl
link-citations: true
editor: 
  markdown: 
    wrap: sentence
---

```{r}
#| include: false
#| echo: false
#| 
library(tidyverse)
library(gert)
library(glue)
#source('scripts/dependencies.R')
knitr::opts_chunk$set(echo = F, message = F, warning = F)

```
<br>

This repository contains data, code, and output for a manuscript with the working title "Attention shifting and inhibition abilities among people from adverse conditions".

Do you want to download or clone the materials for this project? Go to [https://github.com/stefanvermeent/liss_ef_2024](https://github.com/stefanvermeent/liss_ef_2024)

## Directory Structure {#structure}

The names of each folder are intended to be self-explanatory.
There are eight top-level folders to organize the inputs and outputs of this project:

1.  [`Preregistration`](https://stefanvermeent.github.io/liss_ef_2024/preregistration/README.html): The preregistration for the main analyses.
3.  [`Manuscript`](https://stefanvermeent.github.io/liss_ef_2024/manuscript/README.html): The manuscript written in Quarto.
4.  [`Supplement`](https://stefanvermeent.github.io/liss_ef_2024/supplement/README.html): a supplemental text (to be submitted with the manuscript) documenting all secondary analyses in detail.
5.  [`Scripts`](https://stefanvermeent.github.io/liss_ef_2024/scripts/README.html): R-scripts that read, analyze, and produce all analysis objects.
6.  [`Data`](https://stefanvermeent.github.io/abcd_ddm/data/README.html): Folder in which real LISS data can be placed to make the analyses fully reproducible. Note that we cannot openly share the raw data on the open repository. 
7.  [`Analysis Objects`](https://stefanvermeent.github.io/liss_ef_2024/analysis_objects/README.html): Folder containing all analysis objects.
8.  [`Codebooks`](https://stefanvermeent.github.io/liss_ef_2024/codebooks/README.html): lists of variable names, labels, and value labels (where applicable).

Click on each of the folders to get more details.

## Overview of project milestones

Below is an overview of all the project milestones, such as first-time data access, submissions, and revisions.
Data access events were automatically captured using custom code, which over the course of this project was collected in the R package `projectlog` [https://stefanvermeent.github.io/projectlog/](https://stefanvermeent.github.io/projectlog/).

```{r}
#| results: 'asis'
#| echo: false

x <- gert::git_log(max = 10000) |> 
  select(commit, author, time, message) |>
  # Some custom coding because we switched to using the projectlog package for the final data_access event
  mutate(message = ifelse(str_detect(message, "first-time access to ivs"), str_c("[[data_access]] ", message), message)) |> 
  mutate(message = ifelse(str_detect(message, "Submission of Stage 2 manuscript"), str_c("[[submission]] ", message), message)) |>
  mutate(message = ifelse(str_detect(message, "Stage 2 accepted manuscript at Developmental Science"), str_c("[[accepted]] ", message), message)) |>
  (function(.) filter(., str_detect(string = .$message, pattern = "^\\[\\[")))() |> 
  separate(message, into = c("message", "data_hash", 'code'), sep = "\n") |> 
  separate(message, into = c("milestone", "description"), sep = "\\]\\]\\s") %>% 
  distinct(data_hash, description, .keep_all = TRUE) %>%
  mutate(
    milestone = map_chr(milestone, function(milestone) {
      milestone |> 
        str_remove("\\[\\[") |>
        str_replace("_", " ") |> 
        str_to_title()
    }),
    data_hash = str_remove(data_hash, "^object_hash ")
  ) |> 
  select(time, commit, milestone, description, code, data_hash) |> 
  mutate(
    code = str_replace_all(code, "\\|\\>", "\\|\\>\\\n"),
    link_to_git = paste0(
       gert::git_remote_list()$url |> str_remove(string = _, pattern = "\\.git$"), "/tree/", commit)
    ) |>
  arrange(desc(time)) |> 
  glue_data("  - **[{time}]({link_to_git}): {description}**
        - **Milestone:** {milestone}
        - **Data MD5 hash**: {data_hash}
        - [Link to code snippet](https\\://github.com/StefanVermeent/abcd_ddm/blob/main/.projectlog/{commit}.R)
        \n
    ")
x

```

### Contact

For questions or comments, feel free to contact me at p.c.s.vermeent@gmail.com.

### 2024-09-12 .projectlog/generate_scripts.R

@@ -1,0 +1,34 @@
+gert::git_log(max = 10000) |> 
+  select(commit, author, time, message) |>
+  (function(.) filter(., str_detect(string = .$message, pattern = "^\\[\\[")))() |> 
+  separate(message, into = c("message", "hash", 'code'), sep = "\n") |> 
+  separate(message, into = c("milestone", "description"), sep = "\\]\\]\\s") |> 
+  mutate(milestone = str_remove(milestone, "\\[\\[")) |> 
+  pwalk(function(commit,author,time,message,hash,milestone,description, code){
+    
+    if(milestone == "data_access"){
+      code <- code |> 
+        str_replace_all(string = _, "\\|> ", "\\|>\n") |> 
+        str_remove(string = _, "^code ") 
+      
+      url <- gert::git_remote_list()$url |> 
+        str_remove(string = _, "\\.git$") |> 
+        paste0(paste0("/commit/", commit))
+      
+    #  url <- "<anonymized repository>"
+      
+      script <-
+        glue(
+          "
+          ### Date: {time}\n
+          ### Description: {description}\n\n
+          ### For more information on this commit, see the README file, or go to {url}\n
+          ### Below is the full code that was used to access the data:\n\n
+          {code}"
+        )
+      
+      writeLines(script, con = paste0(".projectlog/", commit, ".R"))
+    }
+  })
+
+

### 2024-09-12 materials/liss/I_flanker/0_flanker_setup.js

@@ -1,0 +1,62 @@
+//------------------------- Functions
+
+var location_stim = function(up, down){
+
+  // Present flanker stimuli either at the top or bottom of the screen
+  var html = "<style>" +
+              "divbottom{display:flex; justify-content: center; align-items: flex-end; height:300px}" +
+              "divtop{display:flex; justify-content: center; align-items: flex-start; height:300px}" +
+             ".topflanker{font-size:50px;}" +
+             ".bottomflanker{font-size:50px;}" +
+             "</style>";
+
+  if(up != null) {
+    html += "<divtop>";
+    html += "<div class = 'topflanker'>";
+    html += up;
+    html += "</div></div>";
+  }
+
+  if(down != null) {
+    html += "<divbottom>";
+    html += "<div class = 'bottomflanker'>";
+    html += down;
+    html += "</div></div>";
+
+  }
+
+  return html;
+};
+
+
+
+//------------------------- Stimuli
+
+  // Flanker stimuli
+  var congruent_left = "&larr;&larr;&larr;&larr;&larr;"
+  var congruent_right = "&rarr;&rarr;&rarr;&rarr;&rarr;"
+  var incongruent_left = "&rarr;&rarr;&larr;&rarr;&rarr;"
+  var incongruent_right = "&larr;&larr;&rarr;&larr;&larr;"
+
+
+  var flanker_stimuli = [
+    {stimulus: congruent_left},
+    {stimulus: congruent_right},
+    {stimulus: incongruent_left},
+    {stimulus: incongruent_right},
+  ]
+
+
+  // Fixation cross
+  var flanker_fixation = {
+  type: jsPsychHtmlKeyboardResponse,
+  stimulus: '<div style="font-size:60px;">+</div>',
+  choices: 'NO_KEYS',
+  trial_duration: 1000,
+  data: {
+    variable: 'fixation',
+    stimulus: ""
+  }
+}
+
+

### 2024-09-12 materials/liss/I_flanker/2_flanker_procedure.js

@@ -1,0 +1,83 @@
+var flanker_present_arrows = {
+  type: jsPsychHtmlKeyboardResponse,
+  stimulus: function(){
+    return jsPsych.timelineVariable('stim')
+  },
+  choices: ['A', 'L'],
+  data: {
+    variable: 'test',
+    task: 'flanker',
+    stimulus: "",
+    location: function(){
+      return jsPsych.timelineVariable('location')
+    },
+    stimtype: function(){
+      return jsPsych.timelineVariable('stimtype')
+    },
+    correct_response: function(){
+      return jsPsych.timelineVariable('correct_response')
+    }
+  },
+  on_finish: function(data) { {
+    if(jsPsych.pluginAPI.compareKeys(data.response, jsPsych.timelineVariable('correct_response', true))) {
+      data.correct = true;
+    } else {
+      data.correct = false;
+    }
+  }
+  }
+};
+
+
+var flanker_test_procedure01 = {
+  timeline: [flanker_fixation, flanker_present_arrows],
+  timeline_variables: [
+    {location: 'top',    correct_response: 'A', stimtype: 'congruent_left',    stim: location_stim(up='&larr;&larr;&larr;&larr;&larr;', down=null)},
+    {location: 'top',    correct_response: 'L', stimtype: 'congruent_right',   stim: location_stim(up='&rarr;&rarr;&rarr;&rarr;&rarr;', down=null)},
+    {location: 'top',    correct_response: 'A', stimtype: 'incongruent_left',  stim: location_stim(up='&rarr;&rarr;&larr;&rarr;&rarr;', down=null)},
+    {location: 'top',    correct_response: 'L', stimtype: 'incongruent_right', stim: location_stim(up='&larr;&larr;&rarr;&larr;&larr;', down=null)},
+    {location: 'bottom', correct_response: 'A', stimtype: 'congruent_left',    stim: location_stim(up=null, down='&larr;&larr;&larr;&larr;&larr;')},
+    {location: 'bottom', correct_response: 'L', stimtype: 'congruent_right',   stim: location_stim(up=null, down='&rarr;&rarr;&rarr;&rarr;&rarr;')},
+    {location: 'bottom', correct_response: 'A', stimtype: 'incongruent_left',  stim: location_stim(up=null, down='&rarr;&rarr;&larr;&rarr;&rarr;')},
[…]

### 2024-09-12 materials/liss/I_simon/0_simon_setup.js

@@ -1,0 +1,58 @@
+//------------------------- Functions
+
+var location_stim = function(left, right){
+
+  // Present flanker stimuli either at the top or bottom of the screen
+  var html = "<style>" +
+              "divleft{text-align: left; width:300px}" +
+              "divright{text-align: right; width:300px}" +
+             ".lefttarget{font-size:80px;}" +
+             ".righttarget{font-size:80px;}" +
+             "</style>";
+
+  if(left != null) {
+    html += "<divleft>";
+    html += "<div class = 'lefttarget'>";
+    html += left;
+    html += "</div></div>";
+  }
+
+  if(left != null) {
+    html += "<divright>";
+    html += "<div class = 'righttarget'>";
+    html += left;
+    html += "</div></div>";
+
+  }
+
+  return html;
+};
+
+
+
+//------------------------- Stimuli
+
+  // Flanker stimuli
+  var left = "LINKS"
+  var right = "RECHTS"
+
+
+  var simon_stimuli = [
+    {stimulus: left},
+    {stimulus: right},
+  ]
+
+
+  // Fixation cross
+  var simon_fixation = {
+  type: jsPsychHtmlKeyboardResponse,
+  stimulus: '<div style="font-size:60px; width: 600px;">+</div>',
+  choices: 'NO_KEYS',
+  trial_duration: 1000,
+  data: {
+    variable: 'fixation',
+    stimulus: ""
+  }
+}
+
+

### 2024-09-12 materials/liss/I_simon/2_simon_procedure.js

@@ -1,0 +1,144 @@
+var simon_target = {
+  type: jsPsychHtmlKeyboardResponse,
+  stimulus: function(){
+    var stim = ""
+
+    if (jsPsych.timelineVariable('location') == 'left'){
+      stim += "<div class='grid-container'>" +
+                "<div style='font-size: 30px;'>" + jsPsych.timelineVariable('stim') + "</div>" +
+                "<div style='font-size: 60px'>+</div>" +
+                "<div style='font-size: 30px;'></div>" +
+              "</div>"
+    }
+    if (jsPsych.timelineVariable('location') == 'right'){
+      stim += "<div class='grid-container'>" +
+                "<div style='font-size: 30px;'></div>" +
+                "<div style='font-size: 60px'>+</div>" +
+                "<div style='font-size: 30px;'>" + jsPsych.timelineVariable('stim') + "</div>" +
+              "</div>"
+    }
+    return stim
+  },
+  choices: ['A', 'L'],
+  data: {
+    variable: 'test',
+    task: 'simon',
+    stimulus: "",
+    location: function(){
+      return jsPsych.timelineVariable('location')
+    },
+    stimtype: function(){
+      return jsPsych.timelineVariable('stimtype')
+    },
+    correct_response: function(){
+      return jsPsych.timelineVariable('correct_response')
+    }
+  },
+  on_finish: function(data) { {
+    if(jsPsych.pluginAPI.compareKeys(data.response, jsPsych.timelineVariable('correct_response', true))) {
+      data.correct = true;
+    } else {
+      data.correct = false;
+    }
+  }
+  }
+};
+
+
+var simon_test_procedure01 = {
+  timeline: [simon_fixation, simon_target],
+  timeline_variables: [
+    {location: 'left', correct_response: 'L', stimtype: 'incongruent', stim: 'RECHTS'},
+    {location: 'right', correct_response: 'A', stimtype: 'incongruent', stim: 'LINKS'},
+    {location: 'left', correct_response: 'A', stimtype: 'congruent', stim: 'LINKS'},
+    {location: 'right', correct_response: 'A', stimtype: 'incongruent', stim: 'LINKS'},
+    {location: 'right', correct_response: 'L', stimtype: 'congruent', stim: 'RECHTS'},
[…]

### 2024-09-12 materials/liss/P_posner/0_posner_setup.js

@@ -1,0 +1,15 @@
+//------------------------- Functions
+
+  // Fixation cross
+  var posner_fixation = {
+  type: jsPsychHtmlKeyboardResponse,
+  stimulus: '<div style="font-size:60px;">+</div>',
+  choices: 'NO_KEYS',
+  trial_duration: 1000,
+  data: {
+    variable: 'fixation',
+    stimulus: ""
+  }
+}
+
+

### 2024-09-12 materials/liss/P_posner/2_posner_procedure.js

@@ -1,0 +1,143 @@
+var posner_letters = {
+  type: jsPsychHtmlKeyboardResponse,
+  stimulus: function(){
+    var stim = '<p style="font-size:80px;font-weight:bold;">'+jsPsych.timelineVariable('stim1')+jsPsych.timelineVariable('stim2')+'</p>'
+    return stim
+  },
+  choices: ['A', 'L'],
+  data: {
+    stimulus: function(){
+      return jsPsych.timelineVariable('stim1') + jsPsych.timelineVariable('stim2')
+    },
+    variable: 'test',
+    task: 'posner',
+    condition: function(){
+      return jsPsych.timelineVariable('condition')
+    },
+    correct_response: function(){
+      return jsPsych.timelineVariable('correct_response')
+    }
+  },
+  on_finish: function(data) { {
+    if(jsPsych.pluginAPI.compareKeys(data.response, jsPsych.timelineVariable('correct_response', true))) {
+      data.correct = true;
+    } else {
+      data.correct = false;
+    }
+  }
+  }
+};
+
+
+var posner_test_procedure01 = {
+  timeline: [posner_fixation, posner_letters],
+  timeline_variables: [
+    {stim1: 'B', stim2: 'q', condition: 'different', correct_response: 'L'},
+    {stim1: 'Q', stim2: 'q', condition: 'same',      correct_response: 'A'},
+    {stim1: 'A', stim2: 'B', condition: 'different', correct_response: 'L'},
+    {stim1: 'q', stim2: 'q', condition: 'same',      correct_response: 'A'},
+    {stim1: 'a', stim2: 'H', condition: 'different', correct_response: 'L'},
+    {stim1: 'F', stim2: 'H', condition: 'different', correct_response: 'L'},
+    {stim1: 'b', stim2: 'f', condition: 'different', correct_response: 'L'},
+    {stim1: 'a', stim2: 'A', condition: 'same',      correct_response: 'A'},
+    {stim1: 'f', stim2: 'B', condition: 'different', correct_response: 'L'},
+    {stim1: 'Q', stim2: 'q', condition: 'same',      correct_response: 'A'},
+    {stim1: 'h', stim2: 'a', condition: 'different', correct_response: 'L'},
+    {stim1: 'A', stim2: 'q', condition: 'different', correct_response: 'L'},
[…]

### 2024-09-12 materials/liss/S_animacysize/0_animacysize_setup.js

@@ -1,0 +1,17 @@
+
+
+var living     = "<img src='img/living.png' height=100>";
+var nonliving  = "<img src='img/nonliving.png'height=100 style='padding-left: 20px;'>";
+var smaller    = "<img src='img/smaller.png' height=100 style='padding-right: 20px;'>";
+var larger     = "<img src='img/larger.png' height=100 style='padding-right: 20px;'>";
+
+//-------------------- Response prompts
+
+var prompt_living  = "<div style='float:right; text-align: right;' >" + living  +  "</div>";
+var prompt_nonliving  = "<div style='float:left;  text-align: left;'  >" + nonliving + "</div>";
+var prompt_smaller = "<div style='float:left;  text-align: left;'  >" + smaller + "</div>";
+var prompt_larger   = "<div style='float:right; text-align: right;' >" + larger + "</div>";
+
+
+
+// stimuli retrieved from https://osf.io/h3e7w/

### 2024-09-12 materials/liss/S_animacysize/2_animacysize_procedure.js

@@ -1,0 +1,122 @@
+
+var animacysize_test = {
+  type: jsPsychCategorizeHtml,
+  choices: ['A','L'],
+  correct_text: "",
+  incorrect_text: "",
+  feedback_duration: 250,
+  show_stim_with_feedback: false,
+  key_answer: function(){return jsPsych.timelineVariable('key_answer')},
+  stimulus: function(){
+    var stim = ""
+    if(jsPsych.timelineVariable('data')['rule'] == "size") {
+      stim += "<div style = 'text-align: center;'><h1>KLEINER of GROTER?<br></h1><br><br><br><div style = 'font-size:60px'>" + jsPsych.timelineVariable('stimulus') + "</div></div>";
+    }
+    if(jsPsych.timelineVariable('data')['rule'] == "animacy") {
+      stim += "<div style = 'text-align: center;'><h1>LEVEND of NIET-LEVEND?<br></h1><br><br><br><div style = 'font-size:60px'>" + jsPsych.timelineVariable('stimulus') + "</div></div>";
+    }
+        return stim
+      },
+  prompt: "<br><br><br><div style='width: 600px; height:50px;'>" + prompt_living + prompt_smaller + prompt_larger + prompt_nonliving + "</div><br><br>" +
+          "<div style='width: 600px;'><h1 style='float: left; margin:0;'>A</h1><h1 style='float: right; margin:0;'>L</h1></div>",
+  data: {
+    stimulus: function(){return jsPsych.timelineVariable('stimulus')},
+    rule: function(){return jsPsych.timelineVariable('data')['rule']},
+    condition: function(){
+      return jsPsych.timelineVariable('data')['type']
+    },
+    variable: function(){return jsPsych.timelineVariable('data')['variable']},
+    task: function(){return jsPsych.timelineVariable('data')['task']},
+  },
+};
+
+var animacysize_procedure01 = {
+  timeline: [animacysize_test],
+  timeline_variables: [
+   {stimulus: 'gorilla', key_answer: 'L', data: {stimulus: 'gorilla', rule: 'size', type: 'first', variable: 'animacysize1', task: 'animacysize'}},
[…]

### 2024-09-12 materials/liss/S_animacysize/scripts/trialmaker.R

@@ -1,0 +1,155 @@
+set.seed(4785)
+
+
+# Stimulus sets -----------------------------------------------------------
+
+lsm_test_words <- c("egel", "goudvis", "kever", "kikker", "mier",
+                    "mug", "muis", "mus", "rat", "rups", "slak", "spin",
+                    "vlinder", "wesp", "worm", "kuiken")
+
+lbi_test_words <- c( "beer", "dolfijn", "ezel", "geit", "hert", "ijsbeer",
+                     "kameel", "leeuw", "olifant", "ooievaar", "orka",
+                     "schaap", "stier", "tijger", "walvis", "wolf")
+
+
+nlsm_test_words <- c("blik", "bril", "kam", "lepel", "mok", "naald", "pen",
+                     "pil", "ring", "schaar", "sok", "speld", "spijker",
+                     "stift", "vork", "zeep")
+
+
+nlbi_test_words <- c("bus", "caravan", "fiets", "iglo",  "kano", "kanon",
+                     "matras",  "lantaarn", "piano", "raket", "saxofoon",
+                     "tent", "toren", "tractor", "vulkaan", "windmolen")
+
+lsm_test_words <- sample(lsm_test_words, size = 16, replace = F)
+lbi_test_words <- sample(lbi_test_words, size = 16, replace = F)
+nlsm_test_words <- sample(nlsm_test_words, size = 16, replace = F)
+nlbi_test_words <- sample(nlbi_test_words, size = 16, replace = F)
+
+# stimulus set 1
+trials01 <- tibble(
+  type     = c(rep('repeat',16), rep('switch',16)),
+  rule     = c(rep(NA, 32)),
+  stimulus = c(lsm_test_words[1:8], lbi_test_words[1:8], nlsm_test_words[1:8], nlbi_test_words[1:8]),
+  correct  = c(rep('living_smaller', 8), rep("living_larger",8), rep("nonliving_smaller",8), rep("nonliving_larger",8)),
+  variable = "animacysize1",
+  task     = "animacysize",
+  key_answer = NA
+)
+
+trials01 <- trials01 |>
+  mutate(
+    valid1 = ifelse(type == lag(type,1) & type == lag(type,2) & type == lag(type,3), FALSE, TRUE),
[…]

### 2024-09-12 materials/liss/S_colorshape/1_colorshape_practice.js

@@ -1,0 +1,231 @@
+
+//------------------------- Obejects to hold trial information for the practice session
+
+// Shape instructions
+
+var colorshape_welcome = {
+  type: jsPsychInstructions,
+  pages: [
+    "Welkom bij het <b>Kleuren en Vormen</b> spel!<br><br><br><br>"
+  ],
+  show_clickable_nav: true,
+  allow_backward: true,
+  key_forward: -1,
+  key_backward: -1,
+  button_label_next: "verder",
+  data: {variable: 'welcome', task: "colorshape_practice"}
+};
+
+var colorshape_shape_instructions = {
+    type: jsPsychInstructions,
+    pages: [
+        "<p style = 'text-align: center;'>We gaan een spel spelen waarin u wisselt tussen kleuren en vormen.<br><br></p>",
+        "<p style = 'text-align: center;'> We zullen eerst het VORMEN-spel spelen.<br><br>" +
+        "In het VORMEN-spel kiest u het plaatje dat dezelfde VORM is als het plaatje in het midden van het scherm.<br><br></p>",
+        "Als het plaatje een DRIEHOEK is, druk dan op de 'A'-toets op uw toetsenbord:<br><br><br>Voorbeeld:"+
+        "<div style = 'text-align: center;'>" + yellowtriangle + "</div>" +
+        "<br><br><br><div>" + prompt_tri + prompt_circle + "</div><br><br><br>",
+        "Als het plaatje een CIRCEL is, druk dan op de 'L'-toets op uw toetsenbord:<br><br><br>Voorbeeld:"+
+        "<div style = 'text-align: center;'>" + bluecircle + "</div>" +
+        "<br><br><br><div>" + prompt_tri + prompt_circle + "</div><br><br><br>",
+        "<p style = 'text-align: center;'> U gaat nu het VORMEN-spel oefenen.<br><br>" +
+        "Antwoord zo snel als u kunt zonder fouten te maken. Af en toe een fout maken is niet erg. Ga in dat geval gewoon door.<br><br>" +
+        "Klik op 'verder' om het VORMEN-spel te oefenen.</p>"
+    ],
+    show_clickable_nav: true,
+    allow_backward: true,
+    key_forward: -1,
+    key_backward: -1,
+    button_label_next: "verder",
+    button_label_previous: "ga terug",
+    data: {variable: "instructions", task: "colorshape_practice"}
+};
+
+// Shape practice
+
[…]

### 2024-09-12 materials/liss/S_colorshape/2_colorshape_procedure.js

@@ -1,0 +1,114 @@
+
+var colorshape_test = {
+  type: jsPsychCategorizeHtml,
+  choices: ['A','L'],
+  correct_text: "",
+  incorrect_text: "",
+  key_answer: function(){return jsPsych.timelineVariable('key_answer')},
+  stimulus: function(){return jsPsych.timelineVariable('stimulus')},
+  feedback_duration: 250,
+  show_stim_with_feedback: false,
+  prompt: "<div style='width: 600px; height:50px;'>" + prompt_yellow + prompt_tri + prompt_circle + prompt_blue + "</div><br><br>" +
+          "<div style='width: 600px;'><h1 style='float: left; margin:0;'>A</h1><h1 style='float: right; margin:0;'>L</h1></div>",
+  data: {
+    stimulus: function(){return jsPsych.timelineVariable('stim_chr')},
+    rule: function(){return jsPsych.timelineVariable('data')['rule']},
+    condition: function(){return jsPsych.timelineVariable('data')['type']},
+    variable: function(){return jsPsych.timelineVariable('data')['variable']},
+    task: function(){return jsPsych.timelineVariable('data')['task']},
+  }
+}
+
+var colorshape_procedure1 = {
+  timeline: [colorshape_test],
+  timeline_variables: [
+    {stimulus: yellowtriangle_color, stim_chr: 'yellowtriangle', key_answer: 'ArrowLeft', data: {rule: 'color', type: 'first', variable: 'colorshape1', task: 'colorshape'}},
+    {stimulus: yellowcircle_shape, stim_chr: 'yellowcircle', key_answer: 'L', data: {rule: 'shape', type: 'switch', variable: 'colorshape1', task: 'colorshape'}},
+    {stimulus: bluecircle_shape, stim_chr: 'bluecircle', key_answer: 'L', data: {rule: 'shape', type: 'repeat', variable: 'colorshape1', task: 'colorshape'}},
+    {stimulus: bluetriangle_color, stim_chr: 'bluetriangle', key_answer: 'L', data: {rule: 'color', type: 'switch', variable: 'colorshape1', task: 'colorshape'}},
+    {stimulus: bluecircle_color, stim_chr: 'bluecircle', key_answer: 'L', data: {rule: 'color', type: 'repeat', variable: 'colorshape1', task: 'colorshape'}},
[…]

### 2024-09-12 materials/liss/S_colorshape/3_colorshape_end.js

@@ -1,0 +1,10 @@
+//------------------------- End of the task and save data
+
+var colorshape_end = {
+  type: jsPsychHtmlButtonResponse,
+  stimulus: "Goed gedaan!<br><br>" +
+  "U bent nu klaar met het Kleuren en Vormen spel.<br><br>" +
+  "Klik op 'verder' om door te gaan.<br><br>",
+  choices: ['verder'],
+  data: {variable: "colorshape_finish", task: "colorshape"},
+};

### 2024-09-12 materials/liss/S_colorshape/scripts/trialmaker.R

@@ -1,0 +1,160 @@
+set.seed(4785)
+
+# Abstract Stimuli ------------------------------------------------------
+
+# stimulus set 1
+trials01 <- tibble(
+  type     = c('first', sample(c(rep('repeat',16), rep('switch',16)), size = 32, replace = F)),
+  rule     = c('color', rep(NA, 32)),
+  variable = "colorshape1",
+  task     = "colorshape",
+  stimulus = sample(c('yellow_circle', 'yellow_triangle', 'blue_circle', 'blue_triangle'), 33, replace = T),
+  key_answer = NA
+)
+
+trials01 <- trials01 |>
+  mutate(
+    valid1 = ifelse(type == lag(type,1) & type == lag(type,2) & type == lag(type,3), FALSE, TRUE),
+    valid2 = ifelse(stimulus == lag(stimulus,1) & stimulus == lag(stimulus,2), FALSE, TRUE)
+  )
+
+n = 1
+while(any(c(trials01$valid1, trials01$valid2) == FALSE, na.rm = T)) {
+  trials01 <- trials01 |>
+    mutate(
+      type = c('first', sample(c(rep('repeat',16), rep('switch',16)), size = 32, replace = F)),
+      stimulus = sample(c('yellow_circle', 'yellow_triangle', 'blue_circle', 'blue_triangle'), 33, replace = T),
+      valid1 = ifelse(type == lag(type,1) & type == lag(type,2) & type == lag(type,3), FALSE, TRUE),
+      valid2 = ifelse(stimulus == lag(stimulus,1) & stimulus == lag(stimulus,2), FALSE, TRUE)
+    )
+}
+
+n = 1
+while(n<=32){
+  trials01 <- trials01 |>
+    mutate(
+      rule = case_when(
+        is.na(rule) & type == "repeat" & lag(rule, n=1) == 'color' ~ 'color',
+        is.na(rule) & type == "repeat" & lag(rule, n=1) == 'shape' ~ 'shape',
+        is.na(rule) & type == "switch" & lag(rule, n=1) == 'color' ~ 'shape',
+        is.na(rule) & type == "switch" & lag(rule, n=1) == 'shape' ~ 'color',
+        TRUE ~ rule
+      )
+    )
+  n = n+1
+}
+
+trials01 <- trials01 |>
+  mutate(
+    key_answer = case_when(
+      rule == 'color' & str_detect(stimulus, "^blue") ~ 'L',
[…]

### 2024-09-12 materials/liss/S_globallocal2/2_globallocal2_procedure.js

@@ -1,0 +1,149 @@
+// Fixation cross
+var globallocal2_fixation = {
+  type: jsPsychHtmlKeyboardResponse,
+  stimulus: function(){
+    var stim = "<div class='grid-container'>" +
+              "<div></div>" +
+              "<div style='font-size:60px;'>+</div>" +
+              "<div></div>" +
+            "</div>" +
+            "<br><br><br><br><div style='width: 600px;'><h1 style='float: left; font-size: 20px; margin:0;'>" + prompt_square + "<br><br>A</h1><h1 style='float: right; margin:0; font-size: 20px;'><br>" + prompt_rectangle +"<br><br>L</h1></div>"
+  return stim
+  },
+  choices: 'NO_KEYS',
+  trial_duration: function(){
+        return jsPsych.randomization.sampleWithoutReplacement([900, 925, 950, 975, 1000, 1025, 1050, 1075, 1100], 1)[0];
+      },
+  data: {
+    variable: 'fixation',
+    task: "globallocal2",
+    stimulus: function(){return jsPsych.timelineVariable('stim_chr')}
+  }
+}
+
+var globallocal2_target = {
+  type: jsPsychCategorizeHtml,
+  choices: ['A','L'],
+  correct_text: "",
+  incorrect_text: "",
+  key_answer: function(){return jsPsych.timelineVariable('key_answer')},
+  feedback_duration: 1000,
+  show_stim_with_feedback: false,
+  prompt: "<br><br><br><br><div style='width: 600px;'><h1 style='float: left; font-size: 20px; margin:0;'>" + prompt_square + "<br><br>A</h1><h1 style='float: right; margin:0; font-size: 20px;'><br>" + prompt_rectangle +"<br><br>L</h1></div>",
+  stimulus: function(){
+    var stim = ""
+
+    if(jsPsych.timelineVariable('data')['rule'] == 'global'){
+
+      stim += "<div class='grid-container'>" +
+                "<div>" + elephantleft + "</div>" +
+                "<div id='figure' style='visibility:hidden;'>" + jsPsych.timelineVariable('stimulus') + "</div>" +
+                "<div>" + elephantright + "</div>" +
+              "</div>"
+    }
+
+    if(jsPsych.timelineVariable('data')['rule'] == 'local'){
+      stim += "<div class='grid-container'>" +
[…]

### 2024-09-12 materials/liss/S_globallocal2/scripts/trialmaker.R

@@ -1,0 +1,261 @@
+set.seed(478)
+
+# Abstract Stimuli ------------------------------------------------------
+
+# stimulus set 1
+trials01 <- tibble(
+  type     = c('first', sample(c(rep('repeat',16), rep('switch',16)), size = 32, replace = F)),
+  rule     = c('global', rep(NA, 32)),
+  congruency = c('congruent', sample(c(rep('congruent', 16), rep("incongruent", 16)), size = 32, replace = F)),
+  variable = "globallocal2_01",
+  task     = "globallocal",
+  stimulus = c('SQsq', rep(NA, 32)),
+  key_answer = NA
+)
+
+trials01 <- trials01 |>
+  mutate(
+    valid1 = ifelse(type == lag(type,1) & type == lag(type,2) & type == lag(type,3), FALSE, TRUE),
+    valid2 = ifelse(type == "repeat" & type == lag(type,1) & type == lag(type,2) & lag(type,3) == "switch", FALSE, TRUE),
+    valid3 = ifelse(congruency == lag(congruency,1) & congruency == lag(congruency,2) & congruency == lag(congruency,3), FALSE, TRUE)
+  )
+
+n = 1
+while(any(c(trials01$valid1, trials01$valid2, trials01$valid3) == FALSE, na.rm = T)) {
+  trials01 <- trials01 |>
+    mutate(
+      type = c('first', sample(c(rep('repeat',16), rep('switch',16)), size = 32, replace = F)),
+      congruency = c('congruent', sample(c(rep('congruent', 16), rep("incongruent", 16)), size = 32, replace = F)),
+      valid1 = ifelse(type == lag(type,1) & type == lag(type,2) & type == lag(type,3), FALSE, TRUE),
+      valid2 = ifelse(type == "repeat" & type == lag(type,1) & type == lag(type,2) & lag(type,3) == "switch", FALSE, TRUE),
+      valid3 = ifelse(congruency == lag(congruency,1) & congruency == lag(congruency,2) & congruency == lag(congruency,3), FALSE, TRUE)
+    )
+  n = n+1
+  print(n)
+}
+
+n = 1
+while(n<=32){
+  trials01 <- trials01 |>
+    mutate(
+      rule = case_when(
+        is.na(rule) & type == "repeat" & lag(rule, n=1) == 'global' ~ 'global',
[…]

### 2024-09-12 materials/pilot01/I_flanker/0_flanker_setup.js

@@ -1,0 +1,61 @@
+//------------------------- Functions
+
+var location_stim = function(up, down){
+
+  // Present flanker stimuli either at the top or bottom of the screen
+  var html = "<style>" +
+              "divbottom{display:flex; justify-content: center; align-items: flex-end; height:300px}" +
+              "divtop{display:flex; justify-content: center; align-items: flex-start; height:300px}" +
+             ".topflanker{font-size:50px;}" +
+             ".bottomflanker{font-size:50px;}" +
+             "</style>";
+
+  if(up != null) {
+    html += "<divtop>";
+    html += "<div class = 'topflanker'>";
+    html += up;
+    html += "</div></div>";
+  }
+
+  if(down != null) {
+    html += "<divbottom>";
+    html += "<div class = 'bottomflanker'>";
+    html += down;
+    html += "</div></div>";
+
+  }
+
+  return html;
+};
+
+
+
+//------------------------- Stimuli
+
+  // Flanker stimuli
+  var congruent_left = "&larr;&larr;&larr;&larr;&larr;"
+  var congruent_right = "&rarr;&rarr;&rarr;&rarr;&rarr;"
+  var incongruent_left = "&rarr;&rarr;&larr;&rarr;&rarr;"
+  var incongruent_right = "&larr;&larr;&rarr;&larr;&larr;"
+
+
+  var flanker_stimuli = [
+    {stimulus: congruent_left},
+    {stimulus: congruent_right},
+    {stimulus: incongruent_left},
+    {stimulus: incongruent_right},
+  ]
+
+
+  // Fixation cross
+  var flanker_fixation = {
+  type: jsPsychHtmlKeyboardResponse,
+  stimulus: '<div style="font-size:60px;">+</div>',
+  choices: 'NO_KEYS',
+  trial_duration: 1000,
+  data: {
+    variable: 'fixation'
+  }
+}
+
+

### 2024-09-12 materials/pilot01/I_simon/2_simon_procedure.js

@@ -1,0 +1,153 @@
+var simon_target = {
+  type: jsPsychHtmlKeyboardResponse,
+  stimulus: function(){
+    var stim = ""
+
+    if (jsPsych.timelineVariable('location') == 'left'){
+      stim += "<div class='grid-container'>" +
+                "<div style='font-size: 30px;'>" + jsPsych.timelineVariable('stim') + "</div>" +
+                "<div style='font-size: 60px'>+</div>" +
+                "<div style='font-size: 30px;'></div>" +
+              "</div>"
+    }
+    if (jsPsych.timelineVariable('location') == 'right'){
+      stim += "<div class='grid-container'>" +
+                "<div style='font-size: 30px;'></div>" +
+                "<div style='font-size: 60px'>+</div>" +
+                "<div style='font-size: 30px;'>" + jsPsych.timelineVariable('stim') + "</div>" +
+              "</div>"
+    }
+    return stim
+  },
+  choices: ['A', 'L'],
+  data: {
+    variable: 'test',
+    task: 'simon',
+    location: function(){
+      return jsPsych.timelineVariable('location')
+    },
+    stimtype: function(){
+      return jsPsych.timelineVariable('stimtype')
+    },
+    correct_response: function(){
+      return jsPsych.timelineVariable('correct_response')
+    }
+  },
+  on_finish: function(data) { {
+    if(jsPsych.pluginAPI.compareKeys(data.response, jsPsych.timelineVariable('correct_response', true))) {
+      data.correct = true;
+    } else {
+      data.correct = false;
+    }
+  }
+  }
+};
+
+
+var simon_test_procedure01 = {
+  timeline: [simon_fixation, simon_target],
+  timeline_variables: [
+    {location: 'left', correct_response: 'L', stimtype: 'incongruent', stim: 'RECHTS'},
+    {location: 'right', correct_response: 'A', stimtype: 'incongruent', stim: 'LINKS'},
+    {location: 'left', correct_response: 'A', stimtype: 'congruent', stim: 'LINKS'},
+    {location: 'right', correct_response: 'A', stimtype: 'incongruent', stim: 'LINKS'},
+    {location: 'right', correct_response: 'L', stimtype: 'congruent', stim: 'RECHTS'},
[…]

### 2024-09-12 materials/pilot01/I_simon/scripts/trialmaker.R

@@ -1,0 +1,43 @@
+set.seed(4785)
+
+
+trials01 <- expand_grid(
+  location = c("left", "right"),
+  stim     = c("LINKS", "RECHTS"),
+  n        = 1:16
+) |>
+  select(-n) |>
+  mutate(condition = case_when(
+    location == "left" & stim == "LINKS" ~ "congruent",
+    location == "left" & stim == "RECHTS" ~ "incongruent",
+    location == "right" & stim == "LINKS" ~ "incongruent",
+    location == "right" & stim == "RECHTS" ~ "congruent",
+  )) |>
+  mutate(n = sample(1:n(), size = n(), replace=F)) |>
+  arrange(n) |>
+  mutate(correct_response = ifelse(stim == "LINKS", "A", "L"))
+
+
+trials01 <- trials01 |>
+  mutate(
+    valid1 = ifelse(stim == lag(stim,1) & stim == lag(stim,2) & stim == lag(stim,3), FALSE, TRUE),
+    valid2 = ifelse(condition == lag(condition,1) & condition == lag(condition,2) & condition == lag(condition,3), FALSE, TRUE)
+  )
+
+n = 1
+while(any(c(trials01$valid1, trials01$valid2) == FALSE, na.rm = T)) {
+  trials01 <- trials01 |>
+    mutate(n = sample(1:n(), size = n(), replace=F)) |>
+    arrange(n) |>
+    mutate(
+      valid1 = ifelse(stim == lag(stim,1) & stim == lag(stim,2) & stim == lag(stim,3), FALSE, TRUE),
+      valid2 = ifelse(condition == lag(condition,1) & condition == lag(condition,2) & condition == lag(condition,3), FALSE, TRUE),
+    )
+  print(n)
+  n = n+1
+}
+
+glue_data(
+  trials01,
+  "{{location: '{location}', correct_response: '{correct_response}', condition: '{condition}', stim: '{stim}'}},"
+)

### 2024-09-12 materials/pilot01/P_posner/1_posner_instructions.js

@@ -1,0 +1,125 @@
+//-------------------- Welcome
+var posner_welcome = {
+  type: jsPsychInstructions,
+  pages: [
+    "Welkom bij het <b>Vergelijk</b> spel!<br><br><br><br>"
+  ],
+  show_clickable_nav: true,
+  allow_backward: true,
+  key_forward: -1,
+  key_backward: -1,
+  button_label_next: "verder",
+  data: {variable: 'welcome', task: "posner_practice"}
+};
+
+//-------------------- Instructions
+var posner_instructions = {
+  type: jsPsychInstructions,
+  pages: [
+    "<p style = 'text-align: center;'>"+
+      "In dit spel ziet u steeds twee letters naast elkaar.<br><br>" +
+      "Uw taak is steeds om aan te geven of het <strong>dezelfde</strong> of <strong>verschillende</strong> letters zijn.<br><br>" +
+      "Dit is een voorbeeld van twee <i>dezelfde</i> letters: <strong>AA</strong><br>" +
+      "Dit is een voorbeeld van twee <i>verschillende</i> letters: <strong>bq</strong><br>",
+
+      "De letters kunnen zowel <i>hoofdletters</i> als <i>kleine letters</i> zijn.<br>" +
+      "Dit maakt niet uit voor uw keuze.<br><br>" +
+      "<strong>bB</strong> en <strong>bb</strong> zijn allebei voorbeelden van twee <i>dezelfde</i> letters.<br>" +
+      "<strong>aq</strong> en <strong>Aq</strong> zijn allebei voorbeelden van twee <i>verschillende</i> letters.<br>",
+
+      "<p style = 'text-align: center;'>"+
+      "<div style = 'width: 200px; padding: 20px; float: left;'>Als het dezelfde<br>letters zijn,<br>druk dan op de<br>'A'-toets.<br><br><strong>fF</strong></div>" +
+      "<div style = 'width: 200px; padding: 20px; float: right;'>Als het verschillende<br>letters zijn,<br>druk dan op de<br>'L'-toets.<br><br><strong>Qa</strong></div><br><br><br><br>",
+
+     "<p style = 'text-align: center;'>"+
+      "Antwoord zo snel als u kunt zonder fouten te maken.<br>Af en toe een fout maken is niet erg. Ga in dat geval gewoon door.<br><br>" +
+      "Klik op 'verder' om het spel te oefenen.<br><br><br>"
+  ],
+  show_clickable_nav: true,
+  allow_backward: true,
[…]

### 2024-09-12 materials/pilot01/P_posner/scripts/trialmaker.R

@@ -1,0 +1,79 @@
+library(tidyverse)
+library(glue)
+
+set.seed(35845)
+
+trials <- expand_grid(
+  first  = c("A", "B", "F", "H", "Q", "a", "b", "f", "h", "q"),
+  second = c("A", "B", "F", "H", "Q", "a", "b", "f", "h", "q")
+) |>
+  mutate(
+    f_l  = str_to_lower(first),
+    s_l = str_to_lower(second),
+    condition = ifelse(str_to_lower(first) == str_to_lower(second), "same", "different"),
+    key       = ifelse(condition == "same", "A", "L"),
+    remove    = rep(c(0,1), times = 50)
+  ) |>
+  filter(condition == 'same' | (condition == 'different' & remove == 1))
+
+trials <- trials |>
+  bind_rows(trials |> filter(condition == "same")) |>
+  mutate(rownum = sample(1:n(), size = n())) |>
+  arrange(rownum) |>
+  mutate(
+    valid1 = ifelse(
+      ((f_l == lag(f_l,1) | f_l == lag(s_l,1)) &
+      (f_l == lag(f_l,2) | f_l == lag(s_l,2)) &
+      (f_l == lag(f_l,3) | f_l == lag(s_l,3))),
+      FALSE,
+      TRUE
+    ),
+    valid2 = ifelse(
+      ((s_l == lag(f_l,1) | s_l == lag(s_l,1)) &
+       (s_l == lag(f_l,2) | s_l == lag(s_l,2)) &
+       (s_l == lag(f_l,3) | s_l == lag(s_l,3))),
+      FALSE,
+      TRUE
+    ),
+    valid3 = ifelse(
+      condition == lag(condition,1) &
+      condition == lag(condition,2) &
+      condition == lag(condition,3),
+      FALSE,
+      TRUE
+    )
+  )
+
+while(any(c(trials$valid1, trials$valid2, trials$valid3) == FALSE, na.rm = T)) {
+  trials <- trials |>
+    mutate(rownum    = sample(1:n(), size = n())) |>
+  arrange(rownum) |>
+    mutate(
+      valid1 = ifelse(
+        ((f_l == lag(f_l,1) | f_l == lag(s_l,1)) &
+           (f_l == lag(f_l,2) | f_l == lag(s_l,2)) &
+           (f_l == lag(f_l,3) | f_l == lag(s_l,3))),
+        FALSE,
+        TRUE
+      ),
+      valid2 = ifelse(
[…]

### 2024-09-12 materials/pilot01/S_animacysize/0_animacysize_setup.js

@@ -1,0 +1,16 @@
+
+
+var living     = "<img src='S_animacysize/img/living.png' height=100>";
+var nonliving  = "<img src='S_animacysize/img/nonliving.png'height=100 style='padding-left: 20px;'>";
+var smaller    = "<img src='S_animacysize/img/smaller.png' height=100 style='padding-right: 20px;'>";
+var larger     = "<img src='S_animacysize/img/larger.png' height=100 style='padding-right: 20px;'>";
+
+//-------------------- Response prompts
+
+var prompt_living  = "<div style='float:right; text-align: right;' >" + living  +  "</div>";
+var prompt_nonliving  = "<div style='float:left;  text-align: left;'  >" + nonliving + "</div>";
+var prompt_smaller = "<div style='float:left;  text-align: left;'  >" + smaller + "</div>";
+var prompt_larger   = "<div style='float:right; text-align: right;' >" + larger + "</div>";
+
+
+

### 2024-09-12 materials/pilot01/S_animacysize/1_animacysize_practice.js

@@ -1,0 +1,242 @@
+
+//------------------------- Obejects to hold trial information for the practice session
+
+// Shape instructions
+
+var animacysize_welcome = {
+  type: jsPsychInstructions,
+  pages: [
+    "Welkom bij het <b>Dieren en Objecten</b> spel!<br><br><br><br>"
+  ],
+  show_clickable_nav: true,
+  allow_backward: true,
+  key_forward: -1,
+  key_backward: -1,
+  button_label_next: "verder",
+  data: {variable: 'welcome', task: "animacysize_practice"}
+};
+
+var animacysize_animacy_instructions = {
+    type: jsPsychInstructions,
+    pages: [
+        "<p style = 'text-align: center;'>We gaan een spel spelen waarin u steeds bepaald of iets<br><br>" +
+        "<li>LEVEND of NIET-LEVEND is;<br>" +
+        "<li>KLEINER of GROTER is dan een voetbal.<br>",
+
+        "<p style = 'text-align: center;'> We zullen eerst het LEVEND/NIET-LEVEND spel spelen.<br><br>" +
+        "In het LEVEND/NIET-LEVEND spel bepaalt u of een woord verwijst naar iets <strong>LEVENDS of iets NIET-LEVENDS</strong>.<br><br>" +
+        "Als het woord verwijst naar iets dat LEVEND is, druk dan op de <strong>'A'-toets</strong> op uw toetsenbord.<br><br><br>Voorbeeld:<br><br>" +
+        "<div style = 'text-align: center; font-size:60px'>hond</div>" +
+        "<br><br><br><div>" + prompt_living + prompt_nonliving + "</div><br><br><br>",
+
+        "Als het woord verwijst naar iets dat NIET-LEVEND is, druk dan op de <strong>'L'-toets</strong> op uw toetsenbord.<br><br><br>Voorbeeld:<br><br>" +
+        "<div style = 'text-align: center; font-size:60px'>mes</div>" +
+        "<br><br><br><div>" + prompt_living + prompt_nonliving + "</div><br><br><br>",
+
+        "<p style = 'text-align: center;'> U gaat nu het LEVEND/NIET-LEVEND spel oefenen.<br><br>" +
+        "Antwoord zo snel als u kunt zonder fouten te maken. Af en toe een fout maken is niet erg. Ga in dat geval gewoon door.<br><br>" +
+        "Klik op 'verder' om het LEVEND/NIET-LEVEND/spel te oefenen.</p>"
+    ],
+    show_clickable_nav: true,
[…]

### 2024-09-12 materials/pilot01/S_animacysize/3_animacysize_end.js

@@ -1,0 +1,10 @@
+//------------------------- End of the task and save data
+
+var animacysize_end = {
+  type: jsPsychHtmlButtonResponse,
+  stimulus: "Goed gedaan!<br><br>" +
+  "U bent nu klaar met het Dieren en Objecten spel.<br><br>" +
+  "Klik op 'verder' om door te gaan.<br><br>",
+  choices: ['verder'],
+  data: {variable: "end", task: "animacysize"},
+};

### 2024-09-12 materials/pilot01/S_colorshape/2_colorshape_procedure.js

@@ -1,0 +1,116 @@
+
+var colorshape_procedure1 = {
+  type: jsPsychCategorizeHtml,
+  choices: ['A','L'],
+  correct_text: "",
+  incorrect_text: "",
+  feedback_duration: 250,
+  show_stim_with_feedback: false,
+  prompt: "<div style='width: 600px; height:50px;'>" + prompt_yellow + prompt_tri + prompt_circle + prompt_blue + "</div><br><br>" +
+          "<div style='width: 600px;'><h1 style='float: left; margin:0;'>A</h1><h1 style='float: right; margin:0;'>L</h1></div>",
+  timeline: [
+   {stimulus: yellowtriangle_color, key_answer: 'A', data: {rule: 'color', type: 'first', variable: 'test', task: 'colorshape'}},
+   {stimulus: yellowcircle_shape, key_answer: 'L', data: {rule: 'shape', type: 'switch', variable: 'test', task: 'colorshape'}},
+   {stimulus: bluecircle_shape, key_answer: 'L', data: {rule: 'shape', type: 'repeat', variable: 'test', task: 'colorshape'}},
+   {stimulus: bluetriangle_color, key_answer: 'L', data: {rule: 'color', type: 'switch', variable: 'test', task: 'colorshape'}},
+   {stimulus: bluecircle_color, key_answer: 'L', data: {rule: 'color', type: 'repeat', variable: 'test', task: 'colorshape'}},
+   {stimulus: bluecircle_shape, key_answer: 'L', data: {rule: 'shape', type: 'switch', variable: 'test', task: 'colorshape'}},
+   {stimulus: bluetriangle_shape, key_answer: 'A', data: {rule: 'shape', type: 'repeat', variable: 'test', task: 'colorshape'}},
+   {stimulus: bluetriangle_shape, key_answer: 'A', data: {rule: 'shape', type: 'repeat', variable: 'test', task: 'colorshape'}},
+   {stimulus: bluecircle_color, key_answer: 'L', data: {rule: 'color', type: 'switch', variable: 'test', task: 'colorshape'}},
+   {stimulus: yellowtriangle_shape, key_answer: 'A', data: {rule: 'shape', type: 'switch', variable: 'test', task: 'colorshape'}},
+   {stimulus: yellowcircle_shape, key_answer: 'L', data: {rule: 'shape', type: 'repeat', variable: 'test', task: 'colorshape'}},
[…]

### 2024-09-12 materials/pilot01/S_globallocal/0_globallocal_setup.js

@@ -1,0 +1,40 @@
+var preload_globloc01 = {
+    type: jsPsychPreload,
+    images: [
+      'S_globallocal/img/black_e_of_f.png',
+      'S_globallocal/img/black_e_of_p.png',
+      'S_globallocal/img/black_e_of_t.png',
+      'S_globallocal/img/black_h_of_f.png',
+      'S_globallocal/img/black_h_of_p.png',
+      'S_globallocal/img/black_h_of_t.png'
+      ]
+}
+
+var preload_globloc02 = {
+    type: jsPsychPreload,
+    images: [
+      'S_globallocal/img/black_f_of_e.png',
+      'S_globallocal/img/black_f_of_h.png',
+      'S_globallocal/img/black_p_of_e.png',
+      'S_globallocal/img/black_p_of_h.png',
+      'S_globallocal/img/black_t_of_e.png',
+      'S_globallocal/img/black_t_of_h.png'
+      ]
+}
+
+
+
+var Ge_Lf  = "<img src='S_globallocal/img/black_e_of_f.png' height=350px>";
+var Ge_Lp  = "<img src='S_globallocal/img/black_e_of_p.png' height=350px>";
+var Ge_Lt  = "<img src='S_globallocal/img/black_e_of_t.png' height=350px>";
+var Gf_Le  = "<img src='S_globallocal/img/black_f_of_e.png' height=350px>";
+var Gf_Lh  = "<img src='S_globallocal/img/black_f_of_h.png' height=350px>";
+var Gh_Lf  = "<img src='S_globallocal/img/black_h_of_f.png' height=350px>";
+var Gh_Lp  = "<img src='S_globallocal/img/black_h_of_p.png' height=350px>";
+var Gh_Lt  = "<img src='S_globallocal/img/black_h_of_t.png' height=350px>";
+var Gp_Le  = "<img src='S_globallocal/img/black_p_of_e.png' height=350px>";
+var Gp_Lh  = "<img src='S_globallocal/img/black_p_of_h.png' height=350px>";
+var Gt_Le  = "<img src='S_globallocal/img/black_t_of_e.png' height=350px>";
+var Gt_Lh  = "<img src='S_globallocal/img/black_t_of_h.png' height=350px>";
+
+

### 2024-09-12 materials/pilot01/S_globallocal/1_globallocal_instructions.js

@@ -1,0 +1,179 @@
+//-------------------- Welcome
+var globallocal_welcome = {
+  type: jsPsychInstructions,
+  pages: [
+    "Welkom bij de <strong>Figuren</strong> taak!<br><br><br><br>"
+  ],
+  show_clickable_nav: true,
+  allow_backward: true,
+  key_forward: -1,
+  key_backward: -1,
+  button_label_next: "verder",
+  button_label_previous: "ga terug",
+  data: {variable: 'welcome', task: "globallocal_practice"}
+};
+
+//-------------------- Instructions
+var globallocal_instructions = {
+  type: jsPsychInstructions,
+  pages: [
+    "<p style = 'text-align: center;'>" +
+      "In dit spel ziet u steeds een plaatje van een <strong>GROTE letter</strong> die bestaat uit <strong>KLEINE letters</strong>.<br><br>" +
+      "Hieronder ziet u bijvoorbeeld een <strong>GROTE</strong> letter <strong>'H'</strong> die bestaat uit <strong>KLEINE</strong> letters <strong>'P'</strong>:<br><br><br>" +
+      "<img height = 300 src = S_globallocal/img/black_h_of_p.png></img><br><br>",
+
+      "<p style = 'text-align: center;'>" +
+      "De figuren bevatten altijd in ieder geval de letter <strong>'H'</strong> <i>OF</i> de letter <strong>'E'</strong>, maar nooit tegelijkertijd.<br><br>" +
+      "Links ziet u  een voorbeeld van een <strong>GROTE</strong> letter <strong>'F'</strong> die bestaat uit <strong>KLEINE</strong> letters <strong>'E'</strong>.<br>" +
+      "Rechts ziet u een voorbeeld van een <strong>GROTE</strong> letter <strong>'H'</strong> die bestaat uit <strong>KLEINE</strong> letters <strong>'T'</strong>.<br><br>" +
+
+      "<div class='grid-container' style='height:210px'>" +
+        "<div><img style = 'height: 200px; float: left' src = S_globallocal/img/black_f_of_e_example.png></img></div>" +
+        "<div></div>" +
+        "<div><img style = 'height: 200px; float: right' src = S_globallocal/img/black_h_of_t_example.png></img></div>" +
+      "</div><br><br>",
+
[…]

### 2024-09-12 materials/pilot01/S_globallocal/2_globallocal_procedure.js

@@ -1,0 +1,126 @@
+// Fixation cross
+var fixation = {
+  type: jsPsychHtmlKeyboardResponse,
+  stimulus: '<div style="font-size:60px;">+</div>',
+  choices: "NO_KEYS",
+  trial_duration: 500,
+  data: {
+    variable: 'fixation'
+  }
+};
+
+var globallocal_procedure01 = {
+  type: jsPsychCategorizeHtml,
+  choices: ['A','L'],
+  correct_text: "",
+  incorrect_text: "",
+  feedback_duration: 1000,
+  show_stim_with_feedback: false,
+  prompt: "<br><br><div style='width: 600px;'><h1 style='float: left; font-size: 20px; margin:0;'>GROOT<br><br>A</h1><h1 style='float: right; margin:0; font-size: 20px;'>KLEIN<br><br>L</h1></div>",
+  stimulus: function(){return jsPsych.timelineVariable('stimulus')},
+  data: {
+    rule: function(){return jsPsych.timelineVariable('data')['rule']},
+    type: function(){return jsPsych.timelineVariable('data')['type']},
+    variable: function(){return jsPsych.timelineVariable('data')['variable']},
+    task: function(){return jsPsych.timelineVariable('data')['task']},
+  },
+  timeline: [
+    {stimulus: Ge_Lp, key_answer: 'A', data: {rule: 'global', type: 'first', variable: 'test', task: 'globallocal'}},
+    {stimulus: Ge_Lf, key_answer: 'A', data: {rule: 'global', type: 'repeat', variable: 'test', task: 'globallocal'}},
+    {stimulus: Ge_Lp, key_answer: 'A', data: {rule: 'global', type: 'repeat', variable: 'test', task: 'globallocal'}},
+    {stimulus: Gf_Le, key_answer: 'L', data: {rule: 'local', type: 'switch', variable: 'test', task: 'globallocal'}},
+    {stimulus: Gt_Lh, key_answer: 'L', data: {rule: 'local', type: 'repeat', variable: 'test', task: 'globallocal'}},
+    {stimulus: Gp_Lh, key_answer: 'L', data: {rule: 'local', type: 'repeat', variable: 'test', task: 'globallocal'}},
+    {stimulus: Ge_Lp, key_answer: 'A', data: {rule: 'global', type: 'switch', variable: 'test', task: 'globallocal'}},
[…]

### 2024-09-12 materials/pilot01/S_globallocal/scripts/trialmaker.R

@@ -1,0 +1,179 @@
+library(tidyverse)
+library(glue)
+
+check_if_same_stim <- function(data) {
+  1:nrow(data) |>
+    map_chr(function(x) {
+
+      if(x ==1) {
+        return(data$stimulus[x])
+      }
+
+      current = data$stimulus[x]
+      previous = data$stimulus[x-1]
+      subsequent = data$stimulus[x+1]
+
+      global_stims <- c('Ge_Lf', 'Ge_Lp', 'Ge_Lt', 'Gh_Lf', 'Gh_Lp', 'Gh_Lt')
+      local_stims <-  c('Gf_Le', 'Gf_Lh', 'Gp_Le', 'Gp_Lh', 'Gt_Le', 'Gt_Lh')
+
+      if(current == previous) {
+        if(data$rule[x] == "global") {
+          current <- sample(global_stims[!global_stims %in% c(subsequent, current)], size = 1)
+        }
+        if(data$rule[x] == "local") {
+          current <- sample(local_stims[!local_stims %in% c(subsequent, current)], size = 1)
+        }
+      }
+      return(current)
+    })
+}
+
+set.seed(486)
+
+n_block_trials <-  32
+
+stim_vector <- c('Ge_Lf', 'Ge_Lp', 'Ge_Lt', 'Gf_Le', 'Gf_Lh',
+'Gh_Lf', 'Gh_Lp', 'Gh_Lt', 'Gp_Le', 'Gp_Lh', 'Gt_Le',
+'Gt_Lh')
+
+# stimulus set 1
+trials01 <- tibble(
+  type     = c('first', sample(c(rep('repeat',n_block_trials/2), rep('switch',n_block_trials/2)), size = n_block_trials, replace = F)),
+  rule     = c('global', rep(NA, n_block_trials)),
+  variable = "globloc01",
+  task     = "globloc01",
+  key_answer = NA
+)
+
+n = 1
+while(n<=32){
+  trials01 <- trials01 |>
+    mutate(
+      rule = case_when(
+        is.na(rule) & type == "repeat" & lag(rule,1) == "global" ~ "global",
+        is.na(rule) & type == "repeat" & lag(rule,1) == "local" ~ "local",
+        is.na(rule) & type == "switch" & lag(rule,1) == "global" ~ "local",
+        is.na(rule) & type == "switch" & lag(rule,1) == "local" ~ "global",
+        TRUE ~ rule
+      )
+    )
+  n = n+1
+}
+
+trials01 <- trials01 |>
+  mutate(
[…]

### 2024-09-12 materials/pilot01/global_variables.js

@@ -1,0 +1,69 @@
+// FUNCTIONS
+
+ function shuffle(array) {
+  let currentIndex = array.length,  randomIndex;
+
+  // While there remain elements to shuffle.
+  while (currentIndex > 0) {
+
+    // Pick a remaining element.
+    randomIndex = Math.floor(Math.random() * currentIndex);
+    currentIndex--;
+
+    // And swap it with the current element.
+    [array[currentIndex], array[randomIndex]] = [
+      array[randomIndex], array[currentIndex]];
+  }
+
+  return array;
+}
+
+
+// GENERAL TRIALS
+
+// Task feedback during the practice trials.
+var feedback = {
+  type: jsPsychHtmlKeyboardResponse,
+  stimulus: function(){
+
+    var last_trial_rt = jsPsych.data.getLastTrialData().values()[0].rt;
+    var last_trial_correct = jsPsych.data.get().last(1).values()[0].correct;
+
+    if(last_trial_rt > 4000){
+      return "<p style = 'color:red;font-size:40px'>Te langzaam!</p>";
+    } else {
+      if(last_trial_correct){
+        return "<p style = 'color:green;font-size:40px'>Goed!</p>";
+      } else {
+        return "<p style = 'color:red;font-size:40px'>Fout!</p>";
+      }}},
+  trial_duration: 2000,
+  choices: "NO_KEYS",
+  data: {
+    variable: 'feedback'
+  }
+};
+
+
+// General variables
+var fullscreenmode = {
+  type: jsPsychFullscreen,
+  fullscreen_mode: true,
+  message: "<p>De spelletjes zullen verder gaan in <i>volledig scherm modus</i>.<br>We verzoeken u om in <i>volledig scherm modus</i> te blijven tijdens de gehele duur van de spelletjes.</p>",
+  button_label: "Ga verder"
+};
+
+var cursor_off = {
+    type: jsPsychCallFunction,
+    func: function() {
+        document.body.style.cursor= "none";
+    }
+};
+
+var cursor_on = {
+    type: jsPsychCallFunction,
+    func: function() {
+        document.body.style.cursor= "auto";
+    }
+};
+

### 2024-09-12 materials/pilot02/S_globallocal2/scripts/trialmaker.R

@@ -1,0 +1,261 @@
+set.seed(478)
+
+# Abstract Stimuli ------------------------------------------------------
+
+# stimulus set 1
+trials01 <- tibble(
+  type     = c('first', sample(c(rep('repeat',16), rep('switch',16)), size = 32, replace = F)),
+  rule     = c('global', rep(NA, 32)),
+  congruency = c('congruent', sample(c(rep('congruent', 16), rep("incongruent", 16)), size = 32, replace = F)),
+  variable = "globallocal2_01",
+  task     = "globallocal",
+  stimulus = c('SQsq', rep(NA, 32)),
+  key_answer = NA
+)
+
+trials01 <- trials01 |>
+  mutate(
+    valid1 = ifelse(type == lag(type,1) & type == lag(type,2) & type == lag(type,3), FALSE, TRUE),
+    valid2 = ifelse(type == "repeat" & type == lag(type,1) & type == lag(type,2) & lag(type,3) == "switch", FALSE, TRUE),
+    valid3 = ifelse(congruency == lag(congruency,1) & congruency == lag(congruency,2) & congruency == lag(congruency,3), FALSE, TRUE)
+  )
+
+n = 1
+while(any(c(trials01$valid1, trials01$valid2, trials01$valid3) == FALSE, na.rm = T)) {
+  trials01 <- trials01 |>
+    mutate(
+      type = c('first', sample(c(rep('repeat',16), rep('switch',16)), size = 32, replace = F)),
+      congruency = c('congruent', sample(c(rep('congruent', 16), rep("incongruent", 16)), size = 32, replace = F)),
+      valid1 = ifelse(type == lag(type,1) & type == lag(type,2) & type == lag(type,3), FALSE, TRUE),
+      valid2 = ifelse(type == "repeat" & type == lag(type,1) & type == lag(type,2) & lag(type,3) == "switch", FALSE, TRUE),
+      valid3 = ifelse(congruency == lag(congruency,1) & congruency == lag(congruency,2) & congruency == lag(congruency,3), FALSE, TRUE)
+    )
+  n = n+1
+  print(n)
+}
+
+n = 1
+while(n<=32){
+  trials01 <- trials01 |>
+    mutate(
+      rule = case_when(
+        is.na(rule) & type == "repeat" & lag(rule, n=1) == 'global' ~ 'global',
[…]

### 2024-09-12 materials/pilot03/S_animacysize/scripts/trialmaker.R

@@ -1,0 +1,155 @@
+set.seed(4785)
+
+
+# Stimulus sets -----------------------------------------------------------
+
+lsm_test_words <- c("egel", "goudvis", "kever", "kikker", "konijn", "mier",
+                    "mug", "muis", "mus", "rat", "rups", "slak", "spin",
+                    "vlinder", "wesp", "worm")
+
+lbi_test_words <- c( "beer", "dolfijn", "ezel", "geit", "hert", "ijsbeer",
+                     "kameel", "leeuw", "olifant", "ooievaar", "orka",
+                     "schaap", "stier", "tijger", "walvis", "wolf")
+
+
+nlsm_test_words <- c("blik", "bril", "kam", "lepel", "mok", "naald", "pen",
+                     "pil", "ring", "schaar", "sok", "speld", "spijker",
+                     "stift", "vork", "zeep")
+
+
+nlbi_test_words <- c("bus", "caravan", "fiets", "iglo",  "kano", "kanon",
+                     "kraan",  "lantaarn", "piano", "raket", "saxofoon",
+                     "tent", "toren", "tractor", "vulkaan", "windmolen")
+
+lsm_test_words <- sample(lsm_test_words, size = 16, replace = F)
+lbi_test_words <- sample(lbi_test_words, size = 16, replace = F)
+nlsm_test_words <- sample(nlsm_test_words, size = 16, replace = F)
+nlbi_test_words <- sample(nlbi_test_words, size = 16, replace = F)
+
+# stimulus set 1
+trials01 <- tibble(
+  type     = c(rep('repeat',16), rep('switch',16)),
+  rule     = c(rep(NA, 32)),
+  stimulus = c(lsm_test_words[1:8], lbi_test_words[1:8], nlsm_test_words[1:8], nlbi_test_words[1:8]),
+  correct  = c(rep('living_smaller', 8), rep("living_larger",8), rep("nonliving_smaller",8), rep("nonliving_larger",8)),
+  variable = "animacysize1",
+  task     = "animacysize",
+  key_answer = NA
+)
+
+trials01 <- trials01 |>
+  mutate(
+    valid1 = ifelse(type == lag(type,1) & type == lag(type,2) & type == lag(type,3), FALSE, TRUE),
[…]

### 2024-09-12 materials/pilot03/S_colorshape/0_colorshape_setup.js

@@ -1,0 +1,24 @@
+//------------------------- Shape stimuli
+
+var yellowcircle   = "<svg width='250' height='250'><circle cx='125' cy='125' r='100' fill='yellow' stroke='black' stroke-width='4'/></svg>";
+var bluecircle     = "<svg width='250' height='250'><circle cx='125' cy='125' r='100' fill='blue' stroke='black' stroke-width='4'/></svg>";
+var yellowtriangle = "<svg width='250' height='250'><polygon points='25,225 125,25 225,225' fill='yellow' stroke='black' stroke-width='4'/></svg>";
+var bluetriangle   = "<svg width='250' height='250'><polygon points='25,225 125,25 225,225' fill='blue' stroke='black' stroke-width='4'/></svg>";
+
+//------------------------- Prompts to show to aid with responses-
+
+var prompt_circle  = "<svg style='float:right;' width='100' height='100'><circle cx='40' cy='40' r='40' fill='black'/></svg>";
+var prompt_tri     = "<svg style='float:left;' width='100' height='100'><polygon points='10,80 90,80 50,0' fill='black'/></svg>";
+var prompt_yellow  = "<svg style='float:left;' width='100' height='100'><rect width='80' height='80' fill='yellow' stroke='black' stroke-width='4'/></svg>";
+var prompt_blue    = "<svg style='float:right;' width='100' height='100'><rect width='80' height='80' fill='blue' stroke='black' stroke-width='4'/></svg>";
+
+//------------------------- All possible shapes/colors with the rules displayed at the top
+
+var yellowcircle_color   = "<div style = 'text-align: center;'><h1>Kleur</h1>" + yellowcircle   + "</div>";
+var yellowcircle_shape   = "<div style = 'text-align: center;'><h1>Vorm</h1>" + yellowcircle   + "</div>";
+var bluecircle_color     = "<div style = 'text-align: center;'><h1>Kleur</h1>" + bluecircle     + "</div>";
+var bluecircle_shape     = "<div style = 'text-align: center;'><h1>Vorm</h1>" + bluecircle     + "</div>";
+var yellowtriangle_color = "<div style = 'text-align: center;'><h1>Kleur</h1>" + yellowtriangle + "</div>";
+var yellowtriangle_shape = "<div style = 'text-align: center;'><h1>Vorm</h1>" + yellowtriangle + "</div>";
[…]

### 2024-09-12 materials/pilot03/S_globallocal2/0_globallocal2_setup.js

@@ -1,0 +1,53 @@
+var preload_globloc01 = {
+    type: jsPsychPreload,
+    images: [
+      'S_globallocal2/img/SQsq.png',
+      'S_globallocal2/img/SQrec.png',
+      'S_globallocal2/img/RECrec.png',
+      'S_globallocal2/img/RECsq.png',
+      'S_globallocal2/img/mouse_left.png',
+      'S_globallocal2/img/mouse_right.png',
+      'S_globallocal2/img/elephant_left.png',
+      'S_globallocal2/img/elephant_right.png',
+      'S_globallocal2/img/square.png',
+      'S_globallocal2/img/rectangle.png',
+      ]
+}
+
+var preload_globloc02 = {
+    type: jsPsychPreload,
+    images: [
+      'S_globallocal2/img/SQsq_local.png',
+      'S_globallocal2/img/SQrec_local.png',
+      'S_globallocal2/img/RECrec_local.png',
+      'S_globallocal2/img/RECsq_local.png',
+      'S_globallocal2/img/SQsq_global.png',
+      'S_globallocal2/img/SQrec_global.png',
+      'S_globallocal2/img/RECrec_global.png',
+      'S_globallocal2/img/RECsq_global.png',
+      ]
+}
+
+
+var SQsq   = "<img src='S_globallocal2/img/SQsq.png' width=150px>";
+var SQrec  = "<img src='S_globallocal2/img/SQrec.png' width=150px>";
+var RECrec = "<img src='S_globallocal2/img/RECrec.png' width=300px>";
+var RECsq  = "<img src='S_globallocal2/img/RECsq.png' width=300px>";
+
+var mouseleft     = "<img src='S_globallocal2/img/mouse_left.png' height=100px>";
+var mouseright    = "<img src='S_globallocal2/img/mouse_right.png' height=100px>";
+var elephantleft  = "<img src='S_globallocal2/img/elephant_left.png' height=100>";
+var elephantright = "<img src='S_globallocal2/img/elephant_right.png' height=100px>";
+
+var prompt_square    = "<img src='S_globallocal2/img/square.png' width=50px>";
+var prompt_rectangle = "<img src='S_globallocal2/img/rectangle.png' width=100px>";
+
+var SQsq_local   = "<img src='S_globallocal2/img/SQsq_local.png' width=150px>";
+var SQrec_local  = "<img src='S_globallocal2/img/SQrec_local.png' width=150px>";
[…]

### 2024-09-12 materials/pilot03/S_globallocal2/2_globallocal2_procedure.js

@@ -1,0 +1,159 @@
+// Fixation cross
+var globallocal2_fixation = {
+  type: jsPsychHtmlKeyboardResponse,
+  stimulus: function(){
+    var stim = "<div class='grid-container'>" +
+              "<div></div>" +
+              "<div style='font-size:60px;'>+</div>" +
+              "<div></div>" +
+            "</div>" +
+            "<br><br><br><br><div style='width: 600px;'><h1 style='float: left; font-size: 20px; margin:0;'>" + prompt_square + "<br><br>A</h1><h1 style='float: right; margin:0; font-size: 20px;'><br>" + prompt_rectangle +"<br><br>L</h1></div>"
+  return stim
+  },
+  choices: 'NO_KEYS',
+  trial_duration: function(){
+        return jsPsych.randomization.sampleWithoutReplacement([900, 925, 950, 975, 1000, 1025, 1050, 1075, 1100], 1)[0];
+      },
+  data: {
+    variable: 'fixation',
+    task: "globallocal2"
+  }
+}
+
+var globallocal2_target = {
+  type: jsPsychCategorizeHtml,
+  choices: ['A','L'],
+  correct_text: "",
+  incorrect_text: "",
+  key_answer: function(){return jsPsych.timelineVariable('key_answer')},
+  feedback_duration: 1000,
+  show_stim_with_feedback: false,
+  prompt: "<br><br><br><br><div style='width: 600px;'><h1 style='float: left; font-size: 20px; margin:0;'>" + prompt_square + "<br><br>A</h1><h1 style='float: right; margin:0; font-size: 20px;'><br>" + prompt_rectangle +"<br><br>L</h1></div>",
+  stimulus: function(){
+    var stim = ""
+
+    if(jsPsych.timelineVariable('data')['rule'] == 'global'){
+
+      stim += "<div class='grid-container'>" +
+                "<div>" + elephantleft + "</div>" +
+                "<div id='figure' style='visibility:hidden;'>" + jsPsych.timelineVariable('stimulus') + "</div>" +
+                "<div>" + elephantright + "</div>" +
+              "</div>"
+    }
+
+    if(jsPsych.timelineVariable('data')['rule'] == 'local'){
+      stim += "<div class='grid-container'>" +
+                "<div>" + mouseleft + "</div>" +
[…]

### 2024-09-12 scripts/00_dag.R

@@ -1,0 +1,266 @@
+
+# Read data ---------------------------------------------------------------
+
+library(tidyverse)
+library(dagitty)
+library(haven)
+
+load("analysis_objects/wm_study/confirmatory_results.RData")
+
+
+liss_data <- read_csv("data/liss_data/wm_study_data.csv") |>
+  mutate(
+    inr_m = log(inr_m),
+    inr_cv = log(inr_cv)
+  ) |>
+  mutate(
+    across(c(ends_with('cap'), binding_number_acc, updating_number_acc), ~scale(x = .) |> as.numeric()),
+    across(c(inr_m, inr_cv, threat_comp, p_scar_m, p_scar_cv), ~scale(x = .) |> as.numeric()),
+    age = scale(age) |> as.numeric(),
+    edu = scale(edu) |> as.numeric(),
+   # noise = scale(noise) |> as.numeric(),
+    interrupt = case_when(
+      interrupt == 1 ~ 0,
+      interrupt == 2 ~ 1,
+      TRUE ~ interrupt
+    ),
+    leave = case_when(
+      leave == 1 ~ 0,
+      leave == 2 ~ 1,
+      TRUE ~ leave
+    ),
+    age_sq = age^2
+  )
+
+wm_data <- read_sav("data/liss_data/L_Cognitieve_Vaardigheden_v2_1.0p.sav") |>
+  select(nomem_encr, starts_with("VP"), starts_with("VC"), starts_with("PP"), starts_with("PC"), sex = geslacht, ospan_cap, rspan_cap, binding_number_acc, updating_number_acc) %>%
+  mutate(across(c(VP1, VP3, PP1, PP2, PP3, PP4, VC1, VC3, PC1, PC2, PC3, PC4), ~ 8 - .)) |>
+  mutate(
+    VP_m = across(starts_with("VP")) |> rowMeans(na.rm = T),
+    VC_m = across(starts_with("VC")) |> rowMeans(na.rm = T),
+    PP_m = across(starts_with("PP")) |> rowMeans(na.rm = T),
+    PC_m = across(starts_with("PC")) |> rowMeans(na.rm = T),
+    child_adv = across(c(VP_m, PP_m)) |> rowMeans(na.rm = T),
+    curr_adv = across(c(VC_m, PC_m)) |> rowMeans(na.rm = T)
+  ) |>
[…]

### 2024-09-12 scripts/02_analyze_data.R

@@ -1,0 +1,228 @@
+library(tidyverse)
+
+load('data/pilot_data.RData')
+
+
+flanker_clean <- flanker_data |>
+  filter(variable != "end") |>
+  group_by(id, condition) |>
+  mutate(
+    ex_fast = ifelse(rt < 250, TRUE, FALSE),
+    ex_slow = ifelse(scale(rt) |> as.numeric() > 3.2 | rt > 3000, TRUE, FALSE)
+  ) |>
+  group_by(id) |>
+  mutate(
+    ex_chance = ifelse(sum(correct)/n()< .50, TRUE, FALSE)
+  ) |>
+  filter(ex_fast == FALSE & ex_slow == FALSE & ex_chance == FALSE) |>
+  group_by(id, condition) |>
+  summarise(
+    flanker_RT  = mean(rt, na.rm = T)
+  ) |>
+  ungroup() |>
+  pivot_wider(names_from = 'condition', names_prefix = "flanker_", values_from = 'flanker_RT') |>
+  mutate(flanker_diff = flanker_incongruent - flanker_congruent) |>
+  select(id, flanker_incongruent, flanker_congruent, flanker_diff)
+
+
+colorshape_clean <- colorshape_data |>
+  filter(variable != "colorshape_finish", type != "first") |>
+  group_by(id, type) |>
+  mutate(
+    ex_fast = ifelse(rt < 250, TRUE, FALSE),
+    ex_slow = ifelse(scale(rt) |> as.numeric() > 3.2 | rt > 5000, TRUE, FALSE),
+    ex_error = ifelse(!correct, TRUE, FALSE),
+    ex_preverror = ifelse(lag(correct,1) == FALSE, TRUE, FALSE),
+  ) |>
+  group_by(id) |>
+  mutate(
+    ex_chance = ifelse(sum(correct)/n() < .50, TRUE, FALSE)
+  ) |>
+  filter(ex_fast == FALSE & ex_slow == FALSE, ex_error == FALSE, ex_preverror == FALSE, ex_chance == FALSE) |>
+  group_by(id, type) |>
+  summarise(
+    colorshape_RT  = mean(rt, na.rm = T),
+    colorshape_acc = sum(correct) / n() * 100
+  ) |>
+  pivot_wider(names_from = 'type', names_prefix = "colorshape_", values_from = 'colorshape_RT') |>
+  mutate(colorshape_switchcost = colorshape_switch - colorshape_repeat) |>
[…]

### 2024-09-12 scripts/0_functions/ddm_functions.R

@@ -1,0 +1,91 @@
+
+extract_traces <- function(mcmc, chains = 3, iterations = 12000) {
+
+  mcmc |>
+    select(starts_with("mu")) |>
+    mutate(
+      n = rep(1:iterations, chains),
+      chains = rep(1:chains, each = iterations))  |>
+    rename(
+      a  = muAlpha,
+      t0 = muTau,
+      v  = muDelta
+    ) |>
+    pivot_longer(-c(n, chains), names_to = 'parameter', values_to = 'value')
+}
+
+
+
+extract_traces_2con <- function(mcmc, chains = 3, iterations = 1000) {
+
+  mcmc |>
+    select(starts_with("mu")) |>
+    mutate(
+      n = rep(1:iterations, chains),
+      chains = rep(1:chains, each = iterations))  |>
+    rename(
+      a1  = `muAlpha[1]`,
+      a2  = `muAlpha[2]`,
+      t1 = `muTau[1]`,
+      t2 = `muTau[2]`,
+      v1  = `muDelta[1]`,
+      v2  = `muDelta[2]`
+    ) |>
+    pivot_longer(-c(n, chains), names_to = 'parameter', values_to = 'value')
+}
+
+extract_ddm_estimates <- function(mcmc, task_prefix, id_matches) {
+
+  mcmc |>
+    pivot_longer(everything(), names_to = "parameter", values_to = "estimated") |>
+    group_by(parameter) |>
+    summarise(estimated = mean(estimated, na.rm = T)) |>
+    filter(str_detect(parameter, pattern = 'deviance|^mu|^prec', negate = T)) |>
+    separate(col = parameter, into = c('parameter', 'nomem_encr_num'), sep = "\\[") |>
+    mutate(
+      nomem_encr_num = str_remove(nomem_encr_num, pattern = "\\]$")
+    ) |>
+    mutate(
+      nomem_encr_num = as.numeric(nomem_encr_num),
+      parameter = case_when(
+        parameter == 'alpha' ~ paste0(task_prefix, 'a'),
+        parameter == 'tau' ~ paste0(task_prefix, 't'),
+        parameter == 'delta' ~ paste0(task_prefix, 'v')
+      )) |>
+    pivot_wider(names_from = "parameter", values_from = "estimated") |>
+    left_join(id_matches) |>
+    select(-nomem_encr_num)
+}
+
+extract_ddm_estimates_2con <- function(mcmc, task_prefix, id_matches) {
+
+  mcmc |>
[…]

### 2024-09-12 scripts/0_functions/functions_exclusions.R

@@ -1,0 +1,17 @@
+
+gbinom <- function(n, p, low=0, high=n,scale = F) {
+
+   sd <- sqrt(n * p * (1 - p))
+   if(scale && (n > 10)) {
+      low <- max(0, round(n * p - 4 * sd))
+      high <- min(n, round(n * p + 4 * sd))
+   }
+   values <- low:high
+   probs <- dbinom(values, n, p)
+
+   max_chance_performance <- max(values[which(probs > 0.025)])
+
+
+
+   return((max_chance_performance/n)*100)
+}

### 2024-09-12 scripts/0_functions/general-functions.R

@@ -1,0 +1,167 @@
+fix_subject_ids <- function(data) {
+  
+  # Selection of regular expressions that fix general mistakes in subject identifiers.
+  # Specifically, each subject ID should be preceded by 'NDAR_INV' followed by a unique code.
+  
+  data |> 
+    filter(!str_detect(str_to_lower(subj_idx), "test|tes|practice")) |> 
+    mutate(
+      subj_idx = str_replace_all(str_to_lower(subj_idx), "^.*ndar", "ndar"),
+      subj_idx = str_replace_all(str_to_lower(subj_idx), "ndar_inv\\s|^ndar.*inv|ndarinv|ndar[^[:alnum:]]|^inv|^dar_inv|^nadr_inv|bdar_inv|ndra_inv|ndr_inv|ndat_inv", ""),
+      subj_idx = toupper(paste0("NDAR_INV", subj_idx)),
+      subj_idx = str_remove_all(subj_idx, "(-|_)[0-9]$"),
+      subj_idx = str_remove_all(subj_idx, "(-|\\s).*")
+    )
+}
+
+
+
+fix_id_typos <- function(data, reference, id = "subj_idx", distance = 2) {
+  
+  # Function that tries to detect typos using the Levenshtein distance between unmatched ids and the reference list (i.e., the ids found in the summary score df)
+  # The Levenshtein distance counds the number of deletions, insertions and substitutions necessary to turn one string into another.
+  # Thus, a small Levensthein distance (e.g., 1 or 2) in subject IDs likely indicates a typo.
+  
+  data_ids <- data |> 
+    pull(id) |> 
+    unique()
+  
+  reference |> 
+    as_tibble_col(column_name = "original")
+  
+  mismatches <- data_ids[!data_ids %in% reference] %>% 
+    as_tibble_col(column_name = "subj_idx") %>%
+    mutate(
+      similar = pmap(.l = ., function(subj_idx) {
+        
+        reference |> 
+          as_tibble_col(column_name = "ref") |> 
+          mutate(original = subj_idx) |>  
+          mutate(
+            string_dist = stringdist::stringdist(a = original, b = ref, method = 'lv')
+          ) |> 
+          filter(string_dist == min(string_dist))
+      })
+    ) |> 
+    unnest(similar) 
+    
+  mismatches |> select(subj_idx, string_dist) |> 
+    distinct() |> 
+    count(string_dist) |> 
[…]

### 2024-09-12 scripts/0_pilot/01_create_scales.R

@@ -1,0 +1,531 @@
+# Packages ----------------------------------------------------------------
+library(psych)
+library(tidyverse)
+library(qualtRics)
+library(jsonlite)
+library(here)
+library(sjlabelled)
+
+
+# Data --------------------------------------------------------------------
+pilot1_data <-
+  fetch_survey(
+    surveyID = "SV_0Srsyx4Vh2UajhY",
+    verbose  = T,
+    force_request = T,
+    label = F,
+    convert = F,
+    add_var_labels = F
+  ) %>%
+  rename_with(tolower) %>%
+  filter(finished==1, status == 0, `duration (in seconds)` > 0) |>
+  select(prolific_pid, starts_with("data"))
+
+pilot2_data <-
+  fetch_survey(
+    surveyID = "SV_5tp0I8hsA2YGMyG",
+    verbose  = T,
+    force_request = T,
+    label = F,
+    convert = F,
+    add_var_labels = F
+  ) %>%
+  rename_with(tolower) %>%
+  filter(finished==1, status == 0, `duration (in seconds)` > 0) |>
+  select(-session_id) |>
+  select(prolific_pid, starts_with("data"))
+
+pilot3_data <-
+  fetch_survey(
+    surveyID = "SV_3kot0l1n43R9DU2",
+    verbose  = T,
+    force_request = T,
+    label = F,
+    convert = F,
+    add_var_labels = F
+  ) %>%
+  rename_with(tolower) %>%
+  filter(finished==1, status == 0, `duration (in seconds)` > 0) |>
+  select(-session_id) |>
+  select(prolific_pid, starts_with("data"))
+
+pilot_data <- left_join(pilot1_data, pilot2_data) |>
+  bind_rows(pilot3_data |> select(starts_with('data'))) |>
+  mutate(id = 1:n())
+
+
+
+# Self-report -------------------------------------------------------------
+
+
+## Demographics ----
+
+vars01_dems <-
+  pilot_data %>%
+  select(prolific_pid, matches("^dems_"))
+
+
+
+## Admin ----
+
+vars02_admin <-
+  pilot_data %>%
+  select(ends_with("id"))
+
+
+
+# Flanker Task ------------------------------------------------------------
+
+flanker_prac <-
+  pilot_data |>
+  select(id, prolific_pid, data_flanker_prac) |>
+  filter(!is.na(data_flanker_prac)) |>
[…]

### 2024-09-12 scripts/1_study/01_parse.R

@@ -1,0 +1,223 @@
+# 1. Packages -------------------------------------------------------------
+library(psych)
+library(tidyverse)
+library(qualtRics)
+library(jsonlite)
+library(here)
+library(sjlabelled)
+
+
+# 2. Data -----------------------------------------------------------------
+raw_data <- read_delim("data/L_CognitiveAdversity_TASKDATA.csv", delim = ";")
+
+
+raw_data <- raw_data |>
+  mutate(
+    data_simon = case_when(
+      volgorde__1 == '1' ~ JSON1.4,
+      volgorde__2 == '1' ~ JSON2.4,
+      volgorde__3 == '1' ~ JSON3.4,
+      volgorde__4 == '1' ~ JSON4.4,
+      volgorde__5 == '1' ~ JSON5.4,
+      volgorde__6 == '1' ~ JSON6.4
+    ),
+    data_flanker = case_when(
+      volgorde__1 == '2' ~ JSON1.4,
+      volgorde__2 == '2' ~ JSON2.4,
+      volgorde__3 == '2' ~ JSON3.4,
+      volgorde__4 == '2' ~ JSON4.4,
+      volgorde__5 == '2' ~ JSON5.4,
+      volgorde__6 == '2' ~ JSON6.4
+    ),
+    data_globallocal2 = case_when(
+      volgorde__1 == '3' ~ JSON1.4,
+      volgorde__2 == '3' ~ JSON2.4,
+      volgorde__3 == '3' ~ JSON3.4,
+      volgorde__4 == '3' ~ JSON4.4,
+      volgorde__5 == '3' ~ JSON5.4,
+      volgorde__6 == '3' ~ JSON6.4
+    ),
+    data_colorshape = case_when(
+      volgorde__1 == '4' ~ JSON1.4,
+      volgorde__2 == '4' ~ JSON2.4,
+      volgorde__3 == '4' ~ JSON3.4,
+      volgorde__4 == '4' ~ JSON4.4,
+      volgorde__5 == '4' ~ JSON5.4,
+      volgorde__6 == '4' ~ JSON6.4
+    ),
+    data_animacysize = case_when(
+      volgorde__1 == '5' ~ JSON1.4,
+      volgorde__2 == '5' ~ JSON2.4,
+      volgorde__3 == '5' ~ JSON3.4,
+      volgorde__4 == '5' ~ JSON4.4,
+      volgorde__5 == '5' ~ JSON5.4,
[…]

### 2024-09-12 scripts/1_study/02_clean.R

@@ -1,0 +1,363 @@
+library(tidyverse)
+library(sjlabelled)
+
+load("data/raw_data_long.RData")
+source("scripts/0_functions/functions_exclusions.R")
+source("scripts/0_functions/create_codebook.R")
+
+# Fast: < 250 ms
+# very_slow: > 15s
+# slow: > 3SD of intraindividual mean
+# nval_trials: number of participants with fewer than 20 valid trials
+# n_raw: sample size prior to exclusions
+# n_clean: sample size after exclusions
+exclusions <- list()
+
+exclusions$sample$started <- (length(all_complete_ids) + length(all_incomplete_ids))
+exclusions$sample$not_finish <- length(all_incomplete_ids)
+exclusions$sample$finish <- length(all_complete_ids)
+
+# 1. Preregistered exclusions ---------------------------------------------
+
+## 1.1 Flanker task ----
+
+flanker_clean <- flanker_raw |>
+  filter(rt < 10) |> # Remove excessively long RTs
+  group_by(nomem_encr, condition) |>
+  mutate(
+    ex_fast_outlier = ifelse(rt < 0.250, TRUE, FALSE),
+    ex_slow_outlier = ifelse(as.numeric(scale(log(rt))) > 3, TRUE, FALSE)
+  ) |>
+  group_by(nomem_encr) |>
+  mutate(
+    ex_chance_perf  = ifelse(sum(correct==1)/n()*100 < gbinom(64, 0.50), TRUE, FALSE)
+  ) |>
+  ungroup()
+
+exclusions$flanker <-
+  list(
+    fast      = paste(round(sum(flanker_clean$ex_fast_outlier)/nrow(flanker_clean)*100, 1), "%"),
+    very_slow = paste(round(sum(flanker_raw$rt > 10)/nrow(flanker_raw)*100, 2), "%"),
+    slow      = paste(round(sum(flanker_clean$ex_slow_outlier, na.rm = T)/nrow(flanker_clean)*100, 1), "%"),
+    chance    = flanker_clean |> group_by(nomem_encr) |> summarise(n = sum(ex_chance_perf)) |> filter(n > 1) |> nrow(),
+    raw_n     = unique(flanker_clean$nomem_encr) |> length()
+  )
+
+flanker_clean <- flanker_clean |>
+  filter(ex_slow_outlier == FALSE, ex_fast_outlier == FALSE, ex_chance_perf == FALSE) |>
+  group_by(nomem_encr) |>
[…]

### 2024-09-12 scripts/1_study/03_ddm.R

@@ -1,0 +1,506 @@
+load("data/clean_data_long.RData")
+
+library(runjags)
+
+# 1. Models ---------------------------------------------------------------
+
+## 1.1 Two conditions ----
+
+model_2con <- "model {
+ #likelihood function
+ for (t in 1:nTrials) {
+ y[t] ~ dwiener(alpha[condition[t], subject[t]],
+ tau[condition[t], subject[t]],
+ 0.5,
+ delta[condition[t], subject[t]])
+ }
+ for (s in 1:nSubjects) {
+ for (c in 1:nCon) {
+ tau[c, s] ~ dnorm(muTau[c], precTau) T(.0001, 1)
+ delta[c, s] ~ dnorm(muDelta[c] , precDelta) T(-10, 10)
+ alpha[c, s] ~ dnorm(muAlpha[c], precAlpha) T(.1, 5)
+ }
+
+ }
+ #priors
+ for (c in 1:nCon){
+ muTau[c] ~ dunif(.0001, 1)
+ muDelta[c] ~ dunif(-10, 10)
+ muAlpha[c] ~ dunif(.1, 5)
+ }
+
+ precAlpha ~ dgamma(.001, .001)
+ precTau ~ dgamma(.001, .001)
+ precDelta ~ dgamma(.001, .001)
+}"
+
+initfunction_2con <- function(chain){
+  return(list(
+    muAlpha = runif(2, .2, 4.9),
+    muTau = runif(2, .01, .05),
+    muDelta = runif(2, -9.9, 9.9),
+    precAlpha = runif(1, .01, 100),
+    precTau = runif(1, .01, 100),
+    precDelta = runif(1, .01, 100),
+    y = yInit,
+    .RNG.name = "lecuyer::RngStream",
+    .RNG.seed = sample.int(1e10, 1, replace = F)))
+}
+
+## 1.2 One condition ----
+
+model_1con <- "model {
+ #likelihood function
+ for (t in 1:nTrials) {
+ y[t] ~ dwiener(alpha[subject[t]],
+ tau[subject[t]],
+ 0.5,
+ delta[subject[t]])
+ }
+
+ for (s in 1:nSubjects) {
+ tau[s] ~ dnorm(muTau, precTau) T(.0001, 1)
+ delta[s] ~ dnorm(muDelta, precDelta) T(-10, 10)
+ alpha[s] ~ dnorm(muAlpha, precAlpha) T(.1, 5)
+ }
+
[…]

### 2024-09-12 scripts/1_study/04_parse_ddm.R

@@ -1,0 +1,637 @@
+
+# 1. Load objects ------------------------------------------------------------
+
+source("scripts/0_functions/ddm_functions.R")
+
+load("data/clean_data_long.RData")
+
+# Load DDM model objects
+load("analysis_objects/ddm_flanker_mod1.RData")
+load("analysis_objects/ddm_simon_mod1.RData")
+load("analysis_objects/ddm_colorshape_mod1.RData")
+load("analysis_objects/ddm_animacysize_mod1.RData")
+load("analysis_objects/ddm_globallocal_mod1.RData")
+load("analysis_objects/ddm_posner_mod1.RData")
+
+load("data/colorshape_ddm_ml.RData")
+
+# Load ID matches to link DDM IDs to LISS IDs
+
+flanker_id_matches <- flanker_clean |>
+  distinct(nomem_encr) |>
+  mutate(nomem_encr_num = 1:n())
+
+simon_id_matches <- simon_clean |>
+  distinct(nomem_encr) |>
+  mutate(nomem_encr_num = 1:n())
+
+colorshape_id_matches <- colorshape_clean |>
+  distinct(nomem_encr) |>
+  mutate(nomem_encr_num = 1:n())
+
+animacysize_id_matches <- animacysize_clean |>
+  distinct(nomem_encr) |>
+  mutate(nomem_encr_num = 1:n())
+
+globallocal_id_matches <- globallocal_clean |>
+  distinct(nomem_encr) |>
+  mutate(nomem_encr_num = 1:n())
+
+posner_id_matches <- posner_clean |>
+  distinct(nomem_encr) |>
+  mutate(nomem_encr_num = 1:n())
+
+# 1. Extract traces -------------------------------------------------------
+
+
+flanker_mod1_traces <- extract_traces_2con(mcmc_flanker_mod1, chains = 3, iterations = 1000)
+simon_mod1_traces <- extract_traces_2con(mcmc_simon_mod1, chains = 3, iterations = 1000)
+colorshape_mod1_traces <- extract_traces_2con(mcmc_colorshape_mod1, chains = 3, iterations = 1000)
+animacysize_mod1_traces <- extract_traces_2con(mcmc_animacysize_mod1, chains = 3, iterations = 1000)
+globallocal_mod1_traces <- extract_traces_2con(mcmc_globallocal_mod1, chains = 3, iterations = 1000)
[…]

### 2024-09-12 scripts/1_study/05_ivs.R

@@ -1,0 +1,369 @@
+
+iv_data_2024 <- haven::read_sav('data/L_CognitiveAdversity_1.0p.sav')
+iv_data_2023 <- haven::read_sav('data/liss_data/L_Cognitieve_Vaardigheden_v2_1.0p.sav')
+
+
+
+# Libraries ---------------------------------------------------------------
+
+library(tidyverse)
+library(haven)
+library(openxlsx)
+
+# 1. Read data ------------------------------------------------------------
+
+## 1.1 IDs of participants selected for new study ----
+
+pp_ids <- iv_data_2024 |>
+  pull(nomem_encr) |>
+  as.numeric()
+
+## 1.2 Basic demographic variables ----
+
+background_vars <- list.files('data/liss_data/background', pattern = '.sav', full.names = TRUE) |>
+  map(function(x){haven::read_sav(x)})
+
+# The first eight waves do not contain household-level income (but do contain person-level income), so
+# are treated separately
+background_vars01 <- background_vars[1:8] |>
+  map(function(x){
+    x |>
+      as_tibble() |>
+      select(nomem_encr, nohouse_encr, matches("positie"), wave, belbezig, leeftijd, matches('sted'), aantalki, partner, nettoink) |>
+      # Compute household income by hand
+      group_by(wave, nohouse_encr) |>
+      mutate(
+        nettohh_f = sum(nettoink, na.rm = T)
+      ) |>
+      ungroup()
+  })
+
+background_vars02 <- background_vars[9:length(background_vars)] |>
+  map(function(x){
+    x |>
+      as_tibble() |>
+      select(nomem_encr, matches("positie"), nohouse_encr, wave, belbezig, leeftijd, sted, aantalki, partner, nettohh_f)
+  })
+
+
+background_parsed <- bind_rows(background_vars01, background_vars02) |>
+  separate(col = wave, into = c("year", "month"), sep = 4) |>
+  mutate(across(c(year, month), as.character))
+
+write_csv(background_parsed, "data/liss_data/background/background_parsed.csv")
+
+
+## 1.3 Material deprivation ----
+
[…]

### 2024-09-12 scripts/1_study/06_sem.R

@@ -1,0 +1,760 @@
+
+# 1. Load data ------------------------------------------------------------
+
+load("data/ddm_clean.RData")
+iv_data <- read_csv('data/liss_data/full_data.csv') |>
+  filter(nomem_encr %in% unique(ddm_clean$nomem_encr)) |>
+  mutate(across(c(p_scar_m, vict_sum, neigh_thr02_m, threat_comp, age, edu, child_adv), ~scale(.) |> as.numeric()))
+
+ddm_clean <- ddm_clean |>
+  mutate(across(-nomem_encr, ~scale(.) |> as.numeric()))
+
+full_data <- full_join(iv_data, ddm_clean)
+
+# 2. Drift Rate sub-model -------------------------------------------------
+
+## 2.1 Model 1 ----
+
+# Drift rate model 1 (Attention shifting and inhibition separate)
+mod_v1 <-
+  ' # Latent factors
+    ps =~ 1*pos_v_c + flanker_inc_v_c + flanker_con_v_c + simon_inc_v_c + simon_con_v_c + cs_sw_v_c + cs_rep_v_c + gl_sw_v_c + gl_rep_v_c + as_sw_v_c + as_rep_v_c
+    inh =~ 1*flanker_inc_v_c + simon_inc_v_c
+    as =~ 1*cs_sw_v_c + gl_sw_v_c + as_sw_v_c
+
+    # Covariances
+    inh ~~ as + 0*ps
+    as ~~ 0*ps
+
+    flanker_inc_v_c ~~ flanker_con_v_c
+    simon_inc_v_c   ~~ simon_con_v_c
+    cs_sw_v_c       ~~ cs_rep_v_c
+    gl_sw_v_c       ~~ gl_rep_v_c
+    as_sw_v_c       ~~ as_rep_v_c
+
+    #### INTERCEPTS ####
+
+    pos_v_c         ~ 0
+    flanker_inc_v_c ~ 0
+    flanker_con_v_c ~ 0
+    simon_inc_v_c   ~ 0
+    simon_con_v_c   ~ 0
+    cs_sw_v_c       ~ 0
+    cs_rep_v_c      ~ 0
+    gl_sw_v_c       ~ 0
+    gl_rep_v_c      ~ 0
+    as_sw_v_c       ~ 0
+    as_rep_v_c      ~ 0
+
+    as ~~ 0*as
+  '
+
[…]

### 2024-09-12 scripts/1_study/07_exploratory.R

@@ -1,0 +1,186 @@
+
+# 1. Load data ------------------------------------------------------------
+
+load("data/ddm_clean.RData")
+iv_data <- read_csv('data/liss_data/full_data.csv') |>
+  filter(nomem_encr %in% unique(ddm_clean$nomem_encr)) |>
+  mutate(across(c(child_dep, child_thr), ~scale(.) |> as.numeric()))
+
+ddm_clean <- ddm_clean |>
+  mutate(across(-nomem_encr, ~scale(.) |> as.numeric()))
+
+full_data <- full_join(iv_data, ddm_clean)
+
+
+# 5. Deprivation model ----------------------------------------------------
+
+## 5.1 Primary model (without threat exposure as confounder)
+
+mod_expl <-
+  '
+    #### LATENT FACTORS ####
+
+    # General factors #
+    gen_v =~ 1*pos_v_c + flanker_inc_v_c + flanker_con_v_c + simon_inc_v_c + simon_con_v_c + cs_sw_v_c + cs_rep_v_c + gl_sw_v_c + gl_rep_v_c + as_sw_v_c + as_rep_v_c
+    gen_a =~ 1*pos_a_c + flanker_inc_a_c + flanker_con_a_c + simon_inc_a_c + simon_con_a_c + cs_sw_a_c + cs_rep_a_c + gl_sw_a_c + gl_rep_a_c + as_sw_a_c + as_rep_a_c
+    gen_t =~ 1*pos_t_c + flanker_inc_t_c + flanker_con_t_c + simon_inc_t_c + simon_con_t_c + cs_sw_t_c + cs_rep_t_c + gl_sw_t_c + gl_rep_t_c + as_sw_t_c + as_rep_t_c
+
+    # Specific factors #
+    common_v =~ 1*flanker_inc_v_c + simon_inc_v_c + cs_sw_v_c  + gl_sw_v_c + as_sw_v_c
+    common_a =~ 1*flanker_inc_a_c + simon_inc_a_c + cs_sw_a_c  + gl_sw_a_c + as_sw_a_c
+
+
+    #### COVARIANCES ####
+
+    # Between general factors #
+    gen_v ~~ gen_a + gen_t
+    gen_a ~~ gen_t
+
+    # Between latent inhibition drift rate and shifting drift rate #
[…]

### 2024-09-12 scripts/liss/01_create_scales.R

@@ -1,0 +1,377 @@
+# Packages ----------------------------------------------------------------
+library(psych)
+library(tidyverse)
+library(qualtRics)
+library(jsonlite)
+library(here)
+library(sjlabelled)
+
+
+# Data --------------------------------------------------------------------
+test_data <- read_delim("data/L_CognitiveAdversity_4.csv", delim = ";") |>
+  filter(respondent == '-999')
+
+
+test_data <- test_data |>
+  mutate(
+    data_simon = case_when(
+      `volgorde[1]` == '1' ~ JSON1.4,
+      `volgorde[2]` == '1' ~ JSON2.4,
+      `volgorde[3]` == '1' ~ JSON3.4,
+      `volgorde[4]` == '1' ~ JSON4.4,
+      `volgorde[5]` == '1' ~ JSON5.4,
+      `volgorde[6]` == '1' ~ JSON6.4
+    ),
+    data_flanker = case_when(
+      `volgorde[1]` == '2' ~ JSON1.4,
+      `volgorde[2]` == '2' ~ JSON2.4,
+      `volgorde[3]` == '2' ~ JSON3.4,
+      `volgorde[4]` == '2' ~ JSON4.4,
+      `volgorde[5]` == '2' ~ JSON5.4,
+      `volgorde[6]` == '2' ~ JSON6.4
+    ),
+    data_globallocal2 = case_when(
+      `volgorde[1]` == '3' ~ JSON1.4,
+      `volgorde[2]` == '3' ~ JSON2.4,
+      `volgorde[3]` == '3' ~ JSON3.4,
+      `volgorde[4]` == '3' ~ JSON4.4,
+      `volgorde[5]` == '3' ~ JSON5.4,
+      `volgorde[6]` == '3' ~ JSON6.4
+    ),
+    data_colorshape = case_when(
+      `volgorde[1]` == '4' ~ JSON1.4,
+      `volgorde[2]` == '4' ~ JSON2.4,
+      `volgorde[3]` == '4' ~ JSON3.4,
+      `volgorde[4]` == '4' ~ JSON4.4,
+      `volgorde[5]` == '4' ~ JSON5.4,
+      `volgorde[6]` == '4' ~ JSON6.4
+    ),
+    data_animacysize = case_when(
+      `volgorde[1]` == '5' ~ JSON1.4,
+      `volgorde[2]` == '5' ~ JSON2.4,
[…]

### 2024-09-12 scripts/ml.R

@@ -1,0 +1,46 @@
+
+colorshape_clean <-
+  colorshape_clean |>
+  filter(rt > 0.425) |>
+  select(id = nomem_encr, rt, correct, condition)
+
+write_DDM_files(data = colorshape_clean, path = "data/cs", vars = c("rt", "correct", "condition"), task = "cs")
+
+
+fast_dm_settings(task = "cs",
+                 path = "data/cs",
+                 model_version = "_mod1",
+                 method = "ks",
+                 depend = c("depends v condition", "depends t0 condition", "depends a condition"),
+                 format = "TIME RESPONSE condition")
+
+# Compute DDM parameters
+execute_fast_dm(task = "cs", path = "data/cs", model_version = "_mod1")
+
+
+
+# Read DDM results
+colorshape_ddm_ml <- read_DDM(task = "cs", path = "data/cs", model_version = "_mod1") |>
+  select(
+    nomem_encr = id,
+    cs_rep_v      = v_1,
+    cs_sw_v       = v_2,
+    cs_rep_a      = a_1,
+    cs_sw_a       = a_2,
+    cs_rep_t      = t0_1,
+    cs_sw_t       = t0_2,
+    )
+
+save(colorshape_ddm_ml, file = "data/colorshape_ddm_ml.RData")
+
+cs_ddm |>
+  mutate(n = 1:n()) |>
+  select(a_1, a_2, v_1, v_2, t0_1, t0_2) |>
+  pivot_longer(everything(), names_to = 'parameter', values_to = "value") |>
+  ggplot(aes(value)) +
+  geom_histogram() +
+  facet_wrap(~parameter, scales = 'free')
+
+cs_ddm |>
+  select(a_1, a_2, v_1, v_2, t0_1, t0_2) |>
+  cor()

### 2024-09-12 supplement/stage_sup.R

@@ -1,0 +1,76 @@
+load("data/clean_data_long.RData")
+
+# RT histograms
+
+flanker_clean |>
+  ggplot(aes(rt)) +
+  geom_histogram() +
+  facet_wrap(~condition, scales = 'free')
+
+simon_clean |>
+  ggplot(aes(rt)) +
+  geom_histogram()+
+  facet_wrap(~condition, scales = 'free')
+
+colorshape_clean |>
+  ggplot(aes(rt)) +
+  geom_histogram()+
+  facet_wrap(~condition, scales = 'free')
+
+animacysize_clean |>
+  ggplot(aes(rt)) +
+  geom_histogram()+
+  facet_wrap(~condition, scales = 'free')
+
+globallocal_clean |>
+  ggplot(aes(rt)) +
+  geom_histogram()+
+  facet_wrap(~condition, scales = 'free')
+
+posner_clean |>
+  ggplot(aes(rt)) +
+  geom_histogram()+
+  facet_wrap(~condition, scales = 'free')
+
+# Accuracy histograms
+
+flanker_clean |>
+  group_by(nomem_encr, condition) |>
+  summarise(acc = sum(correct)/n()*100) |>
+  ggplot(aes(acc)) +
+  geom_histogram() +
+  facet_wrap(~condition, scales = 'free')
+
+simon_clean |>
+  group_by(nomem_encr, condition) |>
+  summarise(acc = sum(correct)/n()*100) |>
+  ggplot(aes(acc)) +
+  geom_histogram() +
+  facet_wrap(~condition, scales = 'free')
+
+colorshape_clean |>
+  group_by(nomem_encr, condition) |>
+  summarise(acc = sum(correct)/n()*100) |>
+  ggplot(aes(acc)) +
+  geom_histogram() +
+  facet_wrap(~condition, scales = 'free')
+
+animacysize_clean |>
+  group_by(nomem_encr, condition) |>
+  summarise(acc = sum(correct)/n()*100) |>
+  ggplot(aes(acc)) +
+  geom_histogram() +
+  facet_wrap(~condition, scales = 'free')
+
+globallocal_clean |>
+  group_by(nomem_encr, condition) |>
+  summarise(acc = sum(correct)/n()*100) |>
+  ggplot(aes(acc)) +
+  geom_histogram() +
+  facet_wrap(~condition, scales = 'free')
+
+posner_clean |>
+  group_by(nomem_encr) |>
[…]

### 2024-12-17 scripts/0_functions/ddm_functions.R
update scripts

@@ -1,0 +1,91 @@
+
+extract_traces <- function(mcmc, chains = 3, iterations = 12000) {
+
+  mcmc |>
+    select(starts_with("mu")) |>
+    mutate(
+      n = rep(1:iterations, chains),
+      chains = rep(1:chains, each = iterations))  |>
+    rename(
+      a  = muAlpha,
+      t0 = muTau,
+      v  = muDelta
+    ) |>
+    pivot_longer(-c(n, chains), names_to = 'parameter', values_to = 'value')
+}
+
+
+
+extract_traces_2con <- function(mcmc, chains = 3, iterations = 1000) {
+
+  mcmc |>
+    select(starts_with("mu")) |>
+    mutate(
+      n = rep(1:iterations, chains),
+      chains = rep(1:chains, each = iterations))  |>
+    rename(
+      a1  = `muAlpha[1]`,
+      a2  = `muAlpha[2]`,
+      t1 = `muTau[1]`,
+      t2 = `muTau[2]`,
+      v1  = `muDelta[1]`,
+      v2  = `muDelta[2]`
+    ) |>
+    pivot_longer(-c(n, chains), names_to = 'parameter', values_to = 'value')
+}
+
+extract_ddm_estimates <- function(mcmc, task_prefix, id_matches) {
+
+  mcmc |>
+    pivot_longer(everything(), names_to = "parameter", values_to = "estimated") |>
+    group_by(parameter) |>
+    summarise(estimated = mean(estimated, na.rm = T)) |>
+    filter(str_detect(parameter, pattern = 'deviance|^mu|^prec', negate = T)) |>
+    separate(col = parameter, into = c('parameter', 'nomem_encr_num'), sep = "\\[") |>
+    mutate(
+      nomem_encr_num = str_remove(nomem_encr_num, pattern = "\\]$")
+    ) |>
+    mutate(
+      nomem_encr_num = as.numeric(nomem_encr_num),
+      parameter = case_when(
+        parameter == 'alpha' ~ paste0(task_prefix, 'a'),
+        parameter == 'tau' ~ paste0(task_prefix, 't'),
+        parameter == 'delta' ~ paste0(task_prefix, 'v')
+      )) |>
+    pivot_wider(names_from = "parameter", values_from = "estimated") |>
+    left_join(id_matches) |>
+    select(-nomem_encr_num)
+}
+
+extract_ddm_estimates_2con <- function(mcmc, task_prefix, id_matches) {
+
+  mcmc |>
[…]

### 2024-12-17 scripts/0_functions/functions_DDM.R

@@ -1,0 +1,227 @@
+
+
+# DDM fitting helpers -----------------------------------------------------
+
+fast_dm_settings <- function(task, model_version = "", method = "ml", precision = 3, zr = 0.5, d = 0, szr = 0, sv = 0, st0 = 0, p = 0, depend = "", format) {
+
+  depend = str_c(depend, collapse = "\n")
+
+  spec <- tribble(
+    ~param,        ~setting,                                           ~ command,
+    "method",      method,                                             "{param} {setting}",
+    "precision",   as.character(precision),                            "{param} {setting}",
+    "set zr",      as.character(zr),                                   "{param} {setting}",
+    "set d",       as.character(d),                                    "{param} {setting}",
+    "set szr",     as.character(szr),                                  "{param} {setting}",
+    "set sv",      as.character(sv),                                   "{param} {setting}",
+    "set st0",     as.character(st0),                                  "{param} {setting}",
+    "set p",       as.character(p),                                    "{param} {setting}",
+    "",            depend,                                             "{param} {setting}",
+    "format",      format,                                             "{param} {setting}",
+    "",            "load *.dat",                                       "{param} {setting}",
+    "",            glue("log ddm_results_{task}{model_version}.lst"),  "{param} {setting}"
+  ) %>%
+    filter(setting != "") %>%
+    rowwise() %>%
+    transmute(command = glue(command)) %>%
+    glue_data("{command}")
+
+  message("\nThe following model specification was written to the DDM folder:\n\n")
+  print(spec)
+
+  write_file(str_c(spec, collapse = "\n"), here('data', "1_pilot", "DDM", glue("{task}{model_version}_ml.ctl")))
+}
+
+write_DDM_files <- function(data, vars = c("rt", "correct"), task) {
+
+  for (i in unique(data$id)) {
+    # 1. Filter subject i and write individual data files to data folder
+    change_DDM_i <- data %>%
[…]

### 2024-12-17 scripts/0_functions/general-functions.R

@@ -1,0 +1,169 @@
+fix_subject_ids <- function(data) {
+
+  # Selection of regular expressions that fix general mistakes in subject identifiers.
+  # Specifically, each subject ID should be preceded by 'NDAR_INV' followed by a unique code.
+
+  data |>
+    filter(!str_detect(str_to_lower(subj_idx), "test|tes|practice")) |>
+    mutate(
+      subj_idx = str_replace_all(str_to_lower(subj_idx), "^.*ndar", "ndar"),
+      subj_idx = str_replace_all(str_to_lower(subj_idx), "ndar_inv\\s|^ndar.*inv|ndarinv|ndar[^[:alnum:]]|^inv|^dar_inv|^nadr_inv|bdar_inv|ndra_inv|ndr_inv|ndat_inv", ""),
+      subj_idx = toupper(paste0("NDAR_INV", subj_idx)),
+      subj_idx = str_remove_all(subj_idx, "(-|_)[0-9]$"),
+      subj_idx = str_remove_all(subj_idx, "(-|\\s).*")
+    )
+}
+
+
+
+fix_id_typos <- function(data, reference, id = "subj_idx", distance = 2) {
+
+  # Function that tries to detect typos using the Levenshtein distance between unmatched ids and the reference list (i.e., the ids found in the summary score df)
+  # The Levenshtein distance counds the number of deletions, insertions and substitutions necessary to turn one string into another.
+  # Thus, a small Levensthein distance (e.g., 1 or 2) in subject IDs likely indicates a typo.
+
+  data_ids <- data |>
+    pull(id) |>
+    unique()
+
+  reference |>
+    as_tibble_col(column_name = "original")
+
+  mismatches <- data_ids[!data_ids %in% reference] %>%
+    as_tibble_col(column_name = "subj_idx") %>%
+    mutate(
+      similar = pmap(.l = ., function(subj_idx) {
+
+        reference |>
+          as_tibble_col(column_name = "ref") |>
+          mutate(original = subj_idx) |>
+          mutate(
+            string_dist = stringdist::stringdist(a = original, b = ref, method = 'lv')
+          ) |>
+          filter(string_dist == min(string_dist))
+      })
+    ) |>
+    unnest(similar)
+
+  mismatches |> select(subj_idx, string_dist) |>
+    distinct() |>
+    count(string_dist) |>
[…]

### 2024-12-17 scripts/1_study/01_parse.R

@@ -1,0 +1,235 @@
+# 1. Packages -------------------------------------------------------------
+library(psych)
+library(tidyverse)
+library(jsonlite)
+library(here)
+library(sjlabelled)
+
+
+# 2. Data -----------------------------------------------------------------
+raw_data <- read_delim("data/L_CognitiveAdversity_TASKDATA.csv", delim = ";")
+
+
+
+raw_data <- raw_data |>
+  mutate(
+    data_simon = case_when(
+      volgorde__1 == '1' ~ JSON1.4,
+      volgorde__2 == '1' ~ JSON2.4,
+      volgorde__3 == '1' ~ JSON3.4,
+      volgorde__4 == '1' ~ JSON4.4,
+      volgorde__5 == '1' ~ JSON5.4,
+      volgorde__6 == '1' ~ JSON6.4
+    ),
+    data_flanker = case_when(
+      volgorde__1 == '2' ~ JSON1.4,
+      volgorde__2 == '2' ~ JSON2.4,
+      volgorde__3 == '2' ~ JSON3.4,
+      volgorde__4 == '2' ~ JSON4.4,
+      volgorde__5 == '2' ~ JSON5.4,
+      volgorde__6 == '2' ~ JSON6.4
+    ),
+    data_globallocal2 = case_when(
+      volgorde__1 == '3' ~ JSON1.4,
+      volgorde__2 == '3' ~ JSON2.4,
+      volgorde__3 == '3' ~ JSON3.4,
+      volgorde__4 == '3' ~ JSON4.4,
+      volgorde__5 == '3' ~ JSON5.4,
+      volgorde__6 == '3' ~ JSON6.4
+    ),
+    data_colorshape = case_when(
+      volgorde__1 == '4' ~ JSON1.4,
+      volgorde__2 == '4' ~ JSON2.4,
+      volgorde__3 == '4' ~ JSON3.4,
+      volgorde__4 == '4' ~ JSON4.4,
+      volgorde__5 == '4' ~ JSON5.4,
+      volgorde__6 == '4' ~ JSON6.4
+    ),
+    data_animacysize = case_when(
+      volgorde__1 == '5' ~ JSON1.4,
+      volgorde__2 == '5' ~ JSON2.4,
+      volgorde__3 == '5' ~ JSON3.4,
+      volgorde__4 == '5' ~ JSON4.4,
+      volgorde__5 == '5' ~ JSON5.4,
[…]

### 2024-12-17 scripts/1_study/02_clean.R

@@ -1,0 +1,367 @@
+library(tidyverse)
+library(sjlabelled)
+
+load("data/raw_data_long.RData")
+source("scripts/0_functions/functions_exclusions.R")
+source("scripts/0_functions/create_codebook.R")
+
+# Fast: < 250 ms
+# very_slow: > 15s
+# slow: > 3SD of intraindividual mean
+# nval_trials: number of participants with fewer than 20 valid trials
+# n_raw: sample size prior to exclusions
+# n_clean: sample size after exclusions
+exclusions <- list()
+
+exclusions$sample$started <- (length(all_complete_ids) + length(all_incomplete_ids))
+exclusions$sample$not_finish <- length(all_incomplete_ids)
+exclusions$sample$finish <- length(all_complete_ids)
+
+# 1. Preregistered exclusions ---------------------------------------------
+
+## 1.1 Flanker task ----
+
+flanker_clean <- flanker_raw |>
+  filter(rt > 0, rt < 10) |> # Remove excessively long RTs
+  group_by(nomem_encr, condition) |>
+  mutate(
+    ex_fast_outlier = ifelse(rt < 0.250, TRUE, FALSE),
+    ex_slow_outlier = ifelse(as.numeric(scale(log(rt))) > 3, TRUE, FALSE)
+  ) |>
+  group_by(nomem_encr) |>
+  mutate(
+    ex_chance_perf  = ifelse(sum(correct==1)/n()*100 < gbinom(64, 0.50), TRUE, FALSE)
+  ) |>
+  ungroup()
+
+exclusions$flanker <-
+  list(
+    fast      = paste(round(sum(flanker_clean$ex_fast_outlier)/nrow(flanker_clean)*100, 1), "%"),
+    very_slow = paste(round(sum(flanker_raw$rt > 10)/nrow(flanker_raw)*100, 2), "%"),
+    slow      = paste(round(sum(flanker_clean$ex_slow_outlier, na.rm = T)/nrow(flanker_clean)*100, 1), "%"),
+    chance    = flanker_clean |> group_by(nomem_encr) |> summarise(n = sum(ex_chance_perf)) |> filter(n > 1) |> nrow(),
+    raw_n     = unique(flanker_clean$nomem_encr) |> length()
+  )
+
+flanker_clean <- flanker_clean |>
+  filter(ex_slow_outlier == FALSE, ex_fast_outlier == FALSE, ex_chance_perf == FALSE) |>
+  group_by(nomem_encr) |>
[…]

### 2024-12-17 scripts/1_study/03_ddm.R

@@ -1,0 +1,507 @@
+load("data/clean_data_long.RData")
+source("scripts/0_functions/DBDA2E-utilities.R")
+
+library(runjags)
+
+# 1. Models ---------------------------------------------------------------
+
+## 1.1 Two conditions ----
+
+model_2con <- "model {
+ #likelihood function
+ for (t in 1:nTrials) {
+
+  y[t] ~ dwiener(alpha[condition[t], subject[t]],
+                 tau[condition[t], subject[t]],
+                 0.5,
+                 delta[condition[t], subject[t]])
+ }
+
+ for (s in 1:nSubjects) {
+  for (c in 1:nCon) {
+    tau[c, s] ~ dnorm(muTau[c], precTau) T(.0001, 1)
+    alpha[c, s] ~ dnorm(muAlpha[c], precAlpha) T(.1, 5)
+    delta[c, s] ~ dnorm(muDelta[c] , precDelta) T(-10, 10)
+  }
+ }
+
+ #priors
+ for (c in 1:nCon){
+  muTau[c] ~ dunif(.0001, 1)
+  muAlpha[c] ~ dunif(.1, 5)
+  muDelta[c] ~ dunif(-10, 10)
+ }
+
+ precAlpha ~ dgamma(.001, .001)
+ precTau ~ dgamma(.001, .001)
+ precDelta ~ dgamma(.001, .001)
+}"
+
+initfunction_2con <- function(chain){
+  return(list(
+    muAlpha = runif(2, .2, 4.9),
+    muTau = runif(2, .01, .05),
+    muDelta = runif(2, -9.9, 9.9),
+    precAlpha = runif(1, .01, 100),
+    precTau = runif(1, .01, 100),
+    precDelta = runif(1, .01, 100),
+    y = yInit,
+    .RNG.name = "lecuyer::RngStream",
+    .RNG.seed = sample.int(1e10, 1, replace = F)))
+}
+
+## 1.2 One condition ----
+
+model_1con <- "model {
+ #likelihood function
+ for (t in 1:nTrials) {
+ y[t] ~ dwiener(alpha[subject[t]],
+ tau[subject[t]],
+ 0.5,
+ delta[subject[t]])
+ }
+
+ for (s in 1:nSubjects) {
+ tau[s] ~ dnorm(muTau, precTau) T(.0001, 1)
[…]

### 2024-12-17 scripts/1_study/04_parse_ddm.R

@@ -1,0 +1,547 @@
+
+# 1. Load objects ------------------------------------------------------------
+
+source("scripts/0_functions/ddm_functions.R")
+
+load("data/clean_data_long.RData")
+
+# Load DDM model objects
+load("analysis_objects/ddm_flanker_mod1.RData")
+load("analysis_objects/ddm_simon_mod1.RData")
+load("analysis_objects/ddm_colorshape_mod1.RData")
+load("analysis_objects/ddm_animacysize_mod1.RData")
+load("analysis_objects/ddm_globallocal_mod1.RData")
+load("analysis_objects/ddm_posner_mod1.RData")
+
+# Load ID matches to link DDM IDs to LISS IDs
+
+flanker_id_matches <- flanker_clean |>
+  distinct(nomem_encr) |>
+  mutate(nomem_encr_num = 1:n())
+
+simon_id_matches <- simon_clean |>
+  distinct(nomem_encr) |>
+  mutate(nomem_encr_num = 1:n())
+
+colorshape_id_matches <- colorshape_clean |>
+  distinct(nomem_encr) |>
+  mutate(nomem_encr_num = 1:n())
+
+animacysize_id_matches <- animacysize_clean |>
+  distinct(nomem_encr) |>
+  mutate(nomem_encr_num = 1:n())
+
+globallocal_id_matches <- globallocal_clean |>
+  distinct(nomem_encr) |>
+  mutate(nomem_encr_num = 1:n())
+
+posner_id_matches <- posner_clean |>
+  distinct(nomem_encr) |>
+  mutate(nomem_encr_num = 1:n())
+
+# 1. Extract traces -------------------------------------------------------
+
+
+flanker_mod1_traces <- extract_traces_2con(mcmc_flanker_mod1, chains = 3, iterations = 1000)
+simon_mod1_traces <- extract_traces_2con(mcmc_simon_mod1, chains = 3, iterations = 1000)
+colorshape_mod1_traces <- extract_traces_2con(mcmc_colorshape_mod1, chains = 3, iterations = 1000)
+animacysize_mod1_traces <- extract_traces_2con(mcmc_animacysize_mod1, chains = 3, iterations = 1000)
+globallocal_mod1_traces <- extract_traces_2con(mcmc_globallocal_mod1, chains = 3, iterations = 1000)
+posner_mod1_traces <- extract_traces(mcmc_posner_mod1, chains = 3, iterations = 1000)
+
+
+
[…]

### 2024-12-17 scripts/1_study/05_ivs.R

@@ -1,0 +1,395 @@
+
+iv_data_2024 <- haven::read_sav('data/L_CognitiveAdversity_1.0p.sav')
+iv_data_2023 <- haven::read_sav('data/liss_data/L_Cognitieve_Vaardigheden_v2_1.0p.sav')
+
+load("data/exclusions.RData")
+
+# Libraries ---------------------------------------------------------------
+
+library(tidyverse)
+library(haven)
+library(openxlsx)
+
+# 1. Read data ------------------------------------------------------------
+
+## 1.1 Included ids
+
+pp_ids <- exclusions$sample$ids
+
+## 1.2 Basic demographic variables ----
+
+background_vars <- list.files('data/liss_data/background', pattern = '.sav', full.names = TRUE) |>
+  map(function(x){haven::read_sav(x)})
+
+# The first eight waves do not contain household-level income (but do contain person-level income), so
+# are treated separately
+background_vars01 <- background_vars[1:8] |>
+  map(function(x){
+    x |>
+      as_tibble() |>
+      select(nomem_encr, nohouse_encr, matches("positie"), wave, belbezig, leeftijd, matches('sted'), aantalki, partner, nettoink) |>
+      # Compute household income by hand
+      group_by(wave, nohouse_encr) |>
+      mutate(
+        nettohh_f = sum(nettoink, na.rm = T)
+      ) |>
+      ungroup()
+  })
+
+background_vars02 <- background_vars[9:length(background_vars)] |>
+  map(function(x){
+    x |>
+      as_tibble() |>
+      select(nomem_encr, matches("positie"), nohouse_encr, wave, belbezig, leeftijd, sted, aantalki, partner, nettohh_f)
+  })
+
+
+background_parsed <- bind_rows(background_vars01, background_vars02) |>
+  separate(col = wave, into = c("year", "month"), sep = 4) |>
+  mutate(across(c(year, month), as.character))
+
+write_csv(background_parsed, "data/liss_data/background/background_parsed.csv")
+
+
+## 1.3 Material deprivation ----
+
[…]

### 2024-12-17 scripts/1_study/06_sem.R

@@ -1,0 +1,1197 @@
+library(tidyverse)
+library(lavaan)
+
+equivalence_test_ml = function(sesoi, ml_effect, ml_se){
+  p1 = 1 - pnorm( ml_effect, mean = -sesoi, sd = ml_se)
+  p2 =     pnorm( ml_effect, mean =  sesoi, sd = ml_se)
+  return(max(p1, p2))
+}
+# 1. Load data ------------------------------------------------------------
+
+load("data/ddm_clean.RData")
+iv_data <- read_csv('data/liss_data/full_data.csv') |>
+  filter(nomem_encr %in% unique(ddm_clean$nomem_encr)) |>
+  mutate(across(c(p_scar_m, vict_sum, neigh_thr02_m, threat_comp, age, edu, child_adv), ~scale(.) |> as.numeric()))
+
+ddm_clean <- ddm_clean |>
+  mutate(across(-nomem_encr, ~scale(.) |> as.numeric()))
+
+full_data <- full_join(iv_data, ddm_clean)
+
+# 2. Drift Rate sub-model -------------------------------------------------
+
+## 2.1 Model 1 ----
+
+# Drift rate model 1 (Attention shifting and inhibition separate)
+mod_v1 <-
+  ' # Latent factors
+    ps =~ 1*pos_v_c + flanker_inc_v_c + flanker_con_v_c + simon_inc_v_c + simon_con_v_c + cs_sw_v_c + cs_rep_v_c + gl_sw_v_c + gl_rep_v_c + as_sw_v_c + as_rep_v_c
+    inh =~ 1*flanker_inc_v_c + simon_inc_v_c
+    as =~ 1*as_sw_v_c + cs_sw_v_c + gl_sw_v_c
+
+    # Covariances
+    inh ~~ as + 0*ps
+    as ~~ 0*ps
+
+    flanker_inc_v_c ~~ flanker_con_v_c
+    simon_inc_v_c   ~~ simon_con_v_c
+    cs_sw_v_c       ~~ cs_rep_v_c
+    gl_sw_v_c       ~~ gl_rep_v_c
+    as_sw_v_c       ~~ as_rep_v_c
+
+    #### INTERCEPTS ####
+
+    pos_v_c         ~ 0
+    flanker_inc_v_c ~ 0
+    flanker_con_v_c ~ 0
[…]

### 2024-12-17 scripts/1_study/07_exploratory.R

@@ -1,0 +1,1150 @@
+library(tidyverse)
+library(lavaan)
+
+
+# 1. Load data ------------------------------------------------------------
+
+load('analysis_objects/results_confirmatory.RData')
+
+load("data/ddm_clean.RData")
+iv_data <- read_csv('data/liss_data/full_data.csv') |>
+  filter(nomem_encr %in% unique(ddm_clean$nomem_encr)) |>
+  mutate(across(c(child_dep, child_thr), ~scale(.) |> as.numeric()))
+
+ddm_clean <- ddm_clean |>
+  mutate(across(-nomem_encr, ~scale(.) |> as.numeric()))
+
+full_data <- full_join(iv_data, ddm_clean)
+
+data_long <- load("data/clean_data_long.RData")
+
+equivalence_test_ml = function(sesoi, ml_effect, ml_se){
+  p1 = 1 - pnorm( ml_effect, mean = -sesoi, sd = ml_se)
+  p2 =     pnorm( ml_effect, mean =  sesoi, sd = ml_se)
+  return(max(p1, p2))
+}
+
+
+# 2. Associations with childhood threat --------------------------------
+
+## 2.1 Model specification ----
+mod_expl_ch_thr <-
+  '
+    #### LATENT FACTORS ####
+
+    # General factors #
+    gen_v =~ 1*pos_v_c + flanker_inc_v_c + flanker_con_v_c + simon_inc_v_c + simon_con_v_c + cs_sw_v_c + cs_rep_v_c + gl_sw_v_c + gl_rep_v_c + as_sw_v_c + as_rep_v_c
+    gen_a =~ 1*pos_a_c + flanker_inc_a_c + flanker_con_a_c + simon_inc_a_c + simon_con_a_c + cs_sw_a_c + cs_rep_a_c + gl_sw_a_c + gl_rep_a_c + as_sw_a_c + as_rep_a_c
+    gen_t =~ 1*pos_t_c + flanker_inc_t_c + flanker_con_t_c + simon_inc_t_c + simon_con_t_c + cs_sw_t_c + cs_rep_t_c + gl_sw_t_c + gl_rep_t_c + as_sw_t_c + as_rep_t_c
+
+    # Specific factors #
[…]

### 2024-12-17 scripts/liss/01_create_scales.R

@@ -1,377 +1,0 @@
-# Packages ----------------------------------------------------------------
-library(psych)
-library(tidyverse)
-library(qualtRics)
-library(jsonlite)
-library(here)
-library(sjlabelled)
-
-
-# Data --------------------------------------------------------------------
-test_data <- read_delim("data/L_CognitiveAdversity_4.csv", delim = ";") |>
-  filter(respondent == '-999')
-
-
-test_data <- test_data |>
-  mutate(
-    data_simon = case_when(
-      `volgorde[1]` == '1' ~ JSON1.4,
-      `volgorde[2]` == '1' ~ JSON2.4,
-      `volgorde[3]` == '1' ~ JSON3.4,
-      `volgorde[4]` == '1' ~ JSON4.4,
-      `volgorde[5]` == '1' ~ JSON5.4,
-      `volgorde[6]` == '1' ~ JSON6.4
-    ),
-    data_flanker = case_when(
-      `volgorde[1]` == '2' ~ JSON1.4,
-      `volgorde[2]` == '2' ~ JSON2.4,
-      `volgorde[3]` == '2' ~ JSON3.4,
-      `volgorde[4]` == '2' ~ JSON4.4,
-      `volgorde[5]` == '2' ~ JSON5.4,
-      `volgorde[6]` == '2' ~ JSON6.4
-    ),
-    data_globallocal2 = case_when(
-      `volgorde[1]` == '3' ~ JSON1.4,
-      `volgorde[2]` == '3' ~ JSON2.4,
-      `volgorde[3]` == '3' ~ JSON3.4,
-      `volgorde[4]` == '3' ~ JSON4.4,
-      `volgorde[5]` == '3' ~ JSON5.4,
-      `volgorde[6]` == '3' ~ JSON6.4
-    ),
-    data_colorshape = case_when(
-      `volgorde[1]` == '4' ~ JSON1.4,
-      `volgorde[2]` == '4' ~ JSON2.4,
-      `volgorde[3]` == '4' ~ JSON3.4,
-      `volgorde[4]` == '4' ~ JSON4.4,
-      `volgorde[5]` == '4' ~ JSON5.4,
-      `volgorde[6]` == '4' ~ JSON6.4
-    ),
-    data_animacysize = case_when(
-      `volgorde[1]` == '5' ~ JSON1.4,
-      `volgorde[2]` == '5' ~ JSON2.4,
[…]

### 2024-12-18 manuscript/manuscript.qmd
updated manuscript .qmd

--- 2024-06-04 manuscript/manuscript.qmd
@@
   word_document: default
-editor: 
-  markdown: 
-    wrap: sentence
 ---
@@
 library(officer)
+
+source('../scripts/0_functions/general-functions.R')
+
+load("../data/exclusions.RData")
+load("../analysis_objects/n_sessions.RData")
+load("../analysis_objects/results_confirmatory.RData")
+load("../analysis_objects/results_exploratory.RData")
+load("../analysis_objects/rendered_figures.RData")
+load("../analysis_objects/rendered_tables.RData")
 
@@
 
-#### **Attention shifting and inhibition abilities among people living in adverse conditions**
+#### **Adversity is associated with lower general processing speed rather than specific executive functioning abilities**
 
@@
 
-# Corresponding author
+# Author Note
+
+This research was conducted in part using ODISSEI, the Open Data Infrastructure for Social Science and Economic Innovations (<https://ror.org/03m8v6t10>).
+WEF’s contributions have been supported by the Dutch Research Council (V1.Vidi.195.130) and the James S. McDonnell Foundation (https://doi.org/10.37717/220020502).
+
+Stefan Vermeent served as lead for conceptualization, data curation, methodology, formal analysis, and writing the original draft.
+Anna-Lena Schubert served in a supporting role for conceptualization, formal analysis, methodology, and writing-review and editing.
+Willem E.
+Frankenhuis served in a supporting role for conceptualization, methodology, supervision, funding acquisition, and writing-review and editing.
+
+The preregistration, analysis scripts, and study materials are available on the article's Github repository ([https://stefanvermeent.github.io/liss_ef_2024/](https://stefanvermeent.github.io/liss_ef_2024/)).
+In this paper, we make use of data from the LISS panel (Longitudinal Internet studies for the Social Sciences) managed by the non-profit research institute Centerdata (Tilburg University, the Netherlands).
+All datasets are available to researchers in the LISS data archive, after signing a data use agreement ([https://www.dataarchive.lissdata.nl/study-units/view/1](https://www.dataarchive.lissdata.nl/study-units/view/1)).
 
@@
 
-# Data Availability
+We have no conflict of interest to disclose.
 
-All scripts and materials needed to reproduce the findings are available on the article's Github repository ([https://stefanvermeent.github.io/liss_wm_profiles_2023/](https://stefanvermeent.github.io/liss_wm_profiles_2023/)).
-We also include instructions on how to reproduce each step of our analyses. 
-In this paper, we make use of data from the LISS panel (Longitudinal Internet studies for the Social Sciences) managed by the non-profit research institute Centerdata (Tilburg University, the Netherlands).
-All datasets are available in the LISS data archive. 
-Researchers who want to access the data are required to sign a statement confirming that information about individual persons, households, etc., will not be released to others (go to [https://statements.centerdata.nl](https://statements.centerdata.nl) for more information).
+<br>
 
-# Funding statement
-This research was conducted in whole or in part using ODISSEI, the Open Data Infrastructure for Social Science and Economic Innovations(<https://ror.org/03m8v6t10>).
-WEF’s contributions have been supported by the Dutch Research Council (V1.Vidi.195.130) and the James S. McDonnell Foundation (https://doi.org/10.37717/220020502).
+**word count:** 9573
 
-# Disclosures
+{{< pagebreak >}}
 
-We declare no conflicts of interest.
+# Abstract
 
-# Ethics Approval Statement
+Exposure to adversity may impair executive functioning (i.e., deficit frameworks), but could also enhance, or leave intact, specific EF abilities (i.e., adaptation frameworks).
+Both frameworks often use raw performance (e.g., speed) to estimate EF ability. 
+However, this approach (1) conflates different cognitive processes, and (2) generally does not distinguish specific EF abilities from processes that are shared across EF tasks, such as general processing speed.
+Here, we integrate deficit and adaptation frameworks by building bridges with mathematical and cognitive psychology.
+Specifically, we use cognitive modeling (Drift Diffusion Modeling) to isolate different cognitive processes: speed of information accumulation, response caution, and speed of stimulus encoding and response execution.
+We then use structural equation modeling to investigate whether associations between adversity and cognitive processes are task-general or ability-specific.
+We recruited 1061 participants from the Dutch LISS panel.
+Participants completed a basic processing speed task, two inhibition tasks, and three attention-shifting tasks.
+We measured exposure to threat and material deprivation in childhood and adulthood.
+Exposure to threat (but not material deprivation) in adulthood was negatively associated with task-general processing speed.
+After accounting for task-general processes, remaining variance was not related to either inhibition or attention-shifting ability.
+Non-preregistered analyses showed that childhood exposure to material deprivation and threat were negatively associated with (1) general processing speed, and (2) task-specific information accumulation.
+The latter reflected unique features of individual tasks, rather than specific EF abilities.
+Taken together, these results suggest that adversity researchers overestimate associations between adversity and specific EF abilities when analyzing raw performance.
 
-This study was approved by the Ethics Review Board of the Faculty of Social & Behavioural Sciences of Utrecht University (FETC20-490) and the Ethics committee for research in the Sciences and Life Sciences of the University of Amsterdam (FNWI-41_2023).
+# Public significance statement
 
-\pagebreak
+Decades of research suggests that adversity impairs executive functioning abilities, such as inhibition and attention shifting.
+These conclusions are often derived from analyzing raw performance, such as response times or accuracy.
+We use cognitive modeling to reveal that adversity lowers general processing speed, but not specific executive functioning abilities.
+This suggests that adversity research likely overestimates the extent to which adversity is associated with executive functioning abilities.
 
-Exposure to adversity (i.e., prolonged exposure to intense stress) is associated with robust, persisting changes in executive functioning (EF) [@ellis_2022; @farah_2006; @sheridan_2022; @sheridan_2014].
-Given the importance of EF---a set of abilities involved in the top-down, goal-directed regulation of behavior [@miyake_2000; @rey_mermet_2019]---much research has been devoted to understanding which EF abilities are most strongly associated with adversity.
-In some cases, these insights have already been used to screen for early EF deficits, and to develop interventions to help bridge performance gaps [@distefano_2021].
+*Keywords*: adversity, executive functioning, Drift Diffusion Modeling, structural equation modeling
 
-A lot of what we know about the association between adversity and EF stems from analyzing raw performance, most often response times (RT) or accuracy rates.
-In addition, it is common for studies to only include a single task per EF ability.
-For example, previous studies have inferred difficulties in ignoring distractions (inhibition) in people from adversity using a single inhibition task [e.g., @farah_2006; @fields_2021; @mezzacappa_2004; @noble_2005].
-The same thing is true for observed adaptive benefits in the ability to shift attention between tasks (attention shifting) [@fields_2021; @howard_2020; @nweze_2021; @young_2022; @mittal_2015].
-In effect, an important assumption underlying many deficit and adaptation findings is that raw task performance offers a reliable, relatively pure estimate of the EF ability [@white_2022].
+{{< pagebreak >}}
 
-Unfortunately, an increasing number of psychometric investigations into EF do not support this assumption [@krumm_2009; @loffler_2024; @rey_mermet_2019; @rouder_2019; @stahl_2014].
-A central issue is that raw task performance is not task-pure.
-Aside from the EF ability needed to solve the task, a person's RT is also influenced by other factors such as response caution (i.e., their relative emphasis on speed versus accuracy).
-Accounting for these factors using cognitive modeling, a recent study found that children with more household threat were slower on EF tasks partly because they were more cautious [@vermeent_2024a].
+Psychologists have used two main frameworks to explain associations between adversity exposure---such as material deprivation or threat---and executive functioning (EF); one focuses on deficits, the other on adaptation.
+Deficit frameworks predict that adverse experiences impair brain structure and function in ways that undermine social and cognitive abilities [@farah_2006; @mertz_2019; @rosen_2018; @ursache_2016a].
+Adaptation frameworks predict that adverse experiences can lead to the development of intact or enhanced EF abilities, specifically those abilities useful for solving challenges posed by adverse environments [@ellis_2022; @frankenhuis_2013; @frankenhuis_2020].
 
-Aside from different processing stages, performance across EF tasks is also influenced by general processing speed [@frischkorn_2019; @lerche_2020; @loffler_2024].
+Although the two frameworks make opposing predictions in some cases, they share the goal to reveal how adversity influences specific EF abilities [@farah_2006; @ellis_2022; @frankenhuis_2020; @johnson_2021].
+Knowledge about impaired abilities provides valuable targets for interventions designed to bridge achievement gaps.
+Conversely, knowledge about enhanced abilities can be leveraged in school or work contexts, for instance, by redesigning learning settings to match people’s unique strengths.
 
-[Might not be any ]
-Even more problematic are recent studies suggesting that performance on EF tasks taps into little more than general processing speed.
-For example, after accounting for other factors such as response caution, another study found that a common (latent) EF factor was fully explained by performance on really basic processing speed tasks [@loffler_2024].
-A few studies have focused on this distinction in the context of adversity and/or socioeconomical status.
-on a raw performance level, lower SES was found to be negatively correlated with a task-general factor, with additional associations on the task-specific level.
-In another
+Research within both frameworks typically employs the same inferential strategy, by estimating specific EF abilities based on raw performance on EF tasks (e.g., response speed or accuracy).
+For instance, on the one hand, children who live in less favorable environments tend to be slower and less accurate on inhibition tasks---which has been interpreted as an impaired ability to inhibit distractions [@farah_2006; @fields_2021; @mezzacappa_2004; @mittal_2015; @noble_2005].
+On the other hand, some studies report that adolescents and young adults with more exposure to threat and unpredictability might perform better at shifting their attention between different tasks, without sacrificing much performance on each task---which may be a useful adaptation for tracking information in fast changing and dangerous environments [@fields_2021; @howard_2020; @mittal_2015; @nweze_2021; @young_2022].
+In this paper, we will challenge the inferential strategy of estimating EF abilities based on raw performance. 
+We advocate for cognitive modeling as a viable alternative approach to estimating EF abilities.
 
-# # Implications and potential solutions for adversity research
+# Methodological challenges for measuring specific EF abilities
 
-An important implication of this work is that standard assessments---based on raw performance of single EF tasks---may be ill-suited at estimating associations between adversity and specific EF abilities.
-Patterns that may look like deficits (or adaptations) in specific abilities may not be driven by these abilities at all.
-For example, if adversity is associated with slower task-general processing speed, analyses focused on individual EF tasks could make it seem like many different abilities impaired, instead of a single, general process.
+Recent research shows that raw performance on EF tasks does not necessarily capture specific EF abilities, for two related reasons.
+First, raw performance on EF tasks is also influenced by processes other than the ability of interest, such as response caution and the speed of response execution [@hedge_2021; @ratcliff_2008; @stahl_2014].
+Thus, slower (or faster) responses in people with more adversity exposure need not reflect EF ability, but could, for instance, reflect a tendency to respond more cautiously [@vermeent_2024a].
+Second, raw performance on EF tasks is also influenced by general processes, such as basic speed of information processing [@frischkorn_2019; @loffler_2024; @weigard_2021].
+Thus, lower performance on multiple EF tasks may reflect a difference in a general process, rather than specific EF abilities [@bignardi_2024; @vermeent_2024a; @young_2024].
 
-Fortunately, these challenges can be tackled using a combination of cognitive modeling and structural equation modeling.
-As a first step, ability-relevant processing should be separated from ability-irrelevant processes, such as response caution or the time taken to prepare for a task [...].
-One effective way of doing this is through the use of cognitive models like the Drift Diffusion Model [DDM\; @forstmann_2016; @ratcliff_1998; @ratcliff_2008; @wagenmakers_2009].
-The DDM broadly divides the time taken on speeded EF tasks into a decision-making stage and a non-decision-making stage [see Figure 1].
-In the decision-making stage, people accumulate evidence for one of two responses (e.g., left or right button), until they have enough evidence for one response over the other.
-The amount of evidence a person requires can vary between people (as well as between tasks within the same person) [XXX], and is modeled as the distance between the decision boundaries for each respective response.
-Finally, the DDM models two non-decision stages: one at the start of the trial (e.g., task preparations, stimulus encoding) and one at the end of the trial (i.e., response execution).
+The standard inferential strategy of estimating EF abilities from raw performance limits the effectiveness of interventions. 
+For example, some interventions include training modules for specific EF tasks if performance on a task is below some prespecified cut-off [@distefano_2021].
+Performance on inhibition tasks in particular is thought to mediate the association between adversity exposure and achievement [@taylor_2022].
+At face value, this would suggest interventions targeting inhibition ability (e.g., inhibition training, which might involve removing distractions from the environment) could positively impact achievement outcomes.
+However, if impaired inhibition ability is *not* the root cause of performance differences, but rather general processes are (such as the speed of information processing), the impact of such interventions will likely be limited.
+Therefore, adversity research should move beyond the use of raw performance to better understand the associations between adversity exposures and specific EF abilities.
 
-When applied to empirical performance data, the DDM distinguishes between three parameters mapping onto distinct cognitive processes.
-The *drift rate* reflects the rate at which the evidence accumulation process drifts towards the decision boundary, and thus provides a measure of the speed/efficiency of information processing.
-The *boundary separation* reflects the width between the two decision boundaries, and provides a measure of a person's response caution.
-Finally, the *non-decision time* reflects a combination of preparation time (e.g., stimulus encoding) as well as execution time (e.g., time spent pressing the button).
+# Separating EF ability from other decision-making processes
 
-Researchers have argued that the drift rate on EF tasks is a more valid measure of EF ability, at least to the extent to which the tasks are reliable measures of EF ability [@loffler_2024; @vermeent_2024a; @weigard_2021].
-However, drift rates are still influenced at least partly by general speed of processing.
-Theoretically, this should be the part of the variance that is shared across drift rates of all tasks, and of conditions both including an EF component (e.g., incongruent trials) and excluding an EF component (e.g., congruent trials) [@loffler_2024].
-This part of the variance can be estimated on a latent level using SEM.
-Any remaining task-specific variance then contains EF-specific variance.
+Raw performance confounds multiple stages of processing: from initial preparations (e.g., stimulus encoding), to processing task-relevant information, to deciding on and executing a response [@ratcliff_1978; @forstmann_2016; @wagenmakers_2009].
+Individual differences in response times and error rates also depend on how cautiously people make decisions.
+More cautious people tend to respond more slowly to increase their accuracy, which does not necessarily imply lower ability [@voss_2004].
+Thus, understanding associations between adversity and specific EF abilities requires isolating ability-relevant processing (e.g., inhibition, attention-shifting) from ability-irrelevant processes that contribute to response times.
 
-As mentioned above, a recent investigation did not find significant associations between adversity and task-specific EF variances [@vermeent_2024a].
-This could mean one of two things.
-First, task-specific variance may reliably reflect specific EF abilities, and adversity may not be associated with these abilities.
-Second, task-specific variance may *not* reliably reflect specific EF abilities.
-That is, after accounting for general speed of processing, little more than measurement error may remain.
-Testing these hypotheses requires sampling two or more tasks that are supposed to measure the same ability, and using latent modeling to investigate whether they form a consistent latent factor after accounting for general speed of processing.
+Cognitive modeling can distinguish between abilities and ability-irrelevant processes. 
+Here, we focus on the Drift Diffusion Model [DDM\; @forstmann_2016; @ratcliff_1998; @ratcliff_2008; @wagenmakers_2009], which can be applied to many widely used EF tasks.
+The DDM assumes that people accumulate evidence for one of two responses (e.g., left or right button) until they have acquired sufficient evidence to make a decision (see Figure 1).
+When applied to trial-level response times and accuracy data, the DDM estimates four parameters that represent distinct cognitive processes [@voss_2004].
+The *drift rate* reflects the rate at which people accumulate evidence for the correct response; thus, it measures the efficiency of information processing.
+A higher drift rate leads to faster responses and higher accuracy.
+When applied to EF tasks, the drift rate captures individual differences in specific EF abilities [although drift rate may also capture more general processes, see below\; @loffler_2024; @vermeent_2024a; @weigard_2021].
+The *boundary separation* reflects the width between the two decision boundaries; it measures a person's response caution.
+A higher boundary separation leads to slower responses and higher accuracy.
+The *non-decision time* reflects a combination of preparation time (e.g., stimulus encoding) as well as execution time (e.g., time spent pressing the button).
+A larger non-decision time leads to slower responses without a change in accuracy.
+Finally, the *starting-point* allows for initial bias for one of two responses (e.g., happy or angry faces). 
+In this paper, we will not consider the starting point, because people cannot have biases towards correct or incorrect responses.
+
+```{r}
+#| label: Figure1
+#| fig-width: 6.5
+#| fig-height: 7
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure 1.** Visual representation of the Drift Diffusion Model (DDM).
+The DDM assumes that people accumulate evidence for one of two responses until they have acquired sufficient evidence to make a decision.
+Each squiggly line represents the evidence accumulation process on a single trial.
+The thick, horizontal lines represent the decision boundaries corresponding to the correct response (upper boundary) and the incorrect response (lower boundary).
+Applying the DDM to trial-level performance data (response times and accuracies) yields four parameters representing distinct cognitive processes: (1) **drift rate**: Average rate of evidence accumulation across trials; measuring the efficiency of information processing.
+(2) **boundary separation**: The distance between the two decision boundaries; measuring response caution.
+(3) **Non-decision time**: A combination of the preparation speed (e.g., stimulus encoding) and response execution speed (e.g., time spent pressing the button); measuring speed of non-decision processes.
+**starting point**: The starting point of the evidence accumulation process; measuring response bias (not considered here).
+Figure adopted from Vermeent et al.
+(2024), licensed under CC BY 4.0.
+knitr::include_graphics("figures/fig1.png")
+```
+
+# Separating EF ability from general processing speed
+
+Recent studies show that a single general factor accounts for most of the variance in drift rates across multiple tasks [@lerche_2020; @loffler_2024; @weigard_2021].
+This general factor appears to be largely stable across time [@schubert_2016; @weigard_2021].
+It arises also in analyses of tasks thought to measure different abilities [@schmiedek_2007; @schmitz_2016].
+Even if all tasks are designed to measure the same EF ability [@hedge_2021], variance in the general factor is fully explained by general processing speed instead of any specific EF ability [@loffler_2024].
+These findings suggest that performance on standard EF tasks largely reflects basic speed of processing, rather than EF abilities.
+
+Recent findings suggest that the effects of adversity are more general than specific.
+For instance, youth with lower SES and more exposure to adversity show lower task-general raw performance scores across measures of EF, memory, and intelligence, while task-specific performance (after accounting for general variance) is often practically equivalent to zero [@bignardi_2024; @young_2024].
+In addition, youth with more exposure to household threat showed lower task-general processing speed (based on drift rates), while differences in task-specific drift rates were practically equivalent to zero [@vermeent_2024a].
+These findings suggest that previous research may have overestimated the associations between adversity and specific abilities.
+
+A consequence of both sources of measurement impurity is that estimating specific EF abilities requires latent variable models including two or more measures of the same ability; yet, most studies in adversity research include only one task per EF ability.
+This makes this work less suitable to investigate whether adversity is associated with specific EF abilities after accounting for general processes.
+This is important for both deficit-based and adaptation-based interventions, which aim to remediate or leverage impaired or enhanced abilities, respectively.
+Moreover, while deficit frameworks predict impairments in both specific abilities and general processes [@vermeent_2024a], adaptation frameworks predict that specific types of adversity lead to adaptations in specific EF abilities.
+For example, it has been proposed that unpredictable and dangerous environments may positively affect the ability to rapidly shift attention, promoting the detection of sudden threats and seizing fleeting opportunities [@frankenhuis_2020; @mittal_2015].
+In the current research, therefore, we investigate the associations between adversity and two latent EF abilities—--inhibition and attention shifting—--after accounting for task-general processing speed.
+We will use multiple tasks for each ability to obtain better estimates of specific abilities.
 
 # The current study
+
+Using a combination of DDM and structural equation modeling, we address three central questions.
+First, what is the association of adversity exposure in adulthood with general processing speed that is contributing to performance across all cognitive tasks?
+Second, what is the association of adversity exposure in adulthood with inhibition and attention-shifting abilities after accounting for general processing speed?
+Third, what is the association of adversity exposure in adulthood with general and/or EF-specific response caution?
+We focus on two types of adversity: material deprivation (a lack of access to material resources) and threat (the potential for harm imposed by others).
+In previous research, both types of adversity were associated with performance on tasks used to measure EF [e.g., @fields_2021; @schafer_2022; @sheridan_2022; @vermeent_2024a; @young_2022].
+In addition, both are central to recent dimensional models of adversity [@mclaughlin_2021; @sheridan_2014].
+
+We preregistered predictions of deficit and adaptation frameworks.
+Deficit frameworks predict negative associations between adversity and both the general speed of processing as well as specific abilities.
+Adaptation frameworks predict intact or enhanced attention shifting ability, but not intact or enhanced inhibition ability.
+We only interpret intact attention shifting ability as an adaptation if inhibition ability is also lower.
+Intact performance on both abilities, however, would not provide sufficient evidence of adaptation, as it might instead suggest no association between adversity exposure and these abilities.
+In line with previous work [@vermeent_2024a], we further predict that threat exposure is associated with higher task-general response caution.
 
@@
 
-Participants were 1,000 Dutch people randomly drawn from the Longitudinal Internet studies for the Social Sciences (LISS) panel [@scherpenzeel_2011].
-The LISS panel is an invite-only, representative probability sample of the Dutch population consisting of roughly 7,500 individuals across 5,000 households.
-LISS participants complete a yearly core battery of questionnaires about various key domains, among which their financial situation over the past year.
-In addition, they can participate in further monthly studies covering a wide range of different topics.
-The current study was fielded between May and June 2024.
-Participants were able to complete the cognitive tasks across two or more sessions to increase participation rates.
-People were eligible for participation if they (1) were between 18 and 55 years old; (2) if they had completed at least one wave of a study in the LISS archive assessing exposure to crime (see below); (3) if they had agreed to linking their LISS data to government microdata (not relevant here).
-Participants were excluded if they interacted with other browser tabs mid-trial on any of the cognitive tasks (but not during breaks or in-between tasks). 
-If performance on a specific task was at or below chance level, we excluded the data for that task but retained other task data of that participant.
+We tested a total of `r exclusions$sample$finish` participants from the Dutch Longitudinal Internet Studies for the Social Sciences (LISS) panel [@scherpenzeel_2011].
+The LISS panel is an invitation-only, representative sample of the Dutch population consisting of approximately 7,500 individuals across 5,000 households.
+LISS participants complete a yearly core battery of questionnaires about various domains of life, including one's financial situation over the past year.
+In addition, LISS participants have the option to participate in further monthly studies on different topics.
+We estimated power based on @kretzschmar_2019.
+With $\alpha$ = 0.05 and assuming moderate reliability of our cognitive measures, we would have > 90% power to detect small effect sizes ($\beta$ = 0.1) with a sample size ranging between N = 730 and N = 980.
+Therefore, we aimed for a total sample size of N = 1,000.
+
+The current study took place between May and August 2024.
+Participants were able to complete the cognitive tasks during two or more sessions to increase participation rates.
+People were eligible for the study if they were between 18 and 55 years old, and had agreed to linking their LISS data to government microdata (not relevant here, but for a different study).
+We excluded participants who did not have data on any cognitive task.
+The final sample after exclusions consisted of `r exclusions$sample$final` participants (see Table 1).
+
+<br>
+
+```{r}
+#| tab.id: table1
+#| results: markup
+table1 |> 
+  flextable::autofit() |> 
+  add_header_row(
+    values = " ",
+    colwidths = 2
+  ) |> 
+  flextable::compose(
+    i = 1, j = 1, 
+    as_paragraph(as_b("Table 1. "), "Descriptive statistics for the final sample."),
+    part = "header"
+  ) |> 
+  border_remove() |> 
+  border(i = 1, border.bottom = fp_border_default(), part = "header") |>  
+  border(i = 2, border.top = fp_border_default(), part = "header") |>  
+  border(i = 2, border.bottom = fp_border_default(), part = "header") |> 
+  border(i = 14, border.bottom = fp_border_default(), part = "body") |> 
+  autofit()
+```
+
+<br>
 
@@
 
-We preregistered our approach to compute adversity composites based on observed correlations between measures.
-In short, if all measures of an adversity type correlated .60 or higher (indicating "strong" correlations), we calculated a uniformly weighted average.
-If one or more correlations were lower than .60, we used Principal Component Analysis (PCA) to the separate measures and extracted only the first principal component score.
-Perceived scarcity and income-to-needs were considered as two separate types of material deprivation (and unpredictability), regardless of their correlation.
+We preregistered our approach for computing adversity composite scores based on observed correlations between measures.
+If all measures of an adversity type (i.e., material deprivation and threat) correlated .60 or higher (indicating "strong" correlations, as was the case for material deprivation in adulthood), we calculated a uniformly weighted average.
+If one or more correlations were lower than .60 (as was the case for neighborhood threat in adulthood), we applied Principal Component Analysis to the separate measures and extracted only the first principal component score.
+See section 1 of the supplemental materials for frequency distributions of all adversity measures.
 
-### Neighborhood threat
+### Neighborhood threat in adulthood
 
-*Perceived neighborhood crime.* Participants completed the Neighborhood Violence Scale [NVS\; @frankenhuis_2018; @frankenhuis_2020] in the current study.
-The NVS included seven items about current perceived neighborhood threat (e.g., “Where I live, it is important to be able to defend yourself against physical harm”), on a scale of 1 ("Completely disagree") to 7 ("Completely agree").
-All items of the NVS were averaged together and standardized.
-We also included four items on perceived neighborhood threat from the LISS archive (six waves: <https://doi.org/10.17026/dans-zch-j8xt>), in which participant reported how often it happens that they 1) “avoid certain areas in your place of residence because you perceive them as unsafe”, 2) “do not respond to a call at the door because you feel that it is unsafe”, 3) “leave valuable items at home to avoid theft or robbery in the street?”, 4) “make a detour, by car or on foot, to avoid unsafe areas?” on a scale of 1 (“(Almost) never”), 2 (“Sometimes”), or 3 (“Often”).
-We recoded these items so that "(Almost) never" corresponded to zero.
+Following our preregistration, we measured exposure to neighborhood violence in adulthood using two measures of perceived neighborhood crime, and one measure of crime victimization. 
+Participants completed the Neighborhood Violence Scale [@frankenhuis_2018; @frankenhuis_2020].
+This scale includes seven items about current perceived neighborhood threat (e.g., “Where I live, it is important to be able to defend yourself against physical harm”), on a scale of 1 ("Completely disagree") to 7 ("Completely agree").
+These items were averaged and standardized.
+We also included four items on perceived neighborhood threat from the LISS archive (six waves: <https://doi.org/10.17026/dans-zch-j8xt>).
+Participants reported how frequently they: (1) “avoid certain areas in your place of residence because you perceive them as unsafe”, (2) “do not respond to a call at the door because you feel that it is unsafe”, (3) “leave valuable items at home to avoid theft or robbery in the street?”, (4) “make a detour, by car or on foot, to avoid unsafe areas?”.
+The scale of these four items ranged from 1 (“(Almost) never”), 2 (“Sometimes”), to 3 (“Often”).
+We intended to include these items again in the current study, but due to an oversight this did not happen.
+This meant that we did not have data on these items for participants who never participated in any of the six waves described above (missing N = `r exclusions$sample$missing_crime_vict`). 
+We decided to conduct all analyses involving threat based only on this subset of the data (N = `r exclusions$sample$final - exclusions$sample$missing_crime_vict`), because *a priori* power remained above 80% [@kretzschmar_2019].
 We summed the items within each wave, and then calculated an average across waves for which participants had data.
 
-*Crime victimization.* We used seven items from the LISS archive (six waves: <https://doi.org/10.17026/dans-zch-j8xt>) in which participants reported whether or not they were the victim of eight types of crime in the last two years: (1) burglary or attempted burglary; (2) theft from their car; (3) theft of their wallet or purse, handbag, or other personal possession; (4) wreckage of their car or other private property; (5) intimidation by any other means; (6) maltreatment of such serious nature that it required medical attention; (7) maltreatment that did not require medical attention.
-As the last wave occurred in 2018, we included the same items in the current study asking about people's crime exposure in the last two years.
-We computed the total number of distinct crimes that participants were exposed to at any moment in time (a 'variety score'\; @sweeten_2012).
+We computed crime victimization in adulthood based on seven items from the LISS archive (six waves: <https://doi.org/10.17026/dans-zch-j8xt>) in which participants reported whether or not they had been the victim of seven types of crime in the last two years: (1) burglary or attempted burglary; (2) theft from their car; (3) theft of their wallet or purse, handbag, or other personal possession; (4) wreckage of their car or other private property; (5) intimidation by any other means; (6) maltreatment of such serious nature that it required medical attention; and (7) maltreatment that did not require medical attention.
+We also included these items again in the current study, with people answering them about the last two years.
+We computed the total number of distinct crimes that participants were exposed to at any moment in time [a 'variety score'\; @sweeten_2012].
 
-*Neighborhood threat composite.* Correlations among the two measures of perceived neighborhood crime and crime victimization were low.
-Therefore, we used PCA to extract the first principal component score.
-This score accounted for XXX% of the variance in the three measures.
+The correlations between the two measures of perceived neighborhood crime and crime victimization were low (see Table 2).
+Therefore, following our preregistration, we used Principal Component Analysis to extract the first principal component score.
+This score accounted for 23 % of the variance in the three measures.
 
-### Material deprivation
+### Material deprivation in adulthood
 
-*Perceived scarcity (mean).* Measures of perceived scarcity were derived from the LISS archive, using the yearly recurring core study on household and personal income (16 waves: <https://doi.org/10.57990/1gr4-bf42>).
+We derived measures of material deprivation in adulthood from the LISS archive, using the yearly recurring core study on household and personal income (16 waves: <https://doi.org/10.57990/1gr4-bf42>).
 First, participants reported how difficult it currently is to live off the income of their household, on a scale from 0 (very hard) to 10 (very easy). 
-Second, participants reported which of the following best applied to their current financial situation: (1) “we are accumulating debt”; (2) “we are somewhat eating into savings”; (3) “we are just managing to make ends meet”; (4) “we have a little bit of money to spare”; (5) “we have a lot of money to spare”.
-Third, participants reported which of the following applied to their current financial situation (0 = no, 1 = yes): (1) “having trouble making ends meet”; (2) unable to quickly replace things that break”; (3) “having to lend money for necessary expenditures”; (4) “running behind in paying rent/mortgage or general utilities”; (5) “debt collector/bailiff at the door in the last month”; (6) “received financial support from family or friends in the last month”.
-Responses were coded so that higher scores indicated more perceived scarcity.
+Second, participants reported which of the following statements best described their current financial situation: (1) “we are accumulating debt”; (2) “we are somewhat eating into savings”; (3) “we are just managing to make ends meet”; (4) “we have a little bit of money to spare”; (5) “we have a lot of money to spare”.
+Third, participants reported which of the following applied to their current financial situation (0 = no, 1 = yes): (1) “having trouble making ends meet”; (2) "unable to quickly replace things that break”; (3) “having to lend money for necessary expenditures”; (4) “running behind in paying rent/mortgage or general utilities”; (5) “debt collector/bailiff at the door in the last month”; and (6) “received financial support from family or friends in the last month”.
+We recoded responses so that higher scores indicated more perceived scarcity.
 
-We first averaged each measure separately across waves, and then scaled them.
-As all item correlations were > .60, we computed a uniformly weighted average.
+We first reverse-coded and averaged each measure separately across waves, and then scaled them.
+As all item correlations were > .60 (see Table 2), we computed a uniformly weighted average.
 
-*Income-to-needs (mean).* The income-to-needs ratio was derived from monthly self-reported net household income in the LISS archive (<https://doi.org/10.57990/qn3k-as78>).
-Reported income of zero was set to missing, as this could indicate an unwillingness to report on income rather than a true income of zero.
-If net household income data was missing (or zero) for an entire year, we imputed, if possible, the reported yearly net household income from the core study on income (<https://doi.org/10.57990/1gr4-bf42>), dividing by 12 to obtain monthly estimates.
-We then divided the income by the poverty threshold set by Statistics Netherlands (CBS) [@brakel_2023; CBS, personal communication, December 15, 2023].
-For households with more than three children, we used the threshold for a family with three children, as CBS does not calculate thresholds for families larger than three children.
-For the same reason, income of households with more than two adults was compared to the threshold for households with two adults.
-We computed the average income-to-needs ratio across years, reverse coding it so that higher scores indicated households further below the threshold.
+::: {.landscape}
 
-### Unpredictability
+```{r}
+#| tab.id: table2
+#| results: markup
+table2 |> 
+  flextable() |> 
+  flextable::set_header_labels(Variable = "") |> 
+  flextable::autofit() |> 
+  border_remove() %>% 
+ # border(i = 1, border.top = fp_border_default(), part = "header") %>% 
+  border(i = 1, border.bottom = fp_border_default(), part = "header") %>% 
+  border(i = 18, border.bottom = fp_border_default(), part = "body") %>% 
+  add_header_row(
+    values = " ",
+    colwidths = 13
+  ) |> 
+  flextable::compose(
+    i = 1, j = 1, 
+    as_paragraph(as_b("Table 2. "), "Spearman correlations between the main independent variables and covariates."),
+    part = "header"
+  ) |> 
+  add_footer_row(
+    values = " ",
+    colwidths = 13
+  ) |> 
+  autofit()
+```
 
-*Perceived scarcity (SD/mean).* Unpredictability over time was computed using the coefficient of variation, which is the within-person standard deviation across years divided by the mean across years (@key_2017; @liu_2022; @ugarte_2023; @walasek_2024; @young_2020).
-This measure accounts for the fact that the mean and standard deviation tend to be highly positively correlated.
-Unpredictability in perceived scarcity was derived from the same measures as the mean level of perceived scarcity (see above).
-We first computed the standard deviation for each measure separately across years, and then scaled them.
-We then computed a uniformly weighted average, which was divided by the mean level of perceived scarcity across years.
-
-*Income-to-needs (SD/mean).* Unpredictability over time in the income-to-needs ratio was also computed using the coefficient of variation.
-We first computed the within-person standard deviation of the income-to-needs ratio across years, and then divided it by the mean income-to-needs ratio across years.
+:::
 
@@
 
-All cognitive tasks were programmed in jsPsych 7.3 [@deleeuw_2015].
+We programmed six cognitive tasks in jsPsych 7.3 [@deleeuw_2015]: two inhibition tasks, three attention shifting tasks, and one basic processing speed task.
 At the start of the session, participants entered fullscreen mode to avoid distractions from other browser tabs.
 The tasks were presented against a light-gray background.
+See Section 2 of the supplemental materials for information on condition manipulation checks, split-half reliability estimates, and bivariate correlations between tasks.
 
-*Flanker Task.* This task measures inhibition of distractor interference [...]
+*Flanker Task.* This task measures inhibition of distractor interference [@eriksen_1974].
 On each trial, participants saw five arrows side-by-side horizontally, pointing either left or right.
@@
 
-*Simon Task.* This task measures inhibition of prepotent responses [...].
+*Simon Task.* This task measures inhibition of prepotent responses [@simon_1963].
 On each trial, participants saw the word "LEFT" or "RIGHT" (printed in Dutch), presented either on the left or right side of the screen.
 Their task was to press the 'A' key if the word was "LEFT" and the 'L' key if the word was "RIGHT", regardless of the location on the screen.
-on 50% of the trials, the word matched the location (e.g., the word "LEFT" presented on the left side; congruent trials).
+On 50% of the trials, the word matched the location (e.g., the word "LEFT" presented on the left side; congruent trials).
 On the other half, the word did not match the location (e.g., the word "LEFT" presented on the right side; incongruent trials).
@@
 
-*Color-Shape Task.* This task measures the ability to shift attention between different task goals [XXX].
-On each trial, participants saw a square or a circle in the center of the screen that was either blue or yellow.
-Depending on the task rule printed above the stimulus, their task was to classify the stimulus based on it's shape or color.
+*Color-shape Task.* This task measures the ability to shift attention between different tasks [@miyake_2000].
+On each trial, participants saw a square or a circle in the center of the screen.
+This square or circle was either blue or yellow.
+Depending on the task rule printed above the stimulus, their task was to classify the stimulus based on its shape or color.
 On 50% of the trials, the rule was the same as on the preceding trial (repeat trials).
@@
 
-*Animacy-Size Task.* This task measures the ability to shift attention between different task goals [XXX].
-On each trial, participants saw a single noun (in Dutch) referring to an animal or object [adopted from XXX].
-Depending on the task rule printed above the noun, their task was to classify it based on whether it referred to a living or non-living thing (e.g., wasp vs.
+*Animacy-size Task.* This task measures the ability to shift attention between tasks [@arrington_2004].
+On each trial, participants saw a single noun (in Dutch) referring to an animal or object [adopted from @braem_2017].
+Depending on the task rule presented on the screen, their task was to classify the noun based on whether it referred to a living or non-living thing (e.g., wasp vs.
 piano), or whether it referred to something that was smaller or larger than a soccer ball (e.g., ring vs.
@@
 
-*Global-local Task.* This task measures the ability to shift attention between task goals [XXX]
-Stimuli were adapted from [...].
-On each trial, participants saw a large square or rectangle which was composed of 16 small squares or rectangles.
-The stimulus where flanked on both side by a drawing of an elephant or mouse, which was presented 1,000 ms prior to the appearance of the stimulus.
+*Global-local Task.* This task measures the ability to shift attention between tasks.
+We adapted the stimuli from @huizinga_2010.
+On each trial, participants saw a large square or rectangle composed of 16 small squares or rectangles.
+The stimulus was flanked on both side by a drawing of an elephant or mouse, which was presented 1,000 ms prior to the appearance of the stimulus.
 If the stimulus was flanked by the image of an elephant (50% of trials), participants had to indicate whether the global image was a square or rectangle.
@@
 Finally, the stimuli were congruent on 50% of the trials (e.g., large square consisting of small squares) and incongruent on the other half (e.g., large square consisting of small rectangles).
-Congruency, task rule (switch vs. repeat) or focus (global vs. local) where never repeated more than three times in a row.
+Congruency, task rule (switch vs. repeat), or focus (global vs. local) where never repeated more than three times in a row.
 Participants first completed eight practice trials, followed by two test blocks of 32 trials each, for a total of 64 trials.
 
-*Posner Task.* This task measures basic speed of processing.
-On each trial, participants saw two letters in the center of the screen, drawn from the set A, B, F, H, Q, a, b, f, h , q.
+*Posner Task.* This task measures basic speed of processing [@posner_1967].
+On each trial, participants saw two letters in the center of the screen, drawn from the set A, B, F, H, Q, a, b, f, h, and q.
 Their task was to indicate whether the letters were the same (e.g., "AA", "bB") or different (e.g., "AQ", "Fh").
 On 50% of the trials, the letters were the same, and on the other half they were different.
-Participants first completed 8 practice trials, followed by two test blocks of 32 trials each, for a total of 64 trials.
+Participants first completed eight practice trials, followed by two test blocks of 40 trials each, for a total of 80 trials.
 
-### Covariates
+## State anxiety
 
-We preregistered a Directed Acyclic Graph (DAG) to inform the inclusion of covariates to the model.
-Based on this DAG, we decided to control for age (in years), the quadratic effect of age, education level, and childhood adversity.
-Childhood adversity was the unweighted average across two scales measuring childhood threat and deprivation: (1) the Neighborhood Violence Scale [XXX], which consists of seven items (e.g., "Physical fights were common in the neighborhood where I grew up"), and (2) four items measuring access to housing, food, clothing, and medicine.
+Participants reported their state anxiety after each cognitive task; thus, six times in total.
+We measured state anxiety using the shortened version of the State-Trait Anxiety Inventory [@bij_2003; @marteau_1992], which asks participants how calm, tense, upset, relaxed, content, and worried they currently feel, on a scale of 1 ("not al all") to 4 ("very much").
+We recoded (if necessary) and then summed the answers with higher scores reflecting more state anxiety.
+
+## Environmental noise
+
+Participants reported the level of environmental noise after each cognitive task; thus, six times in total.
+We measured environmental noise using a single item, rated on a scale of 1 to 5: "How much noise was there in your environment during the game?"
+
+### Confounds
+
+We used Directed Acyclic Graphs to identify potential confounds of the key estimands, which were the associations between self-reported threat and material deprivation in adulthood with cognitive outcomes.
+A Directed Acyclic Graph is a visual overview of assumptions about how variables are causally related.
+They are graphs consisting of nodes (variables) and directed arrows (causal pathways).
+An arrow between two variables represents the assumption that experimentally manipulating the variable at the origin of the arrow will change the variable at the end of the arrow (but not the other way around).
+Directed Acyclic Graphs aid in the identification of variables that need to be adjusted for in the statistical models (i.e., confounders with arrows to both the main predictor and the outcome), and, equally importantly, which variables should not be adjusted for (i.e., colliders and mediators).
+For a detailed explanation, see @rohrer_2018.
+
+The set of potential confounders consisted of (1) age [@salthouse_2016; @salthouse_2019; @starns_2010]; (2) education [@hofmarcher_2021]; (3) sex [@ning_2023]; (4) childhood adversity exposure (material deprivation and threat combined) [@bos_2009; @goodman_2019]; and (5) potential causal relations between recent material deprivation and recent threat [@bywaters_2016; @lacey_2022; @ning_2023].
+We made these decisions based on previous literature, or, in cases of doubt, by statistically testing support for specific relations using data from a previous LISS study [@vermeent_2024c].
+Specifically, following our preregistration, we used the *dagitty* R package [@ankan_2021; @textor_2016] to test whether independencies between specific variables implied by the Directed Acyclic Graph (conditional on the other causal pathways) were supported by these data (see the preregistration for more details).
+
+Figure 1 depicts the final preregistered Directed Acyclic Graph for material deprivation and threat exposure in adulthood.
+For threat exposure in adulthood, our statistical model controlled for age, sex, childhood adversity, and recent material deprivation.
+We included material deprivation in adulthood based on studies showing that material deprivation tends to precede exposure to adversities such as threat [@bywaters_2016; @lacey_2022; @ning_2023].
+We did not control for education given a lack of theoretical and statistical support for this effect (see preregistration).
+To explore the effect of our assumption that material deprivation tends to precede threat, we also present a secondary model excluding material deprivation in adulthood as a confound.
+As preregistered, we will not base our main conclusions on this secondary model.
+
+For material deprivation in adulthood, our primary statistical model controlled for age, education, sex, and childhood adversity exposure based on previous studies (see above).
+We did not control for threat exposure in adulthood, as, following the Directed Acyclic Graph, we assume that threat exposure mediates the effect of recent material deprivation on cognitive processes.
+If true, including threat exposure in the model would make the association between material deprivation and cognitive processes an indirect effect, rather than a total effect (which was our estimand).
+To explore the effect of our assumption that threat mediates the effect of material deprivation on cognitive processes, we also present a secondary model that includes threat in adulthood as a confound.
+As preregistered, we will not base our main conclusions on this secondary model.
+
+```{r}
+#| label: Figure2
+#| fig-width: 6.5
+#| fig-height: 7
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure 2.** Direct Acyclic Graphs (DAG) depicting our causal assumptions about the main estimands.
+the rectangles (nodes) depict variables, and directed arrows depict assumed causal pathways.
+An arrow between two variables represents the assumption that experimentally manipulating the variable at the origin of the arrow will change the variable at the end of the arrow (but not the other way around).
+The statistical models control for confounding variables (depicted as dashed paths), which have an arrow both to the independent variable and the dependent variable.
+knitr::include_graphics("figures/fig2.png")
+```
+<br>
 
@@
 
-We obtained ethical approval from the first author's institutional ethical board. 
-The study was fielded on the LISS platform, and participants could only complete the study on a laptop or desktop PC.
+We obtained ethical approval from the Ethics Review Board of the Faculty of Social & Behavioral Sciences of Utrecht University (FETC20-490) and the Ethics committee for research in the Sciences and Life Sciences of the University of Amsterdam (FNWI-41_2023).
+The study was implemented on the LISS platform, and participants could only complete the study on a laptop or desktop PC.
 Participants started with the six cognitive tasks, in randomized order.
-After each task, they were asked if they wanted to continue to the next task or conclude the session. 
-If they choose for the latter, they were able to continue with the tasks at another time.
-The cognitive tasks took around 25 minutes to complete.
-After completing all cognitive tasks, participants completed questionnaires about past and present neighborhood threat and material deprivation, exposure to crime victimization in the past two years, trait impulsivity (not considered here) and trait future orientation (not considered here).
-Finally, they answered a few standard LISS evaluation questions about their experiences with the study, with the opportunity to provide written feedback.
+After each task, they indicated their state anxiety and the level of environmental noise.
+Then, participants had the option to either continue to the next task or complete the other tasks at a later point in time. 
+`r round(n_sessions)` % of participants completed all tasks in a single session, and `r 100-round(n_sessions)` % of participants took two or more sessions.
+Jointly, the cognitive tasks took around 25 minutes to complete.
+After finishing all cognitive tasks, participants completed questionnaires about neighborhood threat and material deprivation during childhood and, adulthood exposure to crime victimization in the past two years.
+They also completed questionnaires on impulsivity, self-control, and future orientation, which are not considered here because they fall outside of the scope of the present study.
+As some participants had already completed all of these questionnaires in a different data collection (about six months prior to the current study), we only presented the questionnaires to new participants.
+Finally, all participants answered a few standard LISS evaluation questions about the study, giving them an opportunity to provide written feedback.
 
@@
 
-For all tasks, we removed trials with RTs < 250ms and trials with RTs more than 3.2 SD above the intra-individual log-transformed mean RT.
+For all tasks, we first removed any response times > 10 seconds.
+This step was not preregistered but was necessary given that we did not specify a response time-out.
+Consequently, a small portion of response times lasted up to several minutes, likely reflecting breaks or interruptions (between `r exclusions$posner$very_slow` and `r exclusions$animacysize$very_slow` for all tasks).
+We removed these first to prevent them from biasing our preregistered exclusion of outliers.
+Next, we applied two preregistered exclusion criteria.
+First, we removed trials with response times < 250 ms and trials with response times more than 3.2 SD above the intra-individual log-transformed mean response time.
+Second, if participants performed at chance level on a particular task, we excluded the data for that task only.
+We set the cut-off for chance performance based on accuracy at the 97.5 % tail of the binomial distribution one would obtain if guessing.
 
@@
 
-We used an Hierarchical Bayesian implementation of the DDM (HDDM) [@vandekerckhove_2011; @wiecki_2013].
-The benefit of HDDM over traditional approaches is that the model uses information of the whole sample to inform individual parameter estimation.
-The model will be applied to each task separately.
-Across all tasks, the bias parameter was held constant, as there is no reason for participants to favor one response over the other across trials.
+We used a Hierarchical Bayesian implementation of the DDM [@vandekerckhove_2011; @wiecki_2013] to leverage group-level information for individual parameter estimation.
+We applied this model to each task separately.
 For the Posner Task, we estimated a single drift rate, boundary separation, and non-decision time for each participant.
-For all other tasks, drift rates and non-decision times were estimated separately for congruent (repeat) and incongruent (switch) trials.
-Boundary separation was always fixed across conditions, following the assumption that people cannot anticipate the condition of the next trial, and hence cannot adjust their strategy on a trial-by-trial basis.
+For the other five tasks, we estimated drift rates, boundary separation, and non-decision times separately for congruent (repeat) and incongruent (switch) trials.
 
 The DDM models were fit using the *runjags* package [@denwood_2016].
-Each model was fit with three Markov Chain Monte Carlo (MCMC) chains.
-We used 2,000 burn-in samples and 10,000 additional samples, of which every 10th sample was retained, resulting in a total of 3,000 posterior samples.
-See the supplemental materials for more information about model convergence and fit assessment.
+We fit each model with three Markov Chain Monte Carlo chains.
+We used 2,000 burn-in samples and 10,000 additional samples, retaining every 10th sample, resulting in a total of 3,000 posterior samples.
+Model convergence and fit was good across all DDM models (see section 3 of the supplemental materials).
 
-### SEM
+We accounted for two potential sources of measurement error in the DDM estimates: (1) the level of environmental noise and (2) within-participant differences from mean state anxiety between the tasks (see above).
+We used linear regression to residualize the variance of noise and anxiety out of all the drift diffusion estimates.
+See Section 4 of the supplemental materials for more information.
+
+### Structural equation modeling
+
+We constructed the full structural equation model sequentially.
+First, we optimized the fit of the drift rate, boundary separation, and non-decision time sub-models.
+Second, we combined these three models into a single measurement model.
+Third, we added the regression paths between measures of adversity in adulthood and the latent factors.
+To assess goodness-of-fit, we used the root mean square error of approximation (RMSEA) and the comparative fit index (CFI). 
+CFI values > 0.90 (> 0.95) and RMSEA values < 0.08 (≤ 0.06) were interpreted as acceptable (good) fit.
+
+Each DDM parameter sub-model was a bi-factor model including the parameter estimates of all tasks as manifest variables.
+We estimated a general factor accounting for variance in all manifest variables.
+A second ability-specific inhibition factor loaded on the parameter estimates of incongruent trials of the Flanker and Simon tasks.
+A third specific attention shifting factor loaded on the parameter estimates of switch trials of the Color-shape, Global-local, and Animacy-size tasks.
+These two factors were allowed to covary.
+We also estimated covariances between the conditions of each task.
+For each sub-model, we compared the fit of this initial model to a version with a common EF factor loading on incongruent/switch trials of all tasks.
+We deemed the second model a better fit when we observed a significant chi squared change test and an AIC value difference \> 10.
+
+For the drift rate model, we interpreted the general factor as basic speed of processing, and the ability-specific factors as reflecting  inhibition and attention-shifting ability (or common EF in the case of the second model).
+For the boundary separation and non-decision time models, we interpreted the general factor as general response caution/speed of non-decision processes, and the ability-specific factors as reflecting response caution/speed of non-decision processes specific to conflict trials.
+
+After optimizing each sub-model, we combined them into a joint measurement model.
+We allowed the general latent factors (processing speed, general caution, and general non-decision time) to covary, as well as the ability-specific latent factors (separately for inhibition and attention shifting, unless a common factor was favored).
+
+Finally, we constructed two versions of the final model, one to estimate the association between material deprivation in adulthood and the outcome measures, and one to estimate the association between threat exposure in adulthood and the outcome measures.
+In both models, we regressed the adversity measure on each latent factor, together with the control variables (see the section on confounds for more details).
+We report indirect effects of control variables in the supplemental materials (section 5).
+
+### Inferential criteria
+
+We tested for three types of associations between different types of adversity with EF ability and processing speed (as indexed by
+drift rates): enhancements, impairments, or practical equivalence.
+We defined enhancements as a positive association between adversity and drift rates. 
+We defined impairments as a negative association
+between adversity and drift rate. 
+We defined practical equivalence as standardized regression coefficients that significantly fell between -0.10 and 0.10.
+We tested this using the Two One-Sided T-tests procedure [@lakens_2018].
+This test evaluates whether the obtained effect is significantly larger than the lower bound *and* significantly smaller than the upper bound.
+Thus, it affords conclusions about practical equivalence based on significant *p*-values, rather than (invalidly) inferring the absence of an effect on the basis of non-significant results.
+
+## Transparency and openness
+
+We have reported how we determined our sample size, all data exclusions, and all measures in the study.
+The preregistration, analysis code, and study materials, can be found on the article's GitHub repository (<https://github.com/stefanvermeent/liss_ef_2024>).
+The newly collected data will be available for other researchers in the LISS data archive, after signing a data use agreement (https://lissdata.nl).
+For more information on the LISS variables used, see the article's GitHub repository (<https://github.com/stefanvermeent/liss_ef_2024>).
+Our study complies with Level 2 of the Transparency and Openness Promotion (TOP) guidelines.
+
+# Results
+
+### Structural equation model fit
+
+As preregistered, we started with optimizing fit in structural equation models for each DDM parameter separately before combining them all into a single model.
+For drift rates, we selected the model containing a general processing speed factor (loading on all drift rates) and a common EF factor (loading only on drift rates of incongruent and switch conditions).
+An alternative model with a separate inhibition and attention shifting factor did not converge.
+The model with a common EF factor provided a good fit to the data (CFI = `r fit_v2_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_v2_fitstats[['rmsea.robust']] |> round(2)` [`r fit_v2_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_v2_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]).
+However, the loadings of the common EF factor were small and in opposing directions, and the latent factor's residual variance was non-significant.
+Therefore, we did not estimate associations between adversity and common EF.
+
+For both boundary separation and non-decision time, we selected a model containing only a general factor (loading on all boundary separations/non-decision times), which provided a good fit to the data (boundary separation: CFI = `r fit_a3_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_a3_fitstats[['rmsea.robust']] |> round(2)` [`r fit_a3_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_a3_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]; non-decision time: CFI = `r fit_t3_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_t3_fitstats[['rmsea.robust']] |> round(2)` [`r fit_t3_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_t3_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]).
+The two preregistered models specifying additional latent factors loading only on switching/incongruent conditions did not converge.
+
+The final model combining all three DDM parameters provided a good fit, although it required the addition of covariances between residual variances of manifest DDM parameters of the same task (reflecting shared method variance), CFI = `r fit_meas_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_meas_fitstats[['rmsea.robust']] |> round(2)` [`r fit_meas_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_meas_fitstats[['rmsea.ci.upper.robust']] |> round(2)`].
+See Figure 3 for a visualization and Table 3 for fit statistics of all models.
+
+::: {.landscape}
+
+```{r}
+#| label: Figure3
+#| dpi: 600
+#| out-width: 12in
+#| fig-cap: | 
+#|   **Figure 3.** Final measurement model of the Drift Diffusion parameters.
+Rectangles depict manifest parameter estimates, elipses depict latent factors, unidirectional arrows depict factor loadings, and bidirectional arrows depict correlations.
+Latent factors with non-significant residual variances are greyed out.
+*a* = manifest boundary separation, AS = Attention-shifting task, CS = Color-shape task, GL = Global-local task, Fl = Flanker task, Sm = Simon task, *t* = manifest non-decision time, *v* = manifest drift rate.
+knitr::include_graphics("figures/fig3.png")
+```
+
+:::
+
+```{r}
+#| tab.id: table3
+#| results: markup
+
+table3 |> 
+  add_header_row(
+    values = " ",
+    colwidths = 4
+  ) |> # Add a new header row on top. We can use this new row to add the title
+  flextable::compose(
+    i = 1, j = 1,
+    as_paragraph(as_b("Table 3. "), "Fit statistics of all preregistered structural equation models."),
+    part = "header"
+  ) |>
+  border_remove() |> 
+  border(i = 1, border.bottom = fp_border_default(), part = "header") |>  
+  border(i = 2, border.top = fp_border_default(), part = "header") |> 
+  border(i = 2, border.bottom = fp_border_default(), part = "header") |> 
+  border(i = 15, border.bottom = fp_border_default(), part = "body") |> 
+  add_footer_row(
+    values = " ",
+    colwidths = 4
+  ) |> 
+  add_footer_row(
+    values = " ",
+    colwidths = 4
+  ) |>  
+  flextable::compose(
+    i = 1, j = 1, 
+    as_paragraph(as_i("Note: "), "for each DDM parameter, measurement model 1 contained a task-general latent factor as well as separate latent factors for inhibition and shifting.
+Measurement model 2 contained a task-general latent factor as well as a common EF factor.
+Measurement model 3 contained only a task-general latent factor."),
+    part = "footer"
+  ) |> 
+  autofit()
+
+```
+
+<br>
+
+## Preregistered analyses
+
+Our primary analyses examined the association of exposure to material deprivation and threat in adulthood with task-general processing speed, task-general response caution, and task-general speed of stimulus encoding and/or response execution.
+As the structural equation model analyses reported above showed that EF-specific factors had non-significant variances, we retained them in the model but did not include them as dependent variables in subsequent analyses.
+
+### Material deprivation in adulthood
+
+Exposure to material deprivation in adulthood was not associated with either general processing speed ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_v', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_v', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`), task-general response caution ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_a', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_a', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`), or task-general non-decision time ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_t', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_t', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`) (see Figure 4).
+In addition, none of the effects fell within the region of practical equivalence (all *p*s ≥ `r fit_dep1_eqtests |> pull(eq_pvalue) |> min() |> format_p(pvalue = _, operator = FALSE)`).
+The effects in the secondary model (including expossure to threat in adulthood as a confounder) were comparable.
+
+<br>
+
+```{r}
+#| label: Figure4
+#| fig-width: 6.5
+#| fig-height: 5.5
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure 4.** Standardized (preregistered) associations between exposure to adversity in adulthood with task-general processing speed, response caution, and non-decision time.
+The left panel depicts effects of material deprivation and the right panel depicts effects of threat.
+The top row depicts effects of the primary statistical model (following from our main Directed Acyclic Graph), and the bottom row depicts effects of the secondary statistical model (see Figure 1).
+For material deprivation in adulthood, the primary statistical model included age, education, and childhood adversity as confounders, whereas the secondary model additionally included threat in adulthood as a confounder.
+For threat in adulthood, the primary statistial model included age, childhood adversity, and material deprivation in adulthood, whereas the secondary statistical model did not include material deprivation in adulthood as a confounder.
+The gray area reflects the preregistered area of practical equivalence.
+Effects depicted with a solid point are practically equivalent, and effects depicted with a hollow point are not practically equivalent.
+Standard errors represent 95% confidence intervals.
+Statistical significance (tested against zero) is indicated with asterisks; * *p* < .05, ** *p* < .01, *** *p* < .001.
+fig4
+```
+
+<br>
+
+### Threat in adulthood
+
+Exposure to threat in adulthood was negatively associated with task-general processing speed ($\beta$ = `r fit_thr1_reg_coef |> filter(lhs == 'gen_v', rhs == 'threat_comp') |> pull(est.std) |> formatC(digits = 2, width = 2, format = 'f')`, *p* `r fit_thr1_reg_coef |> filter(lhs == 'gen_v', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`) (see Figure 4).
+People who reported more exposure to threat in adulthood processed information more slowly across tasks.
+Exposure to threat in adulthood was not associated with task-general response caution ($\beta$ = `r fit_thr1_reg_coef |> filter(lhs == 'gen_a', rhs == 'threat_comp') |> pull(est.std) |> round(2)`, *p* `r fit_thr1_reg_coef |> filter(lhs == 'gen_a', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`), or task-general non-decision time ($\beta$ = `r fit_thr1_reg_coef |> filter(lhs == 'gen_t', rhs == 'threat_comp') |> pull(est.std) |> round(2)`, *p* `r fit_thr1_reg_coef |> filter(lhs == 'gen_t', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`).
+The effects in the secondary model (excluding material deprivation in adulthood as a confounder) were comparable, although the association with task-general non-decision time was practically equivalent in the secondary model.
+
+## Non-preregistered analyses
+
+### Associations with childhood adversity
+
+In the first set of non-preregistered analyses, we analyzed how childhood exposure to threat and material deprivation are associated with task-general and ability-specific cognitive processes.
+The rational for these analyses is that we observed significant indirect associations between childhood adversity and general speed of processing in our preregistered analyses.
+Here, we analyzed *direct* effects of childhood threat and material deprivation (i.e., without including exposure to adversity in adulthood).
+We controlled for sex (but not age and education) and in the case of threat we also controlled for material deprivation, but not the other way around (similarly to the preregistered analyses).
+
+Childhood exposure to material deprivation was negatively associated with general processing speed, $\beta$ = `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(est.std) |> round(2)`, SE = `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(se) |> round(2)`, 95% CI = [`r expl_ch_dep_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(ci.lower) |> round(2)`, `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(ci.upper) |> round(2)`], *p* `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(pvalue_adj) |> format_p(pvalue = _)`.
+People who reported higher levels of childhood exposure to material deprivation processed information more slowly across tasks.
+In addition, childhood exposure to material deprivation was positively associated with non-decision time, $\beta$ = `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(est.std) |> round(2)`, SE = `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(se) |> round(2)`, 95% CI = [`r expl_ch_dep_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(ci.lower) |> round(2)`, `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(ci.upper) |> round(2)`], *p* `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(pvalue_adj) |> format_p(pvalue = _)`.
+People who reported higher levels of childhood exposure to material deprivation took more time to encode stimulus information and/or execute responses across tasks.
+Childhood exposure to material deprivation was not associated with general response caution.
+Equivalence tests revealed that none of these associations fell within the region of practical equivalence.
+
+For childhood exposure to threat, we found a significant negative association with general processing speed, $\beta$ = `r expl_ch_thr_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(est.std) |> round(2)`, SE = `r expl_ch_thr_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(se) |> round(2)`, 95% CI = [`r expl_ch_thr_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(ci.lower) |> round(2)`, `r expl_ch_thr_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(ci.upper) |> round(2)`], *p* `r expl_ch_thr_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(pvalue_adj) |> format_p(pvalue = _)`.
+People who reported higher levels of childhood exposure to threat processed information more slowly across tasks.
+We did not find significant associations between childhood exposure to threat and general response caution or general non-decision time.
+Of these associations, equivalence tests revealed that the association with general response caution fell within the region of practical equivalence (*p* `r expl_ch_thr_eqtests |> filter(lhs == "gen_a", rhs == "child_thr") |> pull(eq_pvalue) |> format_p(pvalue = _)`).
+
+```{r}
+#| label: Figure5
+#| fig-width: 6.5
+#| fig-height: 3
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure 5.** Standardized (non-preregistered) associations between childhood exposure to adversity with task-general processing speed, response caution, and non-decision time.
+The left panel depicts effects of material deprivation and the right panel depicts effects of threat.
+The gray area reflects the preregistered area of practical equivalence.
+Effects depicted with a solid point are practically equivalent, and effects depicted with a hollow point are not practically equivalent.
+Standard errors represent 95% confidence intervals.
+Statistical significance (tested against zero) is indicated with asterisks; * *p* < .05, ** *p* < .01, *** *p* < .001.
+fig5
+```
+
+<br>
+
+### Associations with task-specific drift rates
+
+In the second set of non-preregistered analyses, we analyzed how adversity exposure during childhood and adulthood were associated with task-specific residual variances.
+The rationale for these analyses was that while we did not find coherent latent EF factors, all tasks showed significant residual variance after accounting for general speed of processing.
+This implies that, after accounting for general speed of processing, performance is further determined by unique features of individual tasks (i.e., method variance), which may be associated with adversity exposure.
+For this analysis, we were mostly interested in investigating the extent to which task-specific effects reflect EF abilities or task-specific processing, and therefore only focused on drift rates.
+
+We first attempted to fit a model in which all task-specific drift rates were allowed to covary (i.e., across inhibition and attention-shifting tasks).
+However, this model did not converge.
+A model that only estimated covariances between task-specific drift rates of inhibition tasks and attention shifting tasks separately provided a good fit to the data (CFI = `r fit_expl_ts_mod2_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_expl_ts_mod2_fitstats[['rmsea.robust']] |> round(2)` [`r fit_expl_ts_mod2_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_expl_ts_mod2_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]).
+This suggests that task-specific drift rates correlated among inhibition tasks and among attention-shifting tasks, but not between inhibition and attention-shifting tasks.
+
+Figure 6 shows latent correlations between task-specific drift rates after accounting for task-general processing speed.
+The drift rates of repeat and switch conditions within each attention shifting task correlated moderately to strongly, while all correlations between tasks were low.
+This suggests that task-specific drift rates of attention shifting tasks captured features of the tasks that were not shared between tasks.
+In contrast, all correlations between task-specific drift rates of inhibition tasks were low, even among drift rates of the same task.
+Thus, task-specific drift rate did provide consistent measures of specific EF abilities that are shared across tasks.
+
+<br>
+
+```{r}
+#| label: Figure6
+#| fig-width: 6.5
+#| fig-height: 5
+#| dpi: 600
+#| out-width: 7in
+#| fig-cap: | 
+#|   **Figure 6.** Correlations between task-specific drift rates after accounting for task-general processing speed.
+fig6
+```
+<br>
+
+Next, we explored the associations of task-specific drift rates with exposure to material deprivation and threat in both childhood and adulthood, after accounting for task-general processing speed (Figure 7).
+Exposure to material deprivation in adulthood was not associated with task-specific drift rates of any of the tasks.
+Exposure to threat in adulthood was negatively associated all task-specific drift rates in both the repeat and switch condition of the Color-shape task (repeat: $\beta$ = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(est.std) |> round(2)`, SE = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(se) |> round(2)`, 95% CI = [`r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(ci.lower) |> round(2)`, `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(ci.upper) |> round(2)`], *p* `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`; switch: $\beta$ = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_sw_v', rhs == 'threat_comp') |> pull(est.std) |> round(2)`, SE = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_sw_v', rhs == 'threat_comp') |> pull(se) |> round(2)`, 95% CI = [`r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_sw_v', rhs == 'threat_comp') |> pull(ci.lower) |> round(2)`, `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_sw_v', rhs == 'threat_comp') |> pull(ci.upper) |> round(2)`], *p* `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_sw_v', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`).
+None of the other associations were significant.
+
+Childhood exposure to material deprivation was negatively associated with all task-specific drift rates, with effect sizes ranging from $\beta$ = `r expl_ts_ch_dep_reg_coef |> filter(rhs == 'child_dep') |> pull(est.std) |> round(2) |> max()` for the congruent condition of the Simon Task to $\beta$ = `r expl_ts_ch_dep_reg_coef |> filter(rhs == 'child_dep') |> pull(est.std) |> round(2) |> min()` for the incongruent condition of the Global-local Task.
+Childhood exposure to threat was negatively associated with all but the task-specific drift rates of the congruent condition of the Flanker and the Simon task.
+The effect sizes of significant effects ranged from $\beta$ = `r expl_ts_ch_thr_reg_coef |> filter(rhs == 'child_thr') |> pull(est.std) |> round(2) |> max()` for the incongruent condition of the Simon Task to $\beta$ = `r expl_ts_ch_thr_reg_coef |> filter(rhs == 'child_thr') |> pull(est.std) |> round(2) |> min()` for the switch condition of the Posner Task.
+
+To summarize, we did not find evidence that exposure to threat or material deprivation (either in childhood or adulthood) was associated to specific inhibition or attention-shifting ability.
+Rather, we mostly found associations with task-general processes.
+We did find associations between childhood adversity and task-specific drift rates.
+However, correlations among task-specific drift rates were low, even for tasks thought to measure the same EF ability, suggesting that they reflect method variance rather than specific EF abilities.
+
+```{r}
+#| label: Figure7
+#| fig-width: 7
+#| fig-height: 5.5
+#| dpi: 600
+#| out-width: 6.5in
+#| fig-cap: | 
+#|   **Figure 7.** Standardized (non-preregistered) associations between adversity and task-specific drift rates.
+The effects are split out over adversity type (columns) and timing of adversity (rows).
+The gray area reflects the area of practical equivalence.
+Triangles depict effects for repeat/congruent conditions, and circles reflect effects for switch/incongruent conditions.
+Effects depicted with a solid point are practically equivalent, and effects depicted with a hollow point are not practically equivalent.
+Standard errors represent 95% confidence intervals.
+Statistical significance (tested against zero) is indicated with asterisks; * *p* < 0.05, ** *p* < 0.01, *** *p* < 0.001.
+fig7
+```
+<br>
+
+# Discussion
+
+We investigated associations between exposure to two types of adversity in adulthood---material deprivation and threat---with inhibition and attention shifting ability.
+Participants completed two inhibition tasks, three attention shifting tasks, and one basic processing speed task.
+First, we used DDM to separate raw performance into three distinct cognitive processes: (1) speed of evidence accumulation (drift rate), (2) response caution (boundary separation), and (3) speed of stimulus encoding and response execution (non-decision time).
+Finally, we used structural equation modeling to separate variance in each cognitive process into a task-general factor (shared across all tasks) and ability-specific factors.
+
+## Main findings
+
+People with more exposure to threat in adulthood---but not material deprivation---showed lower general processing speed.
+This is consistent with predictions from deficit frameworks [@sheridan_2014; @tucker_drob_2013].
+After accounting for general processing speed, there was no remaining variance that could be attributed to either inhibition ability or attention shifting ability.
+Additionally, the associations between material deprivation and threat in adulthood with both response caution and non-decision time were neither significantly different from zero, nor practically equivalent to zero.
+In other words, from these data, we were neither able to conclude that people with more exposure to adversity differed in their level of response caution and non-decision time, nor that these processes were similar to people with lower levels of adversity exposure.
+
+We found that variance in EF task performance was mostly explained by a task-general factor, not by ability-specific factors.
+This finding aligns with previous research applying DDM to EF tasks. 
+Three recent studies converged on the similar conclusion that performance across different EF tasks is best explained by a task-general drift rate factor [@frischkorn_2019; @hedge_2021; @loffler_2024; @weigard_2021].
+One study found that the task-general drift rate factor fully reflected basic processing speed, rather than common EF processes [@loffler_2024].
+Similarly, another study including several inhibition tasks found that shared variance was mostly explained by basic processing speed and strategy differences, not by inhibition ability [@hedge_2021].
+It is therefore not surprising that we were unable to estimate ability-specific EF factors after accounting for general processing.
+These issues tie into several other criticisms of common EF tasks in the cognitive psychology literature---such as low test-retest reliability---which should be considered when using these tasks to test associations with adversity [@vonbastian_2020].
+
+Our results also align with research showing that associations between adversity and raw performance (e.g., response times, accuracy) on EF tasks are mostly task-general.
+For instance, across three large cohorts, youth from more disadvantaged backgrounds had lower task-general raw performance across a wide range of cognitive tasks.
+After controlling for task-general effects, performance on specific intelligence tasks was largely practically equivalent to zero, and performance on some specific EF tasks even appeared enhanced [@bignardi_2024].
+Finally, previous research applying DDM in the context of adversity found similar (but not identical) results as reported here.
+Youth with more exposure to threat (but not material deprivation) had slower general processing speed across three EF tasks and a basic processing speed task, while most specific abilities remained intact [@vermeent_2024a].
+In contrast to the findings reported here, these youth also showed more response caution.
+It is important to highlight that both studies measured specific EF abilities with only a single task.
+This makes it unclear to which extent the task-specific effects they observed reflected EF abilities (see also our non-preregistered findings discussed below).
+
+## Non-preregistered findings
+
+Additional non-preregistered analyses showed that childhood exposure to both material deprivation and threat were associated with slower general processing speed.
+The effect sizes of these associations were larger than those of recent adversity exposure.
+This finding is striking considering the average age in our sample was 39 years.
+This is consistent with previous literature suggesting that adverse experiences early in life have a lasting effect on brain development [@shonkoff_2012; @nelson_2020a; @nelson_2020b].
+For instance, both experiences of malnutrition and exposure to stress (e.g., as a result of threat) have been linked to structural changes in brain networks involved in cognitive functioning [@algarin_2017; @polavarapu_2017; @rebello_2018].
+
+A second important non-preregistered finding is that after controlling for task-general processing, adversity exposure was negatively associated with several task-specific drift rates, mostly relating to childhood exposure to threat and deprivation.
+Importantly, the correlations among task-specific drift rates were all small---with the exception of correlations among switch and repeat conditions of the same attention shifting tasks.
+This suggests that these negative associations should not be interpreted as lower ability-specific processing, relating to inhibition and attention shifting ability.
+Instead, it appears to relate to information processing that is much more specific to individual tasks.
+This is consistent with literature showing that a substantial part of the variance in performance on cognitive task consists of method variance resulting from the nature of the task itself, rather than what the task is designed to measure [@barkley_2012; @rey_mermet_2024; @schubert_2016].
+
+## Implications for adversity research and interventions
+
+Our results have two implications for adversity research and interventions, both of which warrant caution against interpreting raw performance on individual EF tasks.
+First, adversity appears to be associated mostly with task-general processes, rather than specific abilities.
+This contrasts with common inferences in both the deficit and adaptation literature, which tend to explain performance differences in terms of specific EF abilities, such as inhibition [@farah_2006; @fields_2021; @mezzacappa_2004; @mittal_2015; @noble_2005] and attention shifting [@fields_2021; @howard_2020; @nweze_2021; @young_2022; @mittal_2015].
+Moreover, our results show that accounting for task-general processing does not guarantees that task-specific associations with adversity reflect differences in EF.
+Instead, negative associations with adversity appear to be partly driven by specific features of individual tasks.
+A general recommendation for adversity research is therefore that studies should include (1) multiple EF tasks to quantify task-general processing speed, and (2) ideally two or more tasks measuring the same EF ability to quantify ability-specific processing.
+DDM and SEM can then be used to account for task-general and task-specific processes.
+For interventions, our results imply that individual EF tasks, or batteries of EF tasks each of which measures a different ability, have limited value as screening tools; and, that training performance on such tasks is unlikely to transfer to broader, sustainable changes in EF abilities.
+
+Our findings complicate research that aims to understand how EF abilities develop in adverse environments.
+We agree with others that results like ours do not necessarily mean that EF abilities do not exist [e.g., @loffler_2024], or that adversity is not associated with differences in EF.
+Rather, it points to issues with how we *measure* EF abilities.
+It is difficult to provide a straightforward solution for this problem, which is still a much-debated topic in cognitive psychology [@vonbastian_2020].
+Part of the solution may lie in adjusting the content of EF tasks, for instance, by using real-world content rather than abstract content [@young_2022].
+By increasing familiarity, relevance, and/or valence, such content might increase the engagement of EF abilities [@niebaum_2023].
+Another solution could be to use adaptive testing procedures, which appear to improve estimation of EF abilities [@draheim_2021b].
+However, more fundamental changes might be necessary as well, such as developing new paradigms that align more closely to people's everyday experiences, goals, and challenges [@doebel_2020; @miller_cotto_2022; @niebaum_2023].
+
+## Strengths, limitations, and future directions
+
+Our study has three main strengths.
+First, we measured inhibition and attention shifting ability using multiple tasks for each EF ability.
+This approach provided two advantages: (1) we could distinguish between general versus specific processes, and (2) we could measure EF abilities without task-specific measurement error.
+Second, we used DDM to distinguish between evidence accumulation (EF ability, general speed of processing), response caution, and non-decision time (stimulus encoding, response execution).
+The DDM allowed us to better understand the cognitive processes driving performance in people with more exposure to adversity.
+Third, we included a large, representative sample of the Dutch population sample with sufficient variance on key adversity measures.
+
+Our study has three main limitations. 
+First, although we included confounding variables based on explicit, theory-guided causal assumptions, the cross-sectional and partly retrospective nature of our data prohibit us from establishing true causal relationships.
+Second, due to time constraints, we only included two inhibition tasks and three attention shifting tasks, limiting our ability to model shared variance between them.
+Future studies could include more EF tasks, and ideally include several basic processing speed tasks.
+Third, participants completed the cognitive tasks online in their own home.
+Despite our controlling for environmental factors that may have disrupted task performance, the home setting may have reduced the reliability of performance measures.
+
+We envision three future directions.
+First, because adversity tends to be associated with task-general processes, it will be important to better understand the nature of these processes.
+To the extent that task-general drift rates reflect general processing speed, it may partly result from lower white matter tract integrity [@fuhrmann_2020; @kievit_2016].
+Such associations with structural brain differences may account for the finding that task-general drift rate is relatively stable over time [@schubert_2016; @weigard_2021b].
+Yet, task-general drift rate may be associated with other factors such as effort, fatigue, or hunger [@weigard_2021b], which may play a bigger role for people from adverse environments [@brose_2012; @sliwinski_2006; @schwabe_2013].
+Thus, future studies may include (a combination of) brain measures or measures of mental states to better understand inter- and intrapersonal differences in task-general drift rates.
+
+Second, future research should investigate why adversity is negatively associated with task-specific drift rates. 
+Our results suggest that task-specific drift rates mostly capture method variance rather than specific EF abilities.
+We currently do not have a clear explanation for their negative associations with adversity.
+One possible explanation might be that the associations reflect people's difficulty with processing abstract content.
+Previous research found that youth with more exposure to adversity performed lower on a working memory updating task (compared with youth with fewer adversity exposures) when task stimuli were abstract geometric shapes.
+However, their performance equalized when the task included real-world content [@young_2022].
+All tasks in our study included abstract content, and each study used different stimuli, which may account for the negative associations with adversity and the lack of correlations between task-specific drift rates.
+However, if so, this would not explain the low correlations between congruent and incongruent conditions of both the Flanker and Simon task, as these conditions do include the same stimuli.
+Nevertheless, research could more systematically test associations between adversity and processing as a function of task content.
+
+Third, future research could design preregistered, well-powered analyses to tease apart the unique effects of childhood adversity and recent adversity on cognitive performance, as well as their combined effects.
+Our understanding of how cognitive outcomes later in life depend on the developmental timing of adversity exposure is still limited [@frankenhuis_2020].
+In this study, we assumed that recent adversity exposure mediates the effect of childhood adversity exposure on cognitive processes, as people who experienced adversity early in life are more likely to also experience adversity later in life (e.g., due to systemic constraints; @hazel_2008).
+Our study focused on adversity in adulthood, the average or summed adversity exposure over the past few years.
+Future research should distinguish such recent adversity from acute stress. 
+For instance, some research suggests that childhood adversity enhances specific EF abilities, but only under situations of acute stress [@mittal_2015; @young_2018].
+Thus, future research should tease apart effects of adversity during childhood and adulthood and study how their effects are moderated by acute stress.
+
+## Constraints on generality
+
+Our research question concerned the association between adversity exposure and performance on EF tasks, using cognitive modeling to investigate whether these associations are driven by task-general processing speed or specific EF abilities.
+Our target population was Dutch adults with a broad range of adversity exposures.
+The LISS panel ensures representation of the Dutch adult population in terms of---among other factors---age, education, and socioeconomic status.
+Our findings may generalize to adults in other Western populations, but not necessarily to adults in non-Western populations [@nketia_2023].
+We used standardized inhibition and attention-shifting tasks that are often used in cognitive psychology. 
+The findings should generalize to other tasks used to measure these abilities, as long as they rely on speeded responses.
+All tasks included abstract stimuli (e.g., geometric shapes, words).
+Results are likely to be different when using more real-world stimuli, or when changing the tasks in other ways to align them with people's lived experiences [@doebel_2020; @miller_cotto_2022; @young_2022].
+Participants completed the study online in their home environment. 
+To reproduce the effects, it is important to carefully instruct participants to limit environmental distractions and complete the study in full-screen mode.
+The results should generalize to more standardized settings (e.g., the lab), but may not generalize to public settings outside of the home environment.
+
+## Conclusion
+
+Adversity research has made important steps in recent years, revealing how exposure to adversity may impair certain EF abilities while enhancing others.
+Further progress in this field hinges on our capacity to accurately measure specific EF abilities.
+We show, consistent with previous research, that this is a more challenging task than typically assumed.
+As a result, adversity researchers likely overestimate the strength of associations between adversity and EF abilities when only analyzing raw performance.
+We present a way forward by building on a combination of cognitive modeling and structural equation modeling. 
+Embracing these techniques, and revising theories in light of their findings, will enhance our understanding of how exposure to adversity shapes EF.
+
+{{< pagebreak >}}
+
+# References
+
+::: {#refs}
+:::

### 2024-12-18 manuscript/masked_manuscript.qmd
add masked manuscript

--- 2024-12-18 manuscript/manuscript.qmd
@@
 
-#### **Adversity is associated with lower general processing speed rather than specific executive functioning abilities**
-
-<br>
-
-#### Stefan Vermeent^1,2^, Anna-Lena Schubert^3^, & Willem E. Frankenhuis^2,4^
-
-#### ^1^ Department of Psychology, Utrecht University, Utrecht, The Netherlands
-
-#### ^2^ Max Planck Institute for the Study of Crime, Security, and Law, Freiburg, Germany
-
-#### ^3^ Department of Psychology, University of Mainz, Mainz, Germany
-
-#### ^4^ Evolutionary and Population Biology, Institute for Biodiversity and Ecosystem Dynamics, University of Amsterdam, Amsterdam, the Netherlands
-
-<br>
-
-# Author Note
-
-This research was conducted in part using ODISSEI, the Open Data Infrastructure for Social Science and Economic Innovations (<https://ror.org/03m8v6t10>).
-WEF’s contributions have been supported by the Dutch Research Council (V1.Vidi.195.130) and the James S. McDonnell Foundation (https://doi.org/10.37717/220020502).
-
-Stefan Vermeent served as lead for conceptualization, data curation, methodology, formal analysis, and writing the original draft.
-Anna-Lena Schubert served in a supporting role for conceptualization, formal analysis, methodology, and writing-review and editing.
-Willem E.
-Frankenhuis served in a supporting role for conceptualization, methodology, supervision, funding acquisition, and writing-review and editing.
-
-The preregistration, analysis scripts, and study materials are available on the article's Github repository ([https://stefanvermeent.github.io/liss_ef_2024/](https://stefanvermeent.github.io/liss_ef_2024/)).
-In this paper, we make use of data from the LISS panel (Longitudinal Internet studies for the Social Sciences) managed by the non-profit research institute Centerdata (Tilburg University, the Netherlands).
-All datasets are available to researchers in the LISS data archive, after signing a data use agreement ([https://www.dataarchive.lissdata.nl/study-units/view/1](https://www.dataarchive.lissdata.nl/study-units/view/1)).
-
-Correspondence concerning this article should be addressed to Stefan Vermeent, Department of Psychology, Utrecht University, Heidelberglaan 1, 3584 CS Utrecht, The Netherlands.
-Email: p.c.s.vermeent@gmail.com / p.c.s.vermeent@uu.nl.
-
-We have no conflict of interest to disclose.
-
-<br>
-
-**word count:** 9573
-
-{{< pagebreak >}}
-
 # Abstract
@@
 {{< pagebreak >}}
+
+#### **Adversity is associated with lower general processing speed rather than specific executive functioning abilities**
 

### 2024-12-18 manuscript/titlepage.qmd
add masked manuscript

--- 2024-09-12 manuscript/manuscript.qmd
@@
   word_document: default
-editor: 
-  markdown: 
-    wrap: sentence
 ---
 
-```{r include = FALSE}
-library(flextable)
-library(stringr)
-library(dplyr)
-library(tidyr)
-library(tibble)
-library(lavaan)
-library(officer)
-
-source('../scripts/0_functions/general-functions.R')
-
-load("../data/exclusions.RData")
-load("../analysis_objects/results_exploratory.RData")
-
-knitr::opts_chunk$set(
-  echo = F,
-  fig.align = "center",
-  fig.pos = "!t", 
-  out.extra = "",
-  fig.show = "asis",
-  message = FALSE,
-  tab.topcaption = T,
-  warning = FALSE
-)
-
-# set up flextable for tables
-set_flextable_defaults(
-  font.family = "Times", 
-  font.size = 10,
-  font.color = "black",
-  line_spacing = 1,
-  padding.bottom = 1, 
-  padding.top = 1,
-  padding.left = 1,
-  padding.right = 1
-)
-```
-
-#### **Attention shifting and inhibition abilities among people living in adverse conditions**
+#### **Adversity is associated with lower general processing speed rather than specific executive functioning abilities**
 
@@
 
-# Data Availability
+We have no conflict of interest to disclose.
 
-All scripts and materials needed to reproduce the findings are available on the article's Github repository ([https://stefanvermeent.github.io/liss_ef_2024/](https://stefanvermeent.github.io/liss_ef_2024/)).
-In this paper, we make use of data from the LISS panel (Longitudinal Internet studies for the Social Sciences) managed by the non-profit research institute Centerdata (Tilburg University, the Netherlands).
-All datasets are available in the LISS data archive ([https://www.dataarchive.lissdata.nl/study-units/view/1](https://www.dataarchive.lissdata.nl/study-units/view/1)). 
-Researchers who want to access the data are required to sign a statement confirming that information about individual persons, households, etc., will not be released to others (go to [https://statements.centerdata.nl](https://statements.centerdata.nl) for more information).
+<br>
 
-# Funding statement
-This research was conducted in part using ODISSEI, the Open Data Infrastructure for Social Science and Economic Innovations (<https://ror.org/03m8v6t10>).
-WEF’s contributions have been supported by the Dutch Research Council (V1.Vidi.195.130) and the James S. McDonnell Foundation (https://doi.org/10.37717/220020502).
-
-# Disclosures
-
-We declare no conflicts of interest.
-
-# Ethics Approval Statement
-
-This study was approved by the Ethics Review Board of the Faculty of Social & Behavioural Sciences of Utrecht University (FETC20-490) and the Ethics committee for research in the Sciences and Life Sciences of the University of Amsterdam (FNWI-41_2023).
-
-\pagebreak
-
-People who live in adverse conditions tend to show lowered performance on executive functioning tasks [xxx].
-For example, exposure to deprivation has been linked to slower performance on inhibition tasks [@farah_2006; @fields_2021; @mezzacappa_2004; @noble_2005].
-This is typically interpreted as reflecting an impaired ability to maintain attentional focus on a central goal while ignoring distractions.
-These findings fit into a broader deficit framework, which emphasizes the ways in which adversity may hamper cognitive development across many different domains [xxx].
-
-In contrast, adaptation frameworks posit that people from adversity may develop specialized cognitive abilities through developmental adaptations in response to environmental challenges [@ellis_2022; @frankenhuis_2013; @frankenhuis_2020].
-This may result in enhanced or intact performance in specific EF abilities in relation to specific types of adversity.
-Consistent with these predictions, some studies have found intact or even enhanced performance in attention shifting, which is the ability to dynamically switch attention between two tasks [@fields_2021; @howard_2020; @nweze_2021; @young_2022; @mittal_2015].
-Thus, an important aim of both frameworks is to understand how adversity exposure is associated with differences in specific EF abilities.
-
-However, both deficit and adaptation interpretations suffer from a *performance-ability gap*.
-That is, lower task performance (e.g., slower response speed, lower accuracy) does not necessarily reflect lower cognitive ability.
-There are two reasons for this.
-First, raw performance *within* individual EF tasks is influenced by several processes besides the ability of interest, such as response caution and the speed of response execution [@hedge_2021; @ratcliff_2008; @stahl_2014].
-Second, raw performance *across* EF tasks is influenced by general processes, such as basic speed of processing [@loffler_2024].
-It therefore remains an open question to what extent adversity is associated with specific EF abilities after accounting for other sources of individual differences.
-In this study, we use cognitive modeling to more precisely estimate associations between adversity and two EF abilities: inhibition and attention shifting.
-
-## Separating EF ability from other decision-making processes
-
-Although mean RTs are routinely used as a direct proxy for cognitive ability, it is well-known that they confound multiple stages of processing, from initial preparations, to the active processing of task-relevant information, to deciding on and executing a response [@ratcliff_1978; @forstmann_2016; @wagenmakers_2009].
-In addition, people differ in how cautiously they make decisions.
-More cautious people tend to respond more slowly, but this does not necessarily indicate a lower ability.
-Differentiating between response caution and EF abilities on the basis of raw performance is difficult, especially because the trade-off between RTs and accuracy is often subtle [@domingue_2022].
-Thus, investigating how adversity is associated with inhibition and attention shifting requires isolating the factors that contribute to differences in RTs.
-
-One way of doing this is through the use of cognitive models like the Drift Diffusion Model [DDM\; @forstmann_2016; @ratcliff_1998; @ratcliff_2008; @wagenmakers_2009].
-The DDM assumes that people accumulate information for one of two responses (e.g., left or right button), until they have enough evidence for one response over the other (see Figure 1 for a visual overview of the DDM).
-When applied to trial-level RTs and accuracies, the DDM estimates four parameters that map onto theoretically meaningful cognitive processes.
-The *drift rate* reflects the rate at which the evidence accumulation process drifts towards the decision boundary, and thus provides a measure of the speed/efficiency of information processing.
-A higher drift rate leads to faster responses and higher accuracy (all else being equal).
-It is generally assumed that individual differences in EF abilities are captured in the drift rate (although not exclusively, see below\; @loffler_2024; @vermeent_2024a; @weigard_2021).
-The *boundary separation* reflects the width between the two decision boundaries, and provides a measure of a person's response caution.
-The *non-decision time* reflects a combination of preparation time (e.g., stimulus encoding) as well as execution time (e.g., time spent pressing the button).
-Finally, the *starting-point* (not considered here) allows for initial bias for one of two responses.
-
-# Accounting for task-impurity
-
-The EF literature has a long history of accounting for task impurity by using structural equation modeling (SEM) [@miyake_2000].
-Specific features of EF tasks may give rise to individual differences that do not purely reflect the EF ability.
-For example, content differences (e.g., numeric vs. figural, abstract vs. concrete) have been shown to affect cognitive performance (e.g., @duquennois_2022; @lerche_2020; @young_2022).
-SEM enables separating such method variance and variance associated with non-executive abilities from domain-specific variance (e.g., inhibition, shifting) (@miyake_2012).
-This requires including at least two or more different tasks that measure the same EF ability.
-The most widely used latent factor model of EF distinguishes between inhibition, attention shifting, and working memory updating, although several studies were unable to recover this factor structure, especially with regard to inhibition as a unitary and separate factor [@karr_2018; @rey_mermet_2018; @stahl_2014; @vonbastian_2020].
-
-Studies that combine SEM with cognitive modeling have found that shared variance across EF tasks largely reflects processing speed.
-One study was able to replicate the three-factor structure of EF when using raw performance measures.
-However, when modeling drift rates, all measures loaded on a single latent factor, which was fully explained by processing speed [@loffler_2024].
-Another investigation using an extension of the DDM for inhibition tasks found that shared variance across nine different tasks mostly reflected processing speed and strategy differences, not inhibition ability [@hedge_2021].
-
-This poses a challenge for adversity research, which often investigates associations between adversity and EF abilities based on a small set of tasks.
-Two recent studies suggest that this may overestimate the extent to which adversity lowers specific EF abilities.
-First, socioeconomic status (which is correlated with, but not the same as adversity) was mostly negatively associated with a task-general factor based on raw performance, while task-specific associations for the most part did not differ meaningfully from zero [@bignardi_2024].
-Second, household threat (but not material deprivation) was negatively associated with task-general drift rates, but not with task-specific variance on three EF tasks [@vermeent_2024a].
-However, both studies used data sets that only included a single task per ability, which may have contributed to the lack of task-specific associations.
-In this study, we built on this work to investigate how adversity is associated with differences in task-general speed, shifting and inhibition ability, and response caution.
-
-# The current study
-
-We analyzed performance on two inhibition tasks, three attention shifting tasks, and one basic processing speed task in relation to adversity exposure in a sample of adults.
-Using a combination of DDM and SEM, our aim was to answer three central questions.
-First, how is adversity associated with general speed of processing that is shared across all cognitive tasks?
-Second, how is adversity associated with inhibition and attention shifting ability after accounting for general speed of processing?
-Third, how is adversity associated with a general tendency for response caution (i.e., across all tasks and conditions), as well as with response caution associated with EF performance (i.e., in task conditions including an EF component)?
-We focused on two self-reported types of adversity: material deprivation (a lack of access to material resources) and threat (the potential for harm imposed by others).
-Both forms of adversity have been associated with both lowered and enhanced performance in EF [e.g., @fields_2021; @schafer_2022; @sheridan_2022; @vermeent_2024a; @young_2022], and are central to recent dimensional models of adversity [@mclaughlin_2021; @sheridan_2014].
-
-We preregistered potential data patterns that we considered consistent with hypotheses generated from deficit and adaptation frameworks.
-Any negative association between adversity and either general speed of processing or domain-specific processing would be consistent with deficit frameworks.
-Associations with response caution would not be consistent with deficit frameworks, as they reflect strategies rather than cognitive abilities.
-Two patterns would be consistent with adaptation frameworks. 
-First, if adversity is positively associated with attention shifting-related processing, regardless of the association with inhibition-related processing.
-Second, if adversity is associated with intact attention shifting-related processing *and* negatively associated with inhibition-related processing.
-If adversity is associated with intact performance in both abilities, this would suggest that adversity exposure is unrelated to the development of these abilities.
-
-# Methods
-
-## Participants
-
-The study was completed by a total of `r exclusions$sample$finish` participants who were drawn from the Dutch Longitudinal Internet Studies for the Social Sciences (LISS) panel [@scherpenzeel_2011].
-The LISS panel is an invite-only, representative probability sample of the Dutch population consisting of roughly 7,500 individuals across 5,000 households.
-LISS participants complete a yearly core battery of questionnaires about various key domains, among which their financial situation over the past year.
-In addition, they can participate in further monthly studies covering a wide range of different topics.
-The current study was fielded between May and August 2024.
-Participants were able to complete the cognitive tasks across two or more sessions to increase participation rates.
-People were eligible for participation if they were between 18 and 55 years old, and if they had agreed to linking their LISS data to government microdata (not relevant here).
-If performance on a specific task was at or below chance level, we excluded the data for that task but retained other task data of that participant.
-The final sample after exclusions consisted of `r exclusions$sample$final` participants.
-
-## Adversity measures
-
-We preregistered our approach to compute adversity composites based on observed correlations between measures.
-In short, if all measures of an adversity type correlated .60 or higher (indicating "strong" correlations), we calculated a uniformly weighted average.
-If one or more correlations were lower than .60, we used Principal Component Analysis (PCA) to the separate measures and extracted only the first principal component score.
-Perceived scarcity and income-to-needs were considered as two separate types of material deprivation (and unpredictability), regardless of their correlation.
-
-### Neighborhood threat
-
-Participants completed the Neighborhood Violence Scale [NVS\; @frankenhuis_2018; @frankenhuis_2020] in the current study.
-The NVS included seven items about current perceived neighborhood threat (e.g., “Where I live, it is important to be able to defend yourself against physical harm”), on a scale of 1 ("Completely disagree") to 7 ("Completely agree").
-All items of the NVS were averaged together and standardized.
-We also included four items on perceived neighborhood threat from the LISS archive (six waves: <https://doi.org/10.17026/dans-zch-j8xt>), in which participant reported how often it happens that they 1) “avoid certain areas in your place of residence because you perceive them as unsafe”, 2) “do not respond to a call at the door because you feel that it is unsafe”, 3) “leave valuable items at home to avoid theft or robbery in the street?”, 4) “make a detour, by car or on foot, to avoid unsafe areas?” on a scale of 1 (“(Almost) never”), 2 (“Sometimes”), or 3 (“Often”).
-We recoded these items so that "(Almost) never" corresponded to zero.
-We summed the items within each wave, and then calculated an average across waves for which participants had data.
-
-We used seven items from the LISS archive (six waves: <https://doi.org/10.17026/dans-zch-j8xt>) in which participants reported whether or not they were the victim of eight types of crime in the last two years: (1) burglary or attempted burglary; (2) theft from their car; (3) theft of their wallet or purse, handbag, or other personal possession; (4) wreckage of their car or other private property; (5) intimidation by any other means; (6) maltreatment of such serious nature that it required medical attention; (7) maltreatment that did not require medical attention.
-As the last wave occurred in 2018, we included the same items in the current study asking about people's crime exposure in the last two years.
-We computed the total number of distinct crimes that participants were exposed to at any moment in time (a 'variety score'\; @sweeten_2012).
-
-Correlations among the two measures of perceived neighborhood crime and crime victimization were low.
-Therefore, we used PCA to extract the first principal component score.
-This score accounted for XXX% of the variance in the three measures.
-
-### Perceived scarcity
-
-Measures of perceived scarcity were derived from the LISS archive, using the yearly recurring core study on household and personal income (16 waves: <https://doi.org/10.57990/1gr4-bf42>).
-First, participants reported how difficult it currently is to live off the income of their household, on a scale from 0 (very hard) to 10 (very easy). 
-Second, participants reported which of the following best applied to their current financial situation: (1) “we are accumulating debt”; (2) “we are somewhat eating into savings”; (3) “we are just managing to make ends meet”; (4) “we have a little bit of money to spare”; (5) “we have a lot of money to spare”.
-Third, participants reported which of the following applied to their current financial situation (0 = no, 1 = yes): (1) “having trouble making ends meet”; (2) unable to quickly replace things that break”; (3) “having to lend money for necessary expenditures”; (4) “running behind in paying rent/mortgage or general utilities”; (5) “debt collector/bailiff at the door in the last month”; (6) “received financial support from family or friends in the last month”.
-Responses were coded so that higher scores indicated more perceived scarcity.
-
-We first averaged each measure separately across waves, and then scaled them.
-As all item correlations were > .60, we computed a uniformly weighted average.
-
-### Cognitive measures
-
-All cognitive tasks were programmed in jsPsych 7.3 [@deleeuw_2015].
-At the start of the session, participants entered fullscreen mode to avoid distractions from other browser tabs.
-The tasks were presented against a light-gray background.
-
-*Flanker Task.* This task measures inhibition of distractor interference [@eriksen_1974].
-On each trial, participants saw five arrows side-by-side horizontally, pointing either left or right.
-Their task was to indicate the direction of the central arrow.
-The arrows were randomly presented 300 pixels above or below the center of the screen.
-On 50% of the trials, all arrows pointed in the same direction (congruent trials).
-On the other half, the arrows surrounding the central arrow pointed in the opposite direction (incongruent trials).
-Participants first completed eight practice trials, followed by two test blocks of 32 trials each, for a total of 64 trials.
-
-*Simon Task.* This task measures inhibition of prepotent responses [@simon_1969].
-On each trial, participants saw the word "LEFT" or "RIGHT" (printed in Dutch), presented either on the left or right side of the screen.
-Their task was to press the 'A' key if the word was "LEFT" and the 'L' key if the word was "RIGHT", regardless of the location on the screen.
-on 50% of the trials, the word matched the location (e.g., the word "LEFT" presented on the left side; congruent trials).
-On the other half, the word did not match the location (e.g., the word "LEFT" presented on the right side; incongruent trials).
-Participants first completed eight practice trials, followed by two test blocks of 32 trials each, for a total of 64 trials.
-
-*Color-Shape Task.* This task measures the ability to shift attention between different task goals [@miyake_2004].
-On each trial, participants saw a square or a circle in the center of the screen that was either blue or yellow.
-Depending on the task rule printed above the stimulus, their task was to classify the stimulus based on it's shape or color.
-On 50% of the trials, the rule was the same as on the preceding trial (repeat trials).
-On the other half, the rule was different than on the preceding trial (switch trials).
-The same stimulus was never presented more than twice in a row, and there were never more than three repeat or switch trials in a row. 
-Participants first completed eight practice trials, followed by two test blocks of 32 trials each, for a total of 64 trials.
-
-*Animacy-Size Task.* This task measures the ability to shift attention between different task goals [@arrington_2004].
-On each trial, participants saw a single noun (in Dutch) referring to an animal or object [adopted from @braem_2017].
-Depending on the task rule printed above the noun, their task was to classify it based on whether it referred to a living or non-living thing (e.g., wasp vs.
-piano), or whether it referred to something that was smaller or larger than a soccer ball (e.g., ring vs.
-dolphin).
-On 50% of the trials, the rule was the same as on the preceding trial (repeat trials).
-On the other half, the rule was different than on the preceding trial (switch trials).
-There were never more than three repeat or switch trials in a row.
-Participants first completed eight practice trials, followed by two test blocks of 32 trials each, for a total of 64 trials.
-
-*Global-local Task.* This task measures the ability to shift attention between task goals.
-Stimuli were adapted from @huizinga_2010.
-On each trial, participants saw a large square or rectangle which was composed of 16 small squares or rectangles.
-The stimulus where flanked on both side by a drawing of an elephant or mouse, which was presented 1,000 ms prior to the appearance of the stimulus.
-If the stimulus was flanked by the image of an elephant (50% of trials), participants had to indicate whether the global image was a square or rectangle.
-If the stimulus was flanked by the image of a mouse (50% of trials), participants had to indicate whether the local images were squares or rectangles.
-On 50 % of the trials, the rule was the same as on the preceding trial (repeat trials).
-On the other half, the rule was different than on the preceding trial (switch trials).
-Finally, the stimuli were congruent on 50% of the trials (e.g., large square consisting of small squares) and incongruent on the other half (e.g., large square consisting of small rectangles).
-Congruency, task rule (switch vs. repeat) or focus (global vs. local) where never repeated more than three times in a row.
-Participants first completed eight practice trials, followed by two test blocks of 32 trials each, for a total of 64 trials.
-
-*Posner Task.* This task measures basic speed of processing.
-On each trial, participants saw two letters in the center of the screen, drawn from the set A, B, F, H, Q, a, b, f, h , q.
-Their task was to indicate whether the letters were the same (e.g., "AA", "bB") or different (e.g., "AQ", "Fh").
-On 50% of the trials, the letters were the same, and on the other half they were different.
-Participants first completed 8 practice trials, followed by two test blocks of 32 trials each, for a total of 64 trials.
-
-### Covariates
-
-We used Directed Acyclic Graphs (DAG) to identify potential confounders of the key estimands (i.e., the association between self-reported threat and deprivation with cognitive outcomes).
-A DAG is a visual overview of our assumptions about how variables are causally related.
-They are graphs consisting of nodes (variables) and directed arrows (causal pathways).
-An arrow between two variables represents the assumption that experimentally manipulating the variable at the origin of the arrow will change the variable at the end of the arrow (but not the other way around).
-DAGs help in the identification of variables that need to be adjusted for in the statistical models (i.e., confounders with arrows to both the main predictor and the outcome), and, importantly, which variables should not be adjusted for (i.e., colliders and mediators).
-For more information on DAGs, see @rohrer_2018.
-
-We preregistered separate DAGs for threat and deprivation.
-As potential confounders of the effect of threat and deprivation on cognitive performance, we examined (1) age [@salthouse_2016; @salthouse_2019; @starns_2010], (2) education [@hofmarcher_2021], (3) sex [@ning_2023], and (4) childhood adversity exposure (material deprivation and threat combined) [@bos_2009; @goodman_2019].
-We also examined evidence for potential causal associations between threat and deprivation.
-Examinations were based on previous literature and/or our own assumptions.
-
-In the case of threat exposure, we were theoretically less sure about the potential effects of age and education.
-Therefore, we also statistically tested support for parts of the DAG for threat involving age, education, and deprivation using data from a previous LISS study [@vermeent_2024c], using working memory capacity as the outcome variable in the DAG.
-Specifically, we used the *dagitty* R package [@ankan_2021; @textor_2016] to test whether DAG-implied independencies between key variables (conditional on the other causal pathways) were supported by these data (see the preregistration for more details).
-
-Figure 2A depicts the final preregistered DAGs for material deprivation and threat exposure.
-For material deprivation, we decided to control for age, education, sex, and childhood adversity exposure on theoretical grounds.
-For threat exposure, we decided to control for age, sex, and childhood adversity, but *not* for education.
-We did not find theoretical support for an causal effect of education on threat exposure.
-In addition, our statistical analysis based on previously collected data did not provide evidence against independence between education and threat, suggesting that people with lower education did not report more exposure to threat.
-
-Some recent work suggests that material deprivation precedes exposure to adversities such as threat [@bywaters_2016; @lacey_2022; @ning_2023].
-If so, threat exposure is a mediator of the effect of material deprivation on cognitive performance, and should therefore not be controlled for when testing the total effect of material deprivation.
-However, establishing this causal relationship is difficult in our study for two reasons.
-First, The studies listed above all focused on childhood adversity, while we focus on adversity in adulthood.
-Second, our measures of threat and deprivation were measured at the same time, making it difficult to establish causality.
-Therefore, we will present two sets of results for the associations between material deprivation and threat exposure with cognitive performance.
-For material deprivation, our primary model did *not* include threat exposure as a confounder.
-A secondary model *did* include threat exposure, which we considered this a robustness check to assess the effect of our causal assumption.
-For threat exposure, our primary model included material deprivation as a confounder.
-A secondary model did *not* include material deprivation.
-
-## Procedure
-
-We obtained ethical approval from the first author's institutional ethical board. 
-The study was fielded on the LISS platform, and participants could only complete the study on a laptop or desktop PC.
-Participants started with the six cognitive tasks, in randomized order.
-After each task, they were asked if they wanted to continue to the next task or conclude the session. 
-If they choose for the latter, they were able to continue with the tasks at another time.
-The cognitive tasks took around 25 minutes to complete.
-After completing all cognitive tasks, participants completed questionnaires about past and present neighborhood threat and material deprivation, exposure to crime victimization in the past two years, trait impulsivity (not considered here) and trait future orientation (not considered here).
-Finally, they answered a few standard LISS evaluation questions about their experiences with the study, with the opportunity to provide written feedback.
-
-## Analysis plan
-
-### Data cleaning
-
-For all tasks, we removed trials with RTs < 250ms and trials with RTs more than 3.2 SD above the intra-individual log-transformed mean RT.
-In addition, if participants performed at chance level on a particular task, we excluded the data for that task only.
-The cut-off for chance performance was determined using the accuracy at the 97.5 % tail of a binomial distribution in the case of pure guessing.
-
-### DDM estimation
-
-We used an Hierarchical Bayesian implementation of the DDM (HDDM) [@vandekerckhove_2011; @wiecki_2013].
-The benefit of HDDM over traditional approaches is that the model uses information of the whole sample to inform individual parameter estimation.
-The HDDM was applied to each task separately.
-Across all tasks, the bias parameter was held constant, as this parameter is uninformative when the decision boundaries are coded as 'correct' and 'incorrect' (as opposed to mapping on the response options).
-For the Posner Task, we estimated a single drift rate, boundary separation, and non-decision time for each participant.
-For all other tasks, drift rates, boundary separation, and non-decision times were estimated separately for congruent (repeat) and incongruent (switch) trials.
-
-The DDM models were fit using the *runjags* package [@denwood_2016].
-Each model was fit with three Markov Chain Monte Carlo (MCMC) chains.
-We used 2,000 burn-in samples and 10,000 additional samples, of which every 10th sample was retained, resulting in a total of 3,000 posterior samples.
-See the supplemental materials for more information about model convergence and fit assessment.'
-
-We accounted for two potential sources of measurement error in the DDM estimates: (1) environmental noise and (2) state anxiety.
-Both measures were repeated after each cognitive task.
-We measured environmental noise with a single item, rated on a scale of 1 to 5: "How much noise was there in your environment during the reaction game?"
-We measured state anxiety with the shortened version of the State-Trait Anxiety Inventory [STAI\; @bij_2003; @marteau_1992], which asks participants how calm, tense, upset, relaxed, content, and worried they currently felt, on a scale of 1 ("not al all") to 4 ("very much").
-We recoded (when necessary) and then summed the answers with higher scores reflecting more state anxiety.
-For each task, we will then calculate the deviation from the intraindividual mean anxiety score.
-We residualized variance associated with noise and anxiety out of all the drift diffusion estimates using linear regression.
-
-### SEM
-
-We constructed the final SEM sequentially.
-First, we optimized the fit of the drift rate, boundary separation, and non-decision time sub-models.
-Second, we combined these three models into a single measurement model.
-Third, we added the regression paths between measures of adversity and the latent factors.
-To assess goodness-of-fit, we used the root mean square error of approximation (RMSEA) and the comparative fit index (CFI). 
-CFI values >.0.90 (>0.95) and RMSEA values <0.08 (≤0.06) were interpreted as acceptable (good) fit.
-
-Each sub-model was a bi-factor model, containing as manifest variables the parameter estimates of both conditions for each task.
-Within each sub-model, we estimated covariances between the conditions of each task.
-We estimated a general factor accounting for all manifest variables.
-A second domain-specific inhibition factor loaded on the parameter estimates of incongruent trials of the Flanker and Simon task.
-A third domain-specific Attention Shifting factor loaded on the parameter estimates of switch trials of the Color-shape, Global-local, and Animacy-size task.
-These two factors were allowed to covary.
-For each sub-model, we compared the fit of this model to a version with a common EF factor loading on incongruent/switch trials of all tasks.
-We deemed the second model a better fit when we observed a significant chi squared change test and an AIC value \> 10.
-
-For the drift rate model, we interpreted the general factor as basic speed of processing, and the domain-specific factors as reflecting specific inhibition and attention-shifting ability (or common EF in the case of model 2).
-For the boundary separation/non-decision time models, we interpreted the general factor as general response caution/speed of non-decision processes, and the domain-specific factors as reflecting response caution/speed of non-decision processes specific to conflict trials.
-
-After optimizing each sub-model, we combined them into a single measurement model.
-The general latent factors (processing speed, general caution, and general non-decision time) were allowed to covary, as well as the domain-specific latent factors (separately for inhibition and attention shifting, unless a common factor was favored).
-
-Finally, we constructed two versions of the final model, one to estimate the association between material deprivation and the outcome measures, and one to estimate the association between threat exposure and the outcome measures.
-In both models, the adversity measure was regressed on each latent factor, together with the control variables (see the Control Variables section for more details).
-We do not report regression coefficients of control variables to prevent their over-interpretation.
-
-# Results
-
-## Model fit
-
-### DDM convergence and fit
-
-The DDM converged normally for all tasks except for the Color-shape task, which showed large jumps in the traces.
-We solved this by increasing burnin and the number of posterior samples.
-In addition, model fit was above the preregistered cut-off of *r* = .80 for all tasks (all *r*s > XXX).
-(see the supplemental materials for a full overview of model convergence and fit results).
-
-### SEM
-
-As preregistered, we started with optimizing fit in SEM models for each DDM parameter separately before combining them all into a single model.
-For drift rates, both the model specifying separate inhibition and attention shifting factors (model 1) as well as the model specifying a common EF factor (model 2) provided a good fit to the data (model 1: CFI = `r fit_v1_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_v1_fitstats[['rmsea.robust']] |> round(2)` [`r fit_v1_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_v1_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]; model 2: CFI = `r fit_v2_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_v2_fitstats[['rmsea.robust']] |> round(2)` [`r fit_v2_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_v2_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]).
-Model comparison showed that model 2 led only to a slightly deteriorated fit, which was below the preregistered cut-off (AIC = `r fit_v_compare$AIC |> diff() |> formatC(digits = 2, width = 3, flag = "0", format = 'f')`).
-Therefore, we selected the drift rate model with a general processing speed factor loading on all manifest drift rates and a common EF factor loading on drift rates of conflict trials.
-For boundary separation, both the model specifying separate factors for response caution on conflict trials of inhibition and attention shifting tasks (model 1) as well as the model specifying a common factor for response caution on conflict trials (model 2) provided a good fit to the data (model 1: CFI = `r fit_a1_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_a1_fitstats[['rmsea.robust']] |> round(2)` [`r fit_a1_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_a1_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]; model 2: CFI = `r fit_a2_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_a2_fitstats[['rmsea.robust']] |> round(2)` [`r fit_a2_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_a2_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]).
-Model comparison showed that model 2 led only to a slightly deteriorated fit, which was below the preregistered cut-off (AIC = `r fit_a_compare$AIC |> diff() |> formatC(digits = 2, width = 3, flag = "0", format = 'f')`).
-Therefore, we selected the boundary separation model with a general response caution factor loading on all manifest boundary separations and a common response caution factor loading on boundary separations of conflict trials.
-For non-decision time, the two preregistered models did not converge.
-A third model only including a general non-decision time factor did converge and provided a good fit to the data, CFI = `r fit_t3_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_t3_fitstats[['rmsea.robust']] |> round(2)` [`r fit_t3_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_t3_fitstats[['rmsea.ci.upper.robust']] |> round(2)`].
-The final model combining all three DDM parameters provided a good fit, although it required the addition of covariances between residual variances of manifest DDM parameters of the same task, CFI = `r fit_meas_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_meas_fitstats[['rmsea.robust']] |> round(2)` [`r fit_meas_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_meas_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]
-
-## Preregistered analyses
-
-## Non-preregistered analyses
-
-In our primary models, we found evidence suggesting an (indirect) association between childhood adversity and general speed of processing (note that in our DAGs, the effect of childhood adversity on cognitive processes is mediated by recent adversity).
-To explore potential *direct* effects of childhood adversity on task-general and domain-specific cognitive processes, we fitted an exploratory SEM model in which we evaluated associations between cognitive processes and childhood adversity *without including recent adversity*, and controlling for sex.
-We decided to include childhood exposure to threat and deprivation separately, in contrast to the primary model which included a composite of the two, consistent with the main analyses focusing on recent exposure to threat and deprivation.
-The results are summarised in Figure XX.
-
-For childhood exposure to deprivation, we found a significant negative association with general processing speed, $\beta$ = `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(est.std) |> round(2)`, SE = `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(se) |> round(2)`, 95% CI = [`r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(ci.lower) |> round(2)`, `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(ci.upper) |> round(2)`] *p* `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(pvalue_adj) |> format_p(pvalue = _)`.
-People who reported higher levels of childhood exposure to deprivation processed information more slowly across tasks.
-In addition, we found a significant positive association between childhood exposure to deprivation and non-decision time, $\beta$ = `r expl_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(est.std) |> round(2)`, SE = `r expl_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(se) |> round(2)`, 95% CI = [`r expl_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(ci.lower) |> round(2)`, `r expl_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(ci.upper) |> round(2)`] *p* `r expl_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(pvalue_adj) |> format_p(pvalue = _)`.
-People who reported higher levels of childhood exposure to deprivation took more time encoding stimulus information and/or executing responses across tasks
-We did not find significant associations between childhood exposure to deprivation and general response caution, nor to domain-specific factors.
-Equivalence tests revealed that none of the associations fell within the region of practical equivalence.
-
-For childhood exposure to threat, we found a significant negative association with general processing speed, $\beta$ = `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(est.std) |> round(2)`, SE = `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(se) |> round(2)`, 95% CI = [`r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(ci.lower) |> round(2)`, `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(ci.upper) |> round(2)`] *p* `r expl_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(pvalue_adj) |> format_p(pvalue = _)`.
-People who reported higher levels of childhood exposure to threat processed information more slowly across tasks.
-We did not find significant associations between childhood exposure to threat and general response caution or general non-decision time, nor to domain-specific factors.
-Equivalence tests revealed that only the association between childhood exposure to threat and general response caution fell within the region of practical equivalence (*p* `r expl_eqtests |> filter(lhs == "gen_a", rhs == "child_thr") |> pull(eq_pvalue) |> format_p(pvalue = _)`).
-
-# Discussion
-
-The goal of this study was to investigate how exposure to two types of adversity---material deprivation and threat---in adulthood is associated to inhibition and attention shifting ability.
-Participants completed two inhibition tasks, three attention shifting tasks, and one basic processing speed task.
-First, we used DDM to
-
-\pagebreak
-
-## References
-
-::: {#refs}
-:::
+**word count:** 9573

### 2024-12-18 materials/liss/S_colorshape/2_colorshape_procedure.js
fix bug in colorshape procedure script

@@ -22,37 +22,37 @@ var colorshape_test = {
 var colorshape_procedure1 = {
   timeline: [colorshape_test],
   timeline_variables: [
-    {stimulus: yellowtriangle_color, stim_chr: 'yellowtriangle', key_answer: 'ArrowLeft', data: {rule: 'color', type: 'first', variable: 'colorshape1', task: 'colorshape'}},
+    {stimulus: yellowtriangle_color, stim_chr: 'yellowtriangle', key_answer: 'A', data: {rule: 'color', type: 'first', variable: 'colorshape1', task: 'colorshape'}},
     {stimulus: yellowcircle_shape, stim_chr: 'yellowcircle', key_answer: 'L', data: {rule: 'shape', type: 'switch', variable: 'colorshape1', task: 'colorshape'}},
     {stimulus: bluecircle_shape, stim_chr: 'bluecircle', key_answer: 'L', data: {rule: 'shape', type: 'repeat', variable: 'colorshape1', task: 'colorshape'}},
     {stimulus: bluetriangle_color, stim_chr: 'bluetriangle', key_answer: 'L', data: {rule: 'color', type: 'switch', variable: 'colorshape1', task: 'colorshape'}},
     {stimulus: bluecircle_color, stim_chr: 'bluecircle', key_answer: 'L', data: {rule: 'color', type: 'repeat', variable: 'colorshape1', task: 'colorshape'}},
     {stimulus: bluecircle_shape, stim_chr: 'bluecircle', key_answer: 'L', data: {rule: 'shape', type: 'switch', variable: 'colorshape1', task: 'colorshape'}},
-    {stimulus: bluetriangle_shape, stim_chr: 'bluetriangle', key_answer: 'ArrowLeft', data: {rule: 'shape', type: 'repeat', variable: 'colorshape1', task: 'colorshape'}},
-    {stimulus: bluetriangle_shape, stim_chr: 'bluetriangle', key_answer: 'ArrowLeft', data: {rule: 'shape', type: 'repeat', variable: 'colorshape1', task: 'colorshape'}},
+    {stimulus: bluetriangle_shape, stim_chr: 'bluetriangle', key_answer: 'A', data: {rule: 'shape', type: 'repeat', variable: 'colorshape1', task: 'colorshape'}},
[…]

### 2024-12-18 supplement/supplement.qmd
add supplemental materials

---
bibliography: ../manuscript/references.bib
csl: ../manuscript/apa.csl
format: 
  docx:
    reference-doc: reference-doc.docx
    toc: true
    toc-location: left
    toc-title: Table of Contents
output:
  officedown::rdocx_document:
    page_margins:
      bottom: 1
      footer: 0
      gutter: 0
      header: 0.5
      left: 1
      right: 1
      top: 1
    plots:
      align: center
      caption:
        pre: 'Figure '
        sep: '. '
        style: Image Caption
    tables:
      caption:
        pre: 'Table '
        sep: '. '
        style: Table Caption
  pdf_document: default
  word_document: default
editor: 
  markdown: 
    wrap: sentence
---

```{r}
#| include: false
#| echo: false
library(flextable)
library(tidyverse)

load("../analysis_objects/task_reliability.RData")
load("../analysis_objects/hddm_model_fit")
load("../analysis_objects/env_effects.RData")
load("../analysis_objects/iv_histograms.RData")
load("../analysis_objects/confounder_coefficients.RData")
load("../analysis_objects/ddm_correlations.RData")

knitr::opts_chunk$set(
  echo = F,
  fig.align = "center",
  fig.pos = "!t", 
  out.extra = "",
  fig.show = "asis",
  message = FALSE,
  tab.topcaption = T,
  warning = FALSE
)

# set up flextable for tables
set_flextable_defaults(
  font.family = "Times", 
  font.size = 10,
  font.color = "black",
  line_spacing = 1,
  padding.bottom = 1, 
  padding.top = 1,
  padding.left = 1,
  padding.right = 1
)
```

# Relevance of supplemental materials

This file contains the supplementary materials for the article entitled "Adversity is associated with lower general processing speed rather than specific executive functioning abilities", which was submitted for consideration at **Journal of Experimental Psychology: General**.
The supplemental materials include supporting information relating to more detailed descriptive statistics, model convergence and fit, and details on confounders.

# Section 1. Descriptive statistics of adversity measures

Figures S1-S3 present histograms of each separate adversity measure, as well as the composite adversity measures used in the analyses.
See the main text for more information on how the composites were calculated.

<br>

```{r}
#| label: FigureS1
#| fig-width: 6.5
#| fig-height: 5
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure S1.** Histograms of material deprivation measures
p_scar_m_combn_hist
```

<br>

```{r}
#| label: FigureS2
#| fig-width: 6.5
#| fig-height: 5
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure S2.** Histograms of threat measures
threat_combn_hist
```

<br>

```{r}
#| label: FigureS3
#| fig-width: 6.5
#| fig-height: 5
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure S3.** Histograms of threat measures
child_adv_combn_hist
```

<br>

# Section 2. Additional information on cognitive tasks

## Distributions of response times and error rates

Figure S4 and S5 show the distributions of mean response times and mean error rate for each cognitive task.

<br>

```{r}
#| label: FigureS4
#| fig-width: 6.5
#| fig-height: 7
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure S4.** Distributions of response times for all tasks.
supp_fig_rt_hist +
  theme_classic()
```

<br>

```{r}
#| label: FigureS5
#| fig-width: 6.5
#| fig-height: 7
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure S5.** Distributions of error rates for all tasks.
supp_fig_acc_hist +
  theme_classic()
```

<br>

## Condition manipulation checks

Table S1 presents manipulation checks for response times for the inhibition and attention-shifting tasks.
The manipulation checks were based on mean log-transformed response times using paired-sample *t*-tests.
For each task, we tested whether there was a significant difference in mean log-transformed response time between the congruent (repeat) condition and the incongruent (switch) condition.
All tasks showed a significant difference in the expected direction, with participants on average being faster on the congruent (repeat) condition compared to the incongruent (switch) condition (all *p*s < .001).

<br>

```{r}
#| tab.id: tableS1
#| results: markup
man_checks |> 
  flextable() |> 
  add_header_row(
    values = " ",
    colwidths = 4
  ) |> 
  flextable::compose(
    i = 1, j = 1,
    as_paragraph(as_b("Table S1. "), "Log-transformed response time differences across conditions for each task."),
    part = "header"
  ) |> 
  flextable::align(i = 2, align = "center", part = "header") |> 
  border_remove() |> 
  border(i = 2, border.top = fp_border_default(), part = "header") |>  
  border(i = 2, border.bottom = fp_border_default(), part = "header") |> 
  border(i = 5, border.bottom = fp_border_default(), part = "body") |> 
  autofit()

```

<br>

## Split-half reliability

For each cognitive task, we calculated split-half reliabilities using the `splithalf` package [@parsons_2021].
We calculated split-half reliabilities for mean response times and error rates, separately for each condition (see Table S1).
The split-half reliability of mean response times was high across all tasks and conditions.
For error rates, the reliability estimates were generally lower, which is likely due to ceiling effects (see also Figure S5).

<br>

```{r}
#| tab.id: tableS2
#| results: markup
sh_rel_rt_table |> 
  mutate(empty = "") |> 
  select(task, `sh_ci_congruent/repeat`, `sh_ci_incongruent/switch`, empty, `sb_ci_congruent/repeat`, `sb_ci_incongruent/switch`) |> 
  flextable() |> 
  flextable::set_header_labels(values = c("Task", "Congruent/Repeat", "Incongruent/Switch", "", "Congruent/Repeat", "Incongruent/Switch")) |> 
  add_header_row(values = c("", "Split-half reliability", "", "", "Spearman-Brown corrected", "")) |> 
  flextable::merge_at(i = 1, j = 2:3, part = "header") |> 
  flextable::merge_at(i = 1, j = 5:6, part = "header") |> 
  add_header_row(
    values = " ",
    colwidths = 6
  ) |> 
  flextable::compose(
    i = 1, j = 1,
    as_paragraph(as_b("Table S2. "), "Split-half reliabilities of mean response times for all cognitive tasks."),
    part = "header"
  ) |> 
  flextable::align(i = 2, align = "center", part = "header") |> 
  border_remove() |> 
  border(i = 2, border.top = fp_border_default(), part = "header") |>  
  border(i = 2, j = c(2:3, 5:6), border.bottom = fp_border_default(), part = "header") |> 
  border(i = 3, j = c(1:6), border.bottom = fp_border_default(), part = "header") |> 
  border(i = 6, border.bottom = fp_border_default(), part = "body") |> 
  autofit()
```

<br>

```{r}
#| tab.id: tableS3
#| results: markup
sh_rel_acc_table |> 
  mutate(empty = "") |> 
  select(task, `sh_ci_congruent/repeat`, `sh_ci_incongruent/switch`, empty, `sb_ci_congruent/repeat`, `sb_ci_incongruent/switch`) |> 
  flextable() |> 
  flextable::set_header_labels(values = c("Task", "Congruent/Repeat", "Incongruent/Switch", "", "Congruent/Repeat", "Incongruent/Switch")) |> 
  add_header_row(values = c("", "Split-half reliability", "", "", "Spearman-Brown corrected", "")) |> 
  flextable::merge_at(i = 1, j = 2:3, part = "header") |> 
  flextable::merge_at(i = 1, j = 5:6, part = "header") |> 
  add_header_row(
    values = " ",
    colwidths = 6
  ) |> 
  flextable::compose(
    i = 1, j = 1,
    as_paragraph(as_b("Table S3. "), "Split-half reliabilities of error rates for all cognitive tasks."),
    part = "header"
  ) |> 
  flextable::align(i = 2, align = "center", part = "header") |> 
  border_remove() |> 
  border(i = 2, border.top = fp_border_default(), part = "header") |>  
  border(i = 2, j = c(2:3, 5:6), border.bottom = fp_border_default(), part = "header") |> 
  border(i = 3, j = c(1:6), border.bottom = fp_border_default(), part = "header") |> 
  border(i = 6, border.bottom = fp_border_default(), part = "body") |> 
  autofit()
```

<br>

# Section 3. Drift Diffusion Modeling

## Model convergence

Figures S6-S11 present the convergence of MCMC chains of the Hierarchical Drift Diffusion Models for all tasks.
The figures should resemble a "fat, hairy caterpillar", which was the case for all tasks.

<br>

```{r}
#| label: FigureS6
#| fig-width: 6.5
#| fig-height: 5
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure S6.** Trace convergence across three chains for the Flanker Task.
flanker_fit_trace
```

<br>

```{r}
#| label: FigureS7
#| fig-width: 6.5
#| fig-height: 5
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure S7.** Trace convergence across three chains for the Simon Task.
simon_fit_trace
```

<br>

```{r}
#| label: FigureS8
#| fig-width: 6.5
#| fig-height: 5
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure S8.** Trace convergence across three chains for the Color-shape Task.
colorshape_fit_trace
```

<br>

```{r}
#| label: FigureS9
#| fig-width: 6.5
#| fig-height: 5
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure S9.** Trace convergence across three chains for the Global-local Task.
globallocal_fit_trace
```

<br>

```{r}
#| label: FigureS10
#| fig-width: 6.5
#| fig-height: 5
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure S10.** Trace convergence across three chains for the Animacy-size Task.
animacysize_fit_trace
```

<br>

```{r}
#| label: FigureS11
#| fig-width: 6.5
#| fig-height: 3
#| dpi: 600
#| out-width: 6in
#| fig-cap: | 
#|   **Figure S11.** Trace convergence across three chains for the Posner Task.
posner_fit_trace
```

<br>

In addition, we calculated the R^ statistic (also known as the Gelman-Rubin statistic) [@gelman_1992].
The R^ calculates the ratio between the variation between MCMC chains and the variation within MCMC chains.
A general guideline is that R^ values should be smaller than 1.1.
All R^ values are presented in Table S4 below.

## Simulation-based model fit assessment

Table S4 presents simulation-based model fit assessments for all tasks.
The simulation procedure was as follows.
First, we used the DDM parameter estimates for each participant and used them to simulate response times and accuracy data (100 trials per participant).
Then, we computed correlations between the simulated and observed response times and accuracies.
In the case of response times, we did so separately at the 25th, 50th, and 75th percentile.
In the case of accuracy, we looked at mean accuracy.
All correlations were > .89, indicating good model fit.

<br>

```{r}
#| tab.id: tableS4
#| results: markup
ddm_fit_table |> 
   add_header_row(
    values = " ",
    colwidths = 7
  ) |> 
  flextable::compose(
    i = 1, j = 1,
    as_paragraph(as_b("Table S4. "), "Model fit assessment"),
    part = "header"
  ) |> 
  border_remove() |> 
  border(i = 1, border.bottom = fp_border_default(), part = "header") |> 
  border(i = 2, border.bottom = fp_border_default(), part = "header") |> 
  border(i = 11, border.bottom = fp_border_default(), part = "body") |> 
  add_footer_row(
    values = " ",
    colwidths = 7
  ) |> 
  add_footer_row(
    values = " ",
    colwidths = 7
  ) |> 
  flextable::compose(
    i = 1, j = 1, 
    as_paragraph(as_i("Note: "), "Simulation-based model fit assessment compared observed and predicted data using 100 simulated trials (accuracy, 25th, 50th, 75th percentile). In addition, We calculated R^ values, which should be below 1.1 to indicate adequate chain convergence."), 
    part = "footer"
  )
```

<br>

# Section 4. Effects of environmental variables

Table S5 presents effects of environmental noise and mean-centered state anxiety on Drift Diffusion parameters.
Effects of environmental noise were mostly small non-significant.
Mean-centered differences in state anxiety were negatively associated with drift rates across all tasks.
In sessions where participants were more anxious than average, their drift rates across all tasks were lower.

<br>

::: {.landscape}

```{r}
#| tab.id: tableS5
#| results: markup
env_fit_table |> 
  add_header_row(
    values = " ",
    colwidths = 9
  ) |> # Add a new header row on top. We can use this new row to add the title
  flextable::compose(
    i = 1, j = 1,
    as_paragraph(as_b("Table S5. "), "Effects of noise and anxiety on Drift Diffusion parameters."),
    part = "header"
  ) |>
  border_remove() |> 
  border(i = 2, border.top = fp_border_default(), part = "header") |> 
  border(i = 3, j = c(2:3, 5:6, 8:9), border.top = fp_border_default(), part = "header") |> 
  border(i = 3, border.bottom = fp_border_default(), part = "header") |> 
  border(i = 33, border.bottom = fp_border_default(), part = "body") |> 
  align(i = 1, align = "left", part = "header") |> 
  add_footer_row(
    values = " ",
    colwidths = 9
  ) %>% 
  flextable::compose(
    i = 1, j = 1, 
    as_paragraph(as_i("Note: "), "State anxiety was mean-centered relative to the overall mean across test sessions. Bound. sep. = boundary separation, Con. = congruent condition, Inc. = incongruent condition, Non-dec. time = non-decision time, Rep. = repeat condition, Sw. = switch condition,"), 
    part = "footer"
  ) |> 
  autofit()
```

:::

<br>

Table S6 presents Spearman correlations between preregistered and non-preregistered adversity measures with environmental noise and mean state anxiety.
Note that S6 includes mean state anxiety, whereas the analyses (and Table S5) include the difference from the grand mean in specific testing sessions.

<br>

```{r}
#| tab.id: tableS6
#| results: markup
env_cor_table |> 
  flextable() |> 
 add_header_row(
    values = " ",
    colwidths = 7
  ) |> # Add a new header row on top. We can use this new row to add the title
  flextable::compose(
    i = 1, j = 1,
    as_paragraph(as_b("Table S6. "), "Bivarate correlations between adversity measures, environmental noise, and mean state anxiety."),
    part = "header"
  ) |>
  border_remove() |> 
  border(i = 2, border.top = fp_border_default(), part = "header") |> 
  border(i = 2, border.bottom = fp_border_default(), part = "header") |> 
  border(i = 6, border.bottom = fp_border_default(), part = "body") |> 
  autofit()
```

<br>

# Section 5. Indirect effects of confounders

Table S7 summarizes the indirect effects of confounders in the confirmatory models: age, education, sex, childhood adversity, and, in the case of threat as dependent variable, material deprivation.
As explained in the section on confounders in the main article, we assume that these confounders are common causes of both adversity exposure in adulthood (the independent variable) and cognitive processes (the dependent variable).
This means that the regression coefficients in Table S7 and S8 should *not* be interpreted as direct effects; rather, they should be interpreted as indirect effects (i.e., the effect at mean levels of adversity exposure).

For both threat and deprivation in adulthood, we found a negative indirect effect of age on task-general drift rate, a positive indirect effect of age on task-general boundary separation, and a positive indirect of age on task-general non-decision time.
Thus, older adults with average levels of adversity exposure in adulthood processed information more slowly across tasks, were generally more cautious, and were generally slower at encoding stimuli and/or executing responses.
In addition, childhood adversity had a negative indirect effect on task-general drift rate.
People with more exposure to childhood adversity, at average levels of adversity exposure in adulthood, processed information more slowly across tasks.
Finally, we found a negative indirect effect of education on task-general drift rate, but not on task-general boundary separation or non-decision time.
People with a higher completed education, at average levels of adversity exposure in adulthood, processed information faster across tasks.
None of the other indirect effects were significant.

<br>

```{r}
confound_coef_confirmatory |> 
  flextable() |> 
  set_header_labels(values = c(
    "Confounder", "\u{03B2}", "SE","95% CI", "p", "",
    "\u{03B2}", "SE","95% CI", "p"
    )) |> 
  add_header_row(values = c("", "Threat in adulthood", "", "", "", "", "Deprivation in adulthood", "", "", "")) |> 
  merge_at(i = 1, j = 2:5, part = "header") |> 
  merge_at(i = 1, j = 7:10, part = "header") |> 
  align(i = 1, j = 2:5, align = "center", part = "header") |> 
  align(i = 1, j = 7:10, align = "center", part = "header") |>

  bold(i = c(1,7,13), j = 1, part = "body") |> 
  padding(i = c(2:6, 8:12, 14:18), j=1, padding.left=15) |> 
  add_header_row(values = " ", colwidths = 10) |> 
  flextable::compose(
    i = 1, j = 1,
    as_paragraph(as_b("Table S7. "), "Standardized indirect effects of confounders in the confirmatory models."),
    part = "header"
  ) |>
  border_remove() |> 
  border(i = 2, border.top = fp_border_default(), part = "header") |>  
  border(i = 2, j = c(2:5, 7:10), border.bottom = fp_border_default(), part = "header") |> 
  border(i = 3, border.bottom = fp_border_default(), part = "header") |> 
  border(i = 18, border.bottom = fp_border_default(), part = "body") |> 
  autofit()
```

<br>

Table S8 summarizes the indirect effects of confounders in the exploratory models: sex, and, in the case of childhood threat as dependent variable, childhood material deprivation.
Childhood exposure to deprivation had a negative indirect effect on task-general drift rate and task-general boundary separation.
People with more exposure to childhood deprivation, at average levels of childhood threat, processed information more slowly across tasks, and were generally more cautious.
Sex did not have an indirect effect on either childhood threat or childhood deprivation.

<br>

```{r}
confound_coef_ch |> 
  flextable() |> 
  set_header_labels(values = c(
    "Confounder", "\u{03B2}", "SE","95% CI", "p", "",
    "\u{03B2}", "SE","95% CI", "p"
  )) |> 
  add_header_row(values = c("", "Childhood threat", "", "", "", "", "Childhood deprivation", "", "", "")) |> 
  merge_at(i = 1, j = 2:5, part = "header") |> 
  merge_at(i = 1, j = 7:10, part = "header") |> 
  align(i = 1, j = 2:5, align = "center", part = "header") |> 
  align(i = 1, j = 7:10, align = "center", part = "header") |> 
  bold(i = c(1,4, 7), j = 1, part = "body") |> 
  padding(i = c(2:3, 5:6, 8:9), j=1, padding.left=15) |> 
  add_header_row(values = " ", colwidths = 10) |> 
  flextable::compose(
    i = 1, j = 1,
    as_paragraph(as_b("Table S8. "), "Standardized indirect effects of confounders in the exploratory models of childhood adversity."),
    part = "header"
  ) |>
  border_remove() |> 
  border(i = 2, border.top = fp_border_default(), part = "header") |>  
  border(i = 2, j = c(2:5, 7:10), border.bottom = fp_border_default(), part = "header") |> 
  border(i = 3, border.bottom = fp_border_default(), part = "header") |> 
  border(i = 9, border.bottom = fp_border_default(), part = "body") |> 
  autofit()
```

\pagebreak

# References

::: {#refs}
:::

### 2024-12-18 supplement/stage_sup.R

@@ -1,0 +1,1054 @@
+library(tidyverse)
+library(furrr)
+library(ggsci)
+library(patchwork)
+library(splithalf)
+
+load("data/clean_data_long.RData")
+
+load('analysis_objects/ddm_flanker_mod1_parsed.RData')
+load('analysis_objects/ddm_simon_mod1_parsed.RData')
+load('analysis_objects/ddm_colorshape_mod1_parsed.RData')
+load('analysis_objects/ddm_animacysize_mod1_parsed.RData')
+load('analysis_objects/ddm_globallocal_mod1_parsed.RData')
+load('analysis_objects/ddm_posner_mod1_parsed.RData')
+
+load('analysis_objects/rhat_flanker_mod1.RData')
+load('analysis_objects/rhat_simon_mod1.RData')
+load('analysis_objects/rhat_colorshape_mod1.RData')
+load('analysis_objects/rhat_animacysize_mod1.RData')
+load('analysis_objects/rhat_globallocal_mod1.RData')
+load('analysis_objects/rhat_posner_mod1.RData')
+
+load('analysis_objects/results_confirmatory.RData')
+
+
+load("data/ddm_clean.RData")
+ddm_raw <-  read_csv("data/ddm_raw.csv")
+stai_noise <- haven::read_sav('data/L_CognitiveAdversity_1.0p.sav')
+iv_data <- read_csv('data/liss_data/full_data.csv')
+
+
+theme_set(
+  theme_classic() +
+  theme(
+  axis.text = element_text(size = 10),
+  axis.title = element_text(size = 10),
+  legend.title = element_blank(),
+  legend_text = element_text(size = 10),
+
+  )
+)
+
+
+# 1. IV histograms --------------------------------------------------------
+
+
+hist_data <- iv_data |>
+  select(live_off_income_m, finan_trouble_m, curr_situation_m, p_scar_m,
+         neigh_thr_m, neigh_thr02_m, vict_sum, threat_comp,
+         child_dep, child_thr, child_adv)
+
+## 1.1 Material deprivation ----
+
+live_off_income_m_hist <- hist_data |>
+  select(live_off_income_m) |>
+  ggplot(aes(live_off_income_m)) +
+  geom_histogram() +
+  theme_classic() +
+  scale_x_continuous(breaks = seq(0, 11, 1)) +
+  labs(
[…]

### 2024-12-18 README.Rmd
add all README files

---
title: ""
output: 
  html_document:
    template: assets/template.html
    css: assets/style.css
    keep_md: true
editor_options: 
  markdown: 
    wrap: sentence
---

```{r}
#| include: false
#| echo: false
#| 
library(tidyverse)
library(gert)
library(glue)
#source('scripts/dependencies.R')
knitr::opts_chunk$set(echo = F, message = F, warning = F)

```
<br>

This repository contains data, code, and output for a manuscript entitled "Adversity is associated with lower general processing speed rather than specific executive functioning abilities". This manuscript has been submitted for publication.

Do you want to download or clone the materials for this project? Go to [https://github.com/stefanvermeent/liss_ef_2024](https://github.com/stefanvermeent/liss_ef_2024)

## Directory Structure {#structure}

The names of each folder are intended to be self-explanatory.
There are eight top-level folders to organize the inputs and outputs of this project:

1.  [`preregistration`](https://stefanvermeent.github.io/liss_ef_2024/preregistration/README.html): The preregistration for the main analyses.
2.  [`manuscript`](https://stefanvermeent.github.io/liss_ef_2024/manuscript/README.html): The manuscript written in Quarto.
3.  [`supplement`](https://stefanvermeent.github.io/liss_ef_2024/supplement/README.html): a supplemental text (to be submitted with the manuscript) documenting all secondary analyses in detail.
4.  [`scripts`](https://stefanvermeent.github.io/liss_ef_2024/scripts/README.html): R-scripts that read, analyze, and produce all analysis objects.
5.  [`data`](https://stefanvermeent.github.io/liss_ef_2024/data/README.html): Folder in which real LISS data can be placed to make the analyses fully reproducible. Note that we cannot openly share the raw data on the open repository. See the link for more information on how to include the correct variables from the LISS data archive into this folder. 
6.  [`analysis Objects`](https://stefanvermeent.github.io/liss_ef_2024/analysis_objects/README.html): Folder containing all analysis objects.
7.  [`codebooks`](https://stefanvermeent.github.io/liss_ef_2024/codebooks/README.html): lists of variable names, labels, and value labels (where applicable).

Click on each of the folders to get more details.

## Steps to reproduce our results

There are two ways to reproduce our results.

1.  **Rerender manuscript using analysis objects**. This method uses the result objects that we saved under `analysis_objects/`. Analysis objects are the outcomes of the analysis scripts. Using this method, you can render the manuscript out-of-the-box, without having to rerun any of the analysis scripts under `scripts/`. 
2.  **Rerun all data cleaning and analysis steps**. This method a few steps. First, you need to download the required data from the LISS data archive (<https://lissdata.nl>) and place them in the `data` folder. Second, you need to rerun the analysis scripts under `scripts/liss/`, in the specified order. Please note that `03_ddm.R` involves running six Hierarchical Bayesian Drift Diffusion models. This takes a substantial amount of time, even in a research cloud environment (in our own case, it would take around 2 hours per model).

## Contact

For questions or comments, please feel free to contact me at p.c.s.vermeent@gmail.com.

### 2024-12-18 analysis_objects/README.Rmd
add all README files

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

# Analysis objects

This folder contains all [analysis objects](https://github.com/StefanVermeent/liss_ef_2024/tree/main/analysis_objects), i.e., the outputs of the various data processing scripts.
The analysis objects are loaded by the `manuscript.qmd` and `supplement.qmd` file to fill in the statistics, Tables, and Figures in the manuscript and supplemental materials

### 2024-12-18 materials/README.Rmd
add all README files

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

## Materials {#materials}

This folder contains all the materials of the cognitive tasks.
All tasks were programmed in [JsPsych 7.3.2](https://www.jspsych.org/7.3/).

1.  [`pilot01/`](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials/pilot01): Folder containing the pilot materials for the Flanker, Simon, Posner, Animacy-size, Color-shape, and Global-local tasks. Note that the Global-local task used in this pilot study was different from the one used in the manuscript.
2.  [`pilot02/`](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials/pilot02): Folder containing the pilot materials for the alternative Global-local task (as used in the manuscript).
3.  [`pilot03/`](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials/pilot03): Folder containing the pilot materials for all six tasks used in the manuscript.
4.  [`liss/`](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials/1_liss): Folder containing the final materials of all six tasks as used in the main study.

Within each study folder, each task also has it's own folder.
In each folder, you will find two files:
`0_[task]_setup.js` contains the Javascript code for functions and stimulus set-up.
`1_[task]_instructions.js` contains the Javascript code for all task instructions.
`2_[task]_procedure.js` contains Javascript code defining the stimuli and trials for both practice and test blocks.

In each of these study folders, you will find several additional files and folders:

1.  `index.html`. Script which combines the trial blocks and initiates the tasks.
2.  `img/`. Folder containing stimulus images.
4.  `scripts`. Scripts handling trial rules.

**Run tasks locally.** To run the tasks locally on your device, navigate to any task folder and open the `index.html` file in a browser. 
This will automatically start the cognitive task.

**Run tasks online.** The tasks are hosted online at [https://liss-ef-2024-pilot3.netlify.app/](https://liss-ef-2024-pilot3.netlify.app/).
Navigating to this url will automatically start the tasks.

### 2024-12-18 .projectlog/19cea5dd5cd24f7d6875416ff3ab67e30f0f3308.R

@@ -1,0 +1,13 @@
+### Date: 2024-08-27 11:50:04
+
+### Description: First time access to cognitive task data
+
+
+### For more information on this commit, see the README file, or go to https://github.com/stefanvermeent/liss_ef_2024/commit/19cea5dd5cd24f7d6875416ff3ab67e30f0f3308
+
+### Below is the full code that was used to access the data:
+
+
+readr::read_delim('data/L_CognitiveAdversity_TASKDATA.csv', col_select = NULL,  delim = ';') |>
+ dplyr::filter() |>
+ shuffle(data = _, shuffle_vars = 'NULL', long_format = FALSE, seed = 35356)

### 2024-12-18 scripts/README.Rmd
add script README

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

# Data Processing Scripts

There are four types of R-scripts in this repository, each with a separate folder.

-   [`0_functions/`](https://github.com/StefanVermeent/liss_ef_2024/tree/main/scripts/0_functions): Custom R-functions written for this project.
-   [`0_pilot/`](https://github.com/StefanVermeent/liss_ef_2024/tree/main/scripts/0_pilot): Scripts analyzing the Pilot data.
-   [`1_study/`](https://github.com/StefanVermeent/liss_ef_2024/tree/main/scripts/1_study): The full pipeline for cleaning data, running models, and conducting all analyses.

Each script takes an input(s) and produces output(s).
All outputs are stored in [analysis_objects/](https://github.com/StefanVermeent/liss_ef_2024/tree/main/analysis_objects).
These outputs include list with exclusions, DDM models, and SEM results.

The analysis objects, in turn, are used to stage all the statistics, Tables, and Figures included in the manuscript and supplemental material.
This is done in [`manuscript/scripts/staging.R`](https://github.com/StefanVermeent/liss_ef_2024/tree/main/manuscript/scripts/staging.R) and in [`supplement/stage_sup.R`](https://github.com/StefanVermeent/liss_ef_2024/tree/main/supplement/stage_sup.R).

### 2024-12-18 manuscript/scripts/staging.R

@@ -1,0 +1,472 @@
+library(flextable)
+library(tidyverse)
+library(haven)
+library(patchwork)
+
+source("scripts/0_functions/corr_plot.R")
+
+
+load('analysis_objects/results_confirmatory.RData')
+load('analysis_objects/results_exploratory.RData')
+load('data/exclusions.RData')
+
+full_data <- read_csv('data/liss_data/full_data.csv')
+descr <- read_sav(file = "data/L_CognitiveAdversity_1.0p.sav") |> as_tibble()
+
+
+# 1. Figures -----------------------------------------------------------------
+
+fig4_df <-
+  bind_rows(
+    fit_dep1_reg_coef |>
+    mutate(model = "Primary causal model") |>
+    left_join(fit_dep1_eqtests |> select(lhs, rhs, eq_pvalue)),
+    fit_dep2_reg_coef |>
+      mutate(model = "Secondary causal model") |>
+      left_join(fit_dep2_eqtests |> select(lhs, rhs, eq_pvalue))
+  ) |>
+  bind_rows(
+    fit_thr1_reg_coef |>
+      mutate(model = "Primary causal model") |>
+      left_join(fit_thr1_eqtests |> select(lhs, rhs, eq_pvalue)),
+    fit_thr2_reg_coef |>
+      mutate(model = "Secondary causal model") |>
+      left_join(fit_thr2_eqtests |> select(lhs, rhs, eq_pvalue))
+  ) |>
+  mutate(
+    ddm_parameter = case_when(
+      ddm_parameter == "Drift rate" ~ "Task-general\nprocessing speed",
+      ddm_parameter == "Boundary separation" ~ "Task-general\nresponse caution",
+      ddm_parameter == "Non-decision time" ~ "Task-general\nnon-decision time"
+    )
+  ) |>
+  filter(str_detect(lhs, "gen")) |>
+  ungroup() |>
+  add_row(
+    ddm_parameter = "Filler1",
+    rhs = "p_scar_m",
+    model = c("Primary causal model", "Secondary causal model")
+  ) |>
+  add_row(ddm_parameter = "Filler2",
+          rhs = "p_scar_m",
+          model = c("Primary causal model", "Secondary causal model")
+  ) |>
+  mutate(
[…]

### 2024-12-18 manuscript/masked_manuscript.qmd
anonymized links in masked manuscript

@@
 We have reported how we determined our sample size, all data exclusions, and all measures in the study.
-The preregistration, analysis code, and study materials, can be found on the article's GitHub repository (<https://github.com/stefanvermeent/liss_ef_2024>).
+The preregistration, analysis code, and study materials, can be found on the article's GitHub repository (<https://anonymous.4open.science/w/liss_ef_anon/>).
 The newly collected data will be available for other researchers in the LISS data archive, after signing a data use agreement (https://lissdata.nl).
-For more information on the LISS variables used, see the article's GitHub repository (<https://github.com/stefanvermeent/liss_ef_2024>).
+For more information on the LISS variables used, see the article's GitHub repository (<https://anonymous.4open.science/w/liss_ef_anon/>).
 Our study complies with Level 2 of the Transparency and Openness Promotion (TOP) guidelines.

### 2025-05-29 data/README.qmd
fix repo

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

## Data {#data}

This folder contains all the data that serves as input for the analysis scripts.
LISS data cannot be shared on open repositories.
If you would like to reproduce our results from scratch, you will have to download the necessary LISS data files from the data archive. Below is an overview of the required folders and datafiles:

1.  `data/liss_data/background`: All waves up until (and including) September 2023 (`avars_202309_EN_1.0p.zip`) documented at <https://www.dataarchive.lissdata.nl/study-units/view/322>
2.  `data/liss_data/income`: Wave 1-16 documented at <https://www.dataarchive.lissdata.nl/study-units/view/35>
3.  `data/liss_data/crime`: Wave 1-6 documented at <https://www.dataarchive.lissdata.nl/study-units/view/25>

At this time, the cognitive data collected in the current study are not yet available in the LISS archive. 
I expect them to be made available by LISS in spring 2025.
If you would like to access them before then, please contact me (p.c.s.vermeent@gmail.com) or LISS directly.

If you would like to render our manuscript without rerunning all analysis scripts, you can do so directly using the analysis objects in `analysis_objects/`. These are loaded automatically by the manuscript .qmd file.

### 2025-05-29 manuscript/manuscript.qmd
fix repo

@@
   padding.left = 1,
-  padding.right = 1
+  padding.right = 1,
+  opts_word = list(split = FALSE, keep_with_next = TRUE)
 )
@@
 
-#### **Adversity is associated with lower general processing speed rather than specific executive functioning abilities**
+#### **Adversity is associated with lower general processing speed rather than executive functioning**
 
@@
 
-#### Stefan Vermeent^1,2^, Anna-Lena Schubert^3^, & Willem E. Frankenhuis^2,4^
+#### Stefan Vermeent^1,2,3^, Anna-Lena Schubert^4^, & Willem E. Frankenhuis^1,3^
 
-#### ^1^ Department of Psychology, Utrecht University, Utrecht, The Netherlands
+#### ^1^ Evolutionary and Population Biology, Institute for Biodiversity and Ecosystem Dynamics, University of Amsterdam, Amsterdam, the Netherlands
 
-#### ^2^ Max Planck Institute for the Study of Crime, Security, and Law, Freiburg, Germany
+#### ^2^ Department of Psychology, Utrecht University, Utrecht, The Netherlands
 
-#### ^3^ Department of Psychology, University of Mainz, Mainz, Germany
+#### ^3^ Max Planck Institute for the Study of Crime, Security, and Law, Freiburg, Germany
 
-#### ^4^ Evolutionary and Population Biology, Institute for Biodiversity and Ecosystem Dynamics, University of Amsterdam, Amsterdam, the Netherlands
+#### ^4^ Department of Psychology, University of Mainz, Mainz, Germany
 
@@
 
-Stefan Vermeent served as lead for conceptualization, data curation, methodology, formal analysis, and writing the original draft.
-Anna-Lena Schubert served in a supporting role for conceptualization, formal analysis, methodology, and writing-review and editing.
-Willem E.
-Frankenhuis served in a supporting role for conceptualization, methodology, supervision, funding acquisition, and writing-review and editing.
+This study was approved by the Ethics Review Board of the Faculty of Social & Behavioural Sciences of Utrecht University (FETC20-490) and the Ethics committee for research in the Sciences and Life Sciences of the University of Amsterdam (FNWI-41_2023).
 
-The preregistration, analysis scripts, and study materials are available on the article's Github repository ([https://stefanvermeent.github.io/liss_ef_2024/](https://stefanvermeent.github.io/liss_ef_2024/)).
+All scripts and materials needed to reproduce the findings are available on the article's Github repository ([https://stefanvermeent.github.io/liss_ef_2024/](https://stefanvermeent.github.io/liss_ef_2024/)).
 In this paper, we make use of data from the LISS panel (Longitudinal Internet studies for the Social Sciences) managed by the non-profit research institute Centerdata (Tilburg University, the Netherlands).
-All datasets are available to researchers in the LISS data archive, after signing a data use agreement ([https://www.dataarchive.lissdata.nl/study-units/view/1](https://www.dataarchive.lissdata.nl/study-units/view/1)).
+All datasets are available in the LISS data archive ([https://www.dataarchive.lissdata.nl/study-units/view/1](https://www.dataarchive.lissdata.nl/study-units/view/1)). 
+Researchers who want to access the data are required to sign a statement confirming that information about individual persons, households, etc., will not be released to others (go to [https://statements.centerdata.nl](https://statements.centerdata.nl) for more information).
+The data and ideas in this manuscript were not presented elsewhere.
 
-Correspondence concerning this article should be addressed to Stefan Vermeent, Department of Psychology, Utrecht University, Heidelberglaan 1, 3584 CS Utrecht, The Netherlands.
-Email: p.c.s.vermeent@gmail.com / p.c.s.vermeent@uu.nl.
+SV served as lead for conceptualization, formal analysis, funding acquisition, investigation, methodology, and writing - original draft.
+AS served in a supporting role for supervision, conceptualization, methodology, formal analysis, and writing - review & editing.
+WF served as lead for supervision, and in a supporting role for conceptualization, methodology, formal analysis, and writing - review & editing.
 
-We have no conflict of interest to disclose.
+The authors declare no conflicts of interest.
 
-<br>
+Correspondence concerning this article should be addressed to Stefan Vermeent, Evolutionary and Population Biology, Institute for Biodiversity and Ecosystem Dynamics, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands.
+Email: p.c.s.vermeent@gmail.com.
 
-**word count:** 9573
-
-{{< pagebreak >}}
+\pagebreak
 
@@
 Participants completed a basic processing speed task, two inhibition tasks, and three attention-shifting tasks.
-We measured exposure to threat and material deprivation in childhood and adulthood.
-Exposure to threat (but not material deprivation) in adulthood was negatively associated with task-general processing speed.
+We measured self-reported exposure to threat and material deprivation in childhood and adulthood.
+Exposure to threat (but not deprivation) in adulthood was negatively associated with task-general processing speed.
 After accounting for task-general processes, remaining variance was not related to either inhibition or attention-shifting ability.
-Non-preregistered analyses showed that childhood exposure to material deprivation and threat were negatively associated with (1) general processing speed, and (2) task-specific information accumulation.
+Non-preregistered analyses showed that childhood exposure to deprivation and threat were negatively associated with (1) general processing speed, and (2) task-specific information accumulation.
 The latter reflected unique features of individual tasks, rather than specific EF abilities.
-Taken together, these results suggest that adversity researchers overestimate associations between adversity and specific EF abilities when analyzing raw performance.
+Taken together, these results suggest that adversity researchers may overestimate associations between adversity and specific EF abilities when analyzing raw performance.
 
@@
 These conclusions are often derived from analyzing raw performance, such as response times or accuracy.
-We use cognitive modeling to reveal that adversity lowers general processing speed, but not specific executive functioning abilities.
-This suggests that adversity research likely overestimates the extent to which adversity is associated with executive functioning abilities.
+We use cognitive modeling to reveal that self-reported adversity is associated with general processing speed, but not specific executive functioning abilities.
+This suggests that adversity research may overestimate the extent to which adversity is associated with executive functioning abilities.
 
@@
 
+Most research in this field has focused on cognitive abilities in childhood, or associations between early-life adversity and cognitive abilities up to early adulthood. 
+However, adaptations to early adversity may persist into adulthood, and people may also be able to develop cognitive adaptations in response to adversity exposure later in life. 
+First, there is evidence that early-life adversity is associated with both cognitive deficits and adaptations in middle and late adulthood. 
+Women who reported a history of sexual abuse before the age of 18 showed better delayed recall and better visual reasoning at age 50-64 [@feeney_2013]
+Similarly, self-reported material deprivation in early adulthood was associated with better delayed recall and verbal fluency in middle adulthood [@kunzi_2023]. 
+In contrast, higher levels of self-reported childhood abuse have also been linked to lower cognitive performance in middle- to late adulthood (@roberts_2022). 
+Second, people may also develop cognitive adaptations in response to adversity experienced later in life. 
+People who reported higher levels of material deprivation in middle adulthood showed a smaller decline in delayed recall compared to people with lower levels of material deprivation [@kunzi_2023]. 
+These findings highlight the importance of studying the effects of adversity across the lifespan.
+
+# Methodological challenges for measuring specific EF abilities
+
 Research within both frameworks typically employs the same inferential strategy, by estimating specific EF abilities based on raw performance on EF tasks (e.g., response speed or accuracy).
@@
 
-# Methodological challenges for measuring specific EF abilities
-
 Recent research shows that raw performance on EF tasks does not necessarily capture specific EF abilities, for two related reasons.
-First, raw performance on EF tasks is also influenced by processes other than the ability of interest, such as response caution and the speed of response execution [@hedge_2021; @ratcliff_2008; @stahl_2014].
+First, raw performance on EF tasks is also influenced by processes other than the ability of interest, such as response caution and the speed of response execution [@hedge_2022; @ratcliff_2008; @stahl_2014].
 Thus, slower (or faster) responses in people with more adversity exposure need not reflect EF ability, but could, for instance, reflect a tendency to respond more cautiously [@vermeent_2024a].
@@
 The DDM assumes that people accumulate evidence for one of two responses (e.g., left or right button) until they have acquired sufficient evidence to make a decision (see Figure 1).
-When applied to trial-level response times and accuracy data, the DDM estimates four parameters that represent distinct cognitive processes [@voss_2004].
+When applied to trial-level response times and accuracy data, the DDM estimates three parameters that represent distinct cognitive processes [@voss_2004].
 The *drift rate* reflects the rate at which people accumulate evidence for the correct response; thus, it measures the efficiency of information processing.
@@
 
-```{r}
-#| label: Figure1
-#| fig-width: 6.5
-#| fig-height: 7
-#| dpi: 600
-#| out-width: 6in
-#| fig-cap: | 
-#|   **Figure 1.** Visual representation of the Drift Diffusion Model (DDM).
-The DDM assumes that people accumulate evidence for one of two responses until they have acquired sufficient evidence to make a decision.
-Each squiggly line represents the evidence accumulation process on a single trial.
-The thick, horizontal lines represent the decision boundaries corresponding to the correct response (upper boundary) and the incorrect response (lower boundary).
-Applying the DDM to trial-level performance data (response times and accuracies) yields four parameters representing distinct cognitive processes: (1) **drift rate**: Average rate of evidence accumulation across trials; measuring the efficiency of information processing.
-(2) **boundary separation**: The distance between the two decision boundaries; measuring response caution.
-(3) **Non-decision time**: A combination of the preparation speed (e.g., stimulus encoding) and response execution speed (e.g., time spent pressing the button); measuring speed of non-decision processes.
-**starting point**: The starting point of the evidence accumulation process; measuring response bias (not considered here).
-Figure adopted from Vermeent et al.
-(2024), licensed under CC BY 4.0.
-knitr::include_graphics("figures/fig1.png")
-```
+How do the DDM parameters relate to executive functioning ability?
+Not all EF tasks require a person to accumulate evidence under time pressure, and some research suggests that EF measurement may be improved by removing speed requirements altogether [e.g., @draheim_2021b].
+However, many common EF tasks do depend on evidence accumulation.
+For instance, the Flanker Task requires accumulating evidence for the arrow direction (while not letting the flanking arrows contaminate this evidence), and attention-shifting tasks require accumulating evidence for the goal that is currently relevant.
+Given the popularity of these tasks in adversity research, associations between adversity exposure and EF have largely been studied using response times. 
+This raises the possibility that these associations are caused by ability-irrelevant processes rather than EF abilities. 
+If performance differences really reflect EF abilities, they should be reflected in drift rates, not boundary separations or non-decision times.
 
@@
 
-Recent studies show that a single general factor accounts for most of the variance in drift rates across multiple tasks [@lerche_2020; @loffler_2024; @weigard_2021].
-This general factor appears to be largely stable across time [@schubert_2016; @weigard_2021].
-It arises also in analyses of tasks thought to measure different abilities [@schmiedek_2007; @schmitz_2016].
-Even if all tasks are designed to measure the same EF ability [@hedge_2021], variance in the general factor is fully explained by general processing speed instead of any specific EF ability [@loffler_2024].
-These findings suggest that performance on standard EF tasks largely reflects basic speed of processing, rather than EF abilities.
+Recent studies show that a single general factor accounts for most of the variance in drift rates across multiple EF tasks [@lerche_2020; @loffler_2024; @weigard_2021].
+This general factor appears to be largely stable across test sessions [@schubert_2016; @weigard_2021], and has been interpreted as reflecting general processing speed [@hedge_2022; @loffler_2024].
+Although general processing speed exerts a strong influence on EF task performance, it is conceptually distinct from EF abilities. 
+Processing speed refers to the rate at which elementary cognitive operations are completed, whereas executive functions involve higher-order cognitive processes that support goal-directed behavior, particularly under conditions of distraction or interference [@mashburn_2024; @vonbastian_2020].
+Thus, to isolate specific EF abilities from raw task performance, it is essential to account for the influence of general processing speed.
+This approach has been used in mathematical and cognitive psychology, but hardly in research on the impact of adversity on EF abilities.
 
-Recent findings suggest that the effects of adversity are more general than specific.
-For instance, youth with lower SES and more exposure to adversity show lower task-general raw performance scores across measures of EF, memory, and intelligence, while task-specific performance (after accounting for general variance) is often practically equivalent to zero [@bignardi_2024; @young_2024].
-In addition, youth with more exposure to household threat showed lower task-general processing speed (based on drift rates), while differences in task-specific drift rates were practically equivalent to zero [@vermeent_2024a].
-These findings suggest that previous research may have overestimated the associations between adversity and specific abilities.
+We are aware of three studies that estimated the associations between adversity and specific cognitive abilities after partialing out general processing.
+These studies suggest that the effects of adversity are more general than specific.
+Two studies showed that youth with lower socioeconomic status and more exposure to adversity show lower task-general raw performance scores across measures of EF, memory, and intelligence, while task-specific performance (after accounting for general variance) is often practically equivalent to zero [@bignardi_2024; @young_2024].
+A third study showed that youth with more exposure to household threat showed lower task-general processing speed (based on drift rates), while differences in task-specific drift rates were practically equivalent to zero [@vermeent_2024a].
+Jointly, these findings suggest that previous research may have overestimated the associations between adversity and specific abilities.
 
-A consequence of both sources of measurement impurity is that estimating specific EF abilities requires latent variable models including two or more measures of the same ability; yet, most studies in adversity research include only one task per EF ability.
+A consequence of measurement impurity is that estimating specific EF abilities requires latent variable models including two or more measures of the same ability.
+Yet, most studies in adversity research include only one task per EF ability.
 This makes this work less suitable to investigate whether adversity is associated with specific EF abilities after accounting for general processes.
 This is important for both deficit-based and adaptation-based interventions, which aim to remediate or leverage impaired or enhanced abilities, respectively. 
-Moreover, while deficit frameworks predict impairments in both specific abilities and general processes [@vermeent_2024a], adaptation frameworks predict that specific types of adversity lead to adaptations in specific EF abilities.
+Moreover, while deficit frameworks predict impairments in both specific abilities and general processes [@vermeent_2024a], adaptation frameworks predict that specific types of adversity lead to adaptations in specific EF abilities only.
 For example, it has been proposed that unpredictable and dangerous environments may positively affect the ability to rapidly shift attention, promoting the detection of sudden threats and seizing fleeting opportunities [@frankenhuis_2020; @mittal_2015].
@@
 Using a combination of DDM and structural equation modeling, we address three central questions.
-First, what is the association of adversity exposure in adulthood with general processing speed that is contributing to performance across all cognitive tasks?
-Second, what is the association of adversity exposure in adulthood with inhibition and attention-shifting abilities after accounting for general processing speed?
-Third, what is the association of adversity exposure in adulthood with general and/or EF-specific response caution?
-We focus on two types of adversity: material deprivation (a lack of access to material resources) and threat (the potential for harm imposed by others).
+First, what is the association of self-reported adversity exposure in adulthood with general processing speed that is contributing to performance across all cognitive tasks?
+Second, what is the association of self-reported adversity exposure in adulthood with inhibition and attention-shifting abilities after accounting for general processing speed?
+Third, what is the association of self-reported adversity exposure in adulthood with general and/or EF-specific response caution?
+We focus on two types of adversity: material deprivation (a lack of access to material resources) and environmental threat (the potential for harm imposed by others).
 In previous research, both types of adversity were associated with performance on tasks used to measure EF [e.g., @fields_2021; @schafer_2022; @sheridan_2022; @vermeent_2024a; @young_2022].
 In addition, both are central to recent dimensional models of adversity [@mclaughlin_2021; @sheridan_2014].
+In contrast to specificity models (which distinguish between separate distinct adversities, such as verbal or physical abuse) and cumulative-risk models (which add-up discrete exposures to various adverse experiences), dimensional models of adversity aim to identify dimensions of adversity, consisting of different types of adversity with similar features [@mclaughlin_2021].
 
 We preregistered predictions of deficit and adaptation frameworks.
-Deficit frameworks predict negative associations between adversity and both the general speed of processing as well as specific abilities.
-Adaptation frameworks predict intact or enhanced attention shifting ability, but not intact or enhanced inhibition ability.
-We only interpret intact attention shifting ability as an adaptation if inhibition ability is also lower.
+From a deficit framework, we predict negative associations between adversity and both the general speed of processing as well as specific abilities. 
+From an adaptation framework, we have three predictions.
+First, we predict a positive association between threat exposure and attention-shifting ability.
+This follows from the hypothesis that violence exposure might lead to broader, more vigilant monitoring of information, which could help to detect threats early on [@frankenhuis_2016; @vermeent_2025].
+We only interpret intact attention shifting ability in association with threat as an adaptation if inhibition ability is also lower. 
 Intact performance on both abilities, however, would not provide sufficient evidence of adaptation, as it might instead suggest no association between adversity exposure and these abilities. 
-In line with previous work [@vermeent_2024a], we further predict that threat exposure is associated with higher task-general response caution.
+Second, we expect that broader, more vigilant monitoring of information would hurt performance on inhibition tasks, as it would result in increased processing of distractions.
+Therefore, we do not predict a positive association between threat and inhibition ability.
+Third, although one could argue that material deprivation might also increase monitoring (of potential resources), no previous work that we are aware of has found a positive association between deprivation and attention shifting [@mittal_2015; @young_2022].
+Thus, we do not predict a positive association between material deprivation and attention-shifting ability (nor inhibition ability).
+Finally, in line with previous work [@vermeent_2024a], we further predict that threat exposure is associated with higher task-general response caution.
 
@@
 In addition, LISS participants have the option to participate in further monthly studies on different topics.
+Each month, participants have the opportunity to update their demographic information (e.g., sex, education level, household composition).
 We estimated power based on @kretzschmar_2019.
@@
 Participants were able to complete the cognitive tasks during two or more sessions to increase participation rates.
-People were eligible for the study if they were between 18 and 55 years old, and had agreed to linking their LISS data to government microdata (not relevant here, but for a different study).
+People were eligible for the study if they were between 16 and 55 years old, and had agreed to linking their LISS data to government microdata (not relevant here, but for a different study).
 We excluded participants who did not have data on any cognitive task.
 The final sample after exclusions consisted of `r exclusions$sample$final` participants (see Table 1).
-
-<br>
-
-```{r}
-#| tab.id: table1
-#| results: markup
-table1 |> 
-  flextable::autofit() |> 
-  add_header_row(
-    values = " ",
-    colwidths = 2
-  ) |> 
-  flextable::compose(
-    i = 1, j = 1, 
-    as_paragraph(as_b("Table 1. "), "Descriptive statistics for the final sample."),
-    part = "header"
-  ) |> 
-  border_remove() |> 
-  border(i = 1, border.bottom = fp_border_default(), part = "header") |>  
-  border(i = 2, border.top = fp_border_default(), part = "header") |>  
-  border(i = 2, border.bottom = fp_border_default(), part = "header") |> 
-  border(i = 14, border.bottom = fp_border_default(), part = "body") |> 
-  autofit()
-```
-
-<br>
 
@@
 
-Following our preregistration, we measured exposure to neighborhood violence in adulthood using two measures of perceived neighborhood crime, and one measure of crime victimization. 
+Following our preregistration, we measured exposure to neighborhood violence in adulthood using three measures: two measures of perceived neighborhood crime, and one measure of crime victimization. 
 Participants completed the Neighborhood Violence Scale [@frankenhuis_2018; @frankenhuis_2020].
@@
 We decided to conduct all analyses involving threat based only on this subset of the data (N = `r exclusions$sample$final - exclusions$sample$missing_crime_vict`), because *a priori* power remained above 80% [@kretzschmar_2019].
-We summed the items within each wave, and then calculated an average across waves for which participants had data.
+We summed the items within each wave, and then calculated an average across all the waves for which we had a participant's data.
 
@@
 Therefore, following our preregistration, we used Principal Component Analysis to extract the first principal component score.
-This score accounted for 23 % of the variance in the three measures.
+This score accounted for 23% of the variance in the three measures.
 
@@
 
-::: {.landscape}
+### Childhood adversity
 
-```{r}
-#| tab.id: table2
-#| results: markup
-table2 |> 
-  flextable() |> 
-  flextable::set_header_labels(Variable = "") |> 
-  flextable::autofit() |> 
-  border_remove() %>% 
- # border(i = 1, border.top = fp_border_default(), part = "header") %>% 
-  border(i = 1, border.bottom = fp_border_default(), part = "header") %>% 
-  border(i = 18, border.bottom = fp_border_default(), part = "body") %>% 
-  add_header_row(
-    values = " ",
-    colwidths = 13
-  ) |> 
-  flextable::compose(
-    i = 1, j = 1, 
-    as_paragraph(as_b("Table 2. "), "Spearman correlations between the main independent variables and covariates."),
-    part = "header"
-  ) |> 
-  add_footer_row(
-    values = " ",
-    colwidths = 13
-  ) |> 
-  autofit()
-```
+Participants reported their exposure to childhood deprivation (prior to age 18) in the new data collection using four items: (1) "My family had enough money to afford the kind of home we all needed"; (2) "My family had enough money to afford the kind of clothing we all needed"; (3) "My family had enough money to afford the kind of food that we all needed"; (4) "My family had enough money to afford the kind of medical care that we all needed", on a scale from 1 (completely not agree) to 7 (completely agree).
+This measure differed from those used to measure exposure to deprivation in adulthood. 
+In addition, participants reported their exposure to childhood threat (prior to age 18) in the new data collection using the Neighborhood Violence Scale ("Crime was common in the neighborhood where I grew up") , on a scale from 1 (completely not agree) to 7 (completely agree) [@frankenhuis_2018; @frankenhuis_2020].
+This measure was also used to measure threat exposure in adulthood.
 
-:::
+When used as covariates in the main analyses (see Confounds section), the measures of childhood deprivation and childhood threat were averaged together into a single measure of childhood adversity. 
+When used as main predictors in the exploratory analyses that we conducted, they were treated separately.
 
@@
 If the stimulus was flanked by the image of a mouse (50% of trials), participants had to indicate whether the local images were squares or rectangles.
-On 50 % of the trials, the rule was the same as on the preceding trial (repeat trials).
+On 50% of the trials, the rule was the same as on the preceding trial (repeat trials).
 On the other half, the rule was different than on the preceding trial (switch trials).
@@
 We measured state anxiety using the shortened version of the State-Trait Anxiety Inventory [@bij_2003; @marteau_1992], which asks participants how calm, tense, upset, relaxed, content, and worried they currently feel, on a scale of 1 ("not al all") to 4 ("very much").
-We recoded (if necessary) and then summed the answers with higher scores reflecting more state anxiety.
+We recoded (if necessary) and then summed the answers, with higher scores reflecting more state anxiety.
 
@@
 Participants reported the level of environmental noise after each cognitive task; thus, six times in total.
-We measured environmental noise using a single item, rated on a scale of 1 to 5: "How much noise was there in your environment during the game?"
+We measured environmental noise using a single item, rated on a scale of 1 ("very little noise") to 5 ("a lot of noise"): "How much noise was there in your environment during the game?".
 
@@
 The set of potential confounders consisted of (1) age [@salthouse_2016; @salthouse_2019; @starns_2010]; (2) education [@hofmarcher_2021]; (3) sex [@ning_2023]; (4) childhood adversity exposure (material deprivation and threat combined) [@bos_2009; @goodman_2019]; and (5) potential causal relations between recent material deprivation and recent threat [@bywaters_2016; @lacey_2022; @ning_2023].
-We made these decisions based on previous literature, or, in cases of doubt, by statistically testing support for specific relations using data from a previous LISS study [@vermeent_2024c].
+We made these decisions based on previous literature, or, in cases of doubt, by statistically testing support for specific relations using data from a previous LISS study [@vermeent_2025].
 Specifically, following our preregistration, we used the *dagitty* R package [@ankan_2021; @textor_2016] to test whether independencies between specific variables implied by the Directed Acyclic Graph (conditional on the other causal pathways) were supported by these data (see the preregistration for more details).
 
-Figure 1 depicts the final preregistered Directed Acyclic Graph for material deprivation and threat exposure in adulthood.
+Figure 2 depicts the final preregistered Directed Acyclic Graph for material deprivation and threat exposure in adulthood.
 For threat exposure in adulthood, our statistical model controlled for age, sex, childhood adversity, and recent material deprivation.
@@
 We did not control for threat exposure in adulthood, as, following the Directed Acyclic Graph, we assume that threat exposure mediates the effect of recent material deprivation on cognitive processes.
-If true, including threat exposure in the model would make the association between material deprivation and cognitive processes an indirect effect, rather than a total effect (which was our estimand).
+If such mediation does occur, including threat exposure in the model would make the association between material deprivation and cognitive processes an indirect effect, rather than a total effect (which was our estimand).
 To explore the effect of our assumption that threat mediates the effect of material deprivation on cognitive processes, we also present a secondary model that includes threat in adulthood as a confound.
 As preregistered, we will not base our main conclusions on this secondary model.
-
-```{r}
-#| label: Figure2
-#| fig-width: 6.5
-#| fig-height: 7
-#| dpi: 600
-#| out-width: 6in
-#| fig-cap: | 
-#|   **Figure 2.** Direct Acyclic Graphs (DAG) depicting our causal assumptions about the main estimands.
-the rectangles (nodes) depict variables, and directed arrows depict assumed causal pathways.
-An arrow between two variables represents the assumption that experimentally manipulating the variable at the origin of the arrow will change the variable at the end of the arrow (but not the other way around).
-The statistical models control for confounding variables (depicted as dashed paths), which have an arrow both to the independent variable and the dependent variable.
-knitr::include_graphics("figures/fig2.png")
-```
-<br>
 
@@
 Then, participants had the option to either continue to the next task or complete the other tasks at a later point in time. 
-`r round(n_sessions)` % of participants completed all tasks in a single session, and `r 100-round(n_sessions)` % of participants took two or more sessions.
+`r n_sessions |> filter(nomem_encr %in% exclusions$sample$ids) |> summarise(x = sum(n_sessions == 1)/n()*100) |> round(1) |> pull(x)`% of participants completed all tasks in a single session, and `r n_sessions |> filter(nomem_encr %in% exclusions$sample$ids) |> summarise(x = sum(n_sessions != 1)/n()*100) |> round(1) |> pull(x)`% of participants took two or more sessions.
+Participants who took two or more sessions did not process information differently in the first session compared to later sessions (based on average standardized drift rates), *t*(`r session_effect$df_error`) = `r session_effect$t`, *p* `r format_p(session_effect$p)`.
+
 Jointly, the cognitive tasks took around 25 minutes to complete.
@@
 A second ability-specific inhibition factor loaded on the parameter estimates of incongruent trials of the Flanker and Simon tasks.
-A third specific attention shifting factor loaded on the parameter estimates of switch trials of the Color-shape, Global-local, and Animacy-size tasks.
+A third specific attention shifting factor loaded on the parameter estimates of switch trials of the Color-shape, Global-local, and Animacy-size Tasks.
 These two factors were allowed to covary.
@@
 For the drift rate model, we interpreted the general factor as basic speed of processing, and the ability-specific factors as reflecting  inhibition and attention-shifting ability (or common EF in the case of the second model).
-For the boundary separation and non-decision time models, we interpreted the general factor as general response caution/speed of non-decision processes, and the ability-specific factors as reflecting response caution/speed of non-decision processes specific to conflict trials.
+For the boundary separation and non-decision time models, we interpreted the general factor as general response caution/speed of non-decision processes, and the ability-specific factors as reflecting response caution/speed of non-decision processes specific to incongruent/switch trials.
 
@@
 
-Finally, we constructed two versions of the final model, one to estimate the association between material deprivation in adulthood and the outcome measures, and one to estimate the association between threat exposure in adulthood and the outcome measures.
-In both models, we regressed the adversity measure on each latent factor, together with the control variables (see the section on confounds for more details).
+Finally, we constructed two versions of the final model, one to estimate the association between material deprivation in adulthood and the outcome measures, and the other to estimate the association between threat exposure in adulthood and the outcome measures.
+In both models, we regressed the adversity measure on each latent factor, together with the control variables (see the section on confounds for details).
 We report indirect effects of control variables in the supplemental materials (section 5).
@@
 
-As preregistered, we started with optimizing fit in structural equation models for each DDM parameter separately before combining them all into a single model.
+As preregistered, we first optimized fit in structural equation models for each DDM parameter separately before combining them all into a single model.
 For drift rates, we selected the model containing a general processing speed factor (loading on all drift rates) and a common EF factor (loading only on drift rates of incongruent and switch conditions).
@@
 
+## Preregistered analyses
+
+Our primary analyses examined the association of exposure to material deprivation and threat in adulthood with: task-general processing speed, task-general response caution, and task-general speed of stimulus encoding and/or response execution.
+As the structural equation model analyses reported above showed that EF-specific factors had non-significant variances, we retained them in the model but did not include them as dependent variables in subsequent analyses.
+
+### Material deprivation in adulthood
+
+Exposure to material deprivation in adulthood was negatively associated with task-general processing speed ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_v', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_v', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`) (see Figure 4).
+People who reported more exposure to deprivation in adulthood processed information more slowly across tasks.
+Exposure to deprivation in adulthood was not associated with task-general response caution ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_a', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_a', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`), or task-general non-decision time ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_t', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_t', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`) (see Figure 4).
+In addition, none of the effects fell within the region of practical equivalence (all *p*s ≥ `r fit_dep1_eqtests |> pull(eq_pvalue) |> min() |> format_p(pvalue = _, operator = FALSE)`).
+The effects in the secondary model (including exposure to threat in adulthood as a confounder) were comparable in size, although the association between material deprivation in adulthood and task-general processing speed did not reach statistical significance.
+
+### Threat in adulthood
+
+Exposure to threat in adulthood was negatively associated with task-general processing speed ($\beta$ = `r fit_thr1_reg_coef |> filter(lhs == 'gen_v', rhs == 'threat_comp') |> pull(est.std) |> formatC(digits = 2, width = 2, format = 'f')`, *p* `r fit_thr1_reg_coef |> filter(lhs == 'gen_v', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`) (see Figure 4).
+People who reported more exposure to threat in adulthood processed information more slowly across tasks.
+Exposure to threat in adulthood was not associated with task-general response caution ($\beta$ = `r fit_thr1_reg_coef |> filter(lhs == 'gen_a', rhs == 'threat_comp') |> pull(est.std) |> round(2)`, *p* `r fit_thr1_reg_coef |> filter(lhs == 'gen_a', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`), or task-general non-decision time ($\beta$ = `r fit_thr1_reg_coef |> filter(lhs == 'gen_t', rhs == 'threat_comp') |> pull(est.std) |> round(2)`, *p* `r fit_thr1_reg_coef |> filter(lhs == 'gen_t', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`).
+The effects in the secondary model (excluding material deprivation in adulthood as a confounder) were comparable, although the association with task-general non-decision time was practically equivalent in the secondary model.
+
+## Non-preregistered analyses
+
+### Associations with childhood adversity
+
+In the first set of non-preregistered analyses, we analyzed how childhood exposure to threat and material deprivation are associated with task-general and ability-specific cognitive processes.
+The rationale for these analyses is that we observed significant indirect associations between childhood adversity and general speed of processing in our preregistered analyses.
+Here, we analyzed *direct* effects of childhood threat and material deprivation (i.e., without including exposure to adversity in adulthood).
+We controlled for sex (but not age and education) and in the case of threat we also controlled for material deprivation, but not the other way around (similarly to the preregistered analyses).
+
+Childhood exposure to material deprivation was negatively associated with general processing speed, $\beta$ = `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(est.std) |> round(2)`, SE = `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(se) |> round(2)`, 95% CI = [`r expl_ch_dep_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(ci.lower) |> round(2)`, `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(ci.upper) |> round(2)`], *p* `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(pvalue_adj) |> format_p(pvalue = _)`.
+People who reported higher levels of childhood exposure to material deprivation processed information more slowly across tasks.
+In addition, childhood exposure to material deprivation was positively associated with non-decision time, $\beta$ = `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(est.std) |> round(2)`, SE = `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(se) |> round(2)`, 95% CI = [`r expl_ch_dep_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(ci.lower) |> round(2)`, `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(ci.upper) |> round(2)`], *p* `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(pvalue_adj) |> format_p(pvalue = _)`.
+People who reported higher levels of childhood exposure to material deprivation took more time to encode stimulus information and/or execute responses across tasks.
+Childhood exposure to material deprivation was not associated with general response caution.
+Equivalence tests revealed that none of these associations fell within the region of practical equivalence.
+
+For childhood exposure to threat, we found a significant negative association with general processing speed, $\beta$ = `r expl_ch_thr_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(est.std) |> round(2)`, SE = `r expl_ch_thr_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(se) |> round(2)`, 95% CI = [`r expl_ch_thr_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(ci.lower) |> round(2)`, `r expl_ch_thr_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(ci.upper) |> round(2)`], *p* `r expl_ch_thr_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(pvalue_adj) |> format_p(pvalue = _)`.
+People who reported higher levels of childhood exposure to threat processed information more slowly across tasks.
+We did not find significant associations between childhood exposure to threat and general response caution or general non-decision time.
+Of these associations, equivalence tests revealed that the association with general response caution fell within the region of practical equivalence (*p* `r expl_ch_thr_eqtests |> filter(lhs == "gen_a", rhs == "child_thr") |> pull(eq_pvalue) |> format_p(pvalue = _)`).
+
+### Associations with task-specific drift rates
+
+In the second set of non-preregistered analyses, we analyzed how adversity exposure during childhood and adulthood were associated with task-specific residual variances.
+The rationale for these analyses was that while we did not find coherent latent EF factors, all tasks showed significant residual variance after accounting for general speed of processing.
+This implies that, after accounting for general speed of processing, performance is further determined by unique features of individual tasks (i.e., method variance), which may be associated with adversity exposure.
+For this analysis, we were mostly interested in investigating the extent to which task-specific effects reflect EF abilities or processes that are unique to individual tasks, and therefore only focused on drift rates.
+
+We first attempted to fit a model in which all task-specific drift rates were allowed to covary (i.e., across inhibition and attention-shifting tasks).
+However, this model did not converge.
+A model that only estimated covariances between task-specific drift rates of inhibition tasks and attention shifting tasks separately provided a good fit to the data (CFI = `r fit_expl_ts_mod2_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_expl_ts_mod2_fitstats[['rmsea.robust']] |> round(2)` [`r fit_expl_ts_mod2_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_expl_ts_mod2_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]).
+This suggests that task-specific drift rates correlated among inhibition tasks and among attention-shifting tasks, but not between inhibition and attention-shifting tasks.
+
+Figure 6 shows latent correlations between task-specific drift rates after accounting for task-general processing speed.
+The drift rates of repeat and switch conditions within each attention-shifting task correlated moderately to strongly, while all correlations between tasks were low.
+This suggests that task-specific drift rates of attention-shifting tasks captured features of the tasks that were not shared between tasks.
+In contrast, all correlations between task-specific drift rates of inhibition tasks were low, even among drift rates of the same task.
+Thus, task-specific drift rate did provide consistent measures of specific EF abilities that are shared across tasks.
+
+Next, we explored the associations of task-specific drift rates with exposure to material deprivation and threat in both childhood and adulthood, after accounting for task-general processing speed (Figure 7).
+Exposure to material deprivation in adulthood was not associated with task-specific drift rates of any of the tasks.
+Exposure to threat in adulthood was negatively associated with the task-specific drift rate in the repeat condition of the Color-shape Task ($\beta$ = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(est.std) |> round(2)`, SE = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(se) |> round(2)`, 95% CI = [`r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(ci.lower) |> round(2)`, `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(ci.upper) |> round(2)`], *p* `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`).
+None of the other associations were significant.
+
+Childhood exposure to material deprivation was negatively associated with all task-specific drift rates, with effect sizes ranging from $\beta$ = `r expl_ts_ch_dep_reg_coef |> filter(rhs == 'child_dep') |> pull(est.std) |> round(2) |> max()` for the congruent condition of the Simon Task to $\beta$ = `r expl_ts_ch_dep_reg_coef |> filter(rhs == 'child_dep') |> pull(est.std) |> round(2) |> min()` for the incongruent condition of the Global-local Task.
+Childhood exposure to threat was negatively associated with all but the task-specific drift rates of the congruent condition of the Flanker and the Simon Task.
+The effect sizes of significant effects ranged from $\beta$ = `r expl_ts_ch_thr_reg_coef |> filter(rhs == 'child_thr') |> pull(est.std) |> round(2) |> max()` for the incongruent condition of the Simon Task to $\beta$ = `r expl_ts_ch_thr_reg_coef |> filter(rhs == 'child_thr') |> pull(est.std) |> round(2) |> min()` for the switch condition of the Posner Task.
+
+To summarize, we did not find evidence that exposure to threat or material deprivation (either in childhood or adulthood) was associated to specific inhibition or attention-shifting ability.
+Rather, we mostly found associations of these adversities with task-general processes.
+We did find associations between childhood adversity and task-specific drift rates.
+However, correlations among task-specific drift rates were low, even for tasks thought to measure the same EF ability, suggesting that they reflect method variance rather than specific EF abilities.
+
+# Discussion
+
+We investigated associations between exposure to two types of adversity in adulthood---material deprivation and threat---with inhibition and attention-shifting ability.
+Participants completed two inhibition tasks, three attention shifting tasks, and one basic processing speed task.
+First, we used DDM to separate raw performance into three distinct cognitive processes: (1) speed of evidence accumulation (drift rate), (2) response caution (boundary separation), and (3) speed of stimulus encoding and response execution (non-decision time).
+Finally, we used structural equation modeling to separate variance in each cognitive process into a task-general factor (shared across all tasks) and ability-specific factors.
+
+## Main findings
+
+People with more exposure to threat in adulthood---but not material deprivation---showed lower general processing speed.
+This is consistent with predictions from deficit frameworks [@sheridan_2014; @tucker_drob_2013].
+After accounting for general processing speed, there was no remaining variance that could be attributed to either inhibition ability or attention-shifting ability.
+Additionally, the associations between material deprivation and threat in adulthood with both response caution and non-decision time were neither significantly different from zero, nor practically equivalent to zero.
+In other words, from these data, we were neither able to conclude that people with more exposure to adversity differed in their level of response caution and non-decision time, nor that these processes were similar to people with lower levels of adversity exposure.
+
+We found that variance in EF task performance was mostly explained by a task-general factor, not by ability-specific factors.
+This finding aligns with previous research applying DDM to EF tasks. 
+Three recent studies converged on the similar conclusion that performance across different EF tasks is best explained by a task-general drift rate factor [@frischkorn_2019; @hedge_2022; @loffler_2024; @weigard_2021].
+One study found that the task-general drift rate factor fully reflected basic processing speed, rather than common EF processes [@loffler_2024].
+Similarly, another study including several inhibition tasks found that shared variance was mostly explained by basic processing speed and strategy differences, not by inhibition ability [@hedge_2022].
+It is therefore not surprising that we were unable to estimate ability-specific EF factors after accounting for general processing.
+These issues tie into several other criticisms of common EF tasks in the cognitive psychology literature---such as low test-retest reliability---which should be considered when using these tasks to test associations with adversity [@vonbastian_2020].
+
+Our results also align with research showing that associations between adversity and raw performance (e.g., response times, accuracy) on EF tasks are mostly task-general.
+For instance, across three large cohorts, youth from more disadvantaged backgrounds had lower task-general raw performance across a wide range of cognitive tasks.
+After controlling for task-general effects, performance on specific intelligence tasks was largely practically equivalent to zero, and performance on some specific EF tasks even appeared enhanced [@bignardi_2024].
+Finally, previous research applying DDM in the context of adversity found similar (but not identical) results as reported here.
+Youth with more exposure to threat (but not material deprivation) had slower general processing speed across three EF tasks and a basic processing speed task, while most specific EF abilities remained intact [@vermeent_2024a].
+In contrast to the findings reported here, these youth also showed more response caution.
+It is important to highlight that both studies measured specific EF abilities with only a single task.
+This makes it unclear to which extent the task-specific effects they observed reflected EF abilities versus method variance (see also our non-preregistered findings discussed below).
+
+## Non-preregistered findings
+
+Additional non-preregistered analyses showed that childhood exposure to both material deprivation and threat were associated with slower general processing speed.
+The effect sizes of these associations were larger than those of recent adversity exposure.
+This finding is striking considering the average age in our sample was 39 years.
+This is consistent with previous literature suggesting that adverse experiences early in life can have a lasting effect on brain development [@shonkoff_2012; @nelson_2020a; @nelson_2020b].
+For instance, both experiences of malnutrition and exposure to stress (e.g., as a result of threat) have been linked to structural changes in brain networks involved in cognitive functioning [@algarin_2017; @polavarapu_2017; @rebello_2018].
+
+A second important non-preregistered finding is that after controlling for task-general processing, adversity exposure was negatively associated with several task-specific drift rates, mostly relating to childhood exposure to threat and deprivation.
+Importantly, the correlations among task-specific drift rates were all small---with the exception of correlations among switch and repeat conditions of the same attention-shifting tasks.
+This suggests that these negative associations should not be interpreted as lower ability-specific processing, relating to inhibition and attention-shifting ability.
+Instead, it appears to relate to information processing that is much more specific to individual tasks.
+This is consistent with literature showing that a substantial part of the variance in performance on cognitive task consists of method variance resulting from the nature of the task itself, rather than what the task is actually designed to measure [@barkley_2012; @rey_mermet_2024; @schubert_2016].
+
+## Implications for adversity research and interventions
+
+Our results have two main implications for adversity research and interventions, both of which warrant caution against interpreting raw performance on individual EF tasks.
+First, adversity appears to be associated mostly with task-general processes, rather than specific abilities.
+This contrasts with common inferences in both the deficit and adaptation literature, which tend to explain performance differences in terms of specific EF abilities, such as inhibition [@farah_2006; @fields_2021; @mezzacappa_2004; @mittal_2015; @noble_2005] and attention shifting [@fields_2021; @howard_2020; @nweze_2021; @young_2022; @mittal_2015].
+Moreover, our results show that accounting for task-general processing does not guarantees that task-specific associations with adversity reflect differences in EF.
+Instead, negative associations with adversity appear to be partly driven by specific features of individual tasks (method variance).
+A general recommendation for adversity research is therefore that studies should include (1) multiple EF tasks to quantify task-general processing speed, and (2) ideally two or more tasks measuring the same EF ability to quantify ability-specific processing.
+DDM and SEM can then be used to account for task-general and task-specific processes.
+For interventions, our results imply that individual EF tasks, or batteries of EF tasks each of which measures a different ability, have limited value as screening tools; and, that training performance on such tasks is unlikely to transfer to broader, sustainable changes in EF abilities.
+
+Second, the dominance of task-general processes is a challenge for adaptation frameworks, which predict that specific types of adversity may enhance or leave intact specific cognitive abilities.
+The lack of task-specific adaptations may have resulted from our use of standardized tasks featuring abstract content.
+Previous research suggests that cognitive adaptations may only be visible when using real-world content rather than abstract content [@young_2022], possibly by increasing task engagement [@niebaum_2023].
+On abstract tasks, such adapted abilities may only be visible under current conditions of stress [@mittal_2015; @young_2018].
+However, previous research also shows that content effects are complex and not always in the same direction.
+For instance, while real-world content equalized working memory performance of youth exposed to violence, it lowered attention-shifting performance for all youth, regardless of adversity exposure.
+Similarly, youth from low socioeconomic backgrounds perform worse on math items involving money, relative to other content [@duquennois_2022; @muskens_2024].
+It is currently unclear how performance on different tasks will be affected by real-world content, and which stimulus features drive these effects.
+While outside the scope of this article, DDM could play a key role in addressing these questions by showing how different types of contexts and content influence distinct cognitive processes.
+
+## Strengths, limitations, and future directions
+
+Our study has three main strengths.
+First, we measured inhibition and attention shifting ability using multiple tasks for each EF ability.
+This approach provided two advantages: (1) we could distinguish between general versus specific processes, and (2) we could measure EF abilities without task-specific measurement error.
+Second, we used DDM to distinguish between evidence accumulation (EF ability, general speed of processing), response caution, and non-decision time (stimulus encoding, response execution).
+The DDM allowed us to better understand the cognitive processes driving performance in people with more exposure to adversity.
+Third, we included a large, representative sample of the Dutch population sample with sufficient variance on key adversity measures.
+
+Our study has four main limitations. 
+First, our study included participants with low to moderate levels of adversity exposure, which was measured via self-reports. 
+Therefore, our study provides a less severe test of our hypotheses than a case-control study in which participants with severe exposure to adversity are compared to a control group. 
+Second, although we included confounding variables based on explicit, theory-guided causal assumptions, the cross-sectional and partly retrospective nature of our adversity data prohibit us from establishing true causal relationships.
+Third, due to time constraints, we only included two inhibition tasks and three attention shifting tasks, limiting our ability to model shared variance between them.
+Future studies could include more EF tasks, and ideally include several basic processing speed tasks.
+Fourth, participants completed the cognitive tasks online in their own home.
+Despite our controlling for environmental factors that may have disrupted task performance, the home setting may have reduced the reliability of performance measures.
+
+We envision three future directions.
+First, because adversity tends to be associated with task-general processes, it will be important to better understand the nature of these processes.
+To the extent that task-general drift rates reflect general processing speed, it may partly result from lower white matter tract integrity [@fuhrmann_2020; @kievit_2016].
+Such associations with structural brain differences may account for the finding that task-general drift rate is largely stable over time [@schubert_2016; @weigard_2021b].
+Yet, task-general drift rate may be associated with other factors such as effort, fatigue, or hunger [@weigard_2021b], which may play a bigger role for people from adverse environments [@brose_2012; @sliwinski_2006; @schwabe_2013].
+Thus, future studies may include (a combination of) brain measures or measures of mental states to better understand inter- as well as intrapersonal differences in task-general drift rates.
+
+Second, future research should investigate why adversity is negatively associated with task-specific drift rates. 
+Our results suggest that task-specific drift rates mostly capture method variance rather than specific EF abilities.
+We currently do not have a clear explanation for their negative associations with adversity.
+One possible explanation might be that the associations reflect people's difficulty with processing abstract content.
+As discussed above, people with more exposure to adversity may be better at processing (stimulus) information that aligns with their daily experiences [@young_2022].
+Alternatively, lower performance on abstract tasks may reflect lower levels of engagement or motivation, rather than inability, to engage EF [@niebaum_2023].
+All tasks in our study included abstract content, and each study used different stimuli, which may account for the negative associations with adversity and the lack of correlations between task-specific drift rates.
+However, if so, this would not explain the low correlations between congruent and incongruent conditions of both the Flanker and Simon task, as these conditions do include the same stimuli.
+Nevertheless, research could more systematically test associations between adversity and processing as a function of task content.
+
+Third, future research could design preregistered, well-powered analyses to tease apart the unique effects of childhood adversity and recent adversity on cognitive performance, as well as their combined effects.
+Our understanding of how cognitive abilities later in life depend on the developmental timing of adversity exposure is still limited [@frankenhuis_2020].
+In this study, we assumed that recent adversity exposure mediates the effect of childhood adversity exposure on cognitive processes, as people who experienced adversity early in life are more likely to also experience adversity later in life (e.g., due to systemic constraints; @hazel_2008).
+Our study focused on adversity in adulthood, the average or summed adversity exposure over the past few years.
+Future research should distinguish such recent adversity from acute stress. 
+For instance, some research suggests that childhood adversity enhances specific EF abilities, but only under situations of acute stress [@mittal_2015; @young_2018].
+Thus, future research should tease apart effects of adversity during childhood and adulthood and examine how their effects are moderated by acute stress.
+
+## Constraints on generality
+
+Our research question concerned the association between adversity exposure and performance on EF tasks, using cognitive modeling to investigate whether these associations are driven by task-general processing speed or specific EF abilities.
+Our target population was Dutch adults with a broad range of adversity exposures.
+The LISS panel ensures representation of the Dutch adult population in terms of---among other factors---age, education, and socioeconomic status.
+Our findings may generalize to adults in other Western populations, but not necessarily to adults in non-Western populations [@nketia_2023].
+We used standardized inhibition and attention-shifting tasks that are often used in cognitive psychology. 
+The findings should generalize to other tasks measuring these abilities, as long as they rely on speeded responses.
+It will require different modeling approaches to understand how adversity is associated with EF abilities that do not require evidence accumulation, or with EF decision processes that unfold over longer timescales (e.g., minutes).
+All tasks included abstract stimuli (e.g., geometric shapes, words).
+Results are likely to be different when using more real-world stimuli, or when changing the tasks in other ways to better align with people's daily experiences [@doebel_2020; @miller_cotto_2022; @young_2022].
+Participants completed the study online in their home environment. 
+To reproduce the effects, it is important to carefully instruct participants to limit environmental distractions and complete the study in full-screen mode.
+The results should generalize to more standardized settings (e.g., the lab), but may not generalize to public settings outside of the home environment.
+
+## Conclusion
+
+Adversity research has made important steps in recent years, revealing how exposure to adversity may impair certain EF abilities while enhancing others.
+Further progress in this field hinges on our capacity to accurately measure specific EF abilities.
+We show, consistent with previous research, that this is a more challenging task than typically assumed.
+As a result, adversity researchers likely overestimate the strength of associations between adversity and EF abilities when relying on analyses of raw performance.
+We present a constructive path forward through a combination of cognitive modeling and structural equation modeling. 
+Embracing these techniques, and revising theories in light of their findings, will enhance our understanding of how exposure to adversity shapes both specific EF abilities and general cognitive processes.
+
+{{< pagebreak >}}
+
+# References
+
+::: {#refs}
+:::
+
+{{< pagebreak >}}
+
+```{r}
+#| tab.id: table1
+#| results: markup
+table1 |> 
+  flextable() |> 
+  add_header_row(
+    values = " ",
+    colwidths = 2
+  ) |> 
+  flextable::compose(
+    i = 1, j = 1, 
+    as_paragraph(as_b("Table 1. "), "Descriptive statistics for the final sample."),
+    part = "header"
+  ) |> 
+  border_remove() |> 
+  border(i = 1, border.bottom = fp_border_default(), part = "header") |>  
+  border(i = 2, border.top = fp_border_default(), part = "header") |>  
+  border(i = 2, border.bottom = fp_border_default(), part = "header") |> 
+  border(i = 17, border.bottom = fp_border_default(), part = "body") |> 
+  set_header_labels(cat = "Category", stat = "Statistic") |>
+  padding(i = c(3:5, 7:14, 16:17), j = 1, padding.left = 10) |>
+  autofit()
+```
+
+{{< pagebreak >}}
+
 ::: {.landscape}
@@
 ```{r}
-#| label: Figure3
-#| dpi: 600
-#| out-width: 12in
-#| fig-cap: | 
-#|   **Figure 3.** Final measurement model of the Drift Diffusion parameters.
-Rectangles depict manifest parameter estimates, elipses depict latent factors, unidirectional arrows depict factor loadings, and bidirectional arrows depict correlations.
-Latent factors with non-significant residual variances are greyed out.
-*a* = manifest boundary separation, AS = Attention-shifting task, CS = Color-shape task, GL = Global-local task, Fl = Flanker task, Sm = Simon task, *t* = manifest non-decision time, *v* = manifest drift rate.
-knitr::include_graphics("figures/fig3.png")
+#| tab.id: table2
+#| results: markup
+table2 |> 
+  flextable() |> 
+  flextable::set_header_labels(Variable = "") |> 
+  flextable::autofit() |> 
+  border_remove() %>% 
+ # border(i = 1, border.top = fp_border_default(), part = "header") %>% 
+  border(i = 1, border.bottom = fp_border_default(), part = "header") %>% 
+  border(i = 20, border.bottom = fp_border_default(), part = "body") %>% 
+  add_header_row(
+    values = " ",
+    colwidths = 15
+  ) |> 
+  flextable::compose(
+    i = 1, j = 1, 
+    as_paragraph(as_b("Table 2. "), "Spearman correlations between the main independent variables and covariates."),
+    part = "header"
+  ) |> 
+  add_footer_row(
+    values = " ",
+    colwidths = 15
+  ) |> 
+  autofit()
 ```
@@
 ```{r}
-#| tab.id: table3
+table3 |> 
+  mutate(empty = "") |> 
+  select(task, condition, rt_m, rt_range, empty, ac_m, ac_range) |> 
+  flextable() |> 
+  set_header_labels(task = "Task", condition = "Condition", rt_m = "Mean (SD)", rt_range = "Range", empty = "", ac_m = "Mean (SD)", ac_range = "Range") |> 
+  add_header_row(values = c("", "", "Response times", "", "", "Accuracy", "")) |> 
+  merge_at(j = c(3:4), i = 1, part = "header") |> 
+  merge_at(j = c(6:7), i = 1, part = "header") |>
+
+  add_header_row(values = " ",colwidths = 7) |> 
+  flextable::compose(
+    i = 1, j = 1, 
+    as_paragraph(as_b("Table 3. "), "Descriptive statistics of response times and accuracy on cognitive tasks."),
+    part = "header"
+  ) |> 
+  align(i = 2, align = "center", part = "header") |> 
+  border_remove() %>% 
+  border(i = 2, border.top = fp_border_default(), part = "header") |> 
+  border(i = 2, j = c(3,6), border.bottom = fp_border_default(), part = "header") |> 
+  border(i = 3, border.bottom = fp_border_default(), part = "header") |> 
+  border(i = 11, border.bottom = fp_border_default(), part = "body") |>  
+  autofit()
+```
+
+{{< pagebreak >}}
+
+```{r}
+#| tab.id: table4
 #| results: markup
 
-table3 |> 
+table4 |> 
+  flextable() |>
+  set_header_labels(values = c("Model", "Chi square", "Robust CFI", "Robust RMSEA")) |>
+  flextable::merge_at(i = 2, j = 2:4) |>
+  flextable::merge_at(i = 4, j = 2:4) |>
+  flextable::merge_at(i = 5, j = 2:4) |>
+  flextable::merge_at(i = 7, j = 2:4) |>
+  flextable::merge_at(i = 8, j = 2:4) |>
+  flextable::align(i = c(2,4,5,7,8), j = 2, align = "center") |>
+  bold(i = c(1, 11), j = 1) |>
+  padding(i = c(2:10, 12:15), j = 1, padding.left = 15) |>
   add_header_row(
@@
     i = 1, j = 1,
-    as_paragraph(as_b("Table 3. "), "Fit statistics of all preregistered structural equation models."),
+    as_paragraph(as_b("Table 4. "), "Fit statistics of all preregistered structural equation models."),
     part = "header"
@@
 
-<br>
+{{< pagebreak >}}
 
-## Preregistered analyses
+```{r}
+#| label: Figure1
+#| fig-width: 6.5
+#| fig-height: 7
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure 1.** Visual representation of the Drift Diffusion Model (DDM).
+The DDM assumes that people accumulate evidence for one of two responses until they have acquired sufficient evidence to make a decision.
+Each squiggly line represents the evidence accumulation process on a single trial.
+The thick, horizontal lines represent the decision boundaries corresponding to the correct response (upper boundary) and the incorrect response (lower boundary).
+Applying the DDM to trial-level performance data (response times and accuracies) yields four parameters representing distinct cognitive processes: (1) **drift rate**: Average rate of evidence accumulation across trials; measuring the efficiency of information processing.
+(2) **boundary separation**: The distance between the two decision boundaries; measuring response caution.
+(3) **Non-decision time**: A combination of the preparation speed (e.g., stimulus encoding) and response execution speed (e.g., time spent pressing the button); measuring speed of non-decision processes.
+**starting point**: The starting point of the evidence accumulation process; measuring response bias (not considered here).
+Figure adopted from Vermeent et al.
+(2024), licensed under CC BY 4.0.
+knitr::include_graphics("figures/fig1.png")
+```
 
-Our primary analyses examined the association of exposure to material deprivation and threat in adulthood with task-general processing speed, task-general response caution, and task-general speed of stimulus encoding and/or response execution.
-As the structural equation model analyses reported above showed that EF-specific factors had non-significant variances, we retained them in the model but did not include them as dependent variables in subsequent analyses.
+{{< pagebreak >}}
 
-### Material deprivation in adulthood
+```{r}
+#| label: Figure2
+#| fig-width: 6.5
+#| fig-height: 7
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure 2.** Direct Acyclic Graphs (DAG) depicting our causal assumptions about the main estimands.
+the rectangles (nodes) depict variables, and directed arrows depict assumed causal pathways.
+An arrow between two variables represents the assumption that experimentally manipulating the variable at the origin of the arrow will change the variable at the end of the arrow (but not the other way around).
+The statistical models control for confounding variables (depicted as dashed paths), which have an arrow both to the independent variable and the dependent variable.
+knitr::include_graphics("figures/fig2.png")
+```
 
-Exposure to material deprivation in adulthood was not associated with either general processing speed ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_v', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_v', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`), task-general response caution ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_a', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_a', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`), or task-general non-decision time ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_t', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_t', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`) (see Figure 4).
-In addition, none of the effects fell within the region of practical equivalence (all *p*s ≥ `r fit_dep1_eqtests |> pull(eq_pvalue) |> min() |> format_p(pvalue = _, operator = FALSE)`).
-The effects in the secondary model (including expossure to threat in adulthood as a confounder) were comparable.
+::: {.landscape}
 
-<br>
+```{r}
+#| label: Figure3
+#| dpi: 600
+#| out-width: 15in
+#| fig-cap: | 
+#|   **Figure 3.** Final measurement model of the Drift Diffusion parameters.
+Rectangles depict manifest parameter estimates, elipses depict latent factors, unidirectional arrows depict factor loadings, and bidirectional arrows depict correlations.
+Latent factors with non-significant residual variances are greyed out.
+*a* = manifest boundary separation, AS = Attention-shifting Task, CS = Color-shape Task, GL = Global-local Task, Fl = Flanker Task, Sm = Simon Task, *t* = manifest non-decision time, *v* = manifest drift rate.
+knitr::include_graphics("figures/fig3.png")
+```
+
+:::
 
@@
 Standard errors represent 95% confidence intervals.
-Statistical significance (tested against zero) is indicated with asterisks; * *p* < .05, ** *p* < .01, *** *p* < .001.
+The asterisks indicate statistical significance (tested against zero); * *p* < .05, ** *p* < .01, *** *p* < .001.
 fig4
@@
 
-<br>
-
-### Threat in adulthood
-
-Exposure to threat in adulthood was negatively associated with task-general processing speed ($\beta$ = `r fit_thr1_reg_coef |> filter(lhs == 'gen_v', rhs == 'threat_comp') |> pull(est.std) |> formatC(digits = 2, width = 2, format = 'f')`, *p* `r fit_thr1_reg_coef |> filter(lhs == 'gen_v', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`) (see Figure 4).
-People who reported more exposure to threat in adulthood processed information more slowly across tasks.
-Exposure to threat in adulthood was not associated with task-general response caution ($\beta$ = `r fit_thr1_reg_coef |> filter(lhs == 'gen_a', rhs == 'threat_comp') |> pull(est.std) |> round(2)`, *p* `r fit_thr1_reg_coef |> filter(lhs == 'gen_a', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`), or task-general non-decision time ($\beta$ = `r fit_thr1_reg_coef |> filter(lhs == 'gen_t', rhs == 'threat_comp') |> pull(est.std) |> round(2)`, *p* `r fit_thr1_reg_coef |> filter(lhs == 'gen_t', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`).
-The effects in the secondary model (excluding material deprivation in adulthood as a confounder) were comparable, although the association with task-general non-decision time was practically equivalent in the secondary model.
-
-## Non-preregistered analyses
-
-### Associations with childhood adversity
-
-In the first set of non-preregistered analyses, we analyzed how childhood exposure to threat and material deprivation are associated with task-general and ability-specific cognitive processes.
-The rational for these analyses is that we observed significant indirect associations between childhood adversity and general speed of processing in our preregistered analyses.
-Here, we analyzed *direct* effects of childhood threat and material deprivation (i.e., without including exposure to adversity in adulthood).
-We controlled for sex (but not age and education) and in the case of threat we also controlled for material deprivation, but not the other way around (similarly to the preregistered analyses).
-
-Childhood exposure to material deprivation was negatively associated with general processing speed, $\beta$ = `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(est.std) |> round(2)`, SE = `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(se) |> round(2)`, 95% CI = [`r expl_ch_dep_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(ci.lower) |> round(2)`, `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(ci.upper) |> round(2)`], *p* `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_dep') |> pull(pvalue_adj) |> format_p(pvalue = _)`.
-People who reported higher levels of childhood exposure to material deprivation processed information more slowly across tasks.
-In addition, childhood exposure to material deprivation was positively associated with non-decision time, $\beta$ = `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(est.std) |> round(2)`, SE = `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(se) |> round(2)`, 95% CI = [`r expl_ch_dep_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(ci.lower) |> round(2)`, `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(ci.upper) |> round(2)`], *p* `r expl_ch_dep_reg_coef |> filter(lhs == 'gen_t', rhs == 'child_dep') |> pull(pvalue_adj) |> format_p(pvalue = _)`.
-People who reported higher levels of childhood exposure to material deprivation took more time to encode stimulus information and/or execute responses across tasks.
-Childhood exposure to material deprivation was not associated with general response caution.
-Equivalence tests revealed that none of these associations fell within the region of practical equivalence.
-
-For childhood exposure to threat, we found a significant negative association with general processing speed, $\beta$ = `r expl_ch_thr_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(est.std) |> round(2)`, SE = `r expl_ch_thr_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(se) |> round(2)`, 95% CI = [`r expl_ch_thr_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(ci.lower) |> round(2)`, `r expl_ch_thr_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(ci.upper) |> round(2)`], *p* `r expl_ch_thr_reg_coef |> filter(lhs == 'gen_v', rhs == 'child_thr') |> pull(pvalue_adj) |> format_p(pvalue = _)`.
-People who reported higher levels of childhood exposure to threat processed information more slowly across tasks.
-We did not find significant associations between childhood exposure to threat and general response caution or general non-decision time.
-Of these associations, equivalence tests revealed that the association with general response caution fell within the region of practical equivalence (*p* `r expl_ch_thr_eqtests |> filter(lhs == "gen_a", rhs == "child_thr") |> pull(eq_pvalue) |> format_p(pvalue = _)`).
+{{< pagebreak >}}
 
@@
 Standard errors represent 95% confidence intervals.
-Statistical significance (tested against zero) is indicated with asterisks; * *p* < .05, ** *p* < .01, *** *p* < .001.
+The asterisks indicate statistical significance (tested against zero); * *p* < .05, ** *p* < .01, *** *p* < .001.
 fig5
@@
 
-<br>
-
-### Associations with task-specific drift rates
-
-In the second set of non-preregistered analyses, we analyzed how adversity exposure during childhood and adulthood were associated with task-specific residual variances.
-The rationale for these analyses was that while we did not find coherent latent EF factors, all tasks showed significant residual variance after accounting for general speed of processing.
-This implies that, after accounting for general speed of processing, performance is further determined by unique features of individual tasks (i.e., method variance), which may be associated with adversity exposure.
-For this analysis, we were mostly interested in investigating the extent to which task-specific effects reflect EF abilities or task-specific processing, and therefore only focused on drift rates.
-
-We first attempted to fit a model in which all task-specific drift rates were allowed to covary (i.e., across inhibition and attention-shifting tasks).
-However, this model did not converge.
-A model that only estimated covariances between task-specific drift rates of inhibition tasks and attention shifting tasks separately provided a good fit to the data (CFI = `r fit_expl_ts_mod2_fitstats[['cfi.robust']] |> round(2)`, RMSEA = `r fit_expl_ts_mod2_fitstats[['rmsea.robust']] |> round(2)` [`r fit_expl_ts_mod2_fitstats[['rmsea.ci.lower.robust']] |> round(2)`, `r fit_expl_ts_mod2_fitstats[['rmsea.ci.upper.robust']] |> round(2)`]).
-This suggests that task-specific drift rates correlated among inhibition tasks and among attention-shifting tasks, but not between inhibition and attention-shifting tasks.
-
-Figure 6 shows latent correlations between task-specific drift rates after accounting for task-general processing speed.
-The drift rates of repeat and switch conditions within each attention shifting task correlated moderately to strongly, while all correlations between tasks were low.
-This suggests that task-specific drift rates of attention shifting tasks captured features of the tasks that were not shared between tasks.
-In contrast, all correlations between task-specific drift rates of inhibition tasks were low, even among drift rates of the same task.
-Thus, task-specific drift rate did provide consistent measures of specific EF abilities that are shared across tasks.
-
-<br>
+{{< pagebreak >}}
 
@@
 ```
-<br>
 
-Next, we explored the associations of task-specific drift rates with exposure to material deprivation and threat in both childhood and adulthood, after accounting for task-general processing speed (Figure 7).
-Exposure to material deprivation in adulthood was not associated with task-specific drift rates of any of the tasks.
-Exposure to threat in adulthood was negatively associated all task-specific drift rates in both the repeat and switch condition of the Color-shape task (repeat: $\beta$ = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(est.std) |> round(2)`, SE = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(se) |> round(2)`, 95% CI = [`r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(ci.lower) |> round(2)`, `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(ci.upper) |> round(2)`], *p* `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`; switch: $\beta$ = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_sw_v', rhs == 'threat_comp') |> pull(est.std) |> round(2)`, SE = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_sw_v', rhs == 'threat_comp') |> pull(se) |> round(2)`, 95% CI = [`r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_sw_v', rhs == 'threat_comp') |> pull(ci.lower) |> round(2)`, `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_sw_v', rhs == 'threat_comp') |> pull(ci.upper) |> round(2)`], *p* `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_sw_v', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`).
-None of the other associations were significant.
-
-Childhood exposure to material deprivation was negatively associated with all task-specific drift rates, with effect sizes ranging from $\beta$ = `r expl_ts_ch_dep_reg_coef |> filter(rhs == 'child_dep') |> pull(est.std) |> round(2) |> max()` for the congruent condition of the Simon Task to $\beta$ = `r expl_ts_ch_dep_reg_coef |> filter(rhs == 'child_dep') |> pull(est.std) |> round(2) |> min()` for the incongruent condition of the Global-local Task.
-Childhood exposure to threat was negatively associated with all but the task-specific drift rates of the congruent condition of the Flanker and the Simon task.
-The effect sizes of significant effects ranged from $\beta$ = `r expl_ts_ch_thr_reg_coef |> filter(rhs == 'child_thr') |> pull(est.std) |> round(2) |> max()` for the incongruent condition of the Simon Task to $\beta$ = `r expl_ts_ch_thr_reg_coef |> filter(rhs == 'child_thr') |> pull(est.std) |> round(2) |> min()` for the switch condition of the Posner Task.
-
-To summarize, we did not find evidence that exposure to threat or material deprivation (either in childhood or adulthood) was associated to specific inhibition or attention-shifting ability.
-Rather, we mostly found associations with task-general processes.
-We did find associations between childhood adversity and task-specific drift rates.
-However, correlations among task-specific drift rates were low, even for tasks thought to measure the same EF ability, suggesting that they reflect method variance rather than specific EF abilities.
+{{< pagebreak >}}
 
@@
 Standard errors represent 95% confidence intervals.
-Statistical significance (tested against zero) is indicated with asterisks; * *p* < 0.05, ** *p* < 0.01, *** *p* < 0.001.
+The asterisks indicate statistical significance (tested against zero); * *p* < 0.05, ** *p* < 0.01, *** *p* < 0.001.
 fig7
 ```
-<br>
-
-# Discussion
-
-We investigated associations between exposure to two types of adversity in adulthood---material deprivation and threat---with inhibition and attention shifting ability.
-Participants completed two inhibition tasks, three attention shifting tasks, and one basic processing speed task.
-First, we used DDM to separate raw performance into three distinct cognitive processes: (1) speed of evidence accumulation (drift rate), (2) response caution (boundary separation), and (3) speed of stimulus encoding and response execution (non-decision time).
-Finally, we used structural equation modeling to separate variance in each cognitive process into a task-general factor (shared across all tasks) and ability-specific factors.
-
-## Main findings
-
-People with more exposure to threat in adulthood---but not material deprivation---showed lower general processing speed.
-This is consistent with predictions from deficit frameworks [@sheridan_2014; @tucker_drob_2013].
-After accounting for general processing speed, there was no remaining variance that could be attributed to either inhibition ability or attention shifting ability.
-Additionally, the associations between material deprivation and threat in adulthood with both response caution and non-decision time were neither significantly different from zero, nor practically equivalent to zero.
-In other words, from these data, we were neither able to conclude that people with more exposure to adversity differed in their level of response caution and non-decision time, nor that these processes were similar to people with lower levels of adversity exposure.
-
-We found that variance in EF task performance was mostly explained by a task-general factor, not by ability-specific factors.
-This finding aligns with previous research applying DDM to EF tasks. 
-Three recent studies converged on the similar conclusion that performance across different EF tasks is best explained by a task-general drift rate factor [@frischkorn_2019; @hedge_2021; @loffler_2024; @weigard_2021].
-One study found that the task-general drift rate factor fully reflected basic processing speed, rather than common EF processes [@loffler_2024].
-Similarly, another study including several inhibition tasks found that shared variance was mostly explained by basic processing speed and strategy differences, not by inhibition ability [@hedge_2021].
-It is therefore not surprising that we were unable to estimate ability-specific EF factors after accounting for general processing.
-These issues tie into several other criticisms of common EF tasks in the cognitive psychology literature---such as low test-retest reliability---which should be considered when using these tasks to test associations with adversity [@vonbastian_2020].
-
-Our results also align with research showing that associations between adversity and raw performance (e.g., response times, accuracy) on EF tasks are mostly task-general.
-For instance, across three large cohorts, youth from more disadvantaged backgrounds had lower task-general raw performance across a wide range of cognitive tasks.
-After controlling for task-general effects, performance on specific intelligence tasks was largely practically equivalent to zero, and performance on some specific EF tasks even appeared enhanced [@bignardi_2024].
-Finally, previous research applying DDM in the context of adversity found similar (but not identical) results as reported here.
-Youth with more exposure to threat (but not material deprivation) had slower general processing speed across three EF tasks and a basic processing speed task, while most specific abilities remained intact [@vermeent_2024a].
-In contrast to the findings reported here, these youth also showed more response caution.
-It is important to highlight that both studies measured specific EF abilities with only a single task.
-This makes it unclear to which extent the task-specific effects they observed reflected EF abilities (see also our non-preregistered findings discussed below).
-
-## Non-preregistered findings
-
-Additional non-preregistered analyses showed that childhood exposure to both material deprivation and threat were associated with slower general processing speed.
-The effect sizes of these associations were larger than those of recent adversity exposure.
-This finding is striking considering the average age in our sample was 39 years.
-This is consistent with previous literature suggesting that adverse experiences early in life have a lasting effect on brain development [@shonkoff_2012; @nelson_2020a; @nelson_2020b].
-For instance, both experiences of malnutrition and exposure to stress (e.g., as a result of threat) have been linked to structural changes in brain networks involved in cognitive functioning [@algarin_2017; @polavarapu_2017; @rebello_2018].
-
-A second important non-preregistered finding is that after controlling for task-general processing, adversity exposure was negatively associated with several task-specific drift rates, mostly relating to childhood exposure to threat and deprivation.
-Importantly, the correlations among task-specific drift rates were all small---with the exception of correlations among switch and repeat conditions of the same attention shifting tasks.
-This suggests that these negative associations should not be interpreted as lower ability-specific processing, relating to inhibition and attention shifting ability.
-Instead, it appears to relate to information processing that is much more specific to individual tasks.
-This is consistent with literature showing that a substantial part of the variance in performance on cognitive task consists of method variance resulting from the nature of the task itself, rather than what the task is designed to measure [@barkley_2012; @rey_mermet_2024; @schubert_2016].
-
-## Implications for adversity research and interventions
-
-Our results have two implications for adversity research and interventions, both of which warrant caution against interpreting raw performance on individual EF tasks.
-First, adversity appears to be associated mostly with task-general processes, rather than specific abilities.
-This contrasts with common inferences in both the deficit and adaptation literature, which tend to explain performance differences in terms of specific EF abilities, such as inhibition [@farah_2006; @fields_2021; @mezzacappa_2004; @mittal_2015; @noble_2005] and attention shifting [@fields_2021; @howard_2020; @nweze_2021; @young_2022; @mittal_2015].
-Moreover, our results show that accounting for task-general processing does not guarantees that task-specific associations with adversity reflect differences in EF.
-Instead, negative associations with adversity appear to be partly driven by specific features of individual tasks.
-A general recommendation for adversity research is therefore that studies should include (1) multiple EF tasks to quantify task-general processing speed, and (2) ideally two or more tasks measuring the same EF ability to quantify ability-specific processing.
-DDM and SEM can then be used to account for task-general and task-specific processes.
-For interventions, our results imply that individual EF tasks, or batteries of EF tasks each of which measures a different ability, have limited value as screening tools; and, that training performance on such tasks is unlikely to transfer to broader, sustainable changes in EF abilities.
-
-Our findings complicate research that aims to understand how EF abilities develop in adverse environments.
-We agree with others that results like ours do not necessarily mean that EF abilities do not exist [e.g., @loffler_2024], or that adversity is not associated with differences in EF.
-Rather, it points to issues with how we *measure* EF abilities.
-It is difficult to provide a straightforward solution for this problem, which is still a much-debated topic in cognitive psychology [@vonbastian_2020].
-Part of the solution may lie in adjusting the content of EF tasks, for instance, by using real-world content rather than abstract content [@young_2022].
-By increasing familiarity, relevance, and/or valence, such content might increase the engagement of EF abilities [@niebaum_2023].
-Another solution could be to use adaptive testing procedures, which appear to improve estimation of EF abilities [@draheim_2021b].
-However, more fundamental changes might be necessary as well, such as developing new paradigms that align more closely to people's everyday experiences, goals, and challenges [@doebel_2020; @miller_cotto_2022; @niebaum_2023].
-
-## Strengths, limitations, and future directions
-
-Our study has three main strengths.
-First, we measured inhibition and attention shifting ability using multiple tasks for each EF ability.
-This approach provided two advantages: (1) we could distinguish between general versus specific processes, and (2) we could measure EF abilities without task-specific measurement error.
-Second, we used DDM to distinguish between evidence accumulation (EF ability, general speed of processing), response caution, and non-decision time (stimulus encoding, response execution).
-The DDM allowed us to better understand the cognitive processes driving performance in people with more exposure to adversity.
-Third, we included a large, representative sample of the Dutch population sample with sufficient variance on key adversity measures.
-
-Our study has three main limitations. 
-First, although we included confounding variables based on explicit, theory-guided causal assumptions, the cross-sectional and partly retrospective nature of our data prohibit us from establishing true causal relationships.
-Second, due to time constraints, we only included two inhibition tasks and three attention shifting tasks, limiting our ability to model shared variance between them.
-Future studies could include more EF tasks, and ideally include several basic processing speed tasks.
-Third, participants completed the cognitive tasks online in their own home.
-Despite our controlling for environmental factors that may have disrupted task performance, the home setting may have reduced the reliability of performance measures.
-
-We envision three future directions.
-First, because adversity tends to be associated with task-general processes, it will be important to better understand the nature of these processes.
-To the extent that task-general drift rates reflect general processing speed, it may partly result from lower white matter tract integrity [@fuhrmann_2020; @kievit_2016].
-Such associations with structural brain differences may account for the finding that task-general drift rate is relatively stable over time [@schubert_2016; @weigard_2021b].
-Yet, task-general drift rate may be associated with other factors such as effort, fatigue, or hunger [@weigard_2021b], which may play a bigger role for people from adverse environments [@brose_2012; @sliwinski_2006; @schwabe_2013].
-Thus, future studies may include (a combination of) brain measures or measures of mental states to better understand inter- and intrapersonal differences in task-general drift rates.
-
-Second, future research should investigate why adversity is negatively associated with task-specific drift rates. 
-Our results suggest that task-specific drift rates mostly capture method variance rather than specific EF abilities.
-We currently do not have a clear explanation for their negative associations with adversity.
-One possible explanation might be that the associations reflect people's difficulty with processing abstract content.
-Previous research found that youth with more exposure to adversity performed lower on a working memory updating task (compared with youth with fewer adversity exposures) when task stimuli were abstract geometric shapes.
-However, their performance equalized when the task included real-world content [@young_2022].
-All tasks in our study included abstract content, and each study used different stimuli, which may account for the negative associations with adversity and the lack of correlations between task-specific drift rates.
-However, if so, this would not explain the low correlations between congruent and incongruent conditions of both the Flanker and Simon task, as these conditions do include the same stimuli.
-Nevertheless, research could more systematically test associations between adversity and processing as a function of task content.
-
-Third, future research could design preregistered, well-powered analyses to tease apart the unique effects of childhood adversity and recent adversity on cognitive performance, as well as their combined effects.
-Our understanding of how cognitive outcomes later in life depend on the developmental timing of adversity exposure is still limited [@frankenhuis_2020].
-In this study, we assumed that recent adversity exposure mediates the effect of childhood adversity exposure on cognitive processes, as people who experienced adversity early in life are more likely to also experience adversity later in life (e.g., due to systemic constraints; @hazel_2008).
-Our study focused on adversity in adulthood, the average or summed adversity exposure over the past few years.
-Future research should distinguish such recent adversity from acute stress. 
-For instance, some research suggests that childhood adversity enhances specific EF abilities, but only under situations of acute stress [@mittal_2015; @young_2018].
-Thus, future research should tease apart effects of adversity during childhood and adulthood and study how their effects are moderated by acute stress.
-
-## Constraints on generality
-
-Our research question concerned the association between adversity exposure and performance on EF tasks, using cognitive modeling to investigate whether these associations are driven by task-general processing speed or specific EF abilities.
-Our target population was Dutch adults with a broad range of adversity exposures.
-The LISS panel ensures representation of the Dutch adult population in terms of---among other factors---age, education, and socioeconomic status.
-Our findings may generalize to adults in other Western populations, but not necessarily to adults in non-Western populations [@nketia_2023].
-We used standardized inhibition and attention-shifting tasks that are often used in cognitive psychology. 
-The findings should generalize to other tasks used to measure these abilities, as long as they rely on speeded responses.
-All tasks included abstract stimuli (e.g., geometric shapes, words).
-Results are likely to be different when using more real-world stimuli, or when changing the tasks in other ways to align them with people's lived experiences [@doebel_2020; @miller_cotto_2022; @young_2022].
-Participants completed the study online in their home environment. 
-To reproduce the effects, it is important to carefully instruct participants to limit environmental distractions and complete the study in full-screen mode.
-The results should generalize to more standardized settings (e.g., the lab), but may not generalize to public settings outside of the home environment.
-
-## Conclusion
-
-Adversity research has made important steps in recent years, revealing how exposure to adversity may impair certain EF abilities while enhancing others.
-Further progress in this field hinges on our capacity to accurately measure specific EF abilities.
-We show, consistent with previous research, that this is a more challenging task than typically assumed.
-As a result, adversity researchers likely overestimate the strength of associations between adversity and EF abilities when only analyzing raw performance.
-We present a way forward by building on a combination of cognitive modeling and structural equation modeling. 
-Embracing these techniques, and revising theories in light of their findings, will enhance our understanding of how exposure to adversity shapes EF.
-
-{{< pagebreak >}}
-
-# References
-
-::: {#refs}
-:::

### 2025-05-29 manuscript/masked_manuscript.qmd
fix repo

@@
 In this paper, we will not consider the starting point, because people cannot have biases towards correct or incorrect responses.
+
+How do the DDM parameters relate to executive functioning ability?
+Not all EF tasks require a person to accumulate evidence under time pressure, and some research suggests that EF measurement may be improved by removing speed requirements altogether [e.g., @draheim_2021b].
+However, many common EF tasks do depend on evidence accumulation.
+For instance, the Flanker Task requires accumulating evidence for the arrow direction (while not letting the flanking arrows contaminate this evidence), and attention-shifting tasks require accumulating evidence for the goal that is currently relevant.
+Given the popularity of these tasks in adversity research, associations between adversity exposure and EF have largely been studied using response times. 
+This raises the possibility that these associations are caused by ability-irrelevant processes rather than EF abilities. 
+If performance differences really reflect EF abilities, they should be reflected in drift rates, not boundary separations or non-decision times.
 
@@
 Participants were able to complete the cognitive tasks during two or more sessions to increase participation rates.
-People were eligible for the study if they were between 18 and 55 years old, and had agreed to linking their LISS data to government microdata (not relevant here, but for a different study).
+People were eligible for the study if they were between 16 and 55 years old, and had agreed to linking their LISS data to government microdata (not relevant here, but for a different study).
 We excluded participants who did not have data on any cognitive task.
@@
 As all item correlations were > .60 (see Table 2), we computed a uniformly weighted average.
+
+### Childhood adversity
+
+Participants reported their exposure to childhood deprivation (prior to age 18) in the new data collection using four items: (1) "My family had enough money to afford the kind of home we all needed"; (2) "My family had enough money to afford the kind of clothing we all needed"; (3) "My family had enough money to afford the kind of food that we all needed"; (4) "My family had enough money to afford the kind of medical care that we all needed", on a scale from 1 (completely not agree) to 7 (completely agree).
+This measure differed from those used to measure exposure to deprivation in adulthood. 
+In addition, participants reported their exposure to childhood threat (prior to age 18) in the new data collection using the Neighborhood Violence Scale ("Crime was common in the neighborhood where I grew up") , on a scale from 1 (completely not agree) to 7 (completely agree) [@frankenhuis_2018; @frankenhuis_2020].
+This measure was also used to measure threat exposure in adulthood.
+
+When used as covariates in the main analyses (see Confounds section), the measures of childhood deprivation and childhood threat were averaged together into a single measure of childhood adversity. 
+When used as main predictors in the exploratory analyses that we conducted, they were treated separately.
 
@@
   border(i = 1, border.bottom = fp_border_default(), part = "header") %>% 
-  border(i = 18, border.bottom = fp_border_default(), part = "body") %>% 
+  border(i = 20, border.bottom = fp_border_default(), part = "body") %>% 
   add_header_row(
     values = " ",
-    colwidths = 13
+    colwidths = 15
   ) |> 
@@
     values = " ",
-    colwidths = 13
+    colwidths = 15
   ) |> 
@@
 
+{{< pagebreak >}}
+
+```{r}
+table3 |> 
+  mutate(empty = "") |> 
+  select(task, condition, rt_m, rt_range, empty, ac_m, ac_range) |> 
+  flextable() |> 
+  set_header_labels(task = "Task", condition = "Condition", rt_m = "Mean (SD)", rt_range = "Range", empty = "", ac_m = "Mean (SD)", ac_range = "Range") |> 
+  add_header_row(values = c("", "", "Response times", "", "", "Accuracy", "")) |> 
+  merge_at(j = c(3:4), i = 1, part = "header") |> 
+  merge_at(j = c(6:7), i = 1, part = "header") |>
+
+  add_header_row(values = " ",colwidths = 7) |> 
+  flextable::compose(
+    i = 1, j = 1, 
+    as_paragraph(as_b("Table 3. "), "Descriptive statistics of response times and accuracy on cognitive tasks."),
+    part = "header"
+  ) |> 
+  align(i = 2, align = "center", part = "header") |> 
+  border_remove() %>% 
+  border(i = 2, border.top = fp_border_default(), part = "header") |> 
+  border(i = 2, j = c(3,6), border.bottom = fp_border_default(), part = "header") |> 
+  border(i = 3, border.bottom = fp_border_default(), part = "header") |> 
+  border(i = 11, border.bottom = fp_border_default(), part = "body") |>  
+  autofit()
+```
+
 ## State anxiety
@@
 We did not control for threat exposure in adulthood, as, following the Directed Acyclic Graph, we assume that threat exposure mediates the effect of recent material deprivation on cognitive processes.
-If true, including threat exposure in the model would make the association between material deprivation and cognitive processes an indirect effect, rather than a total effect (which was our estimand).
+If such mediation does occur, including threat exposure in the model would make the association between material deprivation and cognitive processes an indirect effect, rather than a total effect (which was our estimand).
 To explore the effect of our assumption that threat mediates the effect of material deprivation on cognitive processes, we also present a secondary model that includes threat in adulthood as a confound.
@@
 #| dpi: 600
-#| out-width: 12in
+#| out-width: 15in
 #| fig-cap: | 
@@
 Latent factors with non-significant residual variances are greyed out.
-*a* = manifest boundary separation, AS = Attention-shifting task, CS = Color-shape task, GL = Global-local task, Fl = Flanker task, Sm = Simon task, *t* = manifest non-decision time, *v* = manifest drift rate.
+*a* = manifest boundary separation, AS = Attention-shifting Task, CS = Color-shape Task, GL = Global-local Task, Fl = Flanker Task, Sm = Simon Task, *t* = manifest non-decision time, *v* = manifest drift rate.
 knitr::include_graphics("figures/fig3.png")
@@
 ```{r}
-#| tab.id: table3
+#| tab.id: table4
 #| results: markup
 
-table3 |> 
+table4 |> 
   add_header_row(
@@
 
-Our primary analyses examined the association of exposure to material deprivation and threat in adulthood with task-general processing speed, task-general response caution, and task-general speed of stimulus encoding and/or response execution.
+Our primary analyses examined the association of exposure to material deprivation and threat in adulthood with: task-general processing speed, task-general response caution, and task-general speed of stimulus encoding and/or response execution.
 As the structural equation model analyses reported above showed that EF-specific factors had non-significant variances, we retained them in the model but did not include them as dependent variables in subsequent analyses.
@@
 
-Exposure to material deprivation in adulthood was not associated with either general processing speed ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_v', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_v', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`), task-general response caution ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_a', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_a', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`), or task-general non-decision time ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_t', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_t', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`) (see Figure 4).
+Exposure to material deprivation in adulthood was negatively associated with task-general processing speed ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_v', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_v', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`) (see Figure 4).
+People who reported more exposure to deprivation in adulthood processed information more slowly across tasks.
+Exposure to deprivation in adulthood was not associated with task-general response caution ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_a', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_a', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`), or task-general non-decision time ($\beta$ = `r fit_dep1_reg_coef |> filter(lhs == 'gen_t', rhs == 'p_scar_m') |> pull(est.std) |> round(2)`, *p* `r fit_dep1_reg_coef |> filter(lhs == 'gen_t', rhs == 'p_scar_m') |> pull(pvalue_adj) |> format_p(pvalue = _)`) (see Figure 4).
 In addition, none of the effects fell within the region of practical equivalence (all *p*s ≥ `r fit_dep1_eqtests |> pull(eq_pvalue) |> min() |> format_p(pvalue = _, operator = FALSE)`).
-The effects in the secondary model (including expossure to threat in adulthood as a confounder) were comparable.
+The effects in the secondary model (including exposure to threat in adulthood as a confounder) were comparable in size, although the association between material deprivation in adulthood and task-general processing speed did not reach statistical significance.
 
@@
 In the first set of non-preregistered analyses, we analyzed how childhood exposure to threat and material deprivation are associated with task-general and ability-specific cognitive processes.
-The rational for these analyses is that we observed significant indirect associations between childhood adversity and general speed of processing in our preregistered analyses.
+The rationale for these analyses is that we observed significant indirect associations between childhood adversity and general speed of processing in our preregistered analyses.
 Here, we analyzed *direct* effects of childhood threat and material deprivation (i.e., without including exposure to adversity in adulthood).
@@
 This implies that, after accounting for general speed of processing, performance is further determined by unique features of individual tasks (i.e., method variance), which may be associated with adversity exposure.
-For this analysis, we were mostly interested in investigating the extent to which task-specific effects reflect EF abilities or task-specific processing, and therefore only focused on drift rates.
+For this analysis, we were mostly interested in investigating the extent to which task-specific effects reflect EF abilities or processes that are unique to individual tasks, and therefore only focused on drift rates.
 
@@
 Figure 6 shows latent correlations between task-specific drift rates after accounting for task-general processing speed.
-The drift rates of repeat and switch conditions within each attention shifting task correlated moderately to strongly, while all correlations between tasks were low.
-This suggests that task-specific drift rates of attention shifting tasks captured features of the tasks that were not shared between tasks.
+The drift rates of repeat and switch conditions within each attention-shifting task correlated moderately to strongly, while all correlations between tasks were low.
+This suggests that task-specific drift rates of attention-shifting tasks captured features of the tasks that were not shared between tasks.
 In contrast, all correlations between task-specific drift rates of inhibition tasks were low, even among drift rates of the same task.
 Thus, task-specific drift rate did provide consistent measures of specific EF abilities that are shared across tasks.
-
-<br>
 
@@
 Exposure to material deprivation in adulthood was not associated with task-specific drift rates of any of the tasks.
-Exposure to threat in adulthood was negatively associated all task-specific drift rates in both the repeat and switch condition of the Color-shape task (repeat: $\beta$ = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(est.std) |> round(2)`, SE = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(se) |> round(2)`, 95% CI = [`r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(ci.lower) |> round(2)`, `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(ci.upper) |> round(2)`], *p* `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`; switch: $\beta$ = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_sw_v', rhs == 'threat_comp') |> pull(est.std) |> round(2)`, SE = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_sw_v', rhs == 'threat_comp') |> pull(se) |> round(2)`, 95% CI = [`r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_sw_v', rhs == 'threat_comp') |> pull(ci.lower) |> round(2)`, `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_sw_v', rhs == 'threat_comp') |> pull(ci.upper) |> round(2)`], *p* `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_sw_v', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`).
+Exposure to threat in adulthood was negatively associated with the task-specific drift rate in the repeat condition of the Color-shape Task ($\beta$ = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(est.std) |> round(2)`, SE = `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(se) |> round(2)`, 95% CI = [`r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(ci.lower) |> round(2)`, `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(ci.upper) |> round(2)`], *p* `r expl_ts_adult_threat_reg_coef |> filter(lhs == 'sp_cs_rep_v', rhs == 'threat_comp') |> pull(pvalue_adj) |> format_p(pvalue = _)`).
 None of the other associations were significant.
@@
 Childhood exposure to material deprivation was negatively associated with all task-specific drift rates, with effect sizes ranging from $\beta$ = `r expl_ts_ch_dep_reg_coef |> filter(rhs == 'child_dep') |> pull(est.std) |> round(2) |> max()` for the congruent condition of the Simon Task to $\beta$ = `r expl_ts_ch_dep_reg_coef |> filter(rhs == 'child_dep') |> pull(est.std) |> round(2) |> min()` for the incongruent condition of the Global-local Task.
-Childhood exposure to threat was negatively associated with all but the task-specific drift rates of the congruent condition of the Flanker and the Simon task.
+Childhood exposure to threat was negatively associated with all but the task-specific drift rates of the congruent condition of the Flanker and the Simon Task.
 The effect sizes of significant effects ranged from $\beta$ = `r expl_ts_ch_thr_reg_coef |> filter(rhs == 'child_thr') |> pull(est.std) |> round(2) |> max()` for the incongruent condition of the Simon Task to $\beta$ = `r expl_ts_ch_thr_reg_coef |> filter(rhs == 'child_thr') |> pull(est.std) |> round(2) |> min()` for the switch condition of the Posner Task.
@@
 To summarize, we did not find evidence that exposure to threat or material deprivation (either in childhood or adulthood) was associated to specific inhibition or attention-shifting ability.
-Rather, we mostly found associations with task-general processes.
+Rather, we mostly found associations of these adversities with task-general processes.
 We did find associations between childhood adversity and task-specific drift rates.
@@
 Standard errors represent 95% confidence intervals.
-Statistical significance (tested against zero) is indicated with asterisks; * *p* < 0.05, ** *p* < 0.01, *** *p* < 0.001.
+The asterisks indicate statistical significance (tested against zero); * *p* < 0.05, ** *p* < 0.01, *** *p* < 0.001.
 fig7
@@
 
-We investigated associations between exposure to two types of adversity in adulthood---material deprivation and threat---with inhibition and attention shifting ability.
+We investigated associations between exposure to two types of adversity in adulthood---material deprivation and threat---with inhibition and attention-shifting ability.
 Participants completed two inhibition tasks, three attention shifting tasks, and one basic processing speed task.
@@
 This is consistent with predictions from deficit frameworks [@sheridan_2014; @tucker_drob_2013].
-After accounting for general processing speed, there was no remaining variance that could be attributed to either inhibition ability or attention shifting ability.
+After accounting for general processing speed, there was no remaining variance that could be attributed to either inhibition ability or attention-shifting ability.
 Additionally, the associations between material deprivation and threat in adulthood with both response caution and non-decision time were neither significantly different from zero, nor practically equivalent to zero.
@@
 This finding aligns with previous research applying DDM to EF tasks. 
-Three recent studies converged on the similar conclusion that performance across different EF tasks is best explained by a task-general drift rate factor [@frischkorn_2019; @hedge_2021; @loffler_2024; @weigard_2021].
+Three recent studies converged on the similar conclusion that performance across different EF tasks is best explained by a task-general drift rate factor [@frischkorn_2019; @hedge_2022; @loffler_2024; @weigard_2021].
 One study found that the task-general drift rate factor fully reflected basic processing speed, rather than common EF processes [@loffler_2024].
-Similarly, another study including several inhibition tasks found that shared variance was mostly explained by basic processing speed and strategy differences, not by inhibition ability [@hedge_2021].
+Similarly, another study including several inhibition tasks found that shared variance was mostly explained by basic processing speed and strategy differences, not by inhibition ability [@hedge_2022].
 It is therefore not surprising that we were unable to estimate ability-specific EF factors after accounting for general processing.
@@
 Finally, previous research applying DDM in the context of adversity found similar (but not identical) results as reported here.
-Youth with more exposure to threat (but not material deprivation) had slower general processing speed across three EF tasks and a basic processing speed task, while most specific abilities remained intact [@vermeent_2024a].
+Youth with more exposure to threat (but not material deprivation) had slower general processing speed across three EF tasks and a basic processing speed task, while most specific EF abilities remained intact [@vermeent_2024a].
 In contrast to the findings reported here, these youth also showed more response caution.
 It is important to highlight that both studies measured specific EF abilities with only a single task.
-This makes it unclear to which extent the task-specific effects they observed reflected EF abilities (see also our non-preregistered findings discussed below).
+This makes it unclear to which extent the task-specific effects they observed reflected EF abilities versus method variance (see also our non-preregistered findings discussed below).
 
@@
 This finding is striking considering the average age in our sample was 39 years.
-This is consistent with previous literature suggesting that adverse experiences early in life have a lasting effect on brain development [@shonkoff_2012; @nelson_2020a; @nelson_2020b].
+This is consistent with previous literature suggesting that adverse experiences early in life can have a lasting effect on brain development [@shonkoff_2012; @nelson_2020a; @nelson_2020b].
 For instance, both experiences of malnutrition and exposure to stress (e.g., as a result of threat) have been linked to structural changes in brain networks involved in cognitive functioning [@algarin_2017; @polavarapu_2017; @rebello_2018].
@@
 A second important non-preregistered finding is that after controlling for task-general processing, adversity exposure was negatively associated with several task-specific drift rates, mostly relating to childhood exposure to threat and deprivation.
-Importantly, the correlations among task-specific drift rates were all small---with the exception of correlations among switch and repeat conditions of the same attention shifting tasks.
-This suggests that these negative associations should not be interpreted as lower ability-specific processing, relating to inhibition and attention shifting ability.
+Importantly, the correlations among task-specific drift rates were all small---with the exception of correlations among switch and repeat conditions of the same attention-shifting tasks.
+This suggests that these negative associations should not be interpreted as lower ability-specific processing, relating to inhibition and attention-shifting ability.
 Instead, it appears to relate to information processing that is much more specific to individual tasks.
-This is consistent with literature showing that a substantial part of the variance in performance on cognitive task consists of method variance resulting from the nature of the task itself, rather than what the task is designed to measure [@barkley_2012; @rey_mermet_2024; @schubert_2016].
+This is consistent with literature showing that a substantial part of the variance in performance on cognitive task consists of method variance resulting from the nature of the task itself, rather than what the task is actually designed to measure [@barkley_2012; @rey_mermet_2024; @schubert_2016].
 
@@
 
-Our results have two implications for adversity research and interventions, both of which warrant caution against interpreting raw performance on individual EF tasks.
+Our results have two main implications for adversity research and interventions, both of which warrant caution against interpreting raw performance on individual EF tasks.
 First, adversity appears to be associated mostly with task-general processes, rather than specific abilities.
@@
 Moreover, our results show that accounting for task-general processing does not guarantees that task-specific associations with adversity reflect differences in EF.
-Instead, negative associations with adversity appear to be partly driven by specific features of individual tasks.
+Instead, negative associations with adversity appear to be partly driven by specific features of individual tasks (method variance).
 A general recommendation for adversity research is therefore that studies should include (1) multiple EF tasks to quantify task-general processing speed, and (2) ideally two or more tasks measuring the same EF ability to quantify ability-specific processing.
@@
 
-Our findings complicate research that aims to understand how EF abilities develop in adverse environments.
-We agree with others that results like ours do not necessarily mean that EF abilities do not exist [e.g., @loffler_2024], or that adversity is not associated with differences in EF.
-Rather, it points to issues with how we *measure* EF abilities.
-It is difficult to provide a straightforward solution for this problem, which is still a much-debated topic in cognitive psychology [@vonbastian_2020].
-Part of the solution may lie in adjusting the content of EF tasks, for instance, by using real-world content rather than abstract content [@young_2022].
-By increasing familiarity, relevance, and/or valence, such content might increase the engagement of EF abilities [@niebaum_2023].
-Another solution could be to use adaptive testing procedures, which appear to improve estimation of EF abilities [@draheim_2021b].
-However, more fundamental changes might be necessary as well, such as developing new paradigms that align more closely to people's everyday experiences, goals, and challenges [@doebel_2020; @miller_cotto_2022; @niebaum_2023].
+Second, the dominance of task-general processes is a challenge for adaptation frameworks, which predict that specific types of adversity may enhance or leave intact specific cognitive abilities.
+The lack of task-specific adaptations may have resulted from our use of standardized tasks featuring abstract content.
+Previous research suggests that cognitive adaptations may only be visible when using real-world content rather than abstract content [@young_2022], possibly by increasing task engagement [@niebaum_2023].
+On abstract tasks, such adapted abilities may only be visible under current conditions of stress [@mittal_2015; @young_2018].
+However, previous research also shows that content effects are complex and not always in the same direction.
+For instance, while real-world content equalized working memory performance of youth exposed to violence, it lowered attention-shifting performance for all youth, regardless of adversity exposure.
+Similarly, youth from low socioeconomic backgrounds perform worse on math items involving money, relative to other content [@duquennois_2022; @muskens_2024].
+It is currently unclear how performance on different tasks will be affected by real-world content, and which stimulus features drive these effects.
+While outside the scope of this article, DDM could play a key role in addressing these questions by showing how different types of contexts and content influence distinct cognitive processes.
 
@@
 
-Our study has three main limitations. 
-First, although we included confounding variables based on explicit, theory-guided causal assumptions, the cross-sectional and partly retrospective nature of our data prohibit us from establishing true causal relationships.
-Second, due to time constraints, we only included two inhibition tasks and three attention shifting tasks, limiting our ability to model shared variance between them.
+Our study has four main limitations. 
+First, our study included participants with low to moderate levels of adversity exposure, which was measured via self-reports. 
+Therefore, our study provides a less severe test of our hypotheses than a case-control study in which participants with severe exposure to adversity are compared to a control group. 
+Second, although we included confounding variables based on explicit, theory-guided causal assumptions, the cross-sectional and partly retrospective nature of our adversity data prohibit us from establishing true causal relationships.
+Third, due to time constraints, we only included two inhibition tasks and three attention shifting tasks, limiting our ability to model shared variance between them.
 Future studies could include more EF tasks, and ideally include several basic processing speed tasks.
-Third, participants completed the cognitive tasks online in their own home.
+Fourth, participants completed the cognitive tasks online in their own home.
 Despite our controlling for environmental factors that may have disrupted task performance, the home setting may have reduced the reliability of performance measures.
@@
 To the extent that task-general drift rates reflect general processing speed, it may partly result from lower white matter tract integrity [@fuhrmann_2020; @kievit_2016].
-Such associations with structural brain differences may account for the finding that task-general drift rate is relatively stable over time [@schubert_2016; @weigard_2021b].
+Such associations with structural brain differences may account for the finding that task-general drift rate is largely stable over time [@schubert_2016; @weigard_2021b].
 Yet, task-general drift rate may be associated with other factors such as effort, fatigue, or hunger [@weigard_2021b], which may play a bigger role for people from adverse environments [@brose_2012; @sliwinski_2006; @schwabe_2013].
-Thus, future studies may include (a combination of) brain measures or measures of mental states to better understand inter- and intrapersonal differences in task-general drift rates.
+Thus, future studies may include (a combination of) brain measures or measures of mental states to better understand inter- as well as intrapersonal differences in task-general drift rates.
 
@@
 One possible explanation might be that the associations reflect people's difficulty with processing abstract content.
-Previous research found that youth with more exposure to adversity performed lower on a working memory updating task (compared with youth with fewer adversity exposures) when task stimuli were abstract geometric shapes.
-However, their performance equalized when the task included real-world content [@young_2022].
+As discussed above, people with more exposure to adversity may be better at processing (stimulus) information that aligns with their daily experiences [@young_2022].
+Alternatively, lower performance on abstract tasks may reflect lower levels of engagement or motivation, rather than inability, to engage EF [@niebaum_2023].
 All tasks in our study included abstract content, and each study used different stimuli, which may account for the negative associations with adversity and the lack of correlations between task-specific drift rates.
@@
 Third, future research could design preregistered, well-powered analyses to tease apart the unique effects of childhood adversity and recent adversity on cognitive performance, as well as their combined effects.
-Our understanding of how cognitive outcomes later in life depend on the developmental timing of adversity exposure is still limited [@frankenhuis_2020].
+Our understanding of how cognitive abilities later in life depend on the developmental timing of adversity exposure is still limited [@frankenhuis_2020].
 In this study, we assumed that recent adversity exposure mediates the effect of childhood adversity exposure on cognitive processes, as people who experienced adversity early in life are more likely to also experience adversity later in life (e.g., due to systemic constraints; @hazel_2008).
@@
 For instance, some research suggests that childhood adversity enhances specific EF abilities, but only under situations of acute stress [@mittal_2015; @young_2018].
-Thus, future research should tease apart effects of adversity during childhood and adulthood and study how their effects are moderated by acute stress.
+Thus, future research should tease apart effects of adversity during childhood and adulthood and examine how their effects are moderated by acute stress.
 
@@
 We used standardized inhibition and attention-shifting tasks that are often used in cognitive psychology. 
-The findings should generalize to other tasks used to measure these abilities, as long as they rely on speeded responses.
+The findings should generalize to other tasks measuring these abilities, as long as they rely on speeded responses.
+It will require different modeling approaches to understand how adversity is associated with EF abilities that do not require evidence accumulation, or with EF decision processes that unfold over longer timescales (e.g., minutes).
 All tasks included abstract stimuli (e.g., geometric shapes, words).
-Results are likely to be different when using more real-world stimuli, or when changing the tasks in other ways to align them with people's lived experiences [@doebel_2020; @miller_cotto_2022; @young_2022].
+Results are likely to be different when using more real-world stimuli, or when changing the tasks in other ways to better align with people's daily experiences [@doebel_2020; @miller_cotto_2022; @young_2022].
 Participants completed the study online in their home environment. 
@@
 We show, consistent with previous research, that this is a more challenging task than typically assumed.
-As a result, adversity researchers likely overestimate the strength of associations between adversity and EF abilities when only analyzing raw performance.
-We present a way forward by building on a combination of cognitive modeling and structural equation modeling. 
-Embracing these techniques, and revising theories in light of their findings, will enhance our understanding of how exposure to adversity shapes EF.
+As a result, adversity researchers likely overestimate the strength of associations between adversity and EF abilities when relying on analyses of raw performance.
+We present a constructive path forward through a combination of cognitive modeling and structural equation modeling. 
+Embracing these techniques, and revising theories in light of their findings, will enhance our understanding of how exposure to adversity shapes both specific EF abilities and general cognitive processes.
 

### 2025-05-29 manuscript/titlepage.qmd
fix repo

@@
 
-Stefan Vermeent served as lead for conceptualization, data curation, methodology, formal analysis, and writing the original draft.
-Anna-Lena Schubert served in a supporting role for conceptualization, formal analysis, methodology, and writing-review and editing.
-Willem E.
-Frankenhuis served in a supporting role for conceptualization, methodology, supervision, funding acquisition, and writing-review and editing.
+This study was approved by the Ethics Review Board of the Faculty of Social & Behavioural Sciences of Utrecht University (FETC20-490) and the Ethics committee for research in the Sciences and Life Sciences of the University of Amsterdam (FNWI-41_2023).
 
-The preregistration, analysis scripts, and study materials are available on the article's Github repository ([https://stefanvermeent.github.io/liss_ef_2024/](https://stefanvermeent.github.io/liss_ef_2024/)).
+All scripts and materials needed to reproduce the findings are available on the article's Github repository ([https://stefanvermeent.github.io/liss_ef_2024/](https://stefanvermeent.github.io/liss_ef_2024/)).
 In this paper, we make use of data from the LISS panel (Longitudinal Internet studies for the Social Sciences) managed by the non-profit research institute Centerdata (Tilburg University, the Netherlands).
-All datasets are available to researchers in the LISS data archive, after signing a data use agreement ([https://www.dataarchive.lissdata.nl/study-units/view/1](https://www.dataarchive.lissdata.nl/study-units/view/1)).
+All datasets are available in the LISS data archive ([https://www.dataarchive.lissdata.nl/study-units/view/1](https://www.dataarchive.lissdata.nl/study-units/view/1)). 
+Researchers who want to access the data are required to sign a statement confirming that information about individual persons, households, etc., will not be released to others (go to [https://statements.centerdata.nl](https://statements.centerdata.nl) for more information).
+The data and ideas in this manuscript were not presented elsewhere.
 
-Correspondence concerning this article should be addressed to Stefan Vermeent, Department of Psychology, Utrecht University, Heidelberglaan 1, 3584 CS Utrecht, The Netherlands.
-Email: p.c.s.vermeent@gmail.com / p.c.s.vermeent@uu.nl.
+SV served as lead for conceptualization, formal analysis, funding acquisition, investigation, methodology, and writing - original draft.
+AS served in a supporting role for supervision, conceptualization, methodology, formal analysis, and writing - review & editing.
+WF served as lead for supervision, and in a supporting role for conceptualization, methodology, formal analysis, and writing - review & editing.
 
-We have no conflict of interest to disclose.
+The authors declare no conflicts of interest.
 
-<br>
+Correspondence concerning this article should be addressed to Stefan Vermeent, Evolutionary and Population Biology, Institute for Biodiversity and Ecosystem Dynamics, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands.
+Email: p.c.s.vermeent@gmail.com.
 
-**word count:** 9573
+**word count:** 10266

### 2025-05-29 supplement/supplement.qmd
fix repo

@@
 load("../analysis_objects/ddm_correlations.RData")
+load("../analysis_objects/ddm_histograms.RData")
 
@@
 
+## Priors and model settings
+
+We drew individual-level DDM parameters from normal distributions, separately for each condition (except for the Posner task, which only contained one condition).
+The normal distributions were truncated based on reasonable parameter ranges (between -10 and 10 for drift rates, between 0.1 and 5 for boundary separation, and between 0.0001 and 1 for non-decision time).
+The normal distributions were given uninformative priors.
+
 ## Model convergence
@@
   )
+```
+
+<br>
+
+## Distributions of DDM parameters
+
+Figures S12-S14 present histograms of individual estimates of the three DDM parameters.
+With a few exceptions (particularly in non-decision times), the DDM parameters were (approximately) normally distributed.
+
+```{r}
+#| label: FigureS12
+#| fig-width: 6.5
+#| fig-height: 8
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure S12.** Distribution of drift rates.
+hist_v
+```
+
+<br>
+
+```{r}
+#| label: FigureS13
+#| fig-width: 6.5
+#| fig-height: 8
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure S12.** Distribution of boundary separations.
+hist_a
+```
+
+<br>
+
+```{r}
+#| label: FigureS14
+#| fig-width: 6.5
+#| fig-height: 8
+#| dpi: 600
+#| out-width: 6in
+#| fig-cap: | 
+#|   **Figure S14.** Distribution of non-decision times.
+hist_t
 ```

### 2025-05-29 manuscript/scripts/staging.R

@@ -2,13 +2,20 @@ library(flextable)
 library(tidyverse)
 library(haven)
 library(patchwork)
+library(glue)
+library(parameters)
 
 source("scripts/0_functions/corr_plot.R")
 
 
 load('analysis_objects/results_confirmatory.RData')
 load('analysis_objects/results_exploratory.RData')
+load('analysis_objects/results_raw_rts.RData')
+load('analysis_objects/n_sessions.RData')
 load('data/exclusions.RData')
+load('data/ddm_clean.RData')
+load("data/clean_data_mean.RData")
+
 
 full_data <- read_csv('data/liss_data/full_data.csv')
 descr <- read_sav(file = "data/L_CognitiveAdversity_1.0p.sav") |> as_tibble()
@@ -66,7 +73,17 @@ fig4_df <-
       pvalue_adj < 0.01 & pvalue_adj >= 0.001 ~  "**",
       pvalue_adj < 0.001 ~ "***"
     ),
-    sig_pos = ifelse(est.std < 0, ci.lower - 0.03, ci.upper + 0.01)
+    sig_pos = ifelse(est.std < 0, ci.lower - 0.03, ci.upper + 0.01),
+    x = case_when(
+      ddm_parameter == "Task-general\nprocessing speed"  ~ 1.85,
+      ddm_parameter == "Task-general\nresponse caution"  ~ 2.85,
+      ddm_parameter == "Task-general\nnon-decision time" ~ 3.85
+    ),
+    xend = case_when(
+      ddm_parameter == "Task-general\nprocessing speed"  ~ 2.15,
+      ddm_parameter == "Task-general\nresponse caution"  ~ 3.15,
+      ddm_parameter == "Task-general\nnon-decision time" ~ 4.15
+    )
   )
 
 
@@ -80,7 +97,11 @@ fig4 <-
     inherit.aes = T
   ) +
   geom_hline(aes(yintercept = 0), size = 0.7) +
-  geom_errorbar(aes(ymin = ci.lower, ymax = ci.upper, color = ddm_parameter, linetype = .data[['linetype']]), width = 0.2, size = 1) +
+  geom_errorbar(aes(ymin = ci.lower, ymax = ci.upper, color = ddm_parameter, linetype = .data[['linetype']]), width = 0, size = 1) +
+
+  geom_segment(aes(x = x, xend = xend, y = ci.upper, color = ddm_parameter), size = 0.5, linetype = "solid") +
[…]

### 2025-05-29 scripts/1_study/01_parse.R

@@ -9,66 +9,81 @@ library(sjlabelled)
 # 2. Data -----------------------------------------------------------------
 raw_data <- read_delim("data/L_CognitiveAdversity_TASKDATA.csv", delim = ";")
 
-
-
-raw_data <- raw_data |>
-  mutate(
-    data_simon = case_when(
-      volgorde__1 == '1' ~ JSON1.4,
-      volgorde__2 == '1' ~ JSON2.4,
-      volgorde__3 == '1' ~ JSON3.4,
-      volgorde__4 == '1' ~ JSON4.4,
-      volgorde__5 == '1' ~ JSON5.4,
-      volgorde__6 == '1' ~ JSON6.4
-    ),
-    data_flanker = case_when(
-      volgorde__1 == '2' ~ JSON1.4,
-      volgorde__2 == '2' ~ JSON2.4,
-      volgorde__3 == '2' ~ JSON3.4,
-      volgorde__4 == '2' ~ JSON4.4,
-      volgorde__5 == '2' ~ JSON5.4,
-      volgorde__6 == '2' ~ JSON6.4
-    ),
-    data_globallocal2 = case_when(
-      volgorde__1 == '3' ~ JSON1.4,
-      volgorde__2 == '3' ~ JSON2.4,
-      volgorde__3 == '3' ~ JSON3.4,
-      volgorde__4 == '3' ~ JSON4.4,
-      volgorde__5 == '3' ~ JSON5.4,
-      volgorde__6 == '3' ~ JSON6.4
-    ),
-    data_colorshape = case_when(
-      volgorde__1 == '4' ~ JSON1.4,
-      volgorde__2 == '4' ~ JSON2.4,
-      volgorde__3 == '4' ~ JSON3.4,
-      volgorde__4 == '4' ~ JSON4.4,
-      volgorde__5 == '4' ~ JSON5.4,
-      volgorde__6 == '4' ~ JSON6.4
-    ),
-    data_animacysize = case_when(
-      volgorde__1 == '5' ~ JSON1.4,
-      volgorde__2 == '5' ~ JSON2.4,
-      volgorde__3 == '5' ~ JSON3.4,
-      volgorde__4 == '5' ~ JSON4.4,
-      volgorde__5 == '5' ~ JSON5.4,
-      volgorde__6 == '5' ~ JSON6.4
-    ),
-    data_posner = case_when(
-      volgorde__1 == '6' ~ JSON1.4,
[…]

### 2025-05-29 scripts/1_study/02_clean.R

@@ -345,7 +345,70 @@ all_included_ids_n <- all_included_ids |>
 exclusions$sample$ids <- all_included_ids
 exclusions$sample$final <- all_included_ids_n
 
-# 3. Create codebooks
+
+# 3. Calculate raw performance measures -----------------------------------
+
+flanker_clean_mean <- flanker_clean |>
+  group_by(nomem_encr, condition) |>
+  summarize(
+    flanker_rt = mean(rt, na.rm = T),
+    flanker_ac = sum(correct)/n() * 100
+    ) |>
+  mutate(condition = ifelse(condition == 1, "con", "inc")) |>
+  pivot_wider(names_from = "condition", values_from = c("flanker_rt", "flanker_ac"))
+
+simon_clean_mean <- simon_clean |>
+  group_by(nomem_encr, condition) |>
+  summarize(
+    simon_rt = mean(rt, na.rm = T),
+    simon_ac = sum(correct)/n() * 100
+  ) |>
+  mutate(condition = ifelse(condition == 1, "con", "inc")) |>
+  pivot_wider(names_from = "condition", values_from = c("simon_rt", "simon_ac"))
+
+colorshape_clean_mean <- colorshape_clean |>
+  group_by(nomem_encr, condition) |>
+  summarize(
+    colorshape_rt = mean(rt, na.rm = T),
+    colorshape_ac = sum(correct)/n() * 100
+  ) |>
+  mutate(condition = ifelse(condition == 1, "rep", "sw")) |>
+  pivot_wider(names_from = "condition", values_from = c("colorshape_rt", "colorshape_ac"))
+
+globallocal_clean_mean <- globallocal_clean |>
+  group_by(nomem_encr, condition) |>
+  summarize(
+    globallocal_rt = mean(rt, na.rm = T),
+    globallocal_ac = sum(correct)/n() * 100
+  ) |>
+  mutate(condition = ifelse(condition == 1, "rep", "sw")) |>
+  pivot_wider(names_from = "condition", values_from = c("globallocal_rt", "globallocal_ac"))
+
+animacysize_clean_mean <- animacysize_clean |>
+  group_by(nomem_encr, condition) |>
+  summarize(
+    animacysize_rt = mean(rt, na.rm = T),
+    animacysize_ac = sum(correct)/n() * 100
+  ) |>
+  mutate(condition = ifelse(condition == 1, "rep", "sw")) |>
+  pivot_wider(names_from = "condition", values_from = c("animacysize_rt", "animacysize_ac"))
+
[…]

### 2025-05-29 scripts/1_study/04_parse_ddm.R

@@ -477,7 +477,9 @@ stai_noise <- stai_noise |>
     stai_globallocal_diff = stai_globallocal - stai_mean,
     stai_colorshape_diff  = stai_colorshape - stai_mean,
     stai_animacysize_diff = stai_animacysize - stai_mean,
-    stai_posner_diff      = stai_posner - stai_mean
+    stai_posner_diff      = stai_posner - stai_mean,
+
+    one_session           = ifelse(DatumB == DatumE, 0, 1)
   ) |>
   rename(
     noise_simon       = Q1_1,

### 2025-05-29 scripts/1_study/05_ivs.R

@@ -65,15 +65,15 @@ core_income01 <- core_income[1:11] |>
       rename_with(
         .cols = -nomem_encr,
         ~str_replace_all(string = ., pattern = "ci\\d\\d[a-z]_|ci\\d\\d[a-z]", "q_")
-      ) #%>%
+      ) %>%
       #in wave 1 to 11, net income was asked for each employer separately
-    #  mutate(net_inc_y = across(c(q_012, q_021, q_030)) %>% rowSums(na.rm = T))
+      mutate(net_inc_y = across(c(q_012, q_021, q_030)) %>% rowSums(na.rm = T))
   }) |>
   bind_rows() |>
   separate(col = q_m, into = c("year", "month"), sep = 4) |>
-  left_join(background_parsed |> select(nomem_encr, year, month, nohouse_encr))# |>
-#  group_by(year, nohouse_encr) |>
-#  mutate(net_inc_y = sum(net_inc_y, na.rm = T))
+  left_join(background_parsed |> select(nomem_encr, year, month, nohouse_encr)) |>
+  group_by(year, nohouse_encr) |>
+  mutate(net_inc_y = sum(net_inc_y, na.rm = T))
 
 # in wave 12 to 16, net income was asked across all employers
 core_income02 <- core_income[12:16] |>
@@ -81,19 +81,19 @@ core_income02 <- core_income[12:16] |>
     x |>
       sjlabelled::remove_all_labels() |>
       as_tibble() |>
-      select(nomem_encr, matches("(_m|339|244|245|246|247|248|249|250|252)$")) |>
+      select(nomem_encr, matches("(_m|339|244|245|246|247|248|249|250|252|378)$")) |>
       rename_with(
         .cols = -nomem_encr,
         ~str_replace_all(string = ., pattern = "ci\\d\\d[a-z]_|ci\\d\\d[a-z]", "q_")
-      ) #%>%
-      #rename(net_inc_y = q_339)
+      ) %>%
+      rename(net_inc_y = q_339,
+             q_244 = q_378)
   }) |>
   bind_rows() |>
   separate(col = q_m, into = c("year", "month"), sep = 4)
 
 core_income_parsed <- bind_rows(core_income01, core_income02) |>
-  select(-c(q_012, q_021, q_030)) #|>
-#  mutate(net_inc_y = ifelse(net_inc_y <= 0, NA, net_inc_y))
+  select(-c(q_012, q_021, q_030))
 
 # Sanity checks
[…]

### 2025-05-29 scripts/1_study/07_exploratory.R

@@ -7,16 +7,21 @@ library(lavaan)
 load('analysis_objects/results_confirmatory.RData')
 
 load("data/ddm_clean.RData")
+load("data/clean_data_mean.RData")
+
 iv_data <- read_csv('data/liss_data/full_data.csv') |>
   filter(nomem_encr %in% unique(ddm_clean$nomem_encr)) |>
-  mutate(across(c(child_dep, child_thr), ~scale(.) |> as.numeric()))
+  mutate(across(c(child_dep, child_thr, age, p_scar_m, threat_comp, edu, child_adv), ~scale(.) |> as.numeric()))
 
 ddm_clean <- ddm_clean |>
   mutate(across(-nomem_encr, ~scale(.) |> as.numeric()))
 
-full_data <- full_join(iv_data, ddm_clean)
+full_data <- full_join(iv_data, ddm_clean) |>
+  left_join(clean_data_mean |> mutate(across(-nomem_encr, ~scale(.) |> as.numeric())))
 
 data_long <- load("data/clean_data_long.RData")
+
+
 
 equivalence_test_ml = function(sesoi, ml_effect, ml_se){
   p1 = 1 - pnorm( ml_effect, mean = -sesoi, sd = ml_se)
@@ -611,10 +616,11 @@ mod2_expl_ts_adult_threat <-
     as_sw_v_c       ~~ 0*as_sw_v_c
     as_rep_v_c      ~~ 0*as_rep_v_c
 
-    age             ~~ threat_comp + p_scar_m + sex + child_adv
-    threat_comp     ~~ p_scar_m + sex + child_adv
-    p_scar_m        ~~ sex + child_adv
+    age             ~~ p_scar_m + threat_comp + sex + child_adv
+    p_scar_m        ~~ threat_comp + sex + child_adv
+    threat_comp     ~~ sex + child_adv
     sex             ~~ child_adv
+
 
     #### INTERCEPTS ####
 
@@ -978,7 +984,7 @@ expl_ts_ch_thr_reg_coef <- standardizedsolution(fit_expl_ts_ch_thr) |>
   ) |>
   mutate(
     pvalue_adj = p.adjust(pvalue, method = "fdr"),
-    adversity  = "Threat", "Deprivation",
+    adversity  = "Threat",
     timing     = "Childhood"
   ) |>
   ungroup()
[…]

### 2025-05-29 supplement/stage_sup.R

@@ -208,6 +208,7 @@ flanker_rt_hist <- flanker_clean |>
   scale_x_continuous(breaks = seq(0, 5, 0.5)) +
   scale_color_uchicago() +
   scale_fill_uchicago() +
+  theme_classic() +
   labs(
     title = "Flanker task",
     x = "Response time"
@@ -223,6 +224,7 @@ simon_rt_hist <- simon_clean |>
   scale_x_continuous(breaks = seq(0, 5, 0.5)) +
   scale_color_uchicago() +
   scale_fill_uchicago() +
+  theme_classic() +
   labs(
     title = "Simon task",
     x = "Response time"
@@ -238,6 +240,7 @@ colorshape_rt_hist <- colorshape_clean |>
   scale_x_continuous(breaks = seq(0, 5, 1)) +
   scale_color_uchicago() +
   scale_fill_uchicago() +
+  theme_classic() +
   labs(
     title = "Color-shape task",
     x = "Response time"
@@ -253,6 +256,7 @@ animacysize_rt_hist <- animacysize_clean |>
   scale_x_continuous(breaks = seq(0, 5, 1)) +
   scale_color_uchicago() +
   scale_fill_uchicago() +
+  theme_classic() +
   labs(
     title = "Animacy-size task",
     x = "Response time"
@@ -268,6 +272,7 @@ globallocal_rt_hist <- globallocal_clean |>
   scale_x_continuous(breaks = seq(0, 5, 0.5)) +
   scale_color_uchicago() +
   scale_fill_uchicago() +
+  theme_classic() +
   labs(
     title = "Global-local task",
     x = "Response time"
@@ -283,6 +288,7 @@ posner_rt_hist <- posner_clean |>
   scale_fill_uchicago() +
   guides(color = 'none',
         fill = 'none') +
+  theme_classic() +
   labs(
     title = "Posner task",
     x = "Response time"
@@ -304,6 +310,7 @@ flanker_acc_hist <- flanker_clean |>
   geom_density(alpha = 0.5) +
   scale_color_uchicago() +
   scale_fill_uchicago() +
+  theme_classic() +
   labs(
     title = "Flanker task",
     x = "Accuracy"
@@ -318,6 +325,7 @@ simon_acc_hist <- simon_clean |>
   geom_density(alpha = 0.5) +
   scale_color_uchicago() +
   scale_fill_uchicago() +
+  theme_classic() +
   labs(
     title = "Simon task",
     x = "Accuracy"
[…]
